---
title: "CviMVP_LEA_offset"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LEA Genomic offsets

This script is a LEA pipeline for calculating genomic offset. I am using LEA3, which allows for multivariate environmental datasets. I first train the models using LFMM of a large seascape dataset, then calculate offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment. Offset is calculated from the environment of origin to common garden conditions.

## Setup

```{r}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(LEA) # for offsets
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(ggplot2) # for plotting
library(vegan) # for env dat processing
library(dplyr) # for data wrangling
library(stringr) # for data wrangling
library(s2) # mapping
library(rnaturalearth) # mapping
library(rnaturalearthdata) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ggrepel) # mapping
library(terra) # interpolation
library(stars) # interpolation
library(gstat) # interpolation

# data
# env
env_all <- read.csv("data/EnvDat/env_scaled_2025-12-05.csv")
env_rast <- readRDS("data/EnvDat/seascape/raster_interpolated_scaled_env.RDS")
env_sea <- read.env("data/EnvDat/seascape/seascape_scaled_expanded_reord_2025-12-05.env")
env_exp <- read.env("data/EnvDat/exp/exp_scaled_2025-12-08.env")
sea_red <- read.csv("data/EnvDat/seascape/sea_scaled_2025-12-05.csv")[,-1]

# sites
sea_site_full <- read.csv("data/EnvDat/seascape/sea_scaled_expanded_reord_2025-12-05.csv")[,-1]
exp_site_full <- read.csv("data/EnvDat/exp/exp_scaled_2025-12-08.csv")[,-1]

# gen
genoMat <- read.geno("data/GenDat/geno_seascape_2025-08-14.geno")
genoLfmm <- read.lfmm("data/GenDat/geno_seascape_2025-08-14.lfmm")
genoThinMat <- read.geno("data/GenDat/geno_seascape_LDthin_2025-08-14.geno")
genoThinLfmm <- read.lfmm("data/GenDat/geno_seascape_LDthin_2025-08-14.lfmm")

# inds
inds_genomat <- read.csv("data/GenDat/inds_genomat.csv")[,-1]
```


## Data Prep

### Environmental Data Prep

```{r}
# rename columns & rows to be more informative
colnames(env_sea) <- c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "MSX_Prevalence_scaled")

colnames(env_exp) <- c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "MSX_Prevalence_scaled", "lew_sal_quant10_scaled", "lew_sal_quant90_scaled", "lew_temp_quant10_scaled", "lew_temp_quant90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled", "lew_MSX_Prevalence_scaled", "yrk_sal_quant10_scaled", "yrk_sal_quant90_scaled", "yrk_temp_quant10_scaled", "yrk_temp_quant90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled", "yrk_MSX_Prevalence_scaled")

rownames(env_sea) <- sea_site_full$site_name
rownames(env_exp) <- exp_site_full$site_name

# create a dataset with no biotic variables - nb for "no biotic"
env_sea_nb <- env_sea[, c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")]
env_exp_nb <- env_exp[, c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "lew_sal_quant10_scaled", "lew_sal_quant90_scaled", "lew_temp_quant10_scaled", "lew_temp_quant90_scaled", "yrk_sal_quant10_scaled", "yrk_sal_quant90_scaled", "yrk_temp_quant10_scaled", "yrk_temp_quant90_scaled")]

# other env datasets
colnames(env_all) <- c("site_name", "Latitude_manual", "Longitude_manual", "sal_quant10", "sal_quant90", "temp_quant10", "temp_quant90", "Dermo_Prevalence", "Pea_crab", "MSX_Prevalence", "sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "MSX_Prevalence_scaled")

# no msx
env_sea <- env_sea[,c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")]

env_exp <- env_exp[,c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "lew_sal_quant10_scaled", "lew_sal_quant90_scaled", "lew_temp_quant10_scaled", "lew_temp_quant90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled", "yrk_sal_quant10_scaled", "yrk_sal_quant90_scaled", "yrk_temp_quant10_scaled", "yrk_temp_quant90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")]
```


### Genetic Data Prep

```{r}
# name the rows of genomic data
rownames(genoMat) <- rownames(genoLfmm) <- rownames(genoThinMat) <- rownames(genoThinLfmm) <- inds_genomat
```


## Seascape Training

Now that I have prepped and cleaned my data, I will use it to train LFMM models later used to predict genomic offset to some other (in this case experimental) climate. Importantly, I do not need to trim or impute my genomic data, or determine ancestry - Madeline has already done this. The data has been LD thinned, trimmed for MAF, and ancestry analysis has recovered K=2. 
The seascape environmental data was compiled by Madeline, as well as the qPCR results used to calculate pea crab prevalence. The data processing pipeline can be found in "CviMVP_oyster_EnvDatProcess.Rmd"

```{r}
# fit model
mod_lfmm <- lfmm2(input = genoThinLfmm,
                  env = env_sea, 
                  K = 2, # this was number from Madeline's snmf analysis
                  effect.sizes = T) 
mod_lfmm_nb <- lfmm2(input = genoThinLfmm,
                      env = env_sea_nb,
                      K = 2,
                      effect.sizes = T)
```


## Offset Prediction

```{r}
# check env data
colnames(env_all)
colnames(env_exp)

# also want predictions for entire seascape df, so put together this data
exp_sites_data <- do.call("rbind", replicate(nrow(sea_red), unique(env_exp[,c("lew_sal_quant10_scaled",  "lew_sal_quant90_scaled", "lew_temp_quant10_scaled", "lew_temp_quant90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled", "yrk_sal_quant10_scaled",  "yrk_sal_quant90_scaled", "yrk_temp_quant10_scaled", "yrk_temp_quant90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")]), simplify = FALSE))
rownames(exp_sites_data) <- NULL
env_sea_forgo <- cbind(sea_red, exp_sites_data)
env_sea_forgo2 <- as.matrix(env_sea_forgo[,-c(1:3)])
rownames(env_sea_forgo2) <- env_sea_forgo$site_name

# following https://github.com/bcm-uga/SSMPG2025/blob/main/Tutorials/Tutorial%20visualization/Spatial_Prediction_Genomic_Offset.Rmd

# first, I need to get environmental effect sizes from the trained models
B <- mod_lfmm@B
B_nb <- mod_lfmm_nb@B

# not using candidate loci for now
# calculate "genetic gap" for all the experimental sites
gg_lew = NULL
for (i in 1:nrow(env_exp[,c("sal_quant10_scaled",  "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")])){
gg_lew[i] = mean(((env_exp[,c("sal_quant10_scaled",  "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")] - env_exp[,c("lew_sal_quant10_scaled",  "lew_sal_quant90_scaled", "lew_temp_quant10_scaled", "lew_temp_quant90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled")])[i,]  %*% t(B))^2, na.rm = TRUE)
}

gg_yrk = NULL
for (i in 1:nrow(env_exp[,c("sal_quant10_scaled",  "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")])){
gg_yrk[i] = mean(((env_exp[,c("sal_quant10_scaled",  "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")] - env_exp[,c("yrk_sal_quant10_scaled",  "yrk_sal_quant90_scaled", "yrk_temp_quant10_scaled", "yrk_temp_quant90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")])[i,]  %*% t(B))^2, na.rm = TRUE)
}

gg_lew_nb = NULL
for (i in 1:nrow(env_exp_nb[,c("sal_quant10_scaled",  "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")])){
gg_lew_nb[i] = mean(((env_exp_nb[,c("sal_quant10_scaled",  "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")] - env_exp_nb[,c("lew_sal_quant10_scaled",  "lew_sal_quant90_scaled", "lew_temp_quant10_scaled", "lew_temp_quant90_scaled")])[i,]  %*% t(B_nb))^2, na.rm = TRUE)
}

gg_yrk_nb = NULL
for (i in 1:nrow(env_exp_nb[,c("sal_quant10_scaled",  "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")])){
gg_yrk_nb[i] = mean(((env_exp_nb[,c("sal_quant10_scaled",  "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")] - env_exp_nb[,c("yrk_sal_quant10_scaled",  "yrk_sal_quant90_scaled", "yrk_temp_quant10_scaled", "yrk_temp_quant90_scaled")])[i,]  %*% t(B_nb))^2, na.rm = TRUE)
}

# associate predictions with site metadata
gap_df <- env_all[env_all$site_name %in% rownames(env_exp), c("site_name", "Latitude_manual", "Longitude_manual")]
gap_df$lew_go <- gg_lew
gap_df$yrk_go <- gg_yrk
gap_df$lew_go_nb <- gg_lew_nb
gap_df$yrk_go_nb <- gg_yrk_nb

# output results
write.csv(gap_df, paste0("results/GO_results/offsets_LEA_expsites_noMSX_",Sys.Date(),".csv", sep = ""))

# change colnames for env_sea_forgo2 to match models
colnames(env_sea_forgo2)[colnames(env_sea_forgo2) == "salinity_quantile_10_scaled"] <- "sal_quant10_scaled"
colnames(env_sea_forgo2)[colnames(env_sea_forgo2) == "salinity_quantile_90_scaled"] <- "sal_quant90_scaled"
colnames(env_sea_forgo2)[colnames(env_sea_forgo2) == "temp_quantile_10_scaled"] <- "temp_quant10_scaled"
colnames(env_sea_forgo2)[colnames(env_sea_forgo2) == "temp_quantile_90_scaled"] <- "temp_quant90_scaled"

# also for the full seascape sample (just for visualization)
cols_sea <- c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", 
          "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")
cols_lew <- c("lew_sal_quant10_scaled", "lew_sal_quant90_scaled", "lew_temp_quant10_scaled", 
          "lew_temp_quant90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled")
cols_yrk <- c("yrk_sal_quant10_scaled", "yrk_sal_quant90_scaled", "yrk_temp_quant10_scaled", 
          "yrk_temp_quant90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")
cols_seanb <- c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", 
          "temp_quant90_scaled")
cols_lewnb <- c("lew_sal_quant10_scaled", "lew_sal_quant90_scaled", "lew_temp_quant10_scaled", 
          "lew_temp_quant90_scaled")
cols_yrknb <- c("yrk_sal_quant10_scaled", "yrk_sal_quant90_scaled", "yrk_temp_quant10_scaled", 
          "yrk_temp_quant90_scaled")

mat_sea <- as.matrix(env_sea_forgo2[, cols_sea])
mat_lew <- as.matrix(env_sea_forgo2[, cols_lew])
mat_yrk <- as.matrix(env_sea_forgo2[, cols_yrk])
mat_seanb <- as.matrix(env_sea_forgo2[, cols_seanb])
mat_lewnb <- as.matrix(env_sea_forgo2[, cols_lewnb])
mat_yrknb <- as.matrix(env_sea_forgo2[, cols_yrknb])

gg_full_lew = numeric(nrow(env_sea_forgo2))
for (i in 1:nrow(env_sea_forgo2)){
    diff = mat_sea[i,] - mat_lew[i,]
    result = diff %*% t(B)
    gg_full_lew[i] = mean(result^2, na.rm = TRUE)
}

gg_full_yrk = numeric(nrow(env_sea_forgo2))
for (i in 1:nrow(env_sea_forgo2)){
    diff = mat_sea[i,] - mat_yrk[i,]
    gap = diff %*% t(B)
    gg_full_yrk[i] = mean(gap^2, na.rm = TRUE)
}

gg_full_lew_nb = NULL
for (i in 1:nrow(env_sea_forgo2)){
    diff = mat_seanb[i,] - mat_lewnb[i,]
    gap = diff %*% t(B_nb)
    gg_full_yrk[i] = mean(gap^2, na.rm = TRUE)
}

gg_full_yrk_nb = NULL
for (i in 1:nrow(env_sea_forgo2)){
    diff = mat_seanb[i,] - mat_yrknb[i,]
    gap = diff %*% t(B_nb)
    gg_full_yrk[i] = mean(gap^2, na.rm = TRUE)
}

# associate predictions with site metadata
gap_full_df <- sea_red[, c("site_name", "Latitude_manual", "Longitude_manual")]
gap_full_df$lew_go <- gg_full_lew
gap_full_df$yrk_go <- gg_full_yrk
gap_full_df$lew_go_nb <- gg_full_lew_nb
gap_full_df$yrk_go_nb <- gg_full_yrk_nb

# save
write.csv(gap_full_df, paste0("results/GO_results/offsets_LEA_fullseascape_",Sys.Date(),".csv", sep = ""))
```

## Offset Visualization

I start by putting together the population information. 

```{r}
# add the color palette
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME", "LEWISETTA","YORKRIVER"), 
                     order = 1:10, 
                     latitude = c(28.0841, 29.24916667, 30.44116, 37.211, 37.9803, 37.247284, 43.0524, 43.986, 37.9803, 37.247284),
                     longitude = c(-97.2009, -90.9211111, -81.43908, -76.787, -76.4619, -76.499369, -70.91181, -69.55, -76.4619, -76.499369),
                     cols = c("#365C8DFF", "#277F8EFF", "#1FA187FF", "#4AC16DFF", "#440154FF", "#46337EFF", "#9FDA3AFF", "#FDE725FF","deeppink","deeppink"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME","LEWISETTA","YORKRIVER"))

offsets_plotting_lea <- merge(gap_df, popsdf, by = c("site_name"), all = T)

# save popdf object
write.csv(popsdf, "data/IndDat/popsdf.csv")
```

Then, I interpolate genomic offset across the seascape.

```{r}
sea_sf <- st_as_sf(gap_full_df, 
                   coords = c("Longitude_manual","Latitude_manual"),
                   crs = 4326) # for WGS84 projection

# set spherical geometry
sf::sf_use_s2(FALSE)

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin =-102, xmax = -65, ymin = 24, ymax = 47.05))
coast <- ne_coastline(scale = "medium", returnclass = "sf")
coast_crop <- sf::st_crop(coast, c(xmin =-102, xmax = -65, ymin = 24, ymax = 47.05))

# start by downloading and cropping oceans data
oceans110 <- ne_download(scale=110L, type = "ocean", category = "physical")
oceans_crop <- sf::st_crop(oceans110, c(xmin = -102, xmax = -65, ymin = 24, ymax = 47.05))

# check crs
st_crs(oceans_crop)
st_crs(coast_crop)
st_crs(world_crop) # looks good

# define buffer distance
dist <- 0.25 # 0.25º difference, approximately 15 miles

# define buffer region - note we need to specify the layer, as there are lots of different coastlines
buffered_region <- st_buffer(coast_crop, dist)[51,]
plot(buffered_region, col = "lightblue2")

# crop shapefile to region of interest for interpolation
# this is the region we will plot, but we need to interpolate the entire buffered region
int_region <- st_intersection(oceans_crop, buffered_region) 

# define grid across ocean area
grid <- st_make_grid(buffered_region, cellsize = 0.1, what = "polygons")
grid_int <- st_intersection(grid, buffered_region)
grid_int_crop <- sf::st_crop(grid_int, c(xmin = -102, xmax = -65, 
                                         ymin = 24, ymax = 47.05))

# interpolate
interpolated_lew <- idw(lew_go~1, sea_sf, grid_int)
saveRDS(interpolated_lew, "results/lg_results/interpolated_lea_lew.rds")

interpolated_yrk <- idw(yrk_go~1, sea_sf, grid_int)
saveRDS(interpolated_yrk, "results/lg_results/interpolated_lea_yrk.rds")

# empty raster
ext <- ext(st_bbox(interpolated_lew)) # get spatial extent
rast_empty <- rast(ext, resolution = 0.1, crs = "+proj=longlat +datum=WGS84")

# create a raster stack
# set up
sf_ls <- list(interpolated_lew, interpolated_yrk)
rast_ls <- list()

# loop through sf files to rasterize
for (i in seq_along(sf_ls)) {
  rast = st_rasterize(sf_ls[[i]] %>% dplyr::select(var1.pred, geometry))
  rast_ls[[i]] = rast
}

# combine
rast_all <- c(rast_ls)
names(rast_all) <- c("interpolated_lew_go","interpolated_yrk_go")

# save
saveRDS(rast_all, "results/GO_results/lea_interpolated_go.RDS")

# prep seascape sites for plotting
offsets_sea_plotting_lea <- na.omit(gap_full_df)

# try mapping again
map_lew_int_lea <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_lew, mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_sea_plotting_lea, aes(x = Longitude_manual, y = Latitude_manual, 
                                              fill = lew_go), 
             shape = 21, size = 2) +
  geom_point(data = offsets_plotting_lea[offsets_plotting_lea$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=Longitude_manual, y=Latitude_manual, 
                                       fill = lew_go), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.08)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.08)) +
  geom_label_repel(data = offsets_plotting_lea[offsets_plotting_lea$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=Longitude_manual, y=Latitude_manual, 
                                       label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "Genomic Offset to \nLewisetta (LEA)", x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")
(map_lew_int_lea)

lew_int_lea_inset <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_lew, mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.08)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.08)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_sea_plotting_lea, aes(x = Longitude_manual, y = Latitude_manual, 
                                              fill = lew_go), 
             shape = 21, size = 5) +
  geom_point(data = offsets_plotting_lea[offsets_plotting_lea$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, fill = lew_go), 
             shape = 21, size = 7)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 0.08)) + # double check these values
  geom_point(data = offsets_plotting_lea[offsets_plotting_lea$site_name == "LEWISETTA",], 
             aes(x = longitude-0.2, y = latitude), fill = "deeppink", shape = 24, size = 7)+
  geom_label_repel(data = offsets_plotting_lea[offsets_plotting_lea$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.4,
                  fill = "white",
                  size = 9)+
  guides(fill = "none")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
lew_int_lea_inset

map_yrk_int_lea <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_yrk, mapping = aes(col = var1.pred, fill = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_sea_plotting_lea, aes(x = Longitude_manual, y = Latitude_manual, 
                                              fill = lew_go), 
             shape = 21, size = 2) +
  geom_point(data = offsets_plotting_lea[offsets_plotting_lea$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=longitude, y=latitude, fill = yrk_go), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.08)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.08)) +
  geom_label_repel(data = offsets_plotting_lea[offsets_plotting_lea$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=longitude, y=latitude, label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = "right")+
  labs(title = "Genomic Offset to \nYork River (LEA)", x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")
(map_yrk_int_lea)

yrk_int_lea_inset <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_yrk, mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.08)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.08)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_sea_plotting_lea, aes(x = Longitude_manual, y = Latitude_manual, 
                                              fill = lew_go), 
             shape = 21, size = 5) +
  geom_point(data = offsets_plotting_lea[offsets_plotting_lea$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, fill = yrk_go), 
             shape = 21, size = 9)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 0.08)) + # double check these values
  geom_point(data = offsets_plotting_lea[offsets_plotting_lea$site_name == "YORKRIVER",], 
             aes(x = longitude-0.1, y = latitude+0.1), fill = "deeppink", shape = 24, size = 7)+
  geom_label_repel(data = offsets_plotting_lea[offsets_plotting_lea$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.5,
                  fill = "white",
                  size = 9)+
  guides(fill = "none")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
yrk_int_lea_inset
```
