---
title: "CviMVP_oyster_GenDatProcess"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Genetic Data Processing

This script is to process individual level genetic data for the seascape populations by identifying the number and identities of individuals in each dataset and preparing this information and the genomic data into formats needed for downstream analysis.

The seascape data comes from Madeline's seascape study, Eppley, M. G., Bajaj, K., Rumberger, C., Leung, E., Mongillo, N., Small, J., & Lotterhos, K. (2026). Range-wide genetic population structure and environmental adaptation in the eastern oyster (Crassostrea virginica) provides insight for aquaculture. In bioRxiv (p. 2026.03.30.715280). https://doi.org/10.64898/2026.03.30.715280

## Setup

```{r}
# set wd
setwd("~/Documents/Github/MVP_oyster_GO")

# packages
library(dplyr)
library(ggplot2)
library(tidyverse)
library(LEA)
library(stringr)
library(adegenet)
library(hierfstat)
library(SNPfiltR)
library(reshape2)

# data
geno_mat_ld <- readRDS("data/GenDat/20250606_THINNEDSNPs_seascape_labeled.rds")
sea_inds <- readRDS("data/IndDat/20250604_seascape_indsmatrix.rds")
env_sea_sites <- read.csv("data/EnvDat/seascape/seascape_sitenames.csv")[,-1]
```


## Pre-processing

Process data to check for missing data, non-variable genomic sites.

```{r}
# check data
dim(geno_mat_ld) # 746 inds, 115185 snps (~40k less than unthinned)

# any missing data?
length(which(geno_mat_ld == 9)) # nope!

# non-variable genomic sites
nonvarsite_ld <- apply(geno_mat_ld, 2, function(x) all(x == x[1]))
which(nonvarsite_ld == T) # none
```


## Individual Data

```{r}
dim(sea_inds) # 746 inds (presumably in same order), 79 cols with identifying info incl pop

# let's get rid of the empty columns
sea_inds_rm <- sea_inds[, apply(sea_inds, 2, function(x) !any(is.na(x)))] # only 29 columns now

# rownames of genomic data match clean_ID in individual df, and appear to be in the same order

# i need to get rid of inds not in my dataset (the ones with no environmental data)
seaRows_keep <- which((sea_inds_rm$ID_SiteDate %in% env_sea_sites))
geno_mat_ld_keep <- geno_mat_ld[seaRows_keep,]

# refilter for MAF
freqs <- colMeans(geno_mat_ld_keep, na.rm = TRUE) / 2
maf <- pmin(freqs, 1 - freqs)
geno_mat_ld_keep_filtered <- geno_mat_ld_keep[, which(maf > 0.05)] # about 2K snps lost (~1.6%)

# also be sure to put this data in the correct order to match environmental df
# first let's figure out which population everyone is from
genomat_order <- data.frame(clean_ID = rownames(geno_mat_ld_keep_filtered), order_mat = 1:length(rownames(geno_mat_ld_keep_filtered)))
sea_inds_order <- sea_inds_rm[seaRows_keep, c("clean_ID","ID_SiteDate")]
sea_inds_order$order_inds <- 1:length(sea_inds_order$clean_ID)
which(!(sea_inds_order$clean_ID %in% genomat_order$clean_ID))
ordered_sea_gen <- merge(genomat_order, sea_inds_order, by = "clean_ID")
ordered_sea_gen
```


## Genomic Data Formatting

```{r}
# write these matrices as usable files
# for LEA need lfmm format data
geno2lfmm(paste0("data/GenDat/geno_seascape_LDthin_", Sys.Date(), ".geno"))

# for RDA and gf geno matrix should work
write.geno(geno_mat_ld_keep_filtered, paste0("data/GenDat/geno_seascape_LDthin_", Sys.Date(), ".geno"))

# also write out information about the individuals in each row (same inds in ld and not ld mats)
inds_genomat <- rownames(geno_mat_ld_keep_filtered)
write.csv(inds_genomat, "data/GenDat/inds_genomat.csv")
```


## Calculate Basic Genetic Summary Statistics

```{r}
# create genind object
geno_mat_genind <- df2genind(geno_mat_ld_keep_filtered, sep = "", ncode = 1, ploidy = 2, pop = as.factor(sea_inds_order$ID_SiteDate))

geno_mat_genind_exp <- geno_mat_genind[geno_mat_genind@pop %in% c("CB_TX_2022-6-9", "Sister_Lk_LA_2022-8-9", "KP_FGR_FL_2022-8-23", "DWS_CB_VA_2022-10-7", "SR_GB_NH_2022-8-12", "DC_DR_ME_2022-10-25"),]

# format for hierfstat
geno_mat_hier <- genind2hierfstat(geno_mat_genind)
geno_mat_hier[1:10, 1:10]
geno_mat_hier_exp <- genind2hierfstat(geno_mat_genind_exp)

saveRDS(geno_mat_hier, "data/GenDat/geno_mat_hierfstat.RDS")
saveRDS(geno_mat_hier_exp, "data/GenDat/geno_mat_exp_hierfstat.RDS")

# read back in fst & stats
sea_fst <- readRDS("results/gen/seascape_pairwise_2026-04-15.RDS")
exp_stats <- readRDS("results/gen/exp_pops_statistics_2026-04-23.RDS")

# stats
colnames(exp_stats$Ho) <- colnames(exp_stats$Hs) <- colnames(exp_stats$Fis) <- c("W1_TX", "W6-ME", "W4-VA", "W3-FL", "W2-LA", "W5-NH")
ho_pop <- apply(exp_stats$Ho, 2, mean, na.rm = T)
hs_pop <- apply(exp_stats$Hs, 2, mean, na.rm = T)
fis_pop <- apply(exp_stats$Fis, 2, mean, na.rm = T)
exp_pop_stats <- data.frame(site_name = names(fis_pop), ho = ho_pop, hs = hs_pop, fis = fis_pop)

# fst
# long format
sea_fst_long <- melt(sea_fst)

# order correctly
sites_order <- sea_inds %>% distinct(Site_Order, .keep_all = TRUE) %>% select(ID_SiteDate, ID_Site, EstuaryRegion_name, Site_Order)

# keep just sites in analysis
sites_keep <- sites_order[which((sites_order$ID_SiteDate %in% env_sea_sites)),]
sites_keep_ordered <- sites_keep %>% arrange(Site_Order) %>%
  slice(1, 2, n(), 3:(n() - 1))
sites_keep_ordered$ID_SiteDate1 <- sites_keep_ordered$ID_SiteDate2 <- sites_keep_ordered$ID_SiteDate
sites_keep_ordered$Order1 <- sites_keep_ordered$Order2 <- 1:nrow(sites_keep_ordered)

colnames(sea_fst_long) <- c("ID_SiteDate1","ID_SiteDate2","fst")

# just the experimental source pops
sites_exp <- sites_order[which((sites_order$ID_SiteDate %in% c("CB_TX_2022-6-9", "Sister_Lk_LA_2022-8-9", "KP_FGR_FL_2022-8-23", "DWS_CB_VA_2022-10-7", "SR_GB_NH_2022-8-12", "DC_DR_ME_2022-10-25"))),]
sites_exp_ord <- sites_exp %>% arrange(Site_Order)
sites_exp_ord$Order1 <- sites_exp_ord$Order2 <- 1:nrow(sites_exp_ord)

sites_exp_df <- data.frame(ID_SiteDate = c("CB_TX_2022-6-9", "Sister_Lk_LA_2022-8-9", "KP_FGR_FL_2022-8-23", "DWS_CB_VA_2022-10-7", "SR_GB_NH_2022-8-12", "DC_DR_ME_2022-10-25"), site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "W5_NH", "W6_ME"))

sites_exp_ord_df <- merge(sites_exp_df, sites_exp_ord, by = c("ID_SiteDate"))
sites_exp_ord_df$site_name1 <- sites_exp_ord_df$site_name2 <- sites_exp_ord_df$site_name
sites_exp_ord_df$ID_SiteDate1 <- sites_exp_ord_df$ID_SiteDate2 <- sites_exp_ord_df$ID_SiteDate
sites_exp_ord_df <- sites_exp_ord_df %>% arrange(Order1, Order2)
sites_exp_ord_df$site_name1_rename <- gsub("_", "-", sites_exp_ord_df$site_name1)
sites_exp_ord_df$site_name2_rename <- gsub("_", "-", sites_exp_ord_df$site_name2)

# add to fst data
# seascape
order_lookup1 <- sites_keep_ordered %>% select(ID_SiteDate1, Order1)
order_lookup2 <- sites_keep_ordered %>% select(ID_SiteDate2, Order2)

sea_fst_ordered <- sea_fst_long %>%
  left_join(order_lookup1, by = "ID_SiteDate1") %>%
  left_join(order_lookup2, by = "ID_SiteDate2")

# exp
order_lookup1_exp <- sites_exp_ord_df %>% select(ID_SiteDate1, site_name1, Order1)
order_lookup2_exp <- sites_exp_ord_df %>% select(ID_SiteDate2, site_name2, Order2)

sea_fst_long_exp <- sea_fst_long %>% 
  filter(ID_SiteDate1 %in% sites_exp_ord_df$ID_SiteDate & 
         ID_SiteDate2 %in% sites_exp_ord_df$ID_SiteDate)

sea_fst_ordered_exp <- sea_fst_long_exp %>%
  left_join(order_lookup1_exp, by = "ID_SiteDate1") %>%
  left_join(order_lookup2_exp, by = "ID_SiteDate2") %>% 
  arrange(Order1, Order2)
```

## plotting fst

```{r}
# plot fst heatmap
# seascape training
sea_fst_all <- sea_fst_ordered %>% 
  ggplot(aes(x = fct_reorder(ID_SiteDate1, Order1), y = fct_reorder(ID_SiteDate2, Order2), fill = fst)) + 
  geom_tile() + 
  coord_fixed() + 
  scale_fill_viridis_c(option = "mako", direction = -1, na.value = "white", limits = c(0,0.075)) +
#  geom_text(aes(label = round(fst, 3)), size = 2, color = ifelse(sea_fst_ordered$fst > 0.04, "white", "black")) +
  scale_x_discrete(labels = sites_keep_ordered$ID_Site) +
  scale_y_discrete(labels = sites_keep_ordered$ID_Site) +
  labs(title = "Fst of Seascape Training Sites", x = NULL, y = NULL) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  theme_classic()
sea_fst_all

# experimental pops
sea_fst_exp <- sea_fst_ordered_exp %>% 
  ggplot(aes(x = fct_reorder(site_name1, Order1), y = fct_reorder(site_name2, Order2), fill = fst)) + 
  geom_tile() + 
  coord_fixed() + 
  geom_text(aes(label = round(fst, 3)), size = 3, color = ifelse(sea_fst_ordered_exp$fst > 0.04, "white", "black")) +
  scale_fill_viridis_c(option = "mako", direction = -1, na.value = "white", limits = c(0, 0.075)) +
  scale_x_discrete(labels = sites_exp_ord_df$site_name1_rename) +
  scale_y_discrete(labels = sites_exp_ord_df$site_name2_rename) +
  labs(title = "Fst of Wild Experimental Populations", x = NULL, y = NULL, fill = "Fst") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
sea_fst_exp
```

