---
title: "CviMVP_oyster_RDA.Rmd"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RDA Genomic offsets

This script is a RDA-based pipeline for calculating genomic offset. I am using rdadapt, a package developed for genome scans for selection, adaptive index projection on the landscape, estimation of genomic offsets, identification of co-varying modules of adaptive loci, and/or identification of adaptive groups of individual or populations. I first calculate "adaptive index" across seascape sites, then calculate offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment. Offset is calculated from the environment of origin to common garden conditions.

A vignette detailing the rdadapt pipeline is found [here](https://landscape-genomics.github.io/rdadapt/articles/vignette_tutorial.html).

## Setup

```{r}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(rdadapt) # for offsets
library(LEA) # read.geno
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(vegan) # for env dat processing
library(terra) # for env dat processing
library(reshape2) # for env dat processing
library(raster) # for env dat processing
library(dplyr) # for data wrangling
library(plyr) # for data wrangling
library(stringr) # for data wrangling
library(ggplot2) # for data vis
library(viridis) # for data vis
library(rdadapt) # for rda
library(foreach) # for rdadapt functions
library(qvalue) # for rdadapt functions
library(robust) # for rdadapt functions

# data
# env
env_sea <- read.csv("data/EnvDat/seascape/sea_scaled_expanded_reord_2026-01-30.csv")[,-1]
env_exp <- read.csv("data/EnvDat/exp/exp_scaled_2026-01-30.csv")[,-1]
env_sea_forgo <- read.csv("data/EnvDat/seascape/seascape_environment_predict_noPea.csv")

# sites
env_sea_site <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv")[,-1]
env_sea_date <- read.csv("data/EnvDat/seascape/SeascapeSamples - date.csv")[,-1]
env_exp_site <- read.csv("data/EnvDat/exp/exp_sitenames.csv")[,-1]
exp_site_latlon <- read.csv("data/EnvDat/exp/exp_site_info.csv")[,-1]

# gen
genoThinMat <- read.geno("data/GenDat/geno_seascape_LDthin_2026-04-06.geno")

# individuals
seascapeInds <- readRDS("data/IndDat/20250604_seascape_indsmatrix.rds")
inds_genomat <- read.csv("data/GenDat/inds_genomat.csv")[,-1]
```


## Data Checks

I will start by double checking that I have my genomic and environmental data in the correct formats.

```{r}
# env data
##########
# check data
env_sea
env_exp
seascapeInds

# keep just relevant data
seascapeInds_red <- seascapeInds[,c("clean_ID","Label_ind","Key_ind","ID_SiteDate","SiteDate_Key","EstuaryRegion_name","Site_Order","Site_Color","Region","Region_color")]
levels(as.factor(seascapeInds$ID_SiteDate))

# pops list
pops_list <- levels(as.factor(seascapeInds$ID_SiteDate))

# get some lat lons
sea_latlon <- merge(env_sea_site[,c("ID_Site", "Latitude_manual", "Longitude_manual")], env_sea_date[,c("ID_Site", "ID_SiteDate")], by = "ID_Site")

# write out for later
#write.csv(sea_latlon, "data/IndDat/seascape_latlon.csv")

# add latlon to env data
env_sea_complete <- merge(sea_latlon, env_sea, by = c("ID_SiteDate"))
##########

# gen data
##########
# check gen dat
dim(genoThinMat)

# assign meaningful rownames - the inds_genomat vector has individual names in the order they appear in the genotype matrix (as determined in GenDatProcess script), which is also the order of the seascape env matrix
rownames(genoThinMat) <- inds_genomat
##########
```

## Outlier Detection

While I do not intend to use an outlier dataset to calculate adaptive index/offset, I will still go through the process of outlier detection with RDA. This process is also part of training the RDA model for offsets later. First, though, there are a few functions I need from rdadapt, but the package seems to be broken currently as it is still in development.

```{r, include = F}
## rdadapt isn't installing correctly, this is the function i need directly from the git repo

##' 
##' @export
##' 

## Class not exported by vegan, so initialized here
setClass("rda", slots = c("colsum", "tot.chi", "Ybar"
                          , "method", "call"
                          , "pCCA", "CCA", "CA", "inertia"
                          , "regularization", "terms", "terminfo"))

##########################################################################################
setGeneric("rdadapt", def = function(RDA, K, scores) { standardGeneric( "rdadapt") })

##'
##' @rdname rdadapt
##' @export
##'

setMethod('rdadapt', signature(RDA = "rda"), function(RDA, K)
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("v" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$v seems not to exist")
  }
  if (K > ncol(RDA$CCA$v)) {
    K = ncol(RDA$CCA$v)
    warning("\n Not enough RDA axis available, K is set to ncol(RDA$CCA$v)")
  }
  
  ## FUNCTION -----------------------------------------------------------------
  scores <- RDA$CCA$v[, 1:as.numeric(K)]
  return(rdadapt(scores = scores))
})

##'
##' @rdname rdadapt
##' @export
##'

setMethod('rdadapt', signature(scores = "data.frame", RDA = "missing", K = "missing"), function(scores)
{
  scores <- as.matrix(scores)
  return(rdadapt(scores = scores))
})

##'
##' @rdname rdadapt
##' @export
##'

setMethod('rdadapt', signature(scores = "matrix", RDA = "missing", K = "missing"), function(scores)
{
  K <- ncol(scores)
  resscale <- apply(scores, 2, scale)
  resmaha <- covRob(resscale, distance = TRUE, na.action = na.omit, estim = "pairwiseGK")$dist
  lambda <- median(resmaha) / qchisq(0.5, df = K)
  reschi2test <- pchisq(resmaha / lambda, K, lower.tail = FALSE)
  qval <- qvalue(reschi2test)
  q.values_rdadapt <- qval$qvalues
  return(data.frame(p.values = reschi2test, q.values = q.values_rdadapt))
})

library(qvalue)
library(robust)
##########################################################################################

###################################################################################################
##' @name adaptive_index
##' @author Thibaut Capblancq
##' 
##' @title Project adaptive variation across space onto current or future landscapes
##' 
##' @description The \code{adaptive_index} function allows the user to estimate the optimal 
##' adaptive genetic component for any combination of environmental conditions, and project it 
##' onto the landscape.
##' 
##' @param RDA a \code{RDA} model from which to extract loci and environmental variable scores
##' @param K an \code{integer} specifying the number of RDA axes to use for the projection
##' @param env a \code{RasterStack} object or a \code{data.frame} with the environmental 
##' conditions of the sites to predict
##' @param env_mask (\emph{optional, default} \code{NULL}) \cr a \code{Raster} object to 
##' limit the projection to a specific area
##' @param method  (\emph{default} \code{'loadings'}) \cr a \code{character} defining whether 
##' the function is to use weighted averages (scaling type 1, \code{loadings}) or linear 
##' combinations (scaling type 2, \code{predict}) of the projected environmental variables 
##' to predict site scores (i.e., adaptive index)
##' 
##' @return  
##' 
##' A \code{RasterStack} object or a \code{data.frame} containing the projections for the 
##' \code{K} first RDA axes
##' 
##' @details
##' 
##' The \emph{adaptive index} provides an estimate of adaptive genetic similarity or difference 
##' of all pixels on the landscape as a function of the values of the environmental predictors 
##' at that location. 
##' When projected on a map, it allows visualizing the different adaptive gradients across a 
##' species range.
##' 
##' 
##' @keywords RDA, adaptive index, genetic similarity, genetic difference, projection
## 
## @seealso 
## 
## @examples
##' 
##' 
##' @importFrom foreach foreach %do%
##' @importFrom terra crs rast as.data.frame mask
##' @include rda_class.R
##' 
##' @export
##' 
##'
###################################################################################################

setGeneric("adaptive_index", def = function(RDA, K, env, env_mask = NULL, method = "loadings") { standardGeneric("adaptive_index") })

##'
##' @rdname adaptive_index
##' @export
##'

setMethod('adaptive_index', signature(RDA = "rda", env = "SpatRaster"), function(RDA, K, env, env_mask = NULL, method = "loadings")
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("biplot" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$biplot seems not to exist")
  }
  if (K > ncol(RDA$CCA$biplot)) {
    K <- ncol(RDA$CCA$biplot)
    warning("\n Not enough RDA axis available, K is set to ncol(RDA$CCA$biplot)")
  }
  if (!inherits(env, c("SpatRaster", "RasterLayer", "RasterStack"))) {
    stop("\n env must be a 'SpatRaster', 'RasterLayer' or 'RasterStack' object")
  } else if (inherits(env, c("RasterLayer", "RasterStack"))) {
    env <- rast(env)
  }
  if (nlyr(env) != nrow(RDA$CCA$biplot) || any(! names(env) %in% rownames(RDA$CCA$biplot))) {
    stop("\n env must contain same variables as the ones used in RDA (see rownames(RDA$CCA$biplot))")
  }
  if (!is.null(env_mask)) {
    if (!inherits(env_mask, c("SpatRaster", "RasterLayer"))) {
      stop("\n env_mask must be a 'SpatRaster' or 'RasterLayer' object")
    } else if (inherits(env_mask, "RasterLayer")) {
      env_mask <- rast(env_mask)
    }
    if (nlyr(env_mask) > 1) {
      stop("\n env_mask must contain only one layer")
    }
  }
  if (!(method %in% c("loadings", "predict"))) {
    stop("\n method must be 'loadings' or 'predict'")
  }
  
  
  ## FUNCTION -----------------------------------------------------------------
  
  ## Get RDA information -----------------------------------------------------
  RDA_biplot <- RDA$CCA$biplot
  var_names <- rownames(RDA_biplot)
  
  ## Transform environmental raster for prediction
  if (!is.null(env_mask)) { ## Mask with env_mask
    env <- try(mask(env, env_mask))
    if (inherits(env, "try-error")) {
      stop("\n env and env_mask must match in terms of extent, origin, crs")
    }
  }
  env_crs <- crs(env)
  env_df <- as.data.frame(env[[var_names]], xy = TRUE)
  if (any(is.na(env_df))) {
    env_df <- na.exclude(env_df)
    warning("\n env contained NA values which have been removed")
  }
  env_xy <- env_df[, c("x", "y")]
  env_var <- env_df[, var_names]
  
  
  ## Make predictions ---------------------------------------------------------
  if (method == "predict") {
    pred <- predict(RDA, env_var, type = "lc")
  }
  Proj <- foreach(i = 1:K) %do%
    {
      if (method == "loadings") { ## Predict pixels genetic component based on RDA axes
        tmp_df <- data.frame(env_xy, z = as.vector(apply(env_var, 1, function(x) sum(x * RDA_biplot[, i]))))
      } else if (method == "predict") { ## Predict with RDA model and linear combinations
        tmp_df <- data.frame(env_xy, z = as.vector(pred[, i]))
      }
      ras <- rast(tmp_df, type = "xyz", crs = crs(env))
      names(ras) <- paste0("RDA", i)
      return(ras)
    }
  Proj <- rast(Proj)
  
  ## Return predictions for each RDA axis
  return(Proj)
})


##'
##' @rdname adaptive_index
##' @export
##'

setMethod('adaptive_index', signature(RDA = "rda", env = "data.frame", env_mask = "missing"), function(RDA, K, env, method = "loadings")
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("biplot" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$biplot seems not to exist")
  }
  if (K > ncol(RDA$CCA$biplot)) {
    K <- ncol(RDA$CCA$biplot)
    warning("\n Not enough RDA axis available, K is set to ncol(RDA$CCA$biplot)")
  }
  if (ncol(env) != nrow(RDA$CCA$biplot) || any(! colnames(env) %in% rownames(RDA$CCA$biplot))) {
    stop("\n env must contain same variables as the ones used in RDA (see rownames(RDA$CCA$biplot))")
  }
  if (!(method %in% c("loadings", "predict"))) {
    stop("\n method must be 'loadings' or 'predict'")
  }
  
  
  ## FUNCTION -----------------------------------------------------------------
  
  ## Get RDA information -----------------------------------------------------
  RDA_biplot <- RDA$CCA$biplot
  var_names <- rownames(RDA_biplot)
  env_var <- env[, var_names]
  
  ## Make predictions ---------------------------------------------------------
  if (method == "predict") {
    pred <- predict(RDA, env_var, type = "lc")
  }
  AI <- foreach(i = 1:K) %do%
    {
      if (method == "loadings") { ## Predict genetic component based on RDA axes
        tmp_vec <- as.vector(apply(env_var, 1, function(x) sum(x * RDA_biplot[, i])))
      } else if (method == "predict") { ## Predict with RDA model and linear combinations
        tmp_vec <- as.vector(pred[, i])
      }
      return(tmp_vec)
    }
  names(AI) <- paste0("RDA", 1:K)
  AI <- as.data.frame(do.call(cbind, AI))
  
  ## Return predictions for each RDA axis
  return(AI)
})

###################################################################################################
##' @name genomic_offset
##' @author Thibaut Capblancq
##' 
##' @title Predict temporal or geographic genomic offset
##' 
##' @description The \code{genomic_offset} function allows the user to predict genomic offset 
##' from a RDA model. The \code{genomic_offset} function can estimate both temporal or spatial 
##' genomic offset and accommodates raster data or discrete populations.
##' 
##' @param RDA a \code{RDA} model from which to extract loci and environmental variable scores
##' @param K an \code{integer} specifying the number of RDA axes to use for the projection
##' @param env_pres a \code{RasterStack} object or a \code{data.frame} with the environmental 
##' conditions in the present
##' @param env_fut a \code{RasterStack} object or a \code{data.frame} with the environmental 
##' conditions in the future
##' @param env_mask (\emph{optional, default} \code{NULL}) \cr a \code{Raster} object to limit 
##' the projection to a specific area
##' @param method  (\emph{default} \code{'loadings'}) \cr a \code{character} defining whether 
##' the function is to use weighted averages (scaling type 1, \code{loadings}) or linear 
##' combinations (scaling type 2, \code{predict}) of the projected environmental variables 
##' to predict site scores (i.e., adaptive index)
##' 
##' 
##' @return  
##' 
##' A \code{list} containing :
##' \itemize{
##'   \item \code{genomic_offset} : a \code{RasterStack} or a \code{data.frame} containing 
##'   the genomic offset predictions for the \code{K} first RDA axes, as well as the overall 
##'   genomic offset prediction
##'   \item \code{weights} : the weights associated with each RDA axis used for the predictions
##' }
##' 
##' 
##' @details
##' 
##' This RDA-based method to predict \emph{genomic offset} is relatively simple. 
##' RDA is first used to predict the \emph{optimal adaptive genetic composition} for each 
##' environmental pixel under consideration (see \code{\link{adaptive_index}} function), 
##' using both current and future environmental conditions. 
##' The \emph{euclidean distance} between these two predictions in the RDA space provides an 
##' estimate of the change in genetic composition that would be required to track climate change.
##' 
##' 
##' @keywords RDA, genomic offset, adaptive index, projection
## 
##' @seealso \code{\link{adaptive_index}}
## 
## @examples
##' 
##' 
##' @importFrom stats dist
##' @importFrom raster rasterToPoints rasterFromXYZ mask
##' @include rda_class.R
##'  
##' @export
##' 
##'
###################################################################################################

setGeneric("genomic_offset", def = function(RDA, K, env_pres, env_fut, env_mask = NULL, env_gar, method = "loadings") { standardGeneric("genomic_offset") })

##'
##' @rdname genomic_offset
##' @export
##'

setMethod('genomic_offset', signature(RDA = "rda", env_pres = "SpatRaster", env_gar = "missing"), function(RDA, K, env_pres, env_fut, env_mask = NULL, method = "loadings")
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("eig" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$eig seems not to exist")
  }
  if (K > length(RDA$CCA$eig)) {
    K <- length(RDA$CCA$eig)
    warning("\n Not enough RDA axis available, K is set to length(RDA$CCA$eig)")
  }
  
  ## FUNCTION -----------------------------------------------------------------
  
  ## Make predictions
  AI_pres <- adaptive_index(RDA = RDA
                            , K = K
                            , env = env_pres
                            , env_mask = env_mask
                            , method = method)
  AI_fut <- adaptive_index(RDA = RDA
                           , K = K
                           , env = env_fut
                           , env_mask = env_mask
                           , method = method)
  
  ## Single axis genetic offset -----------------------------------------------
  offset <- foreach(i = 1:K) %do%
    {
      ras <- abs(AI_pres[[i]] - AI_fut[[i]])
      names(ras) <- paste0("RDA", i)
      return(ras)
    }
  offset <- rast(offset)
  
  ## Weight current and future adaptive indices based on eigen values of associated axes
  weights <- RDA$CCA$eig / sum(RDA$CCA$eig)
  weights <- weights[1:K]
  AI_pres_w <- AI_pres * weights
  AI_fut_w <- AI_fut * weights
  
  ## Predict a global genetic offset
  offset_global <- offset[[1]]
  offset_global[!is.na(offset_global)] <- sapply(1:ncell(AI_pres_w), function(x) {
    tmp_mat <- rbind(AI_pres_w[x], AI_fut_w[x])
    return(dist(tmp_mat, method = "euclidean"))
  })
  names(offset_global) <- "Global_offset"
  
  
  return(list(Proj_pres = AI_pres,
              Proj_fut = AI_fut,
              Proj_offset = offset,
              Proj_offset_global = offset_global,
              weights = weights))
})


##'
##' @rdname genomic_offset
##' @export
##'

setMethod('genomic_offset', signature(RDA = "rda", env_pres = "data.frame", env_mask = "missing", env_gar = "missing"), function(RDA, K, env_pres, env_fut, method = "loadings")
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("eig" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$eig seems not to exist")
  }
  if (K > length(RDA$CCA$eig)) {
    K <- length(RDA$CCA$eig)
    warning("\n Not enough RDA axis available, K is set to length(RDA$CCA$eig)")
  }
  
  ## FUNCTION -----------------------------------------------------------------
  
  ## Make predictions
  AI_pres <- adaptive_index(RDA = RDA
                            , K = K
                            , env = env_pres
                            , method = method)
  AI_fut <- adaptive_index(RDA = RDA
                           , K = K
                           , env = env_fut
                           , method = method)
  
  ## Single axis genetic offset -----------------------------------------------
  offset <- foreach(i = 1:K) %do%
    {
      vec <- abs(AI_pres[,i] - AI_fut[,i])
      return(vec)
    }
  offset <- as.data.frame(do.call(cbind, offset))
  
  ## Weight offset based on eigen values of associated axes
  weights <- RDA$CCA$eig / sum(RDA$CCA$eig)
  weights <- weights[1:K]
  offset <- data.frame(do.call(cbind, lapply(1:ncol(offset), function(x) offset[,x]*weights[x])))
  colnames(offset) <- paste0("RDA", 1:K)
  
  ## Predict a global genetic offset
  offset$Global <- apply(offset, 1, function(x) sqrt(sum(x*x)))
  
  return(list(Proj_pres = AI_pres,
              Proj_fut = AI_fut,
              genomic_offset = offset,
              weights = weights))
})


##'
##' @rdname genomic_offset
##' @export
##'

setMethod('genomic_offset', signature(RDA = "rda", env_pres = "data.frame", env_mask = "missing", env_fut = "missing"), function(RDA, K, env_pres, env_gar, method = "loadings")
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("eig" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$eig seems not to exist")
  }
  if (K > length(RDA$CCA$eig)) {
    K <- length(RDA$CCA$eig)
    warning("\n Not enough RDA axis available, K is set to length(RDA$CCA$eig)")
  }
  
  ## FUNCTION -----------------------------------------------------------------
  
  ## Make predictions
  AI_pres <- adaptive_index(RDA = RDA
                            , K = K
                            , env = env_pres
                            , method = method)
  
  AI_gar <- adaptive_index(RDA = RDA
                           , K = K
                           , env = env_gar
                           , method = method)
  
  ## Single axis genetic offset -----------------------------------------------
  offset <- foreach(i = 1:K) %do%
    {
      vec <- abs(AI_pres[,i] - AI_gar[,i])
      return(vec)
    }
  offset <- as.data.frame(do.call(cbind, offset))
  
  ## Weight offset based on eigen values of associated axes
  weights <- RDA$CCA$eig / sum(RDA$CCA$eig)
  weights <- weights[1:K]
  offset <- data.frame(do.call(cbind, lapply(1:ncol(offset), function(x) offset[,x]*weights[x])))
  colnames(offset) <- paste0("RDA", 1:K)
  
  ## Predict a global genetic offset
  offset$Global <- apply(offset, 1, function(x) sqrt(sum(x*x)))
  
  return(list(genomic_offset = offset,
              weights = weights))
})
```

Now I can start the process of training models:

```{r}
# reduce size of env data set used for training
env_sea_red <- env_sea[,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "MSX_Prevalence_scaled")]
env_sea_red_nb <- env_sea_red %>% dplyr::select(-"Dermo_Prevalence_scaled", -"MSX_Prevalence_scaled")

# run rda
start <- Sys.time()
RDA_sea <- rda(genoThinMat ~ ., as.data.frame(env_sea_red))
plot(RDA_sea)
screeplot(RDA_sea, main = "Eigenvalues of constrained axes")
end <- Sys.time()
end-start

start <- Sys.time()
RDA_sea_nb <- rda(genoThinMat ~ ., as.data.frame(env_sea_red_nb))
plot(RDA_sea_nb)
screeplot(RDA_sea_nb, main = "Eigenvalues of constrained axes")
end <- Sys.time()
end-start

# run rdadapt with K = 2 for first two axes
rdadapt_sea <- rdadapt(RDA = RDA_sea, K = 2)
rdadapt_sea_nb <- rdadapt(RDA = RDA_sea_nb, K = 2)

# bonferroni correction
# P-values threshold after Bonferroni correction
thres_env <- 0.01 / length(rdadapt_sea$p.values)
thres_env_nb <- 0.01 / length(rdadapt_sea_nb$p.values)

# Identifying the loci that are below the p-value threshold
outliers <- (1:ncol(genoThinMat))[which(rdadapt_sea$p.values < thres_env)]
length(outliers) # 1389 outliers

outliers_nb <- (1:ncol(genoThinMat))[which(rdadapt_sea_nb$p.values < thres_env_nb)]
length(outliers_nb) # 1252 outliers

# Formatting table for ggplot
locus_scores <- as.data.frame(scores(RDA_sea, choices=c(1:2), display="species", scaling="none"))
TAB_loci <- data.frame(names = row.names(locus_scores), locus_scores)
TAB_loci$type <- "Neutral"
TAB_loci$type[outliers] <- "Outliers"
TAB_loci$type <- factor(TAB_loci$type, levels = c("Neutral", "Outliers"))
TAB_loci <- TAB_loci[order(TAB_loci$type),]
TAB_var <- as.data.frame(scores(RDA_sea, choices=c(1,2), display="bp")) 

locus_scores_nb <- as.data.frame(scores(RDA_sea_nb, choices=c(1:2), display="species", scaling="none"))
TAB_loci_nb <- data.frame(names = row.names(locus_scores_nb), locus_scores_nb)
TAB_loci_nb$type <- "Neutral"
TAB_loci_nb$type[outliers_nb] <- "Outliers"
TAB_loci_nb$type <- factor(TAB_loci_nb$type, levels = c("Neutral", "Outliers"))
TAB_loci_nb <- TAB_loci_nb[order(TAB_loci_nb$type),]
TAB_var_nb <- as.data.frame(scores(RDA_sea_nb, choices=c(1,2), display="bp")) 

# Biplot of RDA loci and variables scores
ggplot() +
  geom_hline(yintercept=0, linetype="dashed", color = gray(.80), linewidth=0.6) +
  geom_vline(xintercept=0, linetype="dashed", color = gray(.80), linewidth=0.6) +
  geom_point(data = TAB_loci, aes(x=RDA1*20, y=RDA2*20, colour = type), size = 1.4, alpha = 0.7) +
  scale_color_manual(values = c("gray90", "#EB8055FF")) +
  geom_segment(data = TAB_var, aes(xend=RDA1, yend=RDA2, x=0, y=0), colour="black", size=0.15, linetype=1, arrow=arrow(length = unit(0.02, "npc"))) +
  geom_text(data = TAB_var, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(TAB_var)), size = 2.5, family = "Times") +
  guides(color=guide_legend(title="Locus type")) +
  labs(x = "RDA 1 (73.7%)", y = "RDA 2 (14.2%)", subtitle = "RDA space") +
  theme_bw(base_size = 11, base_family = "Times") +
  theme(panel.grid = element_blank()
        , plot.background = element_blank()
        , panel.background = element_blank())

# plot on abiotic only
ggplot() +
  geom_hline(yintercept=0, linetype="dashed", color = gray(.80), linewidth=0.6) +
  geom_vline(xintercept=0, linetype="dashed", color = gray(.80), linewidth=0.6) +
  geom_point(data = TAB_loci_nb, aes(x=RDA1*20, y=RDA2*20, colour = type), size = 1.4, alpha = 0.7) +
  scale_color_manual(values = c("gray90", "#EB8055FF")) +
  geom_segment(data = TAB_var_nb, aes(xend=RDA1, yend=RDA2, x=0, y=0), colour="black", size=0.15, linetype=1, arrow=arrow(length = unit(0.02, "npc"))) +
  geom_text(data = TAB_var_nb, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(TAB_var_nb)), size = 2.5, family = "Times") +
  guides(color=guide_legend(title="Locus type")) +
  labs(x = "RDA 1 (76.1%)", y = "RDA 2 (15.8%)", subtitle = "RDA space") +
  theme_bw(base_size = 11, base_family = "Times") +
  theme(panel.grid = element_blank()
        , plot.background = element_blank()
        , panel.background = element_blank())
```


## Offset

Now, I calculate genomic offset for all of my populations!

```{r}
# common garden info
lew_env <- env_exp[!duplicated(env_exp$site_name) ,c("lew_salinity_quantile_10_scaled", "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_MSX_Prevalence_scaled")]
yrk_env <- env_exp[!duplicated(env_exp$site_name) ,c("yrk_salinity_quantile_10_scaled", "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_MSX_Prevalence_scaled")]
colnames(lew_env) <- colnames(yrk_env) <- c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "MSX_Prevalence_scaled")

env_exp_red <- env_exp[!duplicated(env_exp$site_name) ,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "MSX_Prevalence_scaled")]

# df for predicting to seascape sites
rownames(env_sea_forgo) <- env_sea_forgo$X
env_sea_forgo_pres <- env_sea_forgo[,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "MSX_Prevalence_scaled")]
env_sea_forgo_lew <- env_sea_forgo[,c("lew_sal_quant10_scaled", "lew_sal_quant90_scaled", "lew_temp_quant10_scaled", "lew_temp_quant90_scaled", "lew_Dermo_Prevalence_scaled", "lew_MSX_Prevalence_scaled")]
env_sea_forgo_yrk <- env_sea_forgo[,c("yrk_sal_quant10_scaled", "yrk_sal_quant90_scaled", "yrk_temp_quant10_scaled", "yrk_temp_quant90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_MSX_Prevalence_scaled")]
colnames(env_sea_forgo_lew) <- colnames(env_sea_forgo_yrk) <- colnames(env_sea_forgo_pres) <- colnames(env_sea_red)

# dataframes with only abiotic variables
lew_env_nb <- lew_env[,c("salinity_quantile_10_scaled","salinity_quantile_90_scaled","temp_quantile_10_scaled","temp_quantile_90_scaled")]
yrk_env_nb <- yrk_env[,c("salinity_quantile_10_scaled","salinity_quantile_90_scaled","temp_quantile_10_scaled","temp_quantile_90_scaled")]
env_exp_red_nb <- env_exp_red[,c("salinity_quantile_10_scaled","salinity_quantile_90_scaled","temp_quantile_10_scaled","temp_quantile_90_scaled")]
env_sea_forgo_pres_nb <- env_sea_forgo_pres[,c("salinity_quantile_10_scaled","salinity_quantile_90_scaled","temp_quantile_10_scaled","temp_quantile_90_scaled")]
env_sea_forgo_lew_nb <- env_sea_forgo_lew[,c("salinity_quantile_10_scaled","salinity_quantile_90_scaled","temp_quantile_10_scaled","temp_quantile_90_scaled")]
env_sea_forgo_yrk_nb <- env_sea_forgo_yrk[,c("salinity_quantile_10_scaled","salinity_quantile_90_scaled","temp_quantile_10_scaled","temp_quantile_90_scaled")]

# Estimate genomic offset
offset_lew <- rdadapt::genomic_offset(RDA = RDA_sea, 
                              K = 2,
                              env_pres = env_exp_red, 
                              env_fut = lew_env, 
                              method = "loadings")
offset_lew_sea <- rdadapt::genomic_offset(RDA = RDA_sea,
                                 K = 2,
                                 env_pres = env_sea_forgo_pres,
                                 env_fut = env_sea_forgo_lew,
                                 method = "loadings")

offset_yrk <- rdadapt::genomic_offset(RDA = RDA_sea, 
                              K = 2, 
                              env_pres = env_exp_red, 
                              env_fut = yrk_env, 
                              method = "loadings")
offset_yrk_sea <- rdadapt::genomic_offset(RDA = RDA_sea, 
                                 K = 2, 
                                 env_pres = env_sea_forgo_pres, 
                                 env_fut = env_sea_forgo_yrk, 
                                 method = "loadings")

# and without biotic environment
offset_lew_nb <- rdadapt::genomic_offset(RDA = RDA_sea_nb, 
                              K = 2,
                              env_pres = env_exp_red_nb, 
                              env_fut = lew_env_nb, 
                              method = "loadings")
offset_lew_sea_nb <- rdadapt::genomic_offset(RDA = RDA_sea_nb,
                                 K = 2,
                                 env_pres = env_sea_forgo_pres_nb,
                                 env_fut = env_sea_forgo_lew_nb,
                                 method = "loadings")

offset_yrk_nb <- rdadapt::genomic_offset(RDA = RDA_sea_nb, 
                              K = 2, 
                              env_pres = env_exp_red_nb, 
                              env_fut = yrk_env_nb, 
                              method = "loadings")
offset_yrk_sea_nb <- rdadapt::genomic_offset(RDA = RDA_sea_nb, 
                                 K = 2, 
                                 env_pres = env_sea_forgo_pres_nb, 
                                 env_fut = env_sea_forgo_yrk_nb, 
                                 method = "loadings")

# reassign the metadata
offsets_rda_df <- data.frame(site_name = env_exp$site_name, lat = env_exp$Latitude_manual, lon = env_exp$Longitude_manual, offset_lew = offset_lew$genomic_offset$Global, offset_yrk = offset_yrk$genomic_offset$Global, offset_lew_nb = offset_lew_nb$genomic_offset$Global, offset_yrk_nb = offset_yrk_nb$genomic_offset$Global)

write.csv(offsets_rda_df, paste0("results/GO_results/offsets_RDA_expSites_",Sys.Date(),".csv", sep = ""))

# now with the seascape
offsets_rda_sea_df <- data.frame(site_name = env_sea_forgo$X, offset_lew_sea = offset_lew_sea$genomic_offset$Global, offset_yrk_sea = offset_yrk_sea$genomic_offset$Global, offset_lew_sea_nb = offset_lew_sea_nb$genomic_offset$Global, offset_yrk_sea_nb = offset_yrk_sea_nb$genomic_offset$Global)
offsets_rda_sea_df
write.csv(offsets_rda_sea_df, paste0("results/GO_results/offsets_RDA_fullseascape_",Sys.Date(),".csv", sep = ""))
```

## Visualization

Now I will plot all genomic offsets. Seascape GOs will be interpolated raster, while the experimental GOs will be overlayed as points. I will also show seascape GOs as points, but smaller than our experimental sites.

```{r}
# add the color palette
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME", "LEWISETTA","YORKRIVER"), 
                     order = 1:10, 
                     latitude = c(28.0841, 29.24916667, 30.44116, 37.211, 37.9803, 37.247284, 43.0524, 43.986, 37.9803, 37.247284),
                     longitude = c(-97.2009, -90.9211111, -81.43908, -76.787, -76.4619, -76.499369, -70.91181, -69.55, -76.4619, -76.499369),
                     cols = c("#365C8DFF", "#277F8EFF", "#1FA187FF", "#4AC16DFF", "#440154FF", "#46337EFF", "#9FDA3AFF", "#FDE725FF","deeppink","deeppink"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME","LEWISETTA","YORKRIVER"))

offsets_plotting <- merge(offsets_rda_df, popsdf, by = c("site_name"), all = T)
```


Interpolated GOs

```{r}
library(sf)
library(terra)
library(stars)
library(gstat)

# add latlong
sea_latlon_mg <- sea_latlon
colnames(sea_latlon_mg) <- c("ID_Site","latitude","longitude","site_name")
offsets_rda_sea_latlon_df <- merge(sea_latlon_mg, offsets_rda_sea_df, by = "site_name", all = T)
offsets_rda_sea_latlon_df[offsets_rda_sea_latlon_df$site_name == "GalBay_TX_2022-6-2",]$latitude <- offsets_rda_sea_latlon_df[offsets_rda_sea_latlon_df$site_name == "Galbay_TX_Grid169_2022-6-2",]$latitude
offsets_rda_sea_latlon_df[offsets_rda_sea_latlon_df$site_name == "GalBay_TX_2022-6-2",]$longitude <- offsets_rda_sea_latlon_df[offsets_rda_sea_latlon_df$site_name == "Galbay_TX_Grid169_2022-6-2",]$longitude
offsets_rda_sea_latlon_df_narm <- na.omit(offsets_rda_sea_latlon_df)

sea_sf_rda <- st_as_sf(offsets_rda_sea_latlon_df_narm, 
                   coords = c("longitude","latitude"),
                   crs = 4326) # for WGS84 projection

# map offsets
library(s2) # mapping
library(rnaturalearth) # mapping
library(rnaturalearthdata) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ggrepel)

# map
# set spherical geometry
sf::sf_use_s2(FALSE)

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin =-102, xmax = -65, ymin = 24, ymax = 47.05))
coast <- ne_coastline(scale = "medium", returnclass = "sf")
coast_crop <- sf::st_crop(coast, c(xmin =-102, xmax = -65, ymin = 24, ymax = 47.05))
world_chesapeake <- sf::st_crop(world, c(xmin =-77.856476 , xmax = -72.762787, ymin = 36.019557, ymax = 39.939163))

# start by downloading and cropping oceans data
oceans110 <- ne_download(scale=110L, type = "ocean", category = "physical")
oceans_crop <- sf::st_crop(oceans110, c(xmin = -102, xmax = -65, ymin = 24, ymax = 47.05))

# check crs
st_crs(oceans_crop)
st_crs(coast_crop)
st_crs(world_crop) # looks good

# define buffer distance
dist <- 0.25 # 0.25º difference, approximately 15 miles

# define buffer region - note we need to specify the layer, as there are lots of different coastlines
buffered_region <- st_buffer(coast_crop, dist)[51,]
plot(buffered_region, col = "lightblue2")

# crop shapefile to region of interest for interpolation
# this is the region we will plot, but we need to interpolate the entire buffered region
int_region <- st_intersection(oceans_crop, buffered_region) 

# define grid across ocean area
grid <- st_make_grid(buffered_region, cellsize = 0.1, what = "polygons")
grid_int <- st_intersection(grid, buffered_region)
grid_int_crop <- sf::st_crop(grid_int, c(xmin = -102, xmax = -65, 
                                         ymin = 24, ymax = 47.05))

# interpolate
interpolated_lew_rda <- idw(offset_lew_sea~1, sea_sf_rda, grid_int)
#saveRDS(interpolated_lew_rda, "results/lg_results/interpolated_rda_lew.rds")

interpolated_yrk_rda <- idw(offset_yrk_sea~1, sea_sf_rda, grid_int)
#saveRDS(interpolated_yrk_rda, "results/lg_results/interpolated_rda_yrk.rds")

# empty raster
ext <- ext(st_bbox(interpolated_lew_rda)) # get spatial extent
rast_empty <- rast(ext, resolution = 0.1, crs = "+proj=longlat +datum=WGS84")

# create a raster stack
# set up
sf_ls <- list(interpolated_lew_rda, interpolated_yrk_rda)
rast_ls <- list()

# loop through sf files to rasterize
for (i in seq_along(sf_ls)) {
  rast = st_rasterize(sf_ls[[i]] %>% dplyr::select(var1.pred, geometry))
  rast_ls[[i]] = rast
}

# combine
rast_all <- c(rast_ls)
names(rast_all) <- c("interpolated_lew_go","interpolated_yrk_go")

# save
saveRDS(rast_all, "results/GO_results/rda_interpolated_go.RDS")

# prep seascape actual sites for plotting
offsets_sea_plotting <- na.omit(offsets_rda_sea_latlon_df)

# try mapping again
map_lew_int_rda <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_lew_rda, mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_sea_plotting, aes(x = longitude, y = latitude, 
                                              fill = offset_lew_sea), 
             shape = 21, size = 2) +
  geom_point(data = offsets_plotting[offsets_plotting$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=longitude, y=latitude, 
                                       fill = offset_lew), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,5)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,5)) +
  geom_label_repel(data = offsets_plotting[offsets_plotting$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=longitude, y=latitude, 
                                       label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "Genomic Offset to \nLewisetta (RDA)", x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")
(map_lew_int_rda)

lew_int_rda_inset <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_lew_rda, mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,5)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,5)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_sea_plotting, aes(x = longitude, y = latitude, 
                                              fill = offset_lew_sea), 
             shape = 21, size = 5) +
  geom_point(data = offsets_plotting[offsets_plotting$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, 
                                       fill = offset_lew), 
             shape = 21, size = 9)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 5)) +
  geom_point(data = offsets_plotting[offsets_plotting$site_name == "LEWISETTA",], 
             aes(x = longitude-0.2, y = latitude), fill = "deeppink", shape = 24, size = 7)+
  geom_label_repel(data = offsets_plotting[offsets_plotting$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.5,
                  fill = "white",
                  size = 9)+
  guides(fill = "none")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
lew_int_rda_inset

map_yrk_int_rda <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_yrk_rda, mapping = aes(col = var1.pred, fill = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_sea_plotting, aes(x = longitude, y = latitude, 
                                              fill = offset_yrk_sea), 
             shape = 21, size = 2) +
  geom_point(data = offsets_plotting[offsets_plotting$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=longitude, y=latitude, fill = offset_yrk), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,5)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,5)) +
  geom_label_repel(data = offsets_plotting[offsets_plotting$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=longitude, y=latitude, 
                                       label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = "right")+
  labs(title = "Genomic Offset to \nYork River (RDA)", x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")
(map_yrk_int_rda)

yrk_int_rda_inset <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_yrk_rda, mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,5)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,5)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_sea_plotting, aes(x = longitude, y = latitude, 
                                              fill = offset_yrk_sea), 
             shape = 21, size = 5) +
  geom_point(data = offsets_plotting[offsets_plotting$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude,
                                       fill = offset_yrk), 
             shape = 21, size = 9)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 5)) + # double check these values
  geom_point(data = offsets_plotting[offsets_plotting$site_name == "YORKRIVER",], 
             aes(x = longitude, y = latitude+0.2), fill = "deeppink", shape = 24, size = 7)+
  geom_label_repel(data = offsets_plotting[offsets_plotting$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.7,
                  fill = "white",
                  size = 9)+
  guides(fill = "none")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
yrk_int_rda_inset
```


