---
title: "CviMVP_oyster_GF"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GF Genomic Offsets

This script is a gradientForest pipeline for calculating genomic offset. I am using gradientForest, which build on randomForest to use predictions of genome-wide allele frequency variation along multivariate environmental gradients to quantify the disruption of genotype-environment associations under environmental change. I first train the models using gradientForest on a large seascape dataset, then predict offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment based in the Chesapeake Bay. Offset is calculated from the environment of origin to common garden conditions.

## Setup

```{r Setup}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(LEA)
library(dplyr) # for data cleaning
library(plyr) # for data cleaning
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(raster) # for environmental pixels
library(ggplot2) # for plotting
library(parallel) # for parallelizing processes
library(doParallel) # for parallelizing processes 
library(gradientForest) # for running gradient forest
library(sf) # plotting
library(rnaturalearth) # plotting
library(maps) # mapping
library(ggspatial) # mapping
library(ggrepel) # mapping
library(terra) # interpolation
library(stars) # interpolation
library(gstat) # interpolation

# data
# env
env_all <- as.data.frame(read.csv("data/EnvDat/env_scaled_2026-01-30.csv")[,-1])
env_sea <- as.data.frame(read.csv("data/EnvDat/seascape/sea_scaled_expanded_reord_2026-01-30.csv")[,-1])
env_exp <- read.csv("data/EnvDat/exp/exp_scaled_2026-01-30.csv")[,-1]

# sites
env_sea_site <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv")[,-1]
env_sea_date <- read.csv("data/EnvDat/seascape/SeascapeSamples - date.csv")[,-1]
env_exp_site <- read.csv("data/EnvDat/exp/exp_sitenames.csv")[,-1]
exp_site_latlon <- read.csv("data/EnvDat/exp/exp_site_info.csv")[,-1]

# gen
genoThinMat <- read.geno("data/GenDat/geno_seascape_LDthin_2026-04-06.geno")

# individuals
seascapeInds <- readRDS("data/IndDat/20250604_seascape_indsmatrix.rds")
inds_genomat <- read.csv("data/GenDat/inds_genomat.csv")[,-1]
```


## Data Prep

Inspect the environmental and genomic data to ensure it's ready for gradientForest. Note that environmental data was pre-processed in the script "CviMVP_oyster_EnvDatPRocess.Rmd" and genomic data was pre-processed in the script "CviMVP_oyster_GenDatProcess.Rmd".

### Individual Data Prep

```{r}
# order this data by genotype info, first by trimming inds to only those in df
seascapeInds_reord <- seascapeInds[seascapeInds$clean_ID %in% inds_genomat,]
seascapeInds_reord$clean_ID <- factor(seascapeInds_reord$clean_ID, levels = inds_genomat)
seascapeInds_reord2 <- seascapeInds_reord[order(seascapeInds_reord$clean_ID),]

# double check
(seascapeInds_reord2$clean_ID == inds_genomat) # looks good!

# get a full list of pops
pops_list_full <- seascapeInds_reord2$ID_SiteDate
```

### Genomic Data Prep

```{r Genomic Data Prep}
# check geno data
dim(genoThinMat)

# rotate df so rows are now columns (we want individuals as columns, alleles as rows)
genoThinMatT <- as.matrix(t(genoThinMat))

# any missing data?
sum((genoThinMatT == 9)) # 0 - good

# add pop data to genomic data
colnames(genoMatT) <- colnames(genoThinMatT) <- pops_list_full

# function to calculate allele frequency @ single locus
calc_freq <- function(x){
  a <- sum(x, na.rm=TRUE)
  b <- (2*length(na.omit(x)))
  a/b
}

# function extending this to calculate allele frequencies across whole pop
calcfreq <- function(a, pop){
  tapply(a, pop, calc_freq)
}

# allele freqs
freqs <- apply((genoMat), 2, calcfreq, pops_list_full)
str(freqs)

freqs_thin <- apply((genoThinMat), 2, calcfreq, pops_list_full)
str(freqs_thin)

# update colnames for gf
colnames(freqs_thin) <- make.names(as.character(1:ncol(genoThinMat)))

# do the same for the full geno matrix, just in case
colnames(genoThinMat) <- make.names(as.character(1:ncol(genoThinMat)))
```


### Environmental Data Prep

```{r Environmental Data Prep}
# check env data
dim(env_sea)
dim(env_exp)
dim(seascapeInds_reord2)

# check seascape env in the same order as the genotype matrix
env_sea$clean_ID == seascapeInds_reord2$clean_ID # in same order

# we also need one just per site
env_sea_pop <- env_sea[!duplicated(env_sea$ID_SiteDate),]

# now experimental data
# no need to order this by any geno data
# should get pop level info 
env_exp_pop <- env_exp[!duplicated(env_exp$site_name),]

# reduced variable set
env_sea_pop <- env_sea[!duplicated(env_sea$ID_SiteDate),]
env_sea_red <- env_sea[,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "MSX_Prevalence_scaled")]
env_sea_red_nb <- env_sea[,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled")]
env_sea_pop_red <- env_sea_pop[,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "MSX_Prevalence_scaled")]
env_sea_pop_red_nb <- env_sea_pop[,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled")]

env_exp_pop_red <- env_exp_pop[,c("site_name","salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled","MSX_Prevalence_scaled")]
env_exp_lew_red <- env_exp_pop[,c("site_name", "lew_salinity_quantile_10_scaled", "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_MSX_Prevalence_scaled")]
env_exp_yrk_red <- env_exp_pop[,c("site_name", "yrk_salinity_quantile_10_scaled", "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_MSX_Prevalence_scaled")]

# change exp colnames to match up
colnames(env_exp_lew_red) <- colnames(env_exp_yrk_red) <- colnames(env_exp_pop_red)
```


## Train Models

To train the models, I will be using the seascape genomic dataset that we processed in the previous chunk, as well as the current environmental data files produced using "CviMVP_oyster_EnvDatProcess.Rmd." This process is very computationally intensive, so has to be submitted as a job to the HPC via the scripts "trainGF.sh" & "trainGF.R"

```{r GO Model Training}
# read in models trained in trainGF.R / trainGF.sh
gf_geno <- readRDS("results/lg_results/gf_geno_2026-04-17.RDS")
gf_geno_nb <- readRDS("results/lg_results/gf_geno_abiotic_2026-04-21.RDS")

# plot bar graphs depicting importance of each spatial and climate variable
plot(gf_geno)
plot(gf_geno_nb)

sink("results/GO_results/gf_importance.txt")
importance(gf_geno)
importance(gf_geno_nb)
sink()
```

For the genotype based model without disease, temp_quant90 is the most important variable, then sal_quant10, then temp_quant10, and finally sal_quant90

For both training runs, it's important to note that I had a significant number of warning messages - specifically:

"Warning in randomForest.default(x = X, y = spec_vec, maxLevel = maxLevel, : The response has five or fewer unique values. Are you sure you want to do regression?"

I think this is coming from non-variable genomic sites, which I could trim in future iterations of this model. Brandon has encountered this issue before and seems to think that the warning is internal to extendedForests (pckg for implementing random forests), and refers to number of unique AFs for a given locus. 


### Predict Offset

Now that we have trained our gradient forest models, we need to use these to predict performance in the mesocosm treatments. This requires no new genomic data, just new environmental data that corresponds to each of the treatments and to the sites these populations originated from. Importantly, this new environmental data set must use the same column names as the data set used for training. We will ensure data compatibility below.

```{r Predict Offset Prep}
# use exp data to predict from/to
env_exp_pop_red
env_exp_lew_red
env_exp_yrk_red

# seascape for visualization
env_sea_pop_red

# pops list experimental groups
pops_list1 <- c("S1_LOLA", "S2_DEBY", "W1_TX", "W2_LA", "W3_FL", "W4_VA", "W5_NH", "W6_ME")
```


Now I can predict offset using these different methods (genotype and allele frequency based).

```{r Predict Offset}
# prediction to home site
pred_geno_home <- predict(gf_geno, env_exp_pop_red[,-1])
pred_geno_home_nb <- predict(gf_geno_nb, env_exp_pop_red[,2:5])

# prediction to common gardens
pred_geno_lew <- predict(gf_geno, env_exp_lew_red[,-1])
pred_geno_yrk <- predict(gf_geno, env_exp_yrk_red[,-1])
pred_geno_lew_nb <- predict(gf_geno_nb, env_exp_lew_red[,2:5])
pred_geno_yrk_nb <- predict(gf_geno_nb, env_exp_yrk_red[,2:5])

# We now compute the euclidean distance between current and new cumulative importance
# need num "individuals" (here populations) for num offsets to calc
nb_ind <- length(pops_list1)

# calculate offsets
euclidean <- function(a, b) sqrt(sum((a - b)^2)) # Euclidean distance for genetic offset

# lew
df_geno_lew <- data.frame() # df
for (i in 1:nb_ind){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home[i,], pred_geno_lew[i,]))
  df_geno_lew <- rbind(df_geno_lew, ec_dist)
}
geno_offset_lew <- cbind(as.data.frame(env_exp_pop[,c("site_name","Latitude_manual","Longitude_manual")]), df_geno_lew) # associate metadata with offset calculation
colnames(geno_offset_lew) <- c("site_name", "lat", "lon", "offset_gf_geno_lew")

df_geno_lew_nb <- data.frame() # df
for (i in 1:nb_ind){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home_nb[i,], pred_geno_lew_nb[i,]))
  df_geno_lew_nb <- rbind(df_geno_lew_nb, ec_dist)
}
geno_offset_lew_nb <- cbind(as.data.frame(env_exp_pop[,c("site_name","Latitude_manual","Longitude_manual")]), df_geno_lew_nb) # associate metadata with offset calculation
colnames(geno_offset_lew_nb) <- c("site_name", "lat", "lon", "offset_gf_geno_lew_nb")

# york
df_geno_yrk <- data.frame() # df
for (i in 1:nb_ind){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home[i,], pred_geno_yrk[i,]))
  df_geno_yrk <- rbind(df_geno_yrk, ec_dist)
}
geno_offset_yrk <- cbind(as.data.frame(env_exp_pop[,c("site_name","Latitude_manual","Longitude_manual")]), df_geno_yrk) # associate metadata with offset calculation
colnames(geno_offset_yrk) <- c("site_name", "lat", "lon", "offset_gf_geno_yrk")

df_geno_yrk_nb <- data.frame() # df
for (i in 1:nb_ind){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home_nb[i,], pred_geno_yrk_nb[i,]))
  df_geno_yrk_nb <- rbind(df_geno_yrk_nb, ec_dist)
}
geno_offset_yrk_nb <- cbind(as.data.frame(env_exp_pop[,c("site_name","Latitude_manual","Longitude_manual")]), df_geno_yrk_nb) # associate metadata with offset calculation
colnames(geno_offset_yrk_nb) <- c("site_name", "lat", "lon", "offset_gf_geno_yrk_nb")

# merge
geno_offsets_b <- merge(geno_offset_lew, geno_offset_yrk, by = c("site_name","lat","lon"))
geno_offsets_nb <- merge(geno_offset_lew_nb, geno_offset_yrk_nb, by = c("site_name","lat","lon"))
offsets_mvp <- merge(geno_offsets_b, geno_offsets_nb, by = c("site_name","lat","lon"))

# write these offsets to results
write.csv(offsets_mvp, paste0("results/GO_results/offsets_GF_expSites_", Sys.Date(), ".csv"))

# also seascape offsets for plotting
# expand seascape common garden env for prediction
env_sea_pop_for_pred <- rbind(unique(env_exp_lew_red[,2:7]), unique(env_exp_yrk_red[,2:7]))
env_sea_pop_for_pred$exp_site <- c("Lewisetta", "YorkRiver")
env_sea_pop_for_pred <- env_sea_pop_for_pred[rep(seq_len(nrow(env_sea_pop_for_pred)), nrow(env_sea_pop_red)),]

# geno
pred_geno_sea <- predict(gf_geno, env_sea_pop_red)
pred_geno_lew_sea <- predict(gf_geno, env_sea_pop_for_pred[env_sea_pop_for_pred$exp_site == "Lewisetta",1:(ncol(env_sea_pop_for_pred) - 1)])
pred_geno_yrk_sea <- predict(gf_geno, env_sea_pop_for_pred[env_sea_pop_for_pred$exp_site == "YorkRiver",-which(names(env_sea_pop_for_pred) == "exp_site")])

# We now compute the euclidean distance between current and new cumulative importance
nb_ind_sea <- nrow(seascapeInds_reord2)

# lew
df_geno_lew_sea <- data.frame() # df
for (i in 1:33){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_sea[i,], pred_geno_lew_sea[i,]))
  df_geno_lew_sea <- rbind(df_geno_lew_sea, ec_dist)
}
geno_offset_lew_sea <- cbind(as.data.frame(env_sea_pop[,c("site_name","Latitude_manual","Longitude_manual")]), df_geno_lew_sea) # associate metadata with offset calculation
colnames(geno_offset_lew_sea) <- c("site_name", "lat", "lon", "offset_gf_geno_lew")

# york
df_geno_yrk_sea <- data.frame()
for (i in 1:33){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_sea[i,], pred_geno_yrk_sea[i,]))
  df_geno_yrk_sea <- rbind(df_geno_yrk_sea, ec_dist)
}
geno_offset_yrk_sea <- cbind(env_sea_pop$site_name, env_sea_pop$Latitude_manual, env_sea_pop$Longitude_manual, df_geno_yrk_sea) # associate metadata with offset calculation, will need to ensure that this list is in the correct order
colnames(geno_offset_yrk_sea) <- c("site_name", "lat", "lon", "offset_gf_geno_yrk")

# df of seascape offsets
seascape_offsets <- merge(geno_offset_lew_sea, geno_offset_yrk_sea, by = c("site_name","lat","lon"))

# save this
write.csv(seascape_offsets, paste0("results/GO_results/offsets_GF_fullseascape_", Sys.Date(), ".csv"))
```

### Plotting

I've got my offsets, and now want to plot them on a map! 

```{r}
# prep data for plotting
offsets_mvp_plot <- offsets_mvp
env_exp_pop_red

popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME", "LEWISETTA","YORKRIVER"), 
                     order = 1:10, 
                     latitude = c(28.0841, 29.24916667, 30.44116, 37.211, 37.9803, 37.247284, 43.0524, 43.986, 37.9803, 37.247284),
                     longitude = c(-97.2009, -90.9211111, -81.43908, -76.787, -76.4619, -76.499369, -70.91181, -69.55, -76.4619, -76.499369),
                     cols = c("#3e4989" ,"#31688e", "#35b779", "#6ece58", "#440154", "#482878", "#fde725", "#b5de2b","slateblue1","slateblue1"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME","LEWISETTA","YORKRIVER"))

popsplot <- merge(popsdf, offsets_mvp_plot, by = "site_name")
popsplot1 <- merge(popsplot, env_exp_pop_red, by = "site_name")

# map
sea_sf <- st_as_sf(seascape_offsets[,c("offset_gf_geno_lew","offset_gf_geno_yrk","lat","lon")], 
                   coords = c("lon","lat"),
                   crs = 4326) # for WGS84 projection

# set spherical geometry
sf::sf_use_s2(FALSE)

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin =-102, xmax = -65, ymin = 24, ymax = 47.05))
coast <- ne_coastline(scale = "medium", returnclass = "sf")
coast_crop <- sf::st_crop(coast, c(xmin =-102, xmax = -65, ymin = 24, ymax = 47.05))

# start by downloading and cropping oceans data
oceans110 <- ne_download(scale=110L, type = "ocean", category = "physical")
oceans_crop <- sf::st_crop(oceans110, c(xmin = -102, xmax = -65, ymin = 24, ymax = 47.05))

# check crs
st_crs(oceans_crop)
st_crs(coast_crop)
st_crs(world_crop) # looks good

# define buffer distance
dist <- 0.25 # 0.25º difference, approximately 15 miles

# define buffer region - note we need to specify the layer, as there are lots of different coastlines
buffered_region <- st_buffer(coast_crop, dist)[51,]
plot(buffered_region, col = "lightblue2")

# crop shapefile to region of interest for interpolation
# this is the region we will plot, but we need to interpolate the entire buffered region
int_region <- st_intersection(oceans_crop, buffered_region) 

# define grid across ocean area
grid <- st_make_grid(buffered_region, cellsize = 0.1, what = "polygons")
grid_int <- st_intersection(grid, buffered_region)
grid_int_crop <- sf::st_crop(grid_int, c(xmin = -102, xmax = -65, 
                                         ymin = 24, ymax = 47.05))

# interpolate
interpolated_lew_geno <- idw(offset_gf_geno_lew~1, sea_sf, grid_int)
#saveRDS(interpolated_lew_geno, "results/lg_results/interpolated_gf_geno_lew.rds")

interpolated_yrk_geno <- idw(offset_gf_geno_yrk~1, sea_sf, grid_int)
#saveRDS(interpolated_yrk_geno, "results/lg_results/interpolated_gf_geno_yrk.rds")

# empty raster
ext <- terra::ext(st_bbox(interpolated_lew_geno)) # get spatial extent
rast_empty <- terra::rast(ext, resolution = 0.1, crs = "+proj=longlat +datum=WGS84")

# create a raster stack
# set up
sf_ls <- list(interpolated_lew_geno, interpolated_yrk_geno)
rast_ls <- list()

# loop through sf files to rasterize
for (i in seq_along(sf_ls)) {
  rast = st_rasterize(sf_ls[[i]] %>% dplyr::select(var1.pred, geometry))
  rast_ls[[i]] = rast
}

# combine
rast_all <- c(rast_ls)
names(rast_all) <- c("interpolated_lew_go_geno","interpolated_yrk_go_geno")

# save object
saveRDS(rast_all, "results/GO_results/gf_interpolated_go.RDS")

# prep seascape sites for plotting
offsets_sea_plotting_gf <- na.omit(seascape_offsets)
offsets_plotting_gf <- na.omit(popsplot1)

# mapping geno models
map_lew_int_gf_geno <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_lew_geno, mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = seascape_offsets, aes(x = lon, y = lat, 
                                              fill = offset_gf_geno_lew), 
             shape = 21, size = 2) +
  geom_point(data = offsets_plotting_gf[offsets_plotting_gf$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=longitude, y=latitude, 
                                       fill = offset_gf_geno_lew), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  geom_label_repel(data = offsets_plotting_gf[offsets_plotting_gf$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=longitude, y=latitude, 
                                       label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "Genomic Offset to \nLewisetta (GF)", x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")
(map_lew_int_gf_geno)

lew_int_gf_geno_inset <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_lew_geno, mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = seascape_offsets, aes(x = lon, y = lat, 
                                              fill = offset_gf_geno_lew), 
             shape = 21, size = 5) +
  geom_point(data = offsets_plotting_gf[offsets_plotting_gf$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, fill = offset_gf_geno_lew), 
             shape = 21, size = 7)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 0.04)) + # double check these values
  geom_point(data = exp_site_latlon[exp_site_latlon$site_name == "LEW",], 
             aes(x = lon-0.2, y = lat), fill = "deeppink", shape = 24, size = 7)+
  geom_label_repel(data = offsets_plotting_gf[offsets_plotting_gf$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.4,
                  fill = "white",
                  size = 9)+
  guides(fill = "none")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
lew_int_gf_geno_inset

map_yrk_int_gf_geno <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_yrk_geno, mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = seascape_offsets, aes(x = lon, y = lat, 
                                              fill = offset_gf_geno_yrk), 
             shape = 21, size = 2) +
  geom_point(data = offsets_plotting_gf[offsets_plotting_gf$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=longitude, y=latitude, 
                                       fill = offset_gf_geno_yrk), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  geom_label_repel(data = offsets_plotting_gf[offsets_plotting_gf$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=longitude, y=latitude, 
                                       label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "Genomic Offset to \nYork River (GF)", x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")
(map_yrk_int_gf_geno)

yrk_int_gf_geno_inset <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_yrk_geno, mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = seascape_offsets, aes(x = lon, y = lat, 
                                              fill = offset_gf_geno_yrk), 
             shape = 21, size = 5) +
  geom_point(data = offsets_plotting_gf[offsets_plotting_gf$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, fill = offset_gf_geno_yrk), 
             shape = 21, size = 7)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 0.04)) + # double check these values
  geom_point(data = exp_site_latlon[exp_site_latlon$site_name == "YRK",], 
             aes(x = lon-0.2, y = lat), fill = "deeppink", shape = 24, size = 7)+
  geom_label_repel(data = offsets_plotting_gf[offsets_plotting_gf$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.4,
                  fill = "white",
                  size = 9)+
  guides(fill = "none")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
yrk_int_gf_geno_inset
```