---
title: "CviMVP_oyster_compareGOexp"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "~/Documents/GitHub/MVP_oyster_GO")
```

# Compare Predicted GO to Experimental Results

This script is for analyzing how well various genomic offset methods predict experimentally obtained fitness. I compare LEA, GF, and RDA offsets to experimental mortality, length, and condition index at the conclusion of the experiment. 

## Setup

```{r}
# setwd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(tidyr)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(plotrix)

# data
# offset predictions
lea_go <- read.csv("results/GO_results/offsets_LEA_expsites_2026-04-06.csv")[,-1]
rda_go <- read.csv("results/GO_results/offsets_RDA_expsites_2026-04-06.csv")[,-1]
gf_go <- read.csv("results/GO_results/offsets_GF_expSites_2026-02-03.csv")[,-1]
env_dist <- read.csv("data/EnvDat/exp/distance_matrix_2026-01-30.csv")[28:35,c("X","Lewisetta","YorkRiver")]
env_dist_nb <- read.csv("data/EnvDat/exp/distance_matrix_abiotic_2026-01-30.csv")[28:35,c("X","Lewisetta","YorkRiver")]

# fitness metrics
mort_bags <- read.csv("results/exp_results/mortality_long_bags_2026-01-15.csv")[,-1]
mort <- read.csv("results/exp_results/mortality_long_popsite_2026-01-15.csv")[,-1]
len <- read.csv("results/exp_results/lengths_bags_event_2025-06-03.csv")[,-1]
```

## Data Prep

Prepare data from offsets & experiment to be directly compared.

```{r}
# inspect data
lea_go
gf_go
rda_go
mort
len
cond

# trim mort
mort_popsite <- mort[!(mort$event == 0), c("bag_site","pop","mean_surv","sd","event")]
colnames(mort_popsite) <- c("bag_site","pop","mean_surv","sd_surv","event")

# also get fitness rankings at each time
# mortality
ord_mort_popsite <- data.frame()
for (i in 1:4) {
  ordt_lew <- mort_popsite[mort_popsite$event == i & mort_popsite$bag_site == "Lewisetta",]
  ordt_lew <- ordt_lew %>% arrange(desc(mean_surv))
  ordt_lew$order_mort <- 1:length(ordt_lew$pop)
  ordt_yrk <- mort_popsite[mort_popsite$event == i & mort_popsite$bag_site == "YorkRiver",]
  ordt_yrk <- ordt_yrk %>% arrange(desc(mean_surv))
  ordt_yrk$order_mort <- 1:length(ordt_yrk$pop)
  ord_mort_popsite <- rbind(ord_mort_popsite, ordt_lew, ordt_yrk)
}
ord_mort_popsite

# do the same with length
# first length needs a pop label and site label
len_bags <- as.data.frame(cbind(len$bag, str_split_fixed(len$bag, "-", 3)[,2:3]))
colnames(len_bags) <- c("bag","pop","bagnum")
len_bags$bagnum <- as.numeric(len_bags$bagnum)
len_bags$site <- ifelse(len_bags$bagnum > 3, "YorkRiver", "Lewisetta")
len_all <- merge(len_bags[,c("bag","pop","site")], len, by = c("bag"))
colnames(len_all) <- c("bag","pop","site","event","length")

len_t1 <- len_all[len_all$event == 1,]
t1_lens <- aggregate(length ~ pop + site, data = len_t1, FUN = mean)
t1_sd <- aggregate(length ~ pop + site, data = len_t1, FUN = sd)
colnames(t1_sd) <- c("pop", "site", "len_sd")
t1_lens_popsite <- merge(t1_lens, t1_sd, by = c("pop", "site"))
t1_lens_popsite$event <- 1
len_t2 <- len_all[len_all$event == 2,]
t2_lens <- aggregate(length ~ pop + site, data = len_t2, FUN = mean)
t2_sd <- aggregate(length ~ pop + site, data = len_t2, FUN = sd)
colnames(t2_sd) <- c("pop", "site", "len_sd")
t2_lens_popsite <- merge(t2_lens, t2_sd, by = c("pop", "site"))
t2_lens_popsite$event <- 2
len_t3 <- len_all[len_all$event == 3,]
t3_lens <- aggregate(length ~ pop + site, data = len_t3, FUN = mean)
t3_sd <- aggregate(length ~ pop + site, data = len_t3, FUN = sd)
colnames(t3_sd) <- c("pop", "site", "len_sd")
t3_lens_popsite <- merge(t3_lens, t3_sd, by = c("pop", "site"))
t3_lens_popsite$event <- 3
len_t4 <- len_all[len_all$event == 4,]
t4_lens <- aggregate(length ~ pop + site, data = len_t4, FUN = mean)
t4_sd <- aggregate(length ~ pop + site, data = len_t4, FUN = sd)
colnames(t4_sd) <- c("pop", "site", "len_sd")
t4_lens_popsite <- merge(t4_lens, t4_sd, by = c("pop", "site"))
t4_lens_popsite$event <- 4

lens_popsite <- rbind(t1_lens_popsite, t2_lens_popsite, t3_lens_popsite, t4_lens_popsite)
colnames(lens_popsite) <- c("pop","bag_site","mean_len","sd_len","event")

# and orders
ord_len_popsite <- data.frame()
for (i in 1:4) {
  ordt_lew <- lens_popsite[lens_popsite$event == i & lens_popsite$bag_site == "Lewisetta",]
  ordt_lew <- ordt_lew %>% arrange(desc(mean_len))
  ordt_lew$order_len <- 1:length(ordt_lew$pop)
  ordt_yrk <- lens_popsite[lens_popsite$event == i & lens_popsite$bag_site == "YorkRiver",]
  ordt_yrk <- ordt_yrk %>% arrange(desc(mean_len))
  ordt_yrk$order_len <- 1:length(ordt_yrk$pop)
  ord_len_popsite <- rbind(ord_len_popsite, ordt_lew, ordt_yrk)
}
ord_len_popsite

# put fitness metrics together
ord_fitness_popsite_mg <- merge(ord_mort_popsite, ord_len_popsite, by = c("pop","bag_site","event"))

# remove the larmix/seedmix pops
ord_fitness_popsite <- ord_fitness_popsite_mg[!ord_fitness_popsite_mg$pop %in% c("LARMIX","SEEDMIX"),]

# need a site name column
ord_fitness_popsite$site_name <- ord_fitness_popsite$pop
ord_fitness_popsite$site_name <- dplyr::recode(ord_fitness_popsite$site_name,"TX" = "W1_TX", "LA" = "W2_LA", "FL" = "W3_FL", "LOLA" = "S1_LOLA", "DEBY" = "S2_DEBY", "JR" = "W4_VA", "NH" = "W5_NH", "ME" = "W6_ME")

# put together the condition data in its own df
ord_fitness_popsite_cond <- merge(orders_cond, orders_gonad, by = c("bag_site", "site_name", "finalTissue_VialLabel", "condition", "gonad", "event"))

# do bag level data as well
colnames(mort_bags) <- c("bag","event","surv")
colnames(len) <- c("bag","event","length")
bagfit <- merge(mort_bags, len, by = c("bag","event"))

# pop info
bagfit$pop <- str_split_fixed(bagfit$bag, "-", 3)[,2]
bagfit

# site info
bagfit$bag_site <- ifelse(str_split_fixed(bagfit$bag, "-", 3)[,3] %in% c(1,2,3), "Lewisetta","YorkRiver")

# add this to other fitness data
bagfullfit <- merge(ord_fitness_popsite, bagfit, by = c("pop","event","bag_site"))

# different df for condition data since only one time point
ord_fitness_popsite_cond$pop <- str_split_fixed(ord_fitness_popsite_cond$site_name, "_", 2)[,2]

# need to align naming conventions to merge
# LEA
# colnames
colnames(lea_go) <- c("site_name", "lat", "lon", "offset_lea_go_lew","offset_lea_go_yrk","offset_lea_go_lew_nb","offset_lea_go_yrk_nb")

# add order
order_lew_lea <- as.data.frame(lea_go[order(lea_go$offset_lea_go_lew, decreasing = T),]$site_name)
colnames(order_lew_lea) <- c("site_name")
order_lew_lea$order_lew_lea <- 1:length(order_lew_lea$site_name)

order_lew_lea_nb <- as.data.frame(lea_go[order(lea_go$offset_lea_go_lew_nb, decreasing = T),]$site_name)
colnames(order_lew_lea_nb) <- c("site_name")
order_lew_lea_nb$order_lew_lea_nb <- 1:length(order_lew_lea_nb$site_name)

order_yrk_lea <- as.data.frame(lea_go[order(lea_go$offset_lea_go_yrk, decreasing = T),]$site_name)
colnames(order_yrk_lea) <- c("site_name")
order_yrk_lea$order_yrk_lea <- 1:length(order_yrk_lea$site_name)

order_yrk_lea_nb <- as.data.frame(lea_go[order(lea_go$offset_lea_go_yrk_nb, decreasing = T),]$site_name)
colnames(order_yrk_lea_nb) <- c("site_name")
order_yrk_lea_nb$order_yrk_lea_nb <- 1:length(order_yrk_lea_nb$site_name)

orders_lea_b <- merge(order_lew_lea, order_yrk_lea, by = c("site_name"))
orders_lea_nb <- merge(order_lew_lea_nb, order_yrk_lea_nb, by = ("site_name"))
orders_lea <- merge(orders_lea_b, orders_lea_nb, by = c("site_name"))
lea_go1 <- merge(lea_go, orders_lea, by = c("site_name"))

# GF
# add order
order_lew_geno <- as.data.frame(gf_go[order(gf_go$offset_gf_geno_lew, decreasing = T),]$site_name)
colnames(order_lew_geno) <- c("site_name")
order_lew_geno$order_lew_geno_gf <- 1:length(order_lew_geno$site_name)

order_lew_geno_nb <- as.data.frame(gf_go[order(gf_go$offset_gf_geno_lew_nb, decreasing = T),]$site_name)
colnames(order_lew_geno_nb) <- c("site_name")
order_lew_geno_nb$order_lew_geno_gf_nb <- 1:length(order_lew_geno_nb$site_name)

order_yrk_geno <- as.data.frame(lea_go[order(gf_go$offset_gf_geno_yrk, decreasing = T),]$site_name)
colnames(order_yrk_geno) <- c("site_name")
order_yrk_geno$order_yrk_geno_gf <- 1:length(order_yrk_geno$site_name)

order_yrk_geno_nb <- as.data.frame(lea_go[order(gf_go$offset_gf_geno_yrk_nb, decreasing = T),]$site_name)
colnames(order_yrk_geno_nb) <- c("site_name")
order_yrk_geno_nb$order_yrk_geno_gf_nb <- 1:length(order_yrk_geno_nb$site_name)

orders_gf_geno_b <- merge(order_lew_geno, order_yrk_geno, by = c("site_name"))
orders_gf_geno_nb <- merge(order_lew_geno_nb, order_yrk_geno_nb, by = c("site_name"))
orders_gf_geno <- merge(orders_gf_geno_b, orders_gf_geno_nb, by = c("site_name"))
gf_go1 <- merge(gf_go, orders_gf_geno, by = c("site_name"))

# RDA
# colnames
colnames(rda_go) <- c("site_name","lat","lon","offset_rda_lew","offset_rda_yrk","offset_rda_lew_nb","offset_rda_yrk_nb")

# add order
order_lew_rda <- as.data.frame(rda_go[order(rda_go$offset_rda_lew, decreasing = T),]$site_name)
colnames(order_lew_rda) <- c("site_name")
order_lew_rda$order_lew_rda <- 1:length(order_lew_rda$site_name)

order_lew_rda_nb <- as.data.frame(rda_go[order(rda_go$offset_rda_lew_nb, decreasing = T),]$site_name)
colnames(order_lew_rda_nb) <- c("site_name")
order_lew_rda_nb$order_lew_rda_nb <- 1:length(order_lew_rda_nb$site_name)

order_yrk_rda <- as.data.frame(rda_go[order(rda_go$offset_rda_yrk, decreasing = T),]$site_name)
colnames(order_yrk_rda) <- c("site_name")
order_yrk_rda$order_yrk_rda <- 1:length(order_yrk_rda$site_name)

order_yrk_rda_nb <- as.data.frame(rda_go[order(rda_go$offset_rda_yrk_nb, decreasing = T),]$site_name)
colnames(order_yrk_rda_nb) <- c("site_name")
order_yrk_rda_nb$order_yrk_rda_nb <- 1:length(order_yrk_rda_nb$site_name)

orders_rda_b <- merge(order_lew_rda, order_yrk_rda, by = c("site_name"))
orders_rda_nb <- merge(order_lew_rda_nb, order_yrk_rda_nb, by = c("site_name"))
orders_rda <- merge(orders_rda_b, orders_rda_nb, by = c("site_name"))
rda_go1 <- merge(rda_go, orders_rda, by = c("site_name"))

# environmental distance
colnames(env_dist) <- c("site_name", "env_dist_lew", "env_dist_yrk")
colnames(env_dist_nb) <- c("site_name", "env_dist_lew_nb", "env_dist_yrk_nb")
env_dist1 <- merge(env_dist, env_dist_nb, by = "site_name")
# add order
order_lew_envdist <- as.data.frame(env_dist1[order(env_dist1$env_dist_lew, decreasing = T),]$site_name)
colnames(order_lew_envdist) <- "site_name"
order_lew_envdist$order_lew_envdist <- 1:length(order_lew_envdist$site_name)

order_lew_envdist_nb <- as.data.frame(env_dist1[order(env_dist1$env_dist_lew_nb, decreasing = T),]$site_name)
colnames(order_lew_envdist_nb) <- "site_name"
order_lew_envdist_nb$order_lew_envdist_nb <- 1:length(order_lew_envdist_nb$site_name)

order_yrk_envdist <- as.data.frame(env_dist1[order(env_dist1$env_dist_yrk, decreasing = T),]$site_name)
colnames(order_yrk_envdist) <- "site_name"
order_yrk_envdist$order_yrk_envdist <- 1:length(order_yrk_envdist$site_name)

order_yrk_envdist_nb <- as.data.frame(env_dist1[order(env_dist1$env_dist_yrk_nb, decreasing = T),]$site_name)
colnames(order_yrk_envdist_nb) <- "site_name"
order_yrk_envdist_nb$order_yrk_envdist_nb <- 1:length(order_yrk_envdist_nb$site_name)

orders_envdist_b <- merge(order_lew_envdist, order_yrk_envdist, by = "site_name")
orders_envdist_nb <- merge(order_lew_envdist_nb, order_yrk_envdist_nb, by = "site_name")
orders_envdist <- merge(orders_envdist_b, orders_envdist_nb, by = "site_name")
env_dist2 <- merge(env_dist1, orders_envdist, by = "site_name")

# put together full go & fitness df
go <- merge(rda_go1, gf_go1, by = c("site_name","lat","lon")) 
go1 <- merge(go, lea_go1, by = c("site_name", "lat", "lon"))
go2 <- merge(go1, env_dist2, by = "site_name")
fullfit <- merge(bagfullfit, go2, by = "site_name")
```


## Compare Offsets to Fitness

Now that offsets and fitness are in a comparable format, look at how well the two match.

### Performance Metric Calculation

Here I calculate the R^2 and pearson's correlation between fitness and GO for each method at each site. I do this with fitness data both at the bag level and summarized at the population level (since offset is population level).

```{r}
# correlations between go/fit at each time point
#################################################
# function for correlations
cor_fit <- function(site, go_method, bio, sel, df) {
  cor_df <- data.frame(method = NA, biotic = NA, sel = NA, bag_site = NA, event = NA, pearson_surv = NA, pearson_surv_pval = NA, spearman_surv = NA, spearman_surv_pval = NA, kendall_surv = NA, kendall_surv_pval = NA, r2_surv = NA, r2_surv_rank = NA, pearson_len = NA, pearson_len_pval = NA, spearman_len = NA, spearman_len_pval = NA, kendall_len = NA, kendall_len_pval = NA, r2_len = NA, r2_len_rank = NA)
  
  # what metric should we use based on inputs
  lookup_metric <- list(
  Lewisetta = list(
    b = list(lea = "offset_lea_go_lew",
             gf_geno = "offset_gf_geno_lew", 
             rda = "offset_rda_lew", 
             env_dist = "env_dist_lew"),
    nb = list(lea = "offset_lea_go_lew_nb",
              gf_geno = "offset_gf_geno_lew_nb", 
              rda = "offset_rda_lew_nb",
              env_dist = "env_dist_lew_nb")),
  YorkRiver = list(
    b = list(lea = "offset_lea_go_yrk",
             gf_geno = "offset_gf_geno_yrk", 
             rda = "offset_rda_yrk", 
             env_dist = "env_dist_yrk"),
    nb = list(lea = "offset_lea_go_yrk_nb",
              gf_geno = "offset_gf_geno_yrk_nb", 
              rda = "offset_rda_yrk_nb", 
              env_dist = "env_dist_yrk_nb")))
  
  lookup_rank <- list(
  Lewisetta = list(
    b = list(lea = "order_lew_lea",
             gf_geno = "order_lew_geno_gf", 
             rda = "order_lew_rda", 
             env_dist = "order_lew_envdist"),
    nb = list(lea = "order_lew_lea_nb", 
              gf_geno = "order_lew_geno_gf_nb", 
              rda = "order_lew_rda_nb", 
              env_dist = "order_lew_envdist_nb")),
  YorkRiver = list(
    b = list(lea = "order_yrk_lea",
             gf_geno = "order_yrk_geno_gf", 
             rda = "order_yrk_rda", 
             env_dist = "order_yrk_envdist"),
    nb = list(lea = "order_yrk_lea_nb", 
              gf_geno = "order_yrk_geno_gf_nb", 
              rda = "order_yrk_rda_nb", 
              env_dist = "order_yrk_envdist_nb")))
  
  go_metric <- lookup_metric[[site]][[bio]][[go_method]]
  go_rank <- lookup_rank[[site]][[bio]][[go_method]]

  for (i in 1:4) {
    if (sel == "sel") {
      fitdat <- df[df$bag_site == (site) & df$event == i,]
    } else {
      fitdat <- df[df$bag_site == (site) & df$event == i & df$site_name %in% c("W1_TX","W2_LA","W3_FL","W4_VA","W5_NH","W6_ME"),]
    }
    
    # Convert to numeric to handle any character/factor columns
    metric_values <- na.omit(as.numeric(fitdat[,go_metric]))
    rank_values <- na.omit(as.numeric(fitdat[,go_rank]))
    surv_values <- na.omit(as.numeric(fitdat$surv))
    surv_ranks <- na.omit(as.numeric(fitdat$order_mort))
    len_values <- na.omit(as.numeric(fitdat$length))
    len_ranks <- na.omit(as.numeric(fitdat$order_len))
    
    # metadata
    cor_df[i,]$method <- as.character(go_method)
    cor_df[i,]$bag_site <- as.character(site)
    cor_df[i,]$event <- i
    cor_df[i,]$biotic <- as.character(bio)
    cor_df[i,]$sel <- as.character(sel)    
    
    # survival
    cor_df[i,]$pearson_surv <- cor.test(metric_values, surv_values, method = "pearson")$estimate
    cor_df[i,]$pearson_surv_pval <- cor.test(metric_values, surv_values, method = "pearson")$p.value
    cor_df[i,]$spearman_surv <- cor.test(metric_values, surv_values, method = "spearman")$estimate
    cor_df[i,]$spearman_surv_pval <- cor.test(metric_values, surv_values, method = "spearman")$p.value
    cor_df[i,]$kendall_surv <- cor.test(metric_values, surv_values, method = "kendall")$estimate
    cor_df[i,]$kendall_surv_pval <- cor.test(metric_values, surv_values, method = "kendall")$p.value
    cor_df[i,]$r2_surv <- summary(lm(surv_values ~ metric_values))$r.squared
    cor_df[i,]$r2_surv_rank <- summary(lm(surv_ranks ~ rank_values))$r.squared

    # length
    cor_df[i,]$pearson_len <- cor.test(metric_values, len_values, method = "pearson")$estimate
    cor_df[i,]$pearson_len_pval <- cor.test(metric_values, len_values, method = "pearson")$p.value
    cor_df[i,]$spearman_len <- cor.test(metric_values, len_values, method = "spearman")$estimate
    cor_df[i,]$spearman_len_pval <- cor.test(metric_values, len_values, method = "spearman")$p.value
    cor_df[i,]$kendall_len <- cor.test(metric_values, len_values, method = "kendall")$estimate
    cor_df[i,]$kendall_len_pval <- cor.test(metric_values, len_values, method = "kendall")$p.value
    cor_df[i,]$r2_len <- summary(lm(len_values ~ metric_values))$r.squared
    cor_df[i,]$r2_len_rank <- summary(lm(len_ranks ~ rank_values))$r.squared
  }
  return(cor_df)
}

# correlations
(cor_lea_lew <- cor_fit("Lewisetta","lea","b", "sel", fullfit))
(cor_lea_yrk <- cor_fit("YorkRiver","lea","b", "sel", fullfit))
(cor_gf_geno_lew <- cor_fit("Lewisetta","gf_geno","b", "sel", fullfit))
(cor_gf_geno_yrk <- cor_fit("YorkRiver","gf_geno","b", "sel", fullfit))
(cor_rda_lew <- cor_fit("Lewisetta","rda","b", "sel", fullfit))
(cor_rda_yrk <- cor_fit("YorkRiver","rda","b", "sel", fullfit))
(cor_envdist_lew <- cor_fit("Lewisetta","env_dist", "b", "sel", fullfit))
(cor_envdist_yrk <- cor_fit("YorkRiver","env_dist", "b", "sel", fullfit))

(cor_lea_lew_nb <- cor_fit("Lewisetta","lea","nb", "sel", fullfit))
(cor_lea_yrk_nb <- cor_fit("YorkRiver","lea","nb", "sel", fullfit))
(cor_gf_geno_lew_nb <- cor_fit("Lewisetta","gf_geno","nb", "sel", fullfit))
(cor_gf_geno_yrk_nb <- cor_fit("YorkRiver","gf_geno","nb", "sel", fullfit))
(cor_rda_lew_nb <- cor_fit("Lewisetta","rda","nb", "sel", fullfit))
(cor_rda_yrk_nb <- cor_fit("YorkRiver","rda","nb", "sel", fullfit))
(cor_envdist_lew_nb <- cor_fit("Lewisetta","env_dist", "nb", "sel", fullfit))
(cor_envdist_yrk_nb <- cor_fit("YorkRiver","env_dist", "nb", "sel", fullfit))

# without selection lines
# correlations
(cor_lea_lew_ns <- cor_fit("Lewisetta","lea","b", "ns", fullfit))
(cor_lea_yrk_ns <- cor_fit("YorkRiver","lea","b", "ns", fullfit))
(cor_gf_geno_lew_ns <- cor_fit("Lewisetta","gf_geno","b", "ns", fullfit))
(cor_gf_geno_yrk_ns <- cor_fit("YorkRiver","gf_geno","b", "ns", fullfit))
(cor_rda_lew_ns <- cor_fit("Lewisetta","rda","b", "ns", fullfit))
(cor_rda_yrk_ns <- cor_fit("YorkRiver","rda","b", "ns", fullfit))
(cor_envdist_lew_ns <- cor_fit("Lewisetta","env_dist", "b", "ns", fullfit))
(cor_envdist_yrk_ns <- cor_fit("YorkRiver","env_dist", "b", "ns", fullfit))

(cor_lea_lew_nb_ns <- cor_fit("Lewisetta","lea","nb", "ns", fullfit))
(cor_lea_yrk_nb_ns <- cor_fit("YorkRiver","lea","nb", "ns", fullfit))
(cor_gf_geno_lew_nb_ns <- cor_fit("Lewisetta","gf_geno","nb", "ns", fullfit))
(cor_gf_geno_yrk_nb_ns <- cor_fit("YorkRiver","gf_geno","nb", "ns", fullfit))
(cor_rda_lew_nb_ns <- cor_fit("Lewisetta","rda","nb", "ns", fullfit))
(cor_rda_yrk_nb_ns <- cor_fit("YorkRiver","rda","nb", "ns", fullfit))
(cor_envdist_lew_nb_ns <- cor_fit("Lewisetta","env_dist", "nb", "ns", fullfit))
(cor_envdist_yrk_nb_ns <- cor_fit("YorkRiver","env_dist", "nb", "ns", fullfit))

corrs_df <- rbind(cor_lea_lew, cor_lea_yrk, cor_gf_geno_lew, cor_gf_geno_yrk, cor_rda_lew, cor_rda_yrk, cor_envdist_lew, cor_envdist_yrk, cor_lea_lew_nb, cor_lea_yrk_nb, cor_gf_geno_lew_nb, cor_gf_geno_yrk_nb, cor_rda_lew_nb, cor_rda_yrk_nb, cor_envdist_lew_nb, cor_envdist_yrk_nb, cor_lea_lew_ns, cor_lea_yrk_ns, cor_gf_geno_lew_ns, cor_gf_geno_yrk_ns, cor_rda_lew_ns, cor_rda_yrk_ns, cor_envdist_lew_ns, cor_envdist_yrk_ns, cor_lea_lew_nb_ns, cor_lea_yrk_nb_ns, cor_gf_geno_lew_nb_ns, cor_gf_geno_yrk_nb_ns, cor_rda_lew_nb_ns, cor_rda_yrk_nb_ns, cor_envdist_lew_nb_ns, cor_envdist_yrk_nb_ns)
rownames(corrs_df) <- NULL
corrs_df$signif_surv <- ifelse(corrs_df$kendall_surv_pval < .05, T, F)
corrs_df$signif_len <- ifelse(corrs_df$kendall_len_pval < .05, T, F)

# adjust pvals
corrs_df$p_adj_surv <- p.adjust(corrs_df$kendall_surv_pval, method = "BH")
corrs_df$p_adj_len <- p.adjust(corrs_df$kendall_len_pval, method = "BH")

# signif for adj
corrs_df$signif_surv_adj <- ifelse(corrs_df$p_adj_surv < .05, T, F)
corrs_df$signif_len_adj <- ifelse(corrs_df$p_adj_len < .05, T, F)

# save this
write.csv(corrs_df, paste0("results/GO_results/go_fit_corrs_",Sys.Date(),".csv"), row.names = F)

# another version without pearson or spearman corrs
corrs_df_kend <- corrs_df %>% dplyr::select(-contains(c("pearson", "spearman")))

# and for only final time point
corrs_df_kend_t4 <- corrs_df_kend[corr_df_kend$event == 4,]

write.csv(corrs_df_kend_t4, paste0("results/GO_results/go_fit_corrs_final_",Sys.Date(),".csv"), row.names = F)
#################################################
```


### Plot GO vs mortality against each other

I'll start by plotting LEA results:

With Selection Lines

```{r}
# function for linetypes
lty_set <- function(method, biotic, sel, site, event) {
    type = ifelse(corrs_df[corrs_df$method == as.character(method) & corrs_df$biotic == as.character(biotic) & corrs_df$sel == as.character(sel) & corrs_df$bag_site == as.character(site) & corrs_df$event == as.numeric(event),]$signif_surv_adj == T, "solid", "dashed")
    return(type)
}

# plot lea vs mort at each site/time point
goFitLEA_lew_plot <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lea_go_lew, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lea_go_lew, y = surv), col = "cadetblue3", fill = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("lea","b","sel","Lewisetta",1)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lea_go_lew, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lea_go_lew, y = surv), col = "cornflowerblue", fill = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("lea","b","sel","Lewisetta",2)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lea_go_lew, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lea_go_lew, y = surv), col = "slateblue", fill = "slateblue", fill = "slateblue", se = T, linetype = lty_set("lea","b","sel","Lewisetta",3)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = surv), col = "purple4", fill = "purple4", fill = "purple4", se = T, linetype = lty_set("lea","b","sel","Lewisetta",4)) +
    labs(title = expression(paste("(A) Predicted ", LEA[offset], " vs Survival at Lewisetta, with Selection Lines")), x = "LEA Offset", y = "Survival") +
    scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
    theme_classic()
goFitLEA_lew_plot

goFitLEA_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_lea_go_yrk, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_lea_go_yrk, y = surv), col = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("lea","b","sel","YorkRiver",1)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_lea_go_yrk, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_lea_go_yrk, y = surv), col = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("lea","b","sel","YorkRiver",2)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_lea_go_yrk, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_lea_go_yrk, y = surv), col = "slateblue", fill = "slateblue", se = T, linetype = lty_set("lea","b","sel","YorkRiver",3)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = surv), col = "purple4", fill = "purple4", se = T, linetype = lty_set("lea","b","sel","YorkRiver",4)) +
  labs(title = expression(paste("(B) Predicted ", LEA[offset], " vs Survival at York River, with Selection Lines")), x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_plot

# plot go lea vs mort at both sites just final time point
goFitLEA_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = surv), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("lea","b","sel","Lewisetta",4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = surv), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("lea","b","sel","YorkRiver",4), se = T) +
  labs(title = expression(paste("(A) Predicted ", LEA[offset], " vs Final Survival, with Selection Lines")), x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotfinalTsurv
```

Without Selection Lines

```{r}
# plot lea vs mort at each site/time point
goFitLEA_lew_plot_nosel <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta" & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_lew, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_lew, y = surv), col = "cadetblue3", fill = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("lea","b","ns","Lewisetta",1)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_lew, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_lew, y = surv), col = "cornflowerblue", fill = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("lea", "b", "ns", "Lewisetta", 2)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_lew, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_lew, y = surv), col = "slateblue", fill = "slateblue", fill = "slateblue", se = T, linetype = lty_set("lea", "b", "ns", "Lewisetta", 3)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_lew, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_lew, y = surv), col = "purple4", fill = "purple4", fill = "purple4", se = T, linetype = lty_set("lea", "b", "ns", "Lewisetta", 4)) +
    labs(title = expression(paste("(A) Predicted ", LEA[offset], " vs Survival at Lewisetta")), x = "LEA Offset", y = "Survival") +
    scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
    theme_classic()
goFitLEA_lew_plot_nosel

goFitLEA_yrk_plot_nosel <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver"  & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_yrk, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_yrk, y = surv), col = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("lea", "b", "ns", "YorkRiver", 1)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_yrk, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_yrk, y = surv), col = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("lea", "b", "ns", "YorkRiver", 2)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_yrk, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_yrk, y = surv), col = "slateblue", fill = "slateblue", se = T, linetype = lty_set("lea", "b", "ns", "YorkRiver", 3)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_yrk, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_yrk, y = surv), col = "purple4", fill = "purple4", se = T, linetype = lty_set("lea", "b", "ns", "YorkRiver", 4)) +
  labs(title = expression(paste("(B) Predicted ", LEA[offset], " vs Survival at York River")), x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_plot_nosel

# plot go lea vs mort at both sites just final time point
goFitLEA_plotfinalTsurv_nosel <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_lew, y = surv), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("lea", "b", "ns", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_lea_go_yrk, y = surv), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("lea", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(A) Predicted ", LEA[offset], " vs Final Survival")), x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotfinalTsurv_nosel
```


Then I'll plot rda results:

With selection lines:

```{r}
# plot rda vs mort at each site/time point
goFitRDA_lew_plot <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_rda_lew, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_rda_lew, y = surv), col = "cadetblue3", fill = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda","b","sel","Lewisetta",1)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_rda_lew, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_rda_lew, y = surv), col = "cornflowerblue", fill = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "sel", "Lewisetta", 2)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_rda_lew, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_rda_lew, y = surv), col = "slateblue", fill = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "sel", "Lewisetta", 3)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = surv), col = "purple4", fill = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "sel", "Lewisetta", 4)) +
    labs(title = expression(paste("(C) Predicted ", RDA[offset], " vs Survival at Lewisetta, with Selection Lines")), x = "RDA Offset", y = "Survival") +
    scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
    theme_classic()
goFitRDA_lew_plot

goFitRDA_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_rda_yrk, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_rda_yrk, y = surv), col = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 1)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_rda_yrk, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver",], aes(x = offset_rda_yrk, y = surv), col = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 2)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_rda_yrk, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_rda_yrk, y = surv), col = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 3)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = surv), col = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 4)) +
  labs(title = expression(paste("(D) Predicted ", RDA[offset], " vs Survival at York River, with Selection Lines")), x = "RDA Offset", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_yrk_plot

# plot go rda vs mort at both sites just final time point
goFitRDA_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = surv), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("rda", "b", "sel", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = surv), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("rda", "b", "sel", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(B) Predicted ", RDA[offset], " vs Final Survival, with Selection Lines")), x = "RDA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitRDA_plotfinalTsurv
```

Without selection lines:

```{r}
# plot rda vs mort at each site/time point
goFitRDA_lew_plot_nosel <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta" & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_lew, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_lew, y = surv), col = "cadetblue3", fill = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda","b","ns","Lewisetta",1)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_lew, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_lew, y = surv), col = "cornflowerblue", fill = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "ns", "Lewisetta", 2)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_lew, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_lew, y = surv), col = "slateblue", fill = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "ns", "Lewisetta", 3)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_lew, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_lew, y = surv), col = "purple4", fill = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "ns", "Lewisetta", 4)) +
    labs(title = expression(paste("(A) Predicted ", RDA[offset], " vs Survival at Lewisetta")), x = "RDA Offset", y = "Survival") +
    scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
    theme_classic()
goFitRDA_lew_plot_nosel

goFitRDA_yrk_plot_nosel <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver"  & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_yrk, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_yrk, y = surv), col = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 1)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_yrk, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_yrk, y = surv), col = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 2)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_yrk, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_yrk, y = surv), col = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 3)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_yrk, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_yrk, y = surv), col = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 4)) +
  labs(title = expression(paste("(B) Predicted ", RDA[offset], " vs Survival at York River")), x = "RDA Offset", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_yrk_plot_nosel

# plot go rda vs mort at both sites just final time point
goFitRDA_plotfinalTsurv_nosel <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_lew, y = surv), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("rda", "b", "ns", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_rda_yrk, y = surv), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("rda", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(A) Predicted ", RDA[offset], " vs Final Survival")), x = "RDA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitRDA_plotfinalTsurv_nosel
```


Then the gradient forest results:

With selection lines

```{r}
# plot gf vs mort at each site/time point
goFitGF_lew_plot <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta" & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv), col = "cadetblue3", fill = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda","b","sel","Lewisetta",1)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv), col = "cornflowerblue", fill = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "sel", "Lewisetta", 2)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv), col = "slateblue", fill = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "sel", "Lewisetta", 3)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv), col = "purple4", fill = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "sel", "Lewisetta", 4)) +
    labs(title = expression(paste("(A) Predicted ", GF[offset], " vs Survival at Lewisetta, with Selection Lines")), x = "GF Offset", y = "Survival") +
    scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
    theme_classic()
goFitGF_lew_plot

goFitGF_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver"  & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv), col = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 1)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv), col = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 2)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv), col = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 3)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv), col = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 4)) +
  labs(title = expression(paste("(B) Predicted ", GF[offset], " vs Survival at York River, with Selection Lines")), x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_yrk_plot

# plot go gf vs mort at both sites just final time point
goFitGF_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("rda", "b", "sel", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("rda", "b", "sel", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(A) Predicted ", GF[offset], " vs Final Survival, with Selection Lines")), x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGF_plotfinalTsurv
```

Without selection lines:

```{r}
# plot gf vs mort at each site/time point
goFitGF_lew_plot_nosel <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta" & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv), col = "cadetblue3", fill = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda","b","ns","Lewisetta",1)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv), col = "cornflowerblue", fill = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "ns", "Lewisetta", 2)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv), col = "slateblue", fill = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "ns", "Lewisetta", 3)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv), col = "purple4", fill = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "ns", "Lewisetta", 4)) +
    labs(title = expression(paste("(A) Predicted ", GF[offset], " vs Survival at Lewisetta")), x = "GF Offset", y = "Survival") +
    scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
    theme_classic()
goFitGF_lew_plot_nosel

goFitGF_yrk_plot_nosel <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver"  & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv), col = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 1)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv), col = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 2)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv), col = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 3)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv), col = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 4)) +
  labs(title = expression(paste("(B) Predicted ", GF[offset], " vs Survival at York River")), x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_yrk_plot_nosel

# plot go gf vs mort at both sites just final time point
goFitGF_plotfinalTsurv_nosel <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_lew, y = surv), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("rda", "b", "ns", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = offset_gf_geno_yrk, y = surv), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("rda", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(A) Predicted ", GF[offset], " vs Final Survival")), x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGF_plotfinalTsurv_nosel
```

And finally, environmental distance:

With selection lines:

```{r}
# plot envdist vs mort at each site/time point
goFitEnvDist_lew_plot <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta" & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv), col = "cadetblue3", fill = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda","b","sel","Lewisetta",1)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv), col = "cornflowerblue", fill = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "sel", "Lewisetta", 2)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv), col = "slateblue", fill = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "sel", "Lewisetta", 3)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv), col = "purple4", fill = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "sel", "Lewisetta", 4)) +
    labs(title = expression(paste("(A) Environmental Distance vs Survival at \nLewisetta, with Selection Lines")), x = "Environmental Distance", y = "Survival") +
    scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
    theme_classic()
goFitEnvDist_lew_plot

goFitEnvDist_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver"  & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv), col = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 1)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv), col = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 2)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv), col = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 3)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv), col = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "sel", "YorkRiver", 4)) +
  labs(title = expression(paste("(B) Environmental Distance vs Survival at \nYork River, with Selection Lines")), x = "Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvDist_yrk_plot

# plot go rda vs mort at both sites just final time point
goFitEnvDist_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("rda", "b", "sel", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("rda", "b", "sel", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(A) Environmental Distance vs Final Survival, with Selection Lines")), x = "Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitEnvDist_plotfinalTsurv
```

Without selection lines:

```{r}
# plot envdist vs mort at each site/time point
goFitEnvDist_lew_plot_nosel <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta" & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv), col = "cadetblue3", fill = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda","b","ns","Lewisetta",1)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv), col = "cornflowerblue", fill = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "ns", "Lewisetta", 2)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv), col = "slateblue", fill = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "ns", "Lewisetta", 3)) +
    geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
    geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv), col = "purple4", fill = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "ns", "Lewisetta", 4)) +
    labs(title = expression(paste("(A) Environmental Distance vs Survival at Lewisetta")), x = "GF Offset", y = "Survival") +
    scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
    theme_classic()
goFitEnvDist_lew_plot_nosel

goFitEnvDist_yrk_plot_nosel <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver"  & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv), col = "cadetblue3", fill = "cadetblue3", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 1)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv), col = "cornflowerblue", fill = "cornflowerblue", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 2)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv), col = "slateblue", fill = "slateblue", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 3)) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv), col = "purple4", fill = "purple4", se = T, linetype = lty_set("rda", "b", "ns", "YorkRiver", 4)) +
  labs(title = expression(paste("(B) Environmental Distance vs Survival at York River")), x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvDist_yrk_plot_nosel

# plot envdist vs mort at both sites just final time point
goFitEnvDist_plotfinalTsurv_nosel <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_lew, y = surv), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("rda", "b", "ns", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4 & !(fullfit$site_name %in% c("S1_LOLA", "S2_DEBY")),], aes(x = env_dist_yrk, y = surv), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("rda", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(A) Environmental Distance vs Final Survival")), x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitEnvDist_plotfinalTsurv_nosel
```


### Plot GO vs length against each other

LEA vs length:

With selection lines: 

```{r}
lty_set <- function(method, biotic, sel, site, event) {
    type = ifelse(corrs_df[corrs_df$method == as.character(method) & corrs_df$biotic == as.character(biotic) & corrs_df$sel == as.character(sel) & corrs_df$bag_site == as.character(site) & corrs_df$event == as.numeric(event),]$signif_len_adj == T, "solid", "dashed")
    return(type)
}

# plot lea vs len at each site/time point
goFitLEA_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lea_go_lew, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lea_go_lew, y = length), col = "cadetblue3", fill = "cadetblue3", fill = "cadetblue3", linetype = lty_set("lea", "b", "sel", "Lewisetta", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lea_go_lew, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lea_go_lew, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("lea", "b", "sel", "Lewisetta", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lea_go_lew, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lea_go_lew, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("lea", "b", "sel", "Lewisetta", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length), col = "purple4", fill = "purple4", linetype = lty_set("lea", "b", "sel", "Lewisetta", 4), se = T) +
  labs(title = expression(paste("(A) Predicted ", LEA[offset], " vs Length at Lewisetta, with Selection Lines")), x = "LEA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_lew_len

goFitLEA_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_lea_go_yrk, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_lea_go_yrk, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("lea", "b", "sel", "YorkRiver", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_lea_go_yrk, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_lea_go_yrk, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("lea", "b", "sel", "YorkRiver", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_lea_go_yrk, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_lea_go_yrk, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("lea", "b", "sel", "YorkRiver", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length), col = "purple4", fill = "purple4", linetype = lty_set("lea", "b", "sel", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(B) Predicted ", LEA[offset], " vs Length at York River, with Selection Lines")), x = "LEA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_len

# plot go lea vs len at both sites just final time point
goFitLEA_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("lea", "b", "sel", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("lea", "b", "sel", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(A) Predicted ", LEA[offset], " vs Final Length, with Selection Lines")), x = "LEA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotfinalTlen
```

Without selection lines:

```{r}
# plot lea vs len at each site/time point
goFitLEA_lew_len_nosel <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lea_go_lew, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lea_go_lew, y = length), col = "cadetblue3", fill = "cadetblue3", fill = "cadetblue3", linetype = lty_set("lea", "b", "ns", "Lewisetta", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lea_go_lew, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lea_go_lew, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("lea", "b", "ns", "Lewisetta", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lea_go_lew, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lea_go_lew, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("lea", "b", "ns", "Lewisetta", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length), col = "purple4", fill = "purple4", linetype = lty_set("lea", "b", "ns", "Lewisetta", 4), se = T) +
  labs(title = expression(paste("(A) Predicted ", LEA[offset], " vs Length at Lewisetta")), x = "LEA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_lew_len_nosel

goFitLEA_yrk_len_nosel <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_lea_go_yrk, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_lea_go_yrk, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("lea", "b", "ns", "YorkRiver", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_lea_go_yrk, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_lea_go_yrk, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("lea", "b", "ns", "YorkRiver", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_lea_go_yrk, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_lea_go_yrk, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("lea", "b", "ns", "YorkRiver", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length), col = "purple4", fill = "purple4", linetype = lty_set("lea", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(B) Predicted ", LEA[offset], " vs Length at York River")), x = "LEA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_len_nosel

# plot go lea vs len at both sites just final time point
goFitLEA_plotfinalTlen_nosel <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("lea", "b", "ns", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("lea", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(A) Predicted ", LEA[offset], " vs Final Length")), x = "LEA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotfinalTlen_nosel
```


RDA GO vs length:

With selection lines:

```{r}
# plot rda vs len at each site/time point
goFitRDA_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_rda_lew, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_rda_lew, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("rda", "b", "sel", "Lewisetta", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_rda_lew, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_rda_lew, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("rda", "b", "sel", "Lewisetta", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_rda_lew, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_rda_lew, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("rda", "b", "sel", "Lewisetta", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length), col = "purple4", fill = "purple4", linetype = lty_set("rda", "b", "sel", "Lewisetta", 4), se = T) +
  labs(title = expression(paste("(C) Predicted ", RDA[offset], " vs Length at Lewisetta, with Selection Lines")), x = "RDA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_lew_len

goFitRDA_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_rda_yrk, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_rda_yrk, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("rda", "b", "sel", "YorkRiver", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_rda_yrk, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_rda_yrk, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("rda", "b", "sel", "YorkRiver", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_rda_yrk, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_rda_yrk, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("rda", "b", "sel", "YorkRiver", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length), col = "purple4", fill = "purple4", linetype = lty_set("rda", "b", "sel", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(D) Predicted ", RDA[offset], " vs Length at York River, with Selection Lines")), x = "RDA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_yrk_len

# plot rda vs len at both sites just final time point
goFitRDA_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("rda", "b", "sel", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("rda", "b", "sel", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(B) Predicted ", RDA[offset], " vs Final Length, with Selection Lines")), x = "RDA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitRDA_plotfinalTlen
```

Without selection lines:

```{r}
# plot rda vs len at each site/time point
goFitRDA_lew_len_nosel <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_rda_lew, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_rda_lew, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("rda", "b", "ns", "Lewisetta", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_rda_lew, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_rda_lew, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("rda", "b", "ns", "Lewisetta", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_rda_lew, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_rda_lew, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("rda", "b", "ns", "Lewisetta", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length), col = "purple4", fill = "purple4", linetype = lty_set("rda", "b", "ns", "Lewisetta", 4), se = T) +
  labs(title = expression(paste("(C) Predicted ", RDA[offset], " vs Length at Lewisetta")), x = "RDA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_lew_len_nosel

goFitRDA_yrk_len_nosel <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_rda_yrk, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_rda_yrk, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("rda", "b", "ns", "YorkRiver", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_rda_yrk, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_rda_yrk, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("rda", "b", "ns", "YorkRiver", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_rda_yrk, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_rda_yrk, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("rda", "b", "ns", "YorkRiver", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length), col = "purple4", fill = "purple4", linetype = lty_set("rda", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(D) Predicted ", RDA[offset], " vs Length at York River")), x = "RDA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_yrk_len_nosel

# plot rda vs len at both sites just final time point
goFitRDA_plotfinalTlen_nosel <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("rda", "b", "ns", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("rda", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(B) Predicted ", RDA[offset], " vs Final Length")), x = "RDA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitRDA_plotfinalTlen_nosel
```


GF geno GO vs length: 

With selection lines:

```{r}
# plot gf vs len at each site/time point
goFitGF_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("gf_geno", "b", "sel", "Lewisetta", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("gf_geno", "b", "sel", "Lewisetta", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("gf_geno", "b", "sel", "Lewisetta", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length), col = "purple4", fill = "purple4", linetype = lty_set("gf_geno", "b", "sel", "Lewisetta", 4), se = T) +
  labs(title = expression(paste("(E) Predicted ", GF[offset], " vs Length at Lewisetta, with Selection Lines")), x = "GF Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_lew_len

goFitGF_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("gf_geno", "b", "sel", "YorkRiver", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("gf_geno", "b", "sel", "YorkRiver", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("gf_geno", "b", "sel", "YorkRiver", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length), col = "purple4", fill = "purple4", linetype = lty_set("gf_geno", "b", "sel", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(F) Predicted ", GF[offset], " vs Length at York River, with Selection Lines")), x = "GF Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_yrk_len

# plot gf vs len at both sites just final time point
goFitGF_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("gf_geno", "b", "sel", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("gf_geno", "b", "sel", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(C) Predicted ", GF[offset], " vs Final Length, with Selection Lines")), x = "GF Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGF_plotfinalTlen
```

Without selection lines:

```{r}
# plot gf vs len at each site/time point
goFitGF_lew_len_nosel <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("gf_geno", "b", "ns", "Lewisetta", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("gf_geno", "b", "ns", "Lewisetta", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("gf_geno", "b", "ns", "Lewisetta", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length), col = "purple4", fill = "purple4", linetype = lty_set("gf_geno", "b", "ns", "Lewisetta", 4), se = T) +
  labs(title = expression(paste("(E) Predicted ", GF[offset], " vs Length at Lewisetta")), x = "GF Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4")) +
  theme_classic()
goFitGF_lew_len_nosel

goFitGF_yrk_len_nosel <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("gf_geno", "b", "ns", "YorkRiver", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("gf_geno", "b", "ns", "YorkRiver", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("gf_geno", "b", "ns", "YorkRiver", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length), col = "purple4", fill = "purple4", linetype = lty_set("gf_geno", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(F) Predicted ", GF[offset], " vs Length at York River")), x = "GF Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_yrk_len_nosel

# plot gf vs len at both sites just final time point
goFitGF_plotfinalTlen_nosel <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("gf_geno", "b", "ns", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("gf_geno", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = expression(paste("(C) Predicted ", GF[offset], " vs Final Length")), x = "GF Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGF_plotfinalTlen_nosel
```


Environmental distance vs length:

With Selection Lines:

```{r}
# plot env dist vs length at each site/time point
goFitEnvDist_len_lew_plot <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("env_dist", "b", "sel", "Lewisetta", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("env_dist", "b", "sel", "Lewisetta", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("env_dist", "b", "sel", "Lewisetta", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length), col = "purple4", fill = "purple4", linetype = lty_set("env_dist", "b", "sel", "Lewisetta", 4), se = T) +
  labs(title = "(G) Environmental Distance vs Length at Lewisetta, with Selection Lines", x = "Environmental Distance", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvDist_len_lew_plot

goFitEnvDist_len_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("env_dist", "b", "sel", "YorkRiver", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("env_dist", "b", "sel", "YorkRiver", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("env_dist", "b", "sel", "YorkRiver", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length), col = "purple4", fill = "purple4", linetype = lty_set("env_dist", "b", "sel", "YorkRiver", 4), se = T) +
  labs(title = "(H) Environmental Distance vs Length at York River, with Selection Lines", x = "Environmental Distance", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvDist_len_yrk_plot

# plot go gf vs len at both sites just final time point
goFitEnvDist_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("env_dist", "b", "sel", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("env_dist", "b", "sel", "YorkRiver", 4), se = T) +
  labs(title = "(D) Environmental Distance vs Final Length, with Selection Lines", x = "Environmental Distance", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitEnvDist_plotfinalTlen
```

Without Selection Lines:

```{r}
# plot env dist vs length at each site/time point
goFitEnvDist_len_lew_plot_nosel <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("env_dist", "b", "ns", "Lewisetta", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("env_dist", "b", "ns", "Lewisetta", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("env_dist", "b", "ns", "Lewisetta", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length), col = "purple4", fill = "purple4", linetype = lty_set("env_dist", "b", "ns", "Lewisetta", 4), se = T) +
  labs(title = "(G) Environmental Distance vs Length at Lewisetta", x = "Environmental Distance", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvDist_len_lew_plot_nosel

goFitEnvDist_len_yrk_plot_nosel <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = length), col = "cadetblue3", fill = "cadetblue3", linetype = lty_set("env_dist", "b", "ns", "YorkRiver", 1), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = length), col = "cornflowerblue", fill = "cornflowerblue", linetype = lty_set("env_dist", "b", "ns", "YorkRiver", 2), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = length), col = "slateblue", fill = "slateblue", linetype = lty_set("env_dist", "b", "ns", "YorkRiver", 3), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length), col = "purple4", fill = "purple4", linetype = lty_set("env_dist", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = "(H) Environmental Distance vs Length at York River", x = "Environmental Distance", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvDist_len_yrk_plot_nosel

# plot go gf vs len at both sites just final time point
goFitEnvDist_plotfinalTlen_nosel <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length), col = "mediumspringgreen", fill = "darkseagreen2", linetype = lty_set("env_dist", "b", "ns", "Lewisetta", 4), se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length), col = "dodgerblue", fill = "dodgerblue", linetype = lty_set("env_dist", "b", "ns", "YorkRiver", 4), se = T) +
  labs(title = "(D) Environmental Distance vs Final Length", x = "Environmental Distance", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitEnvDist_plotfinalTlen_nosel
```


Export all of these plots to pdf:

```{r}
# this version lets me export directly as an editable powerpoint with vector graphics
library(officer) # for powerpoint export/editing
library(rvg) # for converting r graphics to editable vector graphics

doc <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_plotfinalTsurv), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_plotfinalTsurv), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_plotfinalTsurv), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_plotfinalTsurv), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_plotfinalTlen), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_plotfinalTlen), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_plotfinalTlen), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_plotfinalTlen), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_lew_plot), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_yrk_plot), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_lew_plot), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_yrk_plot), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_lew_plot), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_yrk_plot), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_lew_plot), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_yrk_plot), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_lew_len), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_yrk_len), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_lew_len), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_yrk_len), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_lew_len), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_yrk_len), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_len_lew_plot), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_len_yrk_plot), location = ph_location_type(type = "body")) %>% add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_plotfinalTlen), location = ph_location_type(type = "body")) 
print(doc, target = "results/figures/GO_compare_sel.pptx")

doc <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_plotfinalTsurv_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_plotfinalTsurv_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_plotfinalTsurv_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_plotfinalTsurv_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_plotfinalTlen_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_plotfinalTlen_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_plotfinalTlen_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_plotfinalTlen_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_lew_plot_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_yrk_plot_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_lew_plot_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_yrk_plot_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_lew_plot_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_yrk_plot_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_lew_plot_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_yrk_plot_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_lew_len_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitGF_yrk_len_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_lew_len_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_yrk_len_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_lew_len_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitRDA_yrk_len_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_len_lew_plot_nosel), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitEnvDist_len_yrk_plot_nosel), location = ph_location_type(type = "body")) %>% add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = goFitLEA_plotfinalTsurv_nosel), location = ph_location_type(type = "body")) 
print(doc, target = "results/figures/GO_compare_nosel.pptx")

docx <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitLEA_plotfinalTsurv, location = ph_location_type(type = "body"), scale = 1.5) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitRDA_plotfinalTsurv, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitGF_plotfinalTsurv, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitEnvDist_plotfinalTsurv, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitLEA_plotfinalTlen, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitRDA_plotfinalTlen, location = ph_location_type(type = "body"), scale = 1.5) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitGF_plotfinalTlen, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitEnvDist_plotfinalTlen, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitGF_lew_plot, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitGF_yrk_plot, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitLEA_lew_plot, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitLEA_yrk_plot, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitRDA_lew_plot, location = ph_location_type(type = "body"), scale = 1.5) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitRDA_yrk_plot, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitEnvDist_lew_plot, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitEnvDist_yrk_plot, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitGF_lew_len, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitGF_yrk_len, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitLEA_lew_len, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitLEA_yrk_len, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitRDA_lew_len, location = ph_location_type(type = "body"), scale = 1.5) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitRDA_yrk_len, location = ph_location_type(type = "body"), scale = 1.5) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitEnvDist_len_lew_plot, location = ph_location_type(type = "body"), scale = 1.5) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with.gg(value = goFitEnvDist_len_yrk_plot, location = ph_location_type(type = "body"), scale = 1.5)
print(docx, target = "results/figures/GO_compare_gg.pptx")
```

## Compare Offset Methods

Now that I have predictions from different methods, I will look at how offset predictions made by these methods compare to each other and how different their ability is to predict experimental fitness.


### Compare Offsets to Each Other

```{r Offset Comparison}
# create a correlation matrix for offsets
go_forcorrs <- fullfit %>% distinct(site_name, bag_site, event, .keep_all = TRUE) %>% filter(event == 4) %>% dplyr::select(site_name, pop, bag_site, offset_rda_lew, offset_rda_yrk, order_lew_rda, order_yrk_rda, offset_gf_geno_lew, offset_gf_geno_yrk, order_lew_geno_gf, order_yrk_geno_gf, offset_lea_go_lew, offset_lea_go_yrk, order_lew_lea, order_yrk_lea, env_dist_lew, env_dist_yrk, order_lew_envdist, order_yrk_envdist)

# check correlations
library(ggcorrplot)
cor_go_raw <- cor(go_forcorrs[,c("offset_rda_lew", "offset_gf_geno_lew", "offset_lea_go_lew", "env_dist_lew", "offset_rda_yrk", "offset_gf_geno_yrk", "offset_lea_go_yrk", "env_dist_yrk")])
cor_go_raw_pmat <- cor_pmat(go_forcorrs[,c("offset_rda_lew", "offset_gf_geno_lew", "offset_lea_go_lew", "env_dist_lew", "offset_rda_yrk", "offset_gf_geno_yrk", "offset_lea_go_yrk", "env_dist_yrk")])

colnames(cor_go_raw) <- rownames(cor_go_raw) <- c("RDA offset: Lewisetta", "GF offset: Lewisetta", "LEA offset: Lewisetta", "Environmental Distance: Lewisetta", "RDA offset: York River", "GF offset: York River", "LEA offset: York River", "Environmental Distance: York River")

# plot these correlations
# raw go
corrs_go_raw_all <- ggcorrplot(cor_go_raw, 
                               p.mat = cor_go_raw_pmat,
                               insig = "blank",
                               lab = T,
                               colors = c("#6D9EC1", "white", "#E46726")) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Genomic Offset Correlations",
       x = NULL,
       y = NULL)
corrs_go_raw_all
```

Look also at the difference between go models with and without biotic variables.

```{r}
# create a correlation matrix for offsets
go_forcorrs_bio <- fullfit %>% distinct(site_name, bag_site, event, .keep_all = TRUE) %>% filter(event == 4) %>% dplyr::select(site_name, pop, bag_site, offset_rda_lew, offset_rda_yrk, offset_gf_geno_lew, offset_gf_geno_yrk, offset_lea_go_lew, offset_lea_go_yrk, env_dist_lew, env_dist_yrk, offset_rda_lew_nb, offset_rda_yrk_nb, offset_gf_geno_lew_nb, offset_gf_geno_yrk_nb, offset_lea_go_lew_nb, offset_lea_go_yrk_nb, env_dist_lew_nb, env_dist_yrk_nb)
  
# check correlations
cor_go_raw_bio <- cor(go_forcorrs_bio[,c("offset_rda_lew", "offset_gf_geno_lew", "offset_lea_go_lew", "env_dist_lew", "offset_rda_yrk", "offset_gf_geno_yrk", "offset_lea_go_yrk", "env_dist_yrk", "offset_rda_lew_nb", "offset_gf_geno_lew_nb", "offset_lea_go_lew_nb", "env_dist_lew_nb", "offset_rda_yrk_nb", "offset_gf_geno_yrk_nb", "offset_lea_go_yrk_nb", "env_dist_yrk_nb")])

cor_go_raw_pmat_bio <- cor_pmat(go_forcorrs_bio[,c("offset_rda_lew", "offset_gf_geno_lew", "offset_lea_go_lew", "env_dist_lew", "offset_rda_yrk", "offset_gf_geno_yrk", "offset_lea_go_yrk", "env_dist_yrk", "offset_rda_lew_nb", "offset_gf_geno_lew_nb", "offset_lea_go_lew_nb", "env_dist_lew_nb", "offset_rda_yrk_nb", "offset_gf_geno_yrk_nb", "offset_lea_go_yrk_nb", "env_dist_yrk_nb")])


# plot these correlations
# raw go
corrs_go_raw_all_bio <- ggcorrplot(cor_go_raw_bio, 
                               p.mat = cor_go_raw_pmat_bio,
                               insig = "blank",
                               lab = T,
                               colors = c("#6D9EC1", "white", "#E46726")) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Genomic Offset Correlations",
       x = NULL,
       y = NULL)
corrs_go_raw_all_bio

# how different were the offset predictions?
lew_final <- fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]
yrk_final <- fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]

diff_lea_lew <- mean((lew_final$offset_lea_go_lew_nb - lew_final$offset_lea_go_lew) / lew_final$offset_lea_go_lew)
diff_lea_yrk <- mean((yrk_final$offset_lea_go_yrk_nb - yrk_final$offset_lea_go_yrk) / yrk_final$offset_lea_go_yrk)

diff_gf_lew <- mean((lew_final$offset_gf_geno_lew_nb - lew_final$offset_gf_geno_lew) / lew_final$offset_gf_geno_lew)
diff_gf_yrk <- mean((yrk_final$offset_gf_geno_yrk_nb - yrk_final$offset_gf_geno_yrk) / yrk_final$offset_gf_geno_yrk)

diff_rda_lew <- mean((lew_final$offset_rda_lew_nb - lew_final$offset_rda_lew) / lew_final$offset_rda_lew)
diff_rda_yrk <-  mean((yrk_final$offset_rda_yrk_nb - yrk_final$offset_rda_yrk) / yrk_final$offset_rda_yrk)

diff_envdist_lew <- mean((lew_final$env_dist_lew_nb - lew_final$env_dist_lew) / lew_final$env_dist_lew)
diff_envdist_yrk <- mean((yrk_final$env_dist_yrk_nb - yrk_final$env_dist_yrk) / yrk_final$env_dist_yrk)

offset_diffs_bio <- data.frame(method = c("lea","lea","gf","gf","rda","rda","envdist","envdist"), bag_site = rep(c("Lewisetta","YorkRiver"),4), diff_percent = c(diff_lea_lew, diff_lea_yrk, diff_gf_lew, diff_gf_yrk, diff_rda_lew, diff_rda_yrk, diff_envdist_lew, diff_envdist_yrk))

# how different were the offset performances? 
final_corrs_b <- corrs_df[corrs_df$event == 4 & corrs_df$biotic == "b", c("method","bag_site","kendall_surv", "p_adj_surv","signif_surv_adj","kendall_len", "p_adj_len", "signif_len_adj")]
final_corrs_nb <- corrs_df[corrs_df$event == 4 & corrs_df$biotic == "nb", c("method","bag_site","kendall_surv", "p_adj_surv","signif_surv_adj","kendall_len", "p_adj_len", "signif_len_adj")]
colnames(final_corrs_nb) <- c("method","bag_site","kendall_surv_nb", "p_adj_surv_nb","signif_surv_adj_nb","kendall_len_nb", "p_adj_len_nb", "signif_len_adj_nb")
final_corrs <- merge(final_corrs_b, final_corrs_nb, by = c("method", "bag_site"))

final_corrs$diff_surv <- final_corrs$kendall_surv - final_corrs$kendall_surv_nb
final_corrs$diff_len <- final_corrs$kendall_len - final_corrs$kendall_len_nb

write.csv(final_corrs, "results/GO_results/final_correlation_GOfit.csv", row.names = F)

# plot performance differences
disease_vs_no_performance <- ggplot(data = final_corrs) +
  geom_point(data = final_corrs, aes(x = kendall_surv_nb, y = kendall_surv, fill = method, shape = bag_site), size = 4) +
  scale_shape_manual(values = c("Lewisetta" = 21, "YorkRiver" = 24)) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  ylim(-1,0.05)+
  xlim(-1,0.05)+
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Genomic Offset Performance (Survival) with and without \nthe Biotic Landscape", x = "Genomic Offset (Environment Only)", y = "Genomic Offset (Disease & Environment)") +
  theme_classic()
disease_vs_no_performance

disease_vs_no_len <- ggplot(data = final_corrs) +
  geom_point(data = final_corrs, aes(x = kendall_len_nb, y = kendall_len, fill = method, shape = bag_site), size = 4) +
  scale_shape_manual(values = c("Lewisetta" = 21, "YorkRiver" = 24)) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  ylim(-1,0.05)+
  xlim(-1,0.05)+
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Genomic Offset Performance (Length) with and without \nthe Biotic Landscape", x = "Genomic Offset (Environment Only)", y = "Genomic Offset (Disease & Environment)") +
  theme_classic()
disease_vs_no_len
```


## Plotting Fitness vs Environmental Metrics

```{r}
fullfit
#env_exp <- read.csv("data/EnvDat/exp/exp_scaled_2025-12-08.csv")[,-1]
fullfit_env <- merge(fullfit[fullfit$event == 4,], env_exp, by = "site_name")
fullfit_env_nodup_surv <- fullfit_env[!(duplicated(fullfit_env$surv)),]
fullfit_env_nodup_len <- fullfit_env[!(duplicated(fullfit_env$length)),]

# survival
ggplot(data = fullfit_env_nodup_surv, aes(x = temp_quantile_10, y = surv)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Survival as a Function of Temperature", x = "Temperature (10% Quantile)", y = "Survival") +
  theme_classic()

ggplot(data = fullfit_env_nodup_surv, aes(x = temp_quantile_90, y = surv)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Survival as a Function of Temperature", x = "Temperature (90% Quantile)", y = "Survival") +
  theme_classic()

ggplot(data = fullfit_env_nodup_surv, aes(x = salinity_quantile_10, y = surv)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Survival as a Function of Salinity", x = "Salinity (10% Quantile)", y = "Survival") +
  theme_classic()

ggplot(data = fullfit_env_nodup_surv, aes(x = salinity_quantile_90, y = surv)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Survival as a Function of Salinity", x = "Salinity (90% Quantile)", y = "Survival") +
  theme_classic()

ggplot(data = fullfit_env_nodup_surv, aes(x = Dermo_Prevalence, y = surv)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Survival as a Function of Disease (Dermo)", x = "Dermo Prevalence", y = "Survival") +
  theme_classic()

ggplot(data = fullfit_env_nodup_surv, aes(x = MSX_Prevalence, y = surv)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Survival as a Function of Disease (MSX)", x = "MSX Prevalence", y = "Survival") +
  theme_classic()

# now length
ggplot(data = fullfit_env_nodup_len, aes(x = temp_quantile_10, y = length)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Length as a Function of Temperature", x = "Temperature (10% Quantile)", y = "Survival") +
  theme_classic()

ggplot(data = fullfit_env_nodup_len, aes(x = temp_quantile_90, y = length)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Length as a Function of Temperature", x = "Temperature (90% Quantile)", y = "Survival") +
  theme_classic()

ggplot(data = fullfit_env_nodup_len, aes(x = salinity_quantile_10, y = length)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Length as a Function of Salinity", x = "Salinity (10% Quantile)", y = "Survival") +
  theme_classic()

ggplot(data = fullfit_env_nodup_len, aes(x = salinity_quantile_90, y = length)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Length as a Function of Salinity", x = "Salinity (90% Quantile)", y = "Survival") +
  theme_classic()

ggplot(data = fullfit_env_nodup_len, aes(x = Dermo_Prevalence, y = length)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Length as a Function of Disease (Dermo)", x = "Dermo Prevalence", y = "Survival") +
  theme_classic()

ggplot(data = fullfit_env_nodup_len, aes(x = MSX_Prevalence, y = length)) +
  geom_point(shape = 21, size = 4, fill = "lightblue2") +
  geom_smooth(method = "lm", col = "steelblue3") +
  labs(title = "Length as a Function of Disease (MSX)", x = "MSX Prevalence", y = "Survival") +
  theme_classic()
```

