---
title: "CviMVP_oyster_compareGOexp"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "~/Documents/GitHub/MVP_oyster_GO")
```

# Compare Predicted GO to Experimental Results

This script is for analyzing how well various genomic offset methods predict experimentally obtained fitness. I compare LEA, GF, and RDA offsets to experimental mortality, length, and condition index at the conclusion of the experiment. 

## Setup

```{r}
# setwd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(dplyr)
library(tidyr)
library(stringr)
library(vegan)
library(ggplot2)
library(plotrix)

# data
# offset predictions
lea_go <- read.csv("results/GO_results/offsets_LEA_expSites_2025-08-27.csv")[,-1]
rda_go <- read.csv("results/GO_results/offsets_RDA_expsites_2025-08-20.csv")[,-1]
gf_go <- read.csv("results/GO_results/offsets_GF_expSites_2025-08-27.csv")[,-1]

# fitness metrics
mort_bags <- read.csv("results/exp_results/mortality_long_bags_2025-06-06.csv")[,-1]
mort <- read.csv("results/exp_results/mortality_long_popsite_2025-06-06.csv")[,-1]
len <- read.csv("results/exp_results/lengths_bags_event_2025-06-03.csv")[,-1]
```

## Data Prep

Prepare data from offsets & experiment to be directly compared.

```{r}
# inspect data
lea_go
gf_go
rda_go
mort
len

# trim mort
mort_popsite <- mort[!(mort$event == 0), c("bag_site","pop","mean_surv","sd","event")]
colnames(mort_popsite) <- c("bag_site","pop","mean_surv","sd_surv","event")

# also get fitness rankings at each time
# mortality
ord_mort_popsite <- data.frame()
for (i in 1:4) {
  ordt_lew <- mort_popsite[mort_popsite$event == i & mort_popsite$bag_site == "Lewisetta",]
  ordt_lew <- ordt_lew %>% arrange(desc(mean_surv))
  ordt_lew$order_mort <- 1:length(ordt_lew$pop)
  ordt_yrk <- mort_popsite[mort_popsite$event == i & mort_popsite$bag_site == "YorkRiver",]
  ordt_yrk <- ordt_yrk %>% arrange(desc(mean_surv))
  ordt_yrk$order_mort <- 1:length(ordt_yrk$pop)
  ord_mort_popsite <- rbind(ord_mort_popsite, ordt_lew, ordt_yrk)
}
ord_mort_popsite

# do the same with length
# first length needs a pop label and site label
len_bags <- as.data.frame(cbind(len$bag, str_split_fixed(len$bag, "-", 3)[,2:3]))
colnames(len_bags) <- c("bag","pop","bagnum")
len_bags$bagnum <- as.numeric(len_bags$bagnum)
len_bags$site <- ifelse(len_bags$bagnum > 3, "YorkRiver", "Lewisetta")
len_all <- merge(len_bags[,c("bag","pop","site")], len, by = c("bag"))

len_t1 <- len_all[len_all$monitoring_event == 1,]
t1_lens <- aggregate(length ~ pop + site, data = len_t1, FUN = mean)
t1_sd <- aggregate(length ~ pop + site, data = len_t1, FUN = sd)
colnames(t1_sd) <- c("pop", "site", "len_sd")
t1_lens_popsite <- merge(t1_lens, t1_sd, by = c("pop", "site"))
t1_lens_popsite$monitoring_event <- 1
len_t2 <- len_all[len_all$monitoring_event == 2,]
t2_lens <- aggregate(length ~ pop + site, data = len_t2, FUN = mean)
t2_sd <- aggregate(length ~ pop + site, data = len_t2, FUN = sd)
colnames(t2_sd) <- c("pop", "site", "len_sd")
t2_lens_popsite <- merge(t2_lens, t2_sd, by = c("pop", "site"))
t2_lens_popsite$monitoring_event <- 2
len_t3 <- len_all[len_all$monitoring_event == 3,]
t3_lens <- aggregate(length ~ pop + site, data = len_t3, FUN = mean)
t3_sd <- aggregate(length ~ pop + site, data = len_t3, FUN = sd)
colnames(t3_sd) <- c("pop", "site", "len_sd")
t3_lens_popsite <- merge(t3_lens, t3_sd, by = c("pop", "site"))
t3_lens_popsite$monitoring_event <- 3
len_t4 <- len_all[len_all$monitoring_event == 4,]
t4_lens <- aggregate(length ~ pop + site, data = len_t4, FUN = mean)
t4_sd <- aggregate(length ~ pop + site, data = len_t4, FUN = sd)
colnames(t4_sd) <- c("pop", "site", "len_sd")
t4_lens_popsite <- merge(t4_lens, t4_sd, by = c("pop", "site"))
t4_lens_popsite$monitoring_event <- 4

lens_popsite <- rbind(t1_lens_popsite, t2_lens_popsite, t3_lens_popsite, t4_lens_popsite)
colnames(lens_popsite) <- c("pop","bag_site","mean_len","sd_len","event")

# and orders
ord_len_popsite <- data.frame()
for (i in 1:4) {
  ordt_lew <- lens_popsite[lens_popsite$event == i & lens_popsite$bag_site == "Lewisetta",]
  ordt_lew <- ordt_lew %>% arrange(desc(mean_len))
  ordt_lew$order_len <- 1:length(ordt_lew$pop)
  ordt_yrk <- lens_popsite[lens_popsite$event == i & lens_popsite$bag_site == "YorkRiver",]
  ordt_yrk <- ordt_yrk %>% arrange(desc(mean_len))
  ordt_yrk$order_len <- 1:length(ordt_yrk$pop)
  ord_len_popsite <- rbind(ord_len_popsite, ordt_lew, ordt_yrk)
}
ord_len_popsite

# put length and mortality together
ord_fitness_popsite <- merge(ord_mort_popsite, ord_len_popsite, by = c("pop","bag_site","event"))

# need a site name column
ord_fitness_popsite$site_name <- ord_fitness_popsite$pop
map = c("TX"="W1_TX", "LA"="W2_LA", "FL"="W3_FL", "LOLA"="S1_LOLA", "DEBY"="S2_DEBY", "JR"="W4_VA", "NH"="W5_NH", "ME"="W6_ME")
ord_fitness_popsite$site_name <- recode(ord_fitness_popsite$site_name,
                                           "TX"="W1_TX", "LA"="W2_LA", "FL"="W3_FL",
                                           "LOLA"="S1_LOLA", "DEBY"="S2_DEBY", "JR"="W4_VA",
                                           "NH"="W5_NH", "ME"="W6_ME")


# do bag level data as well
colnames(mort_bags) <- c("bag","event","surv")
colnames(len) <- c("bag","event","length")
bagfit <- merge(mort_bags, len, by = c("bag","event"))

# pop info
bagfit$pop <- str_split_fixed(bagfit$bag, "-", 3)[,2]
bagfit

# add this to other fitness data
bagfullfit <- merge(ord_fitness_popsite, bagfit, by = c("pop","event"))

# need to align naming conventions to merge
# LEA
# colnames
colnames(lea_go) <- c("site_name","lew_go_lea","yrk_go_lea","lew_gothin_lea","yrk_gothin_lea")

# add order
order_lew_lea <- as.data.frame(lea_go[order(lea_go$lew_go_lea, decreasing = T),]$site_name)
colnames(order_lew_lea) <- c("site_name")
order_lew_lea$order_lew_lea <- 1:length(order_lew_lea$site_name)
order_yrk_lea <- as.data.frame(lea_go[order(lea_go$yrk_go_lea, decreasing = T),]$site_name)
colnames(order_yrk_lea) <- c("site_name")
order_yrk_lea$order_yrk_lea <- 1:length(order_yrk_lea$site_name)
orders_lea <- merge(order_lew_lea, order_yrk_lea, by = c("site_name"))
lea_go1 <- merge(lea_go, orders_lea, by = c("site_name"))

# GF
# colnames are fine
# add order
order_lew_af <- as.data.frame(gf_go[order(gf_go$offset_gf_af_lew, decreasing = T),]$site_name)
colnames(order_lew_af) <- c("site_name")
order_lew_af$order_lew_af_gf <- 1:length(order_lew_af$site_name)
order_yrk_af <- as.data.frame(lea_go[order(gf_go$offset_gf_af_yrk, decreasing = T),]$site_name)
colnames(order_yrk_af) <- c("site_name")
order_yrk_af$order_yrk_af_gf <- 1:length(order_yrk_af$site_name)
orders_gf <- merge(order_lew_af, order_yrk_af, by = c("site_name"))
gf_go1 <- merge(gf_go, orders_gf, by = c("site_name"))

# RDA
# colnames
colnames(rda_go) <- c("site_name","lat","lon","offset_lew_rda","offset_yrk_rda")

# add order
order_lew_rda <- as.data.frame(rda_go[order(rda_go$offset_lew, decreasing = T),]$site_name)
colnames(order_lew_rda) <- c("site_name")
order_lew_rda$order_lew_rda <- 1:length(order_lew_af$site_name)
order_yrk_rda <- as.data.frame(rda_go[order(rda_go$offset_yrk, decreasing = T),]$site_name)
colnames(order_yrk_rda) <- c("site_name")
order_yrk_rda$order_yrk_rda <- 1:length(order_yrk_rda$site_name)
orders_rda <- merge(order_lew_rda, order_yrk_rda, by = c("site_name"))
rda_go1 <- merge(rda_go, orders_rda, by = c("site_name"))

# put together full go & fitness df
go <- merge(rda_go1, gf_go1, by = "site_name")
go1 <- merge(go, lea_go1, by = "site_name")
fullfit <- merge(bagfullfit, go1, by = "site_name")

# # merge gos & fitness dat
# goFitLEA <- merge(mort_popsite, lea_go, by = "pop")
# goFitGF <- merge(mort_popsite, gf_go, by = "pop")

# # associate fitness data at the bag level as well
# levels(as.factor(mort$pop))
# mort[mort$pop == "JR",]$pop <- "VA"
# goFitLEAbag <- merge(mort, lea_go, by = "pop")
# goFitGFbag <- merge(mort, gf_go, by = "pop")
```


## Compare Offsets to Fitness

Now that offsets and fitness are in a comparable format, look at how well the two match.

### LEA Offset

Start with the bag level mortality data.

```{r}
# correlations between go/mort at each time point
#################################################
# lewisetta
# lea
(cor_lea1_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$lew_go_lea, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$surv)) # -0.1188
(cor_lea2_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$lew_go_lea, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$surv)) # -0.114408
(cor_lea3_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$lew_go_lea, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$surv)) # 0.1155335
(cor_lea4_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$lew_go_lea, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$surv)) # 0.1225753

# gf
(cor_gf1_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$offset_gf_af_lew, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$surv)) # -0.1320573
(cor_gf2_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$offset_gf_af_lew, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$surv)) # -0.1853239
(cor_gf3_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$offset_gf_af_lew, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$surv)) # -0.1934671
(cor_gf4_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$offset_gf_af_lew, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$surv)) # -0.1759728

# rda
(cor_rda1_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$offset_lew_rda, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$surv)) # -0.4708852
(cor_rda2_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$offset_lew_rda, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$surv)) # -0.5056913
(cor_rda3_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$offset_lew_rda, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$surv)) # -0.515159
(cor_rda4_lewm <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$offset_lew_rda, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$surv)) # -0.4586319

# correlations between go/york mort at each time point
# lea
(cor_lea1_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$yrk_go_lea, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$surv)) # 0.09590167
(cor_lea2_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$yrk_go_lea, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$surv)) # 0.08713282
(cor_lea3_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$yrk_go_lea, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$surv)) # -0.118059
(cor_lea4_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$yrk_go_lea, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$surv)) # -0.1314632

# gf
(cor_gf1_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$offset_gf_af_yrk, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$surv)) # -0.08401281
(cor_gf2_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$offset_gf_af_yrk, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$surv)) # -0.1204811
(cor_gf3_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$offset_gf_af_yrk, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$surv)) # -0.2840118
(cor_gf4_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$offset_gf_af_yrk, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$surv)) # -0.2594375

# rda
(cor_rda1_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$offset_yrk_rda, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$surv)) # 0.4217415
(cor_rda2_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$offset_yrk_rda, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$surv)) # 0.4692211
(cor_rda3_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$offset_yrk_rda, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$surv)) # 0.2952187
(cor_rda4_yrkm <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$offset_yrk_rda, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$surv)) # 0.2752647
#################################################

# correlations between go/len at each time point
################################################
# lewisetta
# lea
(cor_lea1_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$lew_go_lea, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$length)) # -0.1602686
(cor_lea2_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$lew_go_lea, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$length)) # 0.08713691
(cor_lea3_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$lew_go_lea, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$length)) # 0.1573283
(cor_lea4_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$lew_go_lea, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$length)) # 0.1036748

# gf
(cor_gf1_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$offset_gf_af_lew, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$length)) # 0.205473
(cor_gf2_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$offset_gf_af_lew, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$length)) # 0.1044639
(cor_gf3_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$offset_gf_af_lew, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$length)) # -0.1849938
(cor_gf4_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$offset_gf_af_lew, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$length)) # -0.129079

# rda
(cor_rda1_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$offset_lew_rda, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$length)) # -0.2086341
(cor_rda2_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$offset_lew_rda, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,]$length)) # -0.3727353
(cor_rda3_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$offset_lew_rda, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,]$length)) # -0.2132694
(cor_rda4_lewl <- cor(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$offset_lew_rda, fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]$length)) # -0.009785706

# correlations between go/york mort at each time point
# lea
(cor_lea1_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$yrk_go_lea, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$length)) # 0.1418931
(cor_lea2_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$yrk_go_lea, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$length)) # -0.1030296
(cor_lea3_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$yrk_go_lea, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$length)) # -0.1687404
(cor_lea4_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$yrk_go_lea, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$length)) # -0.155277

# gf
(cor_gf1_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$offset_gf_af_yrk, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$length)) # 0.2664307
(cor_gf2_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$offset_gf_af_yrk, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$length)) # -0.005909796
(cor_gf3_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$offset_gf_af_yrk, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$length)) # -0.1869383
(cor_gf4_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$offset_gf_af_yrk, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$length)) # -0.02829487

# rda
(cor_rda1_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$offset_yrk_rda, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,]$length)) # 0.2969529
(cor_rda2_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$offset_yrk_rda, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,]$length)) # 0.2129072
(cor_rda3_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$offset_yrk_rda, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,]$length)) # 0.2044589
(cor_rda4_yrkl <- cor(fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$offset_yrk_rda, fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]$length)) # 0.306211
################################################

# compile
#########



# put this all into a df with method/time point/fitness metric/site/corr
corr_df <- data.frame(method = rep(c(rep("LEA", 8), rep("GF", 8), rep("RDA", 8)), 2), metric = c(rep("surv", 24), rep("len", 24)), monitoring_event = rep(c(1:4), 12), bag_site = rep(c(rep("Lewisetta", 4), rep("YorkRiver", 4)), 6), pearson = c(cor_lea1_lewm, cor_lea2_lewm, cor_lea3_lewm, cor_lea4_lewm, cor_lea1_yrkm, cor_lea2_yrkm, cor_lea3_yrkm, cor_lea4_yrkm, cor_gf1_lewm, cor_gf2_lewm, cor_gf3_lewm, cor_gf4_lewm, cor_gf1_yrkm, cor_gf2_yrkm, cor_gf3_yrkm, cor_gf4_yrkm, cor_rda1_lewm, cor_rda2_lewm, cor_rda3_lewm, cor_rda4_lewm, cor_rda1_yrkm, cor_rda2_yrkm, cor_rda3_yrkm, cor_rda4_yrkm, cor_lea1_lewl, cor_lea2_lewl, cor_lea3_lewl, cor_lea4_lewl, cor_lea1_yrkl, cor_lea2_yrkl, cor_lea3_yrkl, cor_lea4_yrkl, cor_gf1_lewl, cor_gf2_lewl, cor_gf3_lewl, cor_gf4_lewl, cor_gf1_yrkl, cor_gf2_yrkl, cor_gf3_yrkl, cor_gf4_yrkl, cor_rda1_lewl, cor_rda2_lewl, cor_rda3_lewl, cor_rda4_lewl, cor_rda1_yrkl, cor_rda2_yrkl, cor_rda3_yrkl, cor_rda4_yrkl))

# save this
write.csv(corr_df, paste0("results/GO_results/go_fit_corrs_",Sys.Date(),".csv"), row.names = F)
#########
```


Plot these data sets against each other

```{r}
# plot lea vs mort at each site/time point
goFitLEA_lew_plotbag <- ggplot(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",]) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t1, y = offset_lew, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t1, y = offset_lew), col = "cadetblue3", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t2, y = offset_lew, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t2, y = offset_lew), col = "cornflowerblue", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t3, y = offset_lew, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t3, y = offset_lew), col = "slateblue", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t4, y = offset_lew, col = "Spring'25")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t4, y = offset_lew), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at Lewisetta", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_lew_plotbag
ggsave(paste0("results/figures/lea_lew_compareBag_",Sys.Date(),".png"))

goFitLEA_yrk_plot <- ggplot(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",]) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t1, y = offset_yrk, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t1, y = offset_yrk), col = "cadetblue3", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t2, y = offset_yrk, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t2, y = offset_yrk), col = "cornflowerblue", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t3, y = offset_yrk, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t3, y = offset_yrk), col = "slateblue", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t4, y = offset_yrk, col = "Spring'25")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t4, y = offset_yrk), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at York River", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_plot
ggsave(paste0("results/figures/lea_york_compareBag_",Sys.Date(),".png"))

# plot go lea vs mort at both sites just final time point
goFitLEA_plotT4bag <- ggplot(data = goFitLEAbag) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = offset_lew, y = t4, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = offset_lew, y = t4), col = "mediumspringgreen", se = T) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = offset_yrk, y = t4, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = offset_yrk, y = t4), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (LEA) vs Mortality at Final Time Point", x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotT4bag
ggsave(paste0("results/figures/lea_yorklew_compareBag_",Sys.Date(),".png"))
```

Now do it at the population level!

Start with the population level mortality data.

```{r}
# correlations between go/lew mort at each time point
# lea
(cor_lea1_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t1, method = "pearson")) # -0.05736393, r2 0.00329062
(cor_lea2_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t2, method = "pearson")) # 0.01456035, r2 0.0002120037
(cor_lea3_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t3, method = "pearson")) # -0.08913129, r2 0.007944386
(cor_lea4_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t4, method = "pearson")) # -0.135897, r2 0.01846799

(Kcor_lea1_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t1, method = "kendall")) # -0.05736393
(Kcor_lea2_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t2, method = "kendall")) # 0.01456035
(Kcor_lea3_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t3, method = "kendal")) # -0.08913129
(Kcor_lea4_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t4, method = "kendall")) # -0.135897

# correlations between go/york mort at each time point
# lea
(cor_lea1_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t1, method = "pearson")) # 0.0400319, r2 0.001602553
(cor_lea2_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t2, method = "pearson")) # -0.01014261, r2 0.0001028726
(cor_lea3_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t3, method = "pearson")) # -0.2889805, r2 0.0835097
(cor_lea4_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t4, method = "pearson")) # -0.3244636, r2 0.1052767

summary(lm(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk ~ goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t4))$r.squared

(KenCor_lea1_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t1, method = "kendall")) # 0.07142857
(KenCor_lea2_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t2, method = "kendall")) # 0.07142857
(KenCor_lea3_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t3, method = "kendall")) # -0.2142857
(KenCor_lea4_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t4, method = "kendall")) # 0

# put this all into a df with method/time point/fitness metric/site/corr
lea_corr_df_pop <- data.frame(method = rep("LEA", 8), metric = rep("surv", 8), monitoring_event = rep(c(1:4),2), bag_site = c(rep("Lewisetta",4 ), rep("YorkRiver", 4)), pearson = c(cor_lea1_lew_pop, cor_lea2_lew_pop, cor_lea3_lew_pop, cor_lea4_lew_pop, cor_lea1_yrk_pop, cor_lea2_yrk_pop, cor_lea3_yrk_pop, cor_lea4_yrk_pop))
```


Plot these data sets against each other

```{r}
# plot lea vs mort at each site/time point
goFitLEA_lew_plot <- ggplot(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",]) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = offset_lew, y = mean_t1, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = offset_lew, y = mean_t1), col = "cadetblue3", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = offset_lew, y = mean_t2, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = offset_lew, y = mean_t2), col = "cornflowerblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = offset_lew, y = mean_t3, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = offset_lew, y = mean_t3), col = "slateblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = offset_lew, y = mean_t4, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = offset_lew, y = mean_t4), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at Lewisetta", x = "Survival", y = "LEA Offset") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23" = "cadetblue3", "Spring'24" = "cornflowerblue", "Fall'24" = "slateblue", "Spring'25" = "purple4"))+
  ylim(0,1) +
  theme_classic()
goFitLEA_lew_plot
ggsave(paste0("results/figures/lea_lew_compare_",Sys.Date(),".png"))

goFitLEA_yrk_plot <- ggplot(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",]) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t1, y = offset_yrk, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t1, y = offset_yrk), col = "cadetblue3", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t2, y = offset_yrk, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t2, y = offset_yrk), col = "cornflowerblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t3, y = offset_yrk, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t3, y = offset_yrk), col = "slateblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_yrk, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_yrk), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at York River", x = "Survival", y = "LEA Offset") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_plot
ggsave(paste0("results/figures/lea_york_compare_",Sys.Date(),".png"))

# plot go lea vs mort at both sites just final time point
goFitLEA_plotT4 <- ggplot(data = goFitLEA) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = offset_lew, y = mean_t4, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = offset_lew, y = mean_t4), col = "mediumspringgreen", se = T) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = offset_yrk, y = mean_t4, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = offset_yrk, y = mean_t4), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (LEA) vs Mortality at Final Time Point", x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotT4
```


### GF Offset

```{r}
# correlations between go/lew mort at each time point
(cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_gf_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t1)) # -0.1931013, r2 0.003633104
(cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_gf_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t2)) # -0.006177896, r2 0.002257078
(cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_gf_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t3)) # 0.04885283, r2 0.003688538
(cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_gf_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t4)) # 0.04589687, r2 0.002099499

summary(lm(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_gf_af_lew ~ goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t1))$r.squared
summary(lm(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_gf_af_lew ~ goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t2))$r.squared
summary(lm(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_gf_af_lew ~ goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t3))$r.squared
summary(lm(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_gf_af_lew ~ goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t4))$r.squared

# correlations between go/yrk mort at each time point
(cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t1)) # 0.2852591
(cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t2)) # 0.318425
(cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t3)) # 0.241053
(cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t4)) # 0.2847928

# really bad correlations, may need to double check offset calculations
# pearson of orders - LEW
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t1) # 0.003129999
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t2) # 0.03004949
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t3) # -0.03736819
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t4) # -0.107596
# pearson of orders - YRK
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t1) # 0.1687206
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t2) # 0.1059771
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t3) # -0.214745
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t4) # -0.2206604
```

Plot these data against each other.

```{r}
# plot lea vs mort at each site/time point
goFitGF_lew_plot <- ggplot(data = goFitGF[goFitGF$bag_site == "Lewisetta",]) +
  geom_point(data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t1, y = offset_af_lew, col = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t1, y = offset_af_lew), fill = "cadetblue3", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t2, y = offset_af_lew, col = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t2, y = offset_af_lew), fill = "cornflowerblue", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t3, y = offset_af_lew, col = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t3, y = offset_af_lew), fill = "slateblue", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_af_lew, col = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_af_lew), fill = "purple4", se = F) +
  labs(title = "Predicted GO (GFaf) vs Mortality at Lewisetta", x = "Survival", y = "GFaf Offset") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_lew_plot
ggsave(paste0("results/figures/gf_lew_compare_",Sys.Date(),".png"))

goFitGF_yrk_plot <- ggplot(data = goFitGF[goFitGF$bag_site == "YorkRiver",]) +
  geom_point(data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t1, y = offset_af_yrk, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t1, y = offset_af_yrk), col = "cadetblue3", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t2, y = offset_af_yrk, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t2, y = offset_af_yrk), col = "cornflowerblue", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t3, y = offset_af_yrk, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t3, y = offset_af_yrk), col = "slateblue", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_af_yrk, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_af_yrk), col = "purple4", se = F) +
  labs(title = "Predicted GO (GFaf) vs Mortality at York River", x = "Survival", y = "GFaf Offset") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_yrk_plot
ggsave(paste0("results/figures/gf_york_compare_",Sys.Date(),".png"))

# plot go lea vs mort at both sites just final time point
goFitGF_plotT4 <- ggplot(data = goFitGF) +
  geom_point(data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_af_lew, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_af_lew), col = "mediumspringgreen", se = T) +
  geom_point(data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_af_yrk, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_af_yrk), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (GFaf) vs Mortality at Final Time Point", x = "Survival", y = "GFaf Offset") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGF_plotT4
```

### Environmental Distance

Does environmental distance between home sites and common garden conditions predict mortality better than genomic offset does? 

```{r}
# read in environmental distance matrix (comes from "EnvDatProcess")
env_dist_mat <- read.csv("data/EnvDat/exp/distance_matrix_2025-06-12.csv")[,-c(1:2)]
rownames(env_dist_mat) <- colnames(env_dist_mat)
env_dist <- read.csv("data/EnvDat/exp/distance_matrix_longer_2025-06-12.csv")[-1]

# check env dist and mortality data
env_dist
mort

# keep just comparisons to experimental sites in dist df
env_dist_exp <- env_dist[env_dist$site1 %in% c("YRK", "LEW"),]
env_dist_exprm <- env_dist_exp[!(env_dist_exp$site2 %in% c("YRK", "LEW")), c("site2", "site1", "value")]
colnames(env_dist_exprm) <- c("site_name","bag_site","euc_dist")
env_dist_exprm$bag_site <- ifelse(env_dist_exprm$bag_site == "LEW", "Lewisetta", "YorkRiver")

# add popsdf
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME", "LARMIX", "SEEDMIX"), 
                     order = 1:10, 
                     cols = c("#3e4989" ,"#31688e", "#26828e", "#440154", "#482878", "#1f9e89", "#35b779", "#6ece58", "#fde725", "#b5de2b"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site_name = c("W1-TX", "W2-LA", "W3-FL", "W4-VA", "S1-LOLA", "S2-DEBY", "W5-NH", "W6-ME", "H1-LARMIX", "H2-SEEDMIX"))
mort_popsdf <- merge(mort, popsdf, by = "pop")

# merge
distMort_mg <- merge(mort_popsdf, env_dist_exprm, by = c("site_name", "bag_site"))

# compare
# lew
cor.test(distMort_mg[distMort_mg$bag_site == "Lewisetta",]$t1, distMort_mg[distMort_mg$bag_site == "Lewisetta",]$euc_dist) # -0.05480737
cor.test(distMort_mg[distMort_mg$bag_site == "Lewisetta",]$t2, distMort_mg[distMort_mg$bag_site == "Lewisetta",]$euc_dist) # -0.05106267
cor.test(distMort_mg[distMort_mg$bag_site == "Lewisetta",]$t3, distMort_mg[distMort_mg$bag_site == "Lewisetta",]$euc_dist) # -0.1423009
cor.test(distMort_mg[distMort_mg$bag_site == "Lewisetta",]$t4, distMort_mg[distMort_mg$bag_site == "Lewisetta",]$euc_dist) # -0.2023333

# yrk
cor.test(distMort_mg[distMort_mg$bag_site == "YorkRiver",]$t1, distMort_mg[distMort_mg$bag_site == "YorkRiver",]$euc_dist) # 0.1967869
cor.test(distMort_mg[distMort_mg$bag_site == "YorkRiver",]$t2, distMort_mg[distMort_mg$bag_site == "YorkRiver",]$euc_dist) # 0.2152596
cor.test(distMort_mg[distMort_mg$bag_site == "YorkRiver",]$t3, distMort_mg[distMort_mg$bag_site == "YorkRiver",]$euc_dist) # 0.1171927
cor.test(distMort_mg[distMort_mg$bag_site == "YorkRiver",]$t4, distMort_mg[distMort_mg$bag_site == "YorkRiver",]$euc_dist) # 0.1491704

summary(lm(distMort_mg[distMort_mg$bag_site == "YorkRiver",]$t4 ~ distMort_mg[distMort_mg$bag_site == "YorkRiver",]$euc_dist))

# visualize
# lewisetta
envdist_mort_plot_lew <- ggplot(data = distMort_mg[distMort_mg$bag_site == "Lewisetta",]) + 
  geom_point(data = distMort_mg[distMort_mg$bag_site == "Lewisetta",], aes(x = euc_dist, y = t1, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distMort_mg[distMort_mg$bag_site == "Lewisetta",], aes(x = euc_dist, y = t1), col = "cadetblue3", se = F) +
  geom_point(data = distMort_mg[distMort_mg$bag_site == "Lewisetta",], aes(x = euc_dist, y = t2, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distMort_mg[distMort_mg$bag_site == "Lewisetta",], aes(x = euc_dist, y = t2), col = "cornflowerblue", se = F) +
  geom_point(data = distMort_mg[distMort_mg$bag_site == "Lewisetta",], aes(x = euc_dist, y = t3, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distMort_mg[distMort_mg$bag_site == "Lewisetta",], aes(x = euc_dist, y = t3), col = "slateblue", se = F) +
  geom_point(data = distMort_mg[distMort_mg$bag_site == "Lewisetta",], aes(x = euc_dist, y = t4, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distMort_mg[distMort_mg$bag_site == "Lewisetta",], aes(x = euc_dist, y = t4), col = "purple4", se = F) +
  labs(title = "Environmental Distance vs Mortality at Lewisetta", x = "Euclidean Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  ylim(0,1)+
  theme_classic()
envdist_mort_plot_lew

# york river
envdist_mort_plot_yrk <- ggplot(data = distMort_mg[distMort_mg$bag_site == "YorkRiver",]) + 
  geom_point(data = distMort_mg[distMort_mg$bag_site == "YorkRiver",], aes(x = euc_dist, y = t1, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distMort_mg[distMort_mg$bag_site == "YorkRiver",], aes(x = euc_dist, y = t1), col = "cadetblue3", se = F) +
  geom_point(data = distMort_mg[distMort_mg$bag_site == "YorkRiver",], aes(x = euc_dist, y = t2, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distMort_mg[distMort_mg$bag_site == "YorkRiver",], aes(x = euc_dist, y = t2), col = "cornflowerblue", se = F) +
  geom_point(data = distMort_mg[distMort_mg$bag_site == "YorkRiver",], aes(x = euc_dist, y = t3, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distMort_mg[distMort_mg$bag_site == "YorkRiver",], aes(x = euc_dist, y = t3), col = "slateblue", se = F) +
  geom_point(data = distMort_mg[distMort_mg$bag_site == "YorkRiver",], aes(x = euc_dist, y = t4, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distMort_mg[distMort_mg$bag_site == "YorkRiver",], aes(x = euc_dist, y = t4), col = "purple4", se = F) +
  labs(title = "Environmental Distance vs Mortality at York River", x = "Euclidean Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  ylim(0,1)+
  theme_classic()
envdist_mort_plot_yrk

# compare
envdist_mort_plotT4 <- ggplot(data = distMort_mg) +
  geom_point(data = distMort_mg[distMort_mg$bag_site == "Lewisetta",], aes(x = euc_dist, y = t4, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distMort_mg[distMort_mg$bag_site == "Lewisetta",], aes(x = euc_dist, y = t4), col = "mediumspringgreen", se = T) +
  geom_point(data = distMort_mg[distMort_mg$bag_site == "YorkRiver",], aes(x = euc_dist, y = t4, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distMort_mg[distMort_mg$bag_site == "YorkRiver",], aes(x = euc_dist, y = t4), col = "dodgerblue", se = T) +
  labs(title = "Environmental Distance vs Mortality at Final Time Point", x = "Euclidean Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  ylim(0,1)+
  theme_classic()
envdist_mort_plotT4
```


## Compare Offset Methods

Now that I have predictions from different methods, I will look at how offset predictions made by these methods compare to each other and how different their ability is to predict experimental fitness.


### Compare Offsets to Each Other

```{r Offset Compilation}
# make a df with all of the offsets in them
offsets_lea_exp <- goFitLEA[,c("pop", "bag_site","offset_lew", "offset_yrk", "order_lew", "order_yrk")]
colnames(offsets_lea_exp) <- c("pop", "bag_site","lea_offset_lew", "lea_offset_yrk", "lea_order_lew", "lea_order_yrk")
offsets_gf_exp <- goFitGF[,c("pop", "bag_site","offset_af_lew", "offset_af_yrk", "order_lew", "order_yrk")]
colnames(offsets_gf_exp) <- c("pop", "bag_site","gf_offset_af_lew", "gf_offset_af_yrk", "gf_order_lew", "gf_order_yrk")
offsets_all <- merge(offsets_lea_exp, offsets_gf_exp, by = c("pop", "bag_site"))
```


```{r Offset Comparison}
# corr gf/lea
cor.test(offsets_all[offsets_all$bag_site == "Lewisetta",]$lea_offset_lew,  offsets_all[offsets_all$bag_site == "Lewisetta",]$gf_offset_af_lew) # 0.8234259, p = 0.02282
cor.test(offsets_all[offsets_all$bag_site == "YorkRiver",]$lea_offset_yrk,  offsets_all[offsets_all$bag_site == "YorkRiver",]$gf_offset_af_yrk) # 0.836363, p = 0.01901

ggplot(data = offsets_all) + 
  geom_point(data = offsets_all[offsets_all$bag_site == "Lewisetta",], aes(x = lea_offset_lew, y = gf_offset_af_lew, fill = pop), pch = 21, size = 3) +
  geom_smooth(method = "lm", data = offsets_all[offsets_all$bag_site == "Lewisetta",], aes(x = lea_offset_lew, y = gf_offset_af_lew), col = "black") +
  labs(title = "LEA vs GF Offsets Lewisetta", x = "LEA Offset", y = "GF Offset", fill = "Broodstock Group") +
  theme_classic()

ggplot(data = offsets_all) + 
  geom_point(data = offsets_all[offsets_all$bag_site == "YorkRiver",], aes(x = lea_offset_yrk, y = gf_offset_af_yrk, fill = pop), pch = 21, size = 3) +
  geom_smooth(method = "lm", data = offsets_all[offsets_all$bag_site == "YorkRiver",], aes(x = lea_offset_yrk, y = gf_offset_af_yrk), col = "black") +
  labs(title = "LEA vs GF Offsets York River", x = "LEA Offset", y = "GF Offset", fill = "Broodstock Group") +
  theme_classic()

# corr gf/rda


# corr lea/rda

```

