---
title: "CviMVP_oyster_compareGOexp"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "~/Documents/GitHub/MVP_oyster_GO")
```

# Compare Predicted GO to Experimental Results

This script is for analyzing how well various genomic offset methods predict experimentally obtained fitness. I compare LEA, GF, and RDA offsets to experimental mortality, length, and condition index at the conclusion of the experiment. 

## Setup

```{r}
# setwd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(dplyr)
library(tidyr)
library(stringr)
library(vegan)
library(ggplot2)
library(plotrix)

# data
# offset predictions
lea_go <- read.csv("results/GO_results/offsets_LEA_expSites_05272025.csv")[,-1]
#rda_go <- 
#gf_go <- 

# fitness metrics
mort <- read.csv("results/mortality_bags_052925.csv")[,-1]
#len <- read.csv("data/ExpDat/")
```

## Data Prep

Prepare data from offsets & experiment to be directly compared.

```{r}
# inspect data
lea_go
mort

# we want to get avg mort per pop per site
mort_popsite <- mort %>% group_by(bag_site,pop) %>%
  summarise(mean_t1 = mean(as.numeric(t1)), 
            sd_t1 = sd(as.numeric(t1)),
            mean_t2 = mean(as.numeric(t2)), 
            sd_t2 = sd(as.numeric(t2)),
            mean_t3 = mean(as.numeric(t3)), 
            sd_t3 = sd(as.numeric(t3)),
            mean_t4 = mean(as.numeric(t4)),
            sd_t4 = sd(as.numeric(t4)),
            .groups = 'drop') %>% as.data.frame()

# also get fitness rankings at each time
# t1
ordt1_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t1")])
ordt1_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t1")])
mort_popsite$ordt1 <- c(ordt1_lew, ordt1_york)
# t2
ordt2_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t2")])
ordt2_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t2")])
mort_popsite$ordt2 <- c(ordt2_lew, ordt2_york)
# t3
ordt3_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t3")])
ordt3_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t3")])
mort_popsite$ordt3 <- c(ordt3_lew, ordt3_york)
# t4
ordt4_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t4")])
ordt4_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t4")])
mort_popsite$ordt4 <- c(ordt4_lew, ordt4_york)

# need to align naming conventions to merge
colnames(lea_go) <- c("pop", "offset_lew", "order_lew", "offset_yrk", "order_yrk")
lea_go$pop <- str_split_fixed(lea_go$pop, "-",2)[,2]
#mort_popsite[mort_popsite$pop == "JR",]$pop <- "VA"

# merge gos & fitness dat
goFit <- merge(mort_popsite, lea_go, by = "pop")
```


## Compare Offsets to Fitness

Now that offsets and fitness are in a comparable format, look at how well the two match.

```{r}
# correlations between go/lew mort at each time point
# lea
corLEAt1lew <- cor(goFit[goFit$bag_site == "Lewisetta",]$offset_lew, goFit[goFit$bag_site == "Lewisetta",]$mean_t1) # 0.07640155
corLEAt2lew <- cor(goFit[goFit$bag_site == "Lewisetta",]$offset_lew, goFit[goFit$bag_site == "Lewisetta",]$mean_t2) # 0.03125308
corLEAt3lew <- cor(goFit[goFit$bag_site == "Lewisetta",]$offset_lew, goFit[goFit$bag_site == "Lewisetta",]$mean_t3) # -0.4493965
corLEAt4lew <- cor(goFit[goFit$bag_site == "Lewisetta",]$offset_lew, goFit[goFit$bag_site == "Lewisetta",]$mean_t4) # -0.4493965
# rda
# gf

# correlations between go/york mort at each time point
# lea
corLEAt1yrk <- cor(goFit[goFit$bag_site == "York River",]$offset_yrk, goFit[goFit$bag_site == "York River",]$mean_t1) # -0.05720262
corLEAt2yrk <- cor(goFit[goFit$bag_site == "York River",]$offset_yrk, goFit[goFit$bag_site == "York River",]$mean_t2) # 0.02890556
corLEAt3yrk <- cor(goFit[goFit$bag_site == "York River",]$offset_yrk, goFit[goFit$bag_site == "York River",]$mean_t3) # -0.4362926
# rda
# gf

# these corrs aren't great, try kendall's tau
cor(goFit[goFit$site == "York River",]$offset_yrk, goFit[goFit$site == "York River",]$mean_t1, method = "kendall") # -0.07142857
cor(goFit[goFit$site == "Lewisetta",]$offset_lew, goFit[goFit$site == "Lewisetta",]$mean_t1, method = "kendall") # 0.2857143

cor(goFit[goFit$site == "York River",]$offset_yrk, goFit[goFit$site == "York River",]$mean_t2, method = "kendall") # 0.07142857
cor(goFit[goFit$site == "Lewisetta",]$offset_lew, goFit[goFit$site == "Lewisetta",]$mean_t2, method = "kendall") # 0.1428571

cor(goFit[goFit$site == "York River",]$offset_yrk, goFit[goFit$site == "York River",]$mean_t3, method = "kendall") # -0.1428571
cor(goFit[goFit$site == "Lewisetta",]$offset_lew, goFit[goFit$site == "Lewisetta",]$mean_t3, method = "kendall") # -0.1428571

# pearson's using rankings
# t1
cor(goFit[goFit$site == "Lewisetta",]$order_lew, goFit[goFit$site == "Lewisetta",]$ordt1, method = "pearson") # 0.03984095
cor(goFit[goFit$site == "York River",]$order_yrk, goFit[goFit$site == "York River",]$ordt1, method = "pearson") # -0.1314684
# t2
cor(goFit[goFit$site == "Lewisetta",]$order_lew, goFit[goFit$site == "Lewisetta",]$ordt2, method = "pearson") # -0.3313972
cor(goFit[goFit$site == "York River",]$order_yrk, goFit[goFit$site == "York River",]$ordt2, method = "pearson") # -0.1428571
# t3
cor(goFit[goFit$site == "Lewisetta",]$order_lew, goFit[goFit$site == "Lewisetta",]$ordt3, method = "pearson") # 0.07647191
cor(goFit[goFit$site == "York River",]$order_yrk, goFit[goFit$site == "York River",]$ordt3, method = "pearson") # -0.8179178

# put this all into a df with method/time point/fitness metric/site/corr

```


Plot these data sets against each other

```{r}
# plot lea vs mort at each site/time point
gofit_lew_plot <- ggplot(data = goFit[goFit$site == "Lewisetta",]) +
  geom_point(data = goFit[goFit$site == "Lewisetta",], aes(x = mean_t1, y = offset_lew, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFit[goFit$site == "Lewisetta",], aes(x = mean_t1, y = offset_lew), col = "cornflowerblue", se = F) +
  geom_point(data = goFit[goFit$site == "Lewisetta",], aes(x = mean_t2, y = offset_lew, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFit[goFit$site == "Lewisetta",], aes(x = mean_t2, y = offset_lew), col = "slateblue", se = F) +
  geom_point(data = goFit[goFit$site == "Lewisetta",], aes(x = mean_t3, y = offset_lew, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFit[goFit$site == "Lewisetta",], aes(x = mean_t3, y = offset_lew), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at Lewisetta", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24"), values = c("Fall'23"="cornflowerblue","Spring'24"="slateblue","Fall'24"="purple4"))+
  theme_classic()
gofit_lew_plot
ggsave("results/lea_lew_compare.png")

gofit_yrk_plot <- ggplot(data = goFit[goFit$site == "York River",]) +
  geom_point(data = goFit[goFit$site == "York River",], aes(x = mean_t1, y = offset_yrk, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFit[goFit$site == "York River",], aes(x = mean_t1, y = offset_yrk), col = "cornflowerblue", se = F) +
  geom_point(data = goFit[goFit$site == "York River",], aes(x = mean_t2, y = offset_yrk, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFit[goFit$site == "York River",], aes(x = mean_t2, y = offset_yrk), col = "slateblue", se = F) +
  geom_point(data = goFit[goFit$site == "York River",], aes(x = mean_t3, y = offset_yrk, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFit[goFit$site == "York River",], aes(x = mean_t3, y = offset_yrk), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at York River", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24"), values = c("Fall'23"="cornflowerblue","Spring'24"="slateblue","Fall'24"="purple4"))+
  theme_classic()
gofit_yrk_plot
ggsave("results/lea_york_compare.png")

# plot go lea vs mort at both sites just final time point
gofit_plotT3 <- ggplot(data = goFit) +
  geom_point(data = goFit[goFit$site == "Lewisetta",], aes(x = mean_t3, y = offset_lew, col = "Lewisetta")) +
  geom_smooth(method = "lm", data = goFit[goFit$site == "Lewisetta",], aes(x = mean_t3, y = offset_lew), col = "mediumspringgreen", se = F) +
  geom_point(data = goFit[goFit$site == "York River",], aes(x = mean_t3, y = offset_lew, col = "York River")) +
  geom_smooth(method = "lm", data = goFit[goFit$site == "York River",], aes(x = mean_t3, y = offset_lew), col = "dodgerblue", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at Final Time Point", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Experimental Site", breaks = c("Lewisetta", "York River"), values = c("Lewisetta"="mediumspringgreen","York River"="dodgerblue"))+
  theme_classic()
gofit_plotT3
```

## Compare Offset Methods

Once I have predictions from different methods, I will look at how offset predictions made by these methods compare to each other and how different their ability is to predict experimental fitness.

```{r}

```

