---
title: "CviMVP_LEA_offset"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LEA Genomic offsets

This script is a LEA pipeline for calculating genomic offset. I am using LEA3, which allows for multivariate environmental datasets. I first train the models using LFMM of a large seascape dataset, then calculate offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment. Offset is calculated from the environment of origin to common garden conditions.

## Setup

```{r}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(LEA) # for offsets
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(ggplot2) # for plotting
library(vegan) # for env dat processing
library(dplyr) # for data wrangling
library(stringr) # for data wrangling

# data
# env
# seascapeEnv <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv") # keep this but also get list of reduced variables
# expEnv <- read.csv("data/EnvDat/exp/envr_summary_stats.csv")[,-1]
# expEnv_env <- read.env("data/EnvDat/exp/exp_scaled_expanded_04222025.env")
# expEnvq_env <- read.env("data/EnvDat/exp/expquant_scaled_expanded_04222025.env")
# york_env <- read.env("data/EnvDat/exp/yrkquant_scaled_expanded_04222025.env")
# lew_env <- read.env("data/EnvDat/exp/lewquant_scaled_expanded_04222025.env")
# 
# expEnv_comp <- read.csv("data/EnvDat/exp/expsitesquants_scaled_04222025.csv")[,-1]

# env
env_all <- read.csv("data/EnvDat/env_scaled_2025-08-13.csv")[,-1]
env_sea <- read.env("data/EnvDat/seascape/seascape_scaled_2025-08-13.env")
env_sea_csv <- read.csv("data/EnvDat/seascape/sea_scaled_expanded_reord_2025-08-14.csv")[,-1]
env_sea_expand <- read.env("data/EnvDat/seascape/seascape_scaled_expanded_reord_2025-08-14.env")
env_exp <- read.env("data/EnvDat/exp/exp_scaled_2025-08-14.env")
env_exp_csv <-  read.csv("data/EnvDat/exp/exp_scaled_expanded_2025-08-14.csv")[,-1]
env_exp_expand <- read.env("data/EnvDat/exp/exp_scaled_expanded_2025-08-14.env")

# sites
env_sea_sites <- read.csv("data/EnvDat/seascape_sitenames.csv")[,-1]
env_exp_sites <- read.csv("data/EnvDat/exp_sitenames.csv")[,-1]

# gen
genoMat <- read.geno("data/GenDat/geno_seascape_2025-08-14.geno")
genoLfmm <- read.lfmm("data/GenDat/geno_seascape_2025-08-14.lfmm")
genoThinMat <- read.geno("data/GenDat/geno_seascape_LDthin_2025-08-14.geno")
genoThinLfmm <- read.lfmm("data/GenDat/geno_seascape_LDthin_2025-08-14.lfmm")

# inds
inds_genomat <- read.csv("data/GenDat/inds_genomat.csv")[,-1]
indsExp <- readRDS("data/IndDat/20240922_experimental_indsmatrix.rds")
indsExpSeq <- readRDS("data/IndDat/20240820_experimentalmatrix.rds")
indsSea <- readRDS("data/IndDat/20250604_seascape_indsmatrix.rds")
```


## Data Prep

### Environmental Data Prep

```{r}
# environmental data prep
# just renaming columns & rows to be more informative
colnames(env_sea) <- colnames(env_sea_expand) <- c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")

colnames(env_exp) <- colnames(env_exp_expand) <- c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "lew_sal_quant10_scaled", "lew_sal_quant90_scaled", "lew_temp_quant10_scaled", "lew_temp_quant90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled", "yrk_sal_quant10_scaled", "yrk_sal_quant90_scaled", "yrk_temp_quant10_scaled", "yrk_temp_quant90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")

rownames(env_sea) <- unique(env_sea_ordered$site_name)
rownames(env_exp) <- env_exp_sites
rownames(env_sea_expand) <- indsSea[indsSea$ID_SiteDate %in% env_sea_sites, "ID_SiteDate"]
rownames(env_exp_expand) <- indsExp[!(indsExp$ID_SiteDate %in% c("DEBY", "LOLA")), "ID_SiteDate"]
```


### Genetic Data Prep

```{r}
# name the rows of genomic data
rownames(genoMat) <- rownames(genoLfmm) <- rownames(genoThinMat) <- rownames(genoThinLfmm) <- inds_genomat

# make sure genotype data is in the same order as individual & environmental data
env_sea_csv$clean_ID %in% inds_genomat
env_sea_csv$clean_ID == inds_genomat # yes, everything is in the correct order
```


## Seascape Training

Now that I have prepped and cleaned my data, I will use it to train LFMM models later used to predict genomic offset to some other (in this case experimental) climate. Importantly, I do not need to trim or impute my genomic data, or determine ancestry - Madeline has already done this. The data has been LD thinned, trimmed for MAF, and ancestry analysis has recovered K=2.
Madeline also compiled the environmental data, which I further processed in the script "CviMVP_oyster_EnvDatProcess."

May decide to train both genotype-based models and allele frequency based models - for now genotype models only.

```{r}
# check the data
genoLfmm[1:10, 1:10]
genoThinLfmm[1:10, 1:10]


# fit model
mod_lfmm <- lfmm2(input = genoLfmm,
                  env = env_sea_expand, 
                  K = 2) # this was number from Madeline's snmf analysis
mod_lfmmthin <- lfmm2(input = genoThinLfmm,
                      env = env_sea_expand,
                      K = 2)

# perform test
pv_lfmm <- lfmm2.test(object = mod_lfmm,
                      input = genoLfmm,
                      env = env_sea_expand, 
                      full = TRUE)
pv_lfmmthin <- lfmm2.test(object = mod_lfmmthin,
                      input = genoThinLfmm,
                      env = env_sea_expand, 
                      full = TRUE)

# pvalues
pv <- pv_lfmm$pvalues
pvthin <- pv_lfmmthin$pvalues

# qvalues
library(qvalue)
qv <- qvalue(pv, fdr.level = 0.01)
qvthin <- qvalue(pvthin, fdr.level = 0.01)

# candidates
candidates <- which(qv$significant)
candidates_pvals <- (pv)[candidates]

candidatesthin <- which(qvthin$significant)
candidates_pvalsthin <- (pvthin)[candidates]

# using FDR to filter out false positives
length(which(qv$qvalues < 0.001)) # 12455
length(which(qvthin$qvalues < 0.001)) # 5430

# identify which SNPs have FDR less than 1%
fdr_cands <- colnames(genoLfmm)[which(qv$qvalues < 0.001)]
fdr_candsthin <- colnames(genoThinLfmm)[which(qv$qvalues < 0.001)]

# df for candidates
cand_df <- data.frame(index = colnames(genoLfmm), pos = 1:length(pv),
                      pvals = pv, qvals = qv$qvalues)
cand_df_thin <- data.frame(index = colnames(genoThinLfmm), pos = 1:length(pvthin),
                      pvals = pvthin, qvals= qvthin$qvalues)

# manhattan plot - unnecessary but just a quick visualization
plot(cand_df$pos, -log10(cand_df$pvals), 
     pch = 19, col = "darkslategray4",
     main = "Candidate SNPs",
     xlab = "Position",
     ylab = "-log10(Pvals)")
points(cand_df$pos[cand_df$qvals < 0.001], 
       -log10(cand_df$pvals[cand_df$qvals < 0.001]),
       pch = 19,
       col = "hotpink")

plot(cand_df_thin$pos, -log10(cand_df_thin$pvals), 
     pch = 19, col = "darkslategray4",
     main = "Candidate SNPs (Thinned)",
     xlab = "Position",
     ylab = "-log10(Pvals)")
points(cand_df_thin$pos[cand_df_thin$qvals < 0.001], 
       -log10(cand_df_thin$pvals[cand_df_thin$qvals < 0.001]),
       pch = 19,
       col = "hotpink")
```


## Offset Prediction

```{r}
# check env data
colnames(env_exp_csv)
colnames(env_exp_expand)
colnames(env_exp_csv)
colnames(env_exp_expand) <- c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "lew_salinity_quantile_10_scaled",  "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled", "yrk_salinity_quantile_10_scaled",  "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")

# calculate genetic gap
gap_lew <- genetic.offset(input = genoLfmm, 
                   env = env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")], 
                   pred.env = env_exp_expand[,c("lew_salinity_quantile_10_scaled",  "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled")], 
                   scale = F, 
                   K = 2)

gap_yrk <- genetic.offset(input = genoLfmm, 
                   env = env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")], 
                   pred.env = env_exp_expand[,c("yrk_salinity_quantile_10_scaled",  "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")], 
                   scale = F, 
                   K = 2)

gap_lewthin <- genetic.offset(input = genoLfmm, 
                   env = env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")], 
                   pred.env = env_exp_expand[,c("lew_salinity_quantile_10_scaled",  "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled")], 
                   scale = F, 
                   K = 2)

gap_yrkthin <- genetic.offset(input = genoLfmm, 
                   env = env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")], 
                   pred.env = env_exp_expand[,c("yrk_salinity_quantile_10_scaled",  "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")], 
                   scale = F, 
                   K = 2)

# THESE PARTS NEED TO BE UPDATED FROM EELGRASS SCRIPT VERSION
# check how well each genetic gap measure correlates with euclidean environmental distance
Delta <- expEnv_env - lew_env
dist_env <- sqrt(rowSums(Delta^2))
plot(dist_env, gap_lew$distance,
     main = "Genetic gap vs Environmental Distance", 
     xlab ="Euclidean distance",  ylab ="sqrt(genetic gap)", 
     cex = .6, col = "darkslategray4", pch = 19)

Delta <- expEnv_env - york_env
dist_env <- sqrt(rowSums(Delta^2))
plot(dist_env, gap_yrk$distance,
     main = "Genetic gap vs Environmental Distance", 
     xlab ="Euclidean distance",  ylab ="sqrt(genetic gap)", 
     cex = .6, col = "darkslategray4", pch = 19)

# Extract genomic offset from genetic gap object
offsets_lew <- gap_lew$offset
offsets_yrk <- gap_yrk$offset

# summary stats on offsets
summary(offsets_lew)
summary(offsets_yrk)

# offset dfs
offsets_df <- cbind(expEnv_comp[,c("site_name","Label_ind")], offsets_lew, offsets_yrk)

# output results
write.csv(offsets_df, paste0("results/GO_results/offsets_LEA_allexpinds_",Sys.Date,".csv", sep = ""))
```


```{r}
# UPDATE FOR OYSTERS WHEN DATA IS IN
# can now calculate offset by region, site, etc
# average offset by site
offset_lew <- aggregate(offsets_df$offsets_lew, by = list(offsets_df$site_name), FUN = mean)
colnames(offset_lew) <- c("site","offset_lew")

offset_yrk <- aggregate(offsets_df$offsets_yrk, by = list(offsets_df$site_name), FUN = mean)
colnames(offset_yrk) <- c("site","offset_yrk")

# order populations by average offset
(ordered_offset_lew <- offset_lew[order(offset_lew$offset_lew),])
ordered_offset_lew$order <- 1:8

(ordered_offset_yrk <- offset_yrk[order(offset_yrk$offset_yrk),])
ordered_offset_yrk$order <- 1:8

# put this together into single df to visualize
offset_sites <- merge(ordered_offset_lew, ordered_offset_yrk, by = "site")
colnames(offset_sites) <- c("site","offset_lew","order_lew","offset_yrk","order_yrk")

# save this and raw offsets as .csv
write.csv(offset_sites, paste0("results/GO_results/offsets_LEA_expSites_",Sys.Date(),".csv", sep = ""))
```


## Offset Visualization

```{r}
# need to add latlongs for each site
sites_geo <- expEnv_comp[,c("site_name","lat.x","lon.x")]
colnames(sites_geo) <- c("site","lat","lon")
offsets_plotting <- merge(sites_geo, offset_sites, by =  c("site"))

# add the color palette
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME", "LARMIX", "SEEDMIX"), 
                     order = 1:10, 
                     cols = c("#3e4989" ,"#31688e", "#26828e", "#1f9e89", "#440154", "#482878", "#35b779", "#6ece58", "#fde725", "#b5de2b"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site = c("W1-TX", "W2-LA", "W3-FL", "W4-VA", "S1-LOLA", "S2-DEBY", "W5-NH", "W6-ME", "H1-LARMIX", "H2-SEEDMIX"))
offsets_plotting_full <- merge(offsets_plotting, popsdf, by = c("site"))

# save popdf object
write.csv(popsdf, "data/IndDat/popsdf.csv")

# try mapping the offsets
library(s2) # mapping
library(rnaturalearth) # mapping
library(rnaturalearthdata) # mapping
library(maps) # mapping
library(ggspatial) # mapping

# map
sf::sf_use_s2(FALSE) # sf package doesn't work

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin =-102 , xmax = -68, ymin = 24, ymax = 47.049130))

# bottom right: 24.353558, -55.373595
# top left: 47.049130, -101.983412

# offset to lew
map_lew <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting_full, aes(x=lon, y=lat, 
                                       fill = offset_lew), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.07))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to Lewisetta", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_lew)

world_chesapeake <- sf::st_crop(world, c(xmin =-77.856476 , xmax = -72.762787, ymin = 36.019557, ymax = 39.939163))

# top left: 39.939163, -77.856476
# bottom right: 36.019557, -72.762787

popout_lew <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting_full, aes(x=lon, y=lat, 
                                       fill = offset_lew), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.07))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
popout_lew

# offset to york
map_york <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting_full, aes(x=lon, y=lat, 
                                       fill = offset_yrk), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.07))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to York River", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_york)

popout_york <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting_full, aes(x=lon, y=lat, 
                                       fill = offset_yrk), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.07))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
popout_york

cor(offsets_plotting_full$offset_lew, offsets_plotting_full$offset_yrk)
plot(x = offsets_plotting_full$offset_lew, y = offsets_plotting_full$offset_yrk)
```


