---
title: "CviMVP_oyster_compareGOexp"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "~/Documents/GitHub/MVP_oyster_GO")
```

# Compare Predicted GO to Experimental Results

This script is for analyzing how well various genomic offset methods predict experimentally obtained fitness. I compare LEA, GF, and RDA offsets to experimental mortality, length, and condition index at the conclusion of the experiment. 

## Setup

```{r}
# setwd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(dplyr)
library(tidyr)
library(stringr)
library(vegan)
library(ggplot2)
library(plotrix)

# data
# offset predictions
lea_go <- read.csv("results/GO_results/offsets_LEA_expSites_05272025.csv")[,-1]
#rda_go <- 
gf_go <- read.csv("results/GO_results/offsetsMVP_2025-06-03.csv")[,-1]

# fitness metrics
mort <- read.csv("results/exp_results/mortality_bags_2025-06-06.csv")[,-1]
#len <- read.csv("data/ExpDat/")
```

## Data Prep

Prepare data from offsets & experiment to be directly compared.

```{r}
# inspect data
lea_go
gf_go
mort

# we want to get avg mort per pop per site
mort_popsite <- mort %>% group_by(bag_site,pop) %>%
  summarise(mean_t1 = mean(as.numeric(t1)), 
            sd_t1 = sd(as.numeric(t1)),
            mean_t2 = mean(as.numeric(t2)), 
            sd_t2 = sd(as.numeric(t2)),
            mean_t3 = mean(as.numeric(t3)), 
            sd_t3 = sd(as.numeric(t3)),
            mean_t4 = mean(as.numeric(t4)),
            sd_t4 = sd(as.numeric(t4)),
            .groups = 'drop') %>% as.data.frame()

# also get fitness rankings at each time
# t1
ordt1_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t1")])
ordt1_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t1")])
mort_popsite$ordt1 <- c(ordt1_lew, ordt1_york)
# t2
ordt2_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t2")])
ordt2_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t2")])
mort_popsite$ordt2 <- c(ordt2_lew, ordt2_york)
# t3
ordt3_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t3")])
ordt3_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t3")])
mort_popsite$ordt3 <- c(ordt3_lew, ordt3_york)
# t4
ordt4_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t4")])
ordt4_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t4")])
mort_popsite$ordt4 <- c(ordt4_lew, ordt4_york)

# need to align naming conventions to merge
# LEA
colnames(lea_go) <- c("pop", "offset_lew", "order_lew", "offset_yrk", "order_yrk")
lea_go$pop <- str_split_fixed(lea_go$pop, "-",2)[,2]
#mort_popsite[mort_popsite$pop == "JR",]$pop <- "VA"

# GF
# add order
gf_go$order_lew <- order(gf_go$offset_af_lew)
gf_go$order_yrk <- order(gf_go$offset_af_yrk)
gf_go$pop <- str_split_fixed(gf_go$site_name,"-",2)[,2]

# mort popsite uses wrong label for JR/VA - call it VA
mort_popsite[mort_popsite$pop == "JR",]$pop <- "VA"

# merge gos & fitness dat
goFitLEA <- merge(mort_popsite, lea_go, by = "pop")
goFitGF <- merge(mort_popsite, gf_go, by = "pop")

# associate fitness data at the bag level as well
levels(as.factor(mort$pop))
mort[mort$pop == "JR",]$pop <- "VA"
goFitLEAbag <- merge(mort, lea_go, by = "pop")
goFitGFbag <- merge(mort, gf_go, by = "pop")
```


## Compare Offsets to Fitness

Now that offsets and fitness are in a comparable format, look at how well the two match.

### LEA Offset

Start with the bag level mortality data.

```{r}
# correlations between go/lew mort at each time point
# lea
(cor_lea1_lew <- cor(goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",]$offset_lew, goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",]$t1)) # -0.04646617
(cor_lea2_lew <- cor(goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",]$offset_lew, goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",]$t2)) # 0.01330168
(cor_lea3_lew <- cor(goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",]$offset_lew, goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",]$t3)) # -0.08660892
(cor_lea4_lew <- cor(goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",]$offset_lew, goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",]$t4)) # -0.1313757

# correlations between go/york mort at each time point
# lea
(cor_lea1_yrk <- cor(goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",]$offset_yrk, goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",]$t1)) # 0.03567268
(cor_lea2_yrk <- cor(goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",]$offset_yrk, goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",]$t2)) # -0.009258749
(cor_lea3_yrk <- cor(goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",]$offset_yrk, goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",]$t3)) # -0.2830991
(cor_lea4_yrk <- cor(goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",]$offset_yrk, goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",]$t4)) # -0.316612

# put this all into a df with method/time point/fitness metric/site/corr
lea_corr_df <- data.frame(method = rep("LEA", 8), metric = rep("surv", 8), monitoring_event = rep(c(1:4),2), bag_site = c(rep("Lewisetta",4 ), rep("YorkRiver", 4)), pearson = c(cor_lea1_lew, cor_lea2_lew, cor_lea3_lew, cor_lea4_lew, cor_lea1_yrk, cor_lea2_yrk, cor_lea3_yrk, cor_lea4_yrk))
```


Plot these data sets against each other

```{r}
# plot lea vs mort at each site/time point
goFitLEA_lew_plotbag <- ggplot(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",]) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t1, y = offset_lew, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t1, y = offset_lew), col = "cadetblue3", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t2, y = offset_lew, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t2, y = offset_lew), col = "cornflowerblue", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t3, y = offset_lew, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t3, y = offset_lew), col = "slateblue", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t4, y = offset_lew, col = "Spring'25")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t4, y = offset_lew), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at Lewisetta", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_lew_plotbag
ggsave(paste0("results/figures/lea_lew_compareBag_",Sys.Date(),".png"))

goFitLEA_yrk_plot <- ggplot(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",]) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t1, y = offset_yrk, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t1, y = offset_yrk), col = "cadetblue3", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t2, y = offset_yrk, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t2, y = offset_yrk), col = "cornflowerblue", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t3, y = offset_yrk, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t3, y = offset_yrk), col = "slateblue", se = F) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t4, y = offset_yrk, col = "Spring'25")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t4, y = offset_yrk), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at York River", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_plot
ggsave(paste0("results/figures/lea_york_compareBag_",Sys.Date(),".png"))

# plot go lea vs mort at both sites just final time point
goFitLEA_plotT4bag <- ggplot(data = goFitLEAbag) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t4, y = offset_lew, col = "Lewisetta")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "Lewisetta",], aes(x = t4, y = offset_lew), col = "mediumspringgreen", se = T) +
  geom_point(data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t4, y = offset_lew, col = "YorkRiver")) +
  geom_smooth(method = "lm", data = goFitLEAbag[goFitLEAbag$bag_site == "YorkRiver",], aes(x = t4, y = offset_lew), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (LEA) vs Mortality at Final Time Point", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotT4bag
ggsave(paste0("results/figures/lea_yorklew_compareBag_",Sys.Date(),".png"))
```

Now do it at the population level!

Start with the population level mortality data.

```{r}
# correlations between go/lew mort at each time point
# lea
(cor_lea1_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t1, method = "pearson")) # -0.05736393, r2 0.00329062
(cor_lea2_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t2, method = "pearson")) # 0.01456035, r2 0.0002120037
(cor_lea3_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t3, method = "pearson")) # -0.08913129, r2 0.007944386
(cor_lea4_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t4, method = "pearson")) # -0.135897, r2 0.01846799

(Kcor_lea1_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t1, method = "kendall")) # -0.05736393
(Kcor_lea2_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t2, method = "kendall")) # 0.01456035
(Kcor_lea3_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t3, method = "kendal")) # -0.08913129
(Kcor_lea4_lew_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t4, method = "kendall")) # -0.135897

# correlations between go/york mort at each time point
# lea
(cor_lea1_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t1, method = "pearson")) # 0.0400319, r2 0.001602553
(cor_lea2_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t2, method = "pearson")) # -0.01014261, r2 0.0001028726
(cor_lea3_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t3, method = "pearson")) # -0.2889805, r2 0.0835097
(cor_lea4_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t4, method = "pearson")) # -0.3244636, r2 0.1052767

summary(lm(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk ~ goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t4))$r.squared

(KenCor_lea1_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t1, method = "kendall")) # 0.07142857
(KenCor_lea2_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t2, method = "kendall")) # 0.07142857
(KenCor_lea3_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t3, method = "kendall")) # -0.2142857
(KenCor_lea4_yrk_pop <- cor.test(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t4, method = "kendall")) # 0

# put this all into a df with method/time point/fitness metric/site/corr
lea_corr_df_pop <- data.frame(method = rep("LEA", 8), metric = rep("surv", 8), monitoring_event = rep(c(1:4),2), bag_site = c(rep("Lewisetta",4 ), rep("YorkRiver", 4)), pearson = c(cor_lea1_lew_pop, cor_lea2_lew_pop, cor_lea3_lew_pop, cor_lea4_lew_pop, cor_lea1_yrk_pop, cor_lea2_yrk_pop, cor_lea3_yrk_pop, cor_lea4_yrk_pop))
```


Plot these data sets against each other

```{r}
# plot lea vs mort at each site/time point
goFitLEA_lew_plot <- ggplot(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",]) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t1, y = offset_lew, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t1, y = offset_lew), col = "cadetblue3", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t2, y = offset_lew, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t2, y = offset_lew), col = "cornflowerblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t3, y = offset_lew, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t3, y = offset_lew), col = "slateblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_lew, col = "Spring'25")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_lew), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at Lewisetta", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_lew_plot
ggsave(paste0("results/figures/lea_lew_compare_",Sys.Date(),".png"))

goFitLEA_yrk_plot <- ggplot(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",]) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t1, y = offset_yrk, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t1, y = offset_yrk), col = "cadetblue3", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t2, y = offset_yrk, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t2, y = offset_yrk), col = "cornflowerblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t3, y = offset_yrk, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t3, y = offset_yrk), col = "slateblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_yrk, col = "Spring'25")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_yrk), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at York River", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_plot
ggsave(paste0("results/figures/lea_york_compare_",Sys.Date(),".png"))

# plot go lea vs mort at both sites just final time point
goFitLEA_plotT4 <- ggplot(data = goFitLEA) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_lew, col = "Lewisetta")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_lew), col = "mediumspringgreen", se = T) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_lew, col = "YorkRiver")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_lew), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (LEA) vs Mortality at Final Time Point", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotT4
```


### GF Offset

```{r}
# correlations between go/lew mort at each time point
(cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t1)) # -0.1931013, r2 0.003633104
(cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t2)) # -0.006177896, r2 0.002257078
(cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t3)) # 0.04885283, r2 0.003688538
(cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t4)) # 0.04589687, r2 0.002099499

summary(lm(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew ~ goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t1))$r.squared
summary(lm(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew ~ goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t2))$r.squared
summary(lm(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew ~ goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t3))$r.squared
summary(lm(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew ~ goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t4))$r.squared

# correlations between go/yrk mort at each time point
(cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t1)) # 0.2852591
(cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t2)) # 0.318425
(cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t3)) # 0.241053
(cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t4)) # 0.2847928

# really bad correlations, may need to double check offset calculations
# pearson of orders - LEW
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t1) # 0.003129999
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t2) # 0.03004949
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t3) # -0.03736819
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t4) # -0.107596
# pearson of orders - YRK
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t1) # 0.1687206
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t2) # 0.1059771
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t3) # -0.214745
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t4) # -0.2206604
```

Plot these data against each other.

```{r}
# plot lea vs mort at each site/time point
goFitGF_lew_plot <- ggplot(data = goFitGF[goFitGF$bag_site == "Lewisetta",]) +
  geom_point(data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t1, y = offset_af_lew, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t1, y = offset_af_lew), col = "cadetblue3", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t2, y = offset_af_lew, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t2, y = offset_af_lew), col = "cornflowerblue", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t3, y = offset_af_lew, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t3, y = offset_af_lew), col = "slateblue", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_af_lew, col = "Spring'25")) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_af_lew), col = "purple4", se = F) +
  labs(title = "Predicted GO (GFaf) vs Mortality at Lewisetta", x = "Survival", y = "GFaf Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_lew_plot
ggsave(paste0("results/figures/gf_lew_compare_",Sys.Date(),".png"))

goFitGF_yrk_plot <- ggplot(data = goFitGF[goFitGF$bag_site == "YorkRiver",]) +
  geom_point(data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t1, y = offset_af_yrk, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t1, y = offset_af_yrk), col = "cadetblue3", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t2, y = offset_af_yrk, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t2, y = offset_af_yrk), col = "cornflowerblue", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t3, y = offset_af_yrk, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t3, y = offset_af_yrk), col = "slateblue", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_af_yrk, col = "Spring'25")) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_af_yrk), col = "purple4", se = F) +
  labs(title = "Predicted GO (GFaf) vs Mortality at York River", x = "Survival", y = "GFaf Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_yrk_plot
ggsave(paste0("results/figures/gf_york_compare_",Sys.Date(),".png"))

# plot go lea vs mort at both sites just final time point
goFitGF_plotT4 <- ggplot(data = goFitGF) +
  geom_point(data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_af_lew, col = "Lewisetta")) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "Lewisetta",], aes(x = mean_t4, y = offset_af_lew), col = "mediumspringgreen", se = F) +
  geom_point(data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_af_yrk, col = "YorkRiver")) +
  geom_smooth(method = "lm", data = goFitGF[goFitGF$bag_site == "YorkRiver",], aes(x = mean_t4, y = offset_af_yrk), col = "dodgerblue", se = F) +
  labs(title = "Predicted GO (GFaf) vs Mortality at Final Time Point", x = "Survival", y = "GFaf Offset") +
  scale_colour_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGF_plotT4
```


## Compare Offset Methods

Now that I have predictions from different methods, I will look at how offset predictions made by these methods compare to each other and how different their ability is to predict experimental fitness.


### Compare Offsets to Each Other

```{r Offset Compilation}
# make a df with all of the offsets in them
offsets_lea_exp <- goFitLEA[,c("pop", "bag_site","offset_lew", "offset_yrk", "order_lew", "order_yrk")]
colnames(offsets_lea_exp) <- c("pop", "bag_site","lea_offset_lew", "lea_offset_yrk", "lea_order_lew", "lea_order_yrk")
offsets_gf_exp <- goFitGF[,c("pop", "bag_site","offset_af_lew", "offset_af_yrk", "order_lew", "order_yrk")]
colnames(offsets_gf_exp) <- c("pop", "bag_site","gf_offset_af_lew", "gf_offset_af_yrk", "gf_order_lew", "gf_order_yrk")
offsets_all <- merge(offsets_lea_exp, offsets_gf_exp, by = c("pop", "bag_site"))
```


```{r Offset Comparison}
# corr gf/lea
cor.test(offsets_all[offsets_all$bag_site == "Lewisetta",]$lea_offset_lew,  offsets_all[offsets_all$bag_site == "Lewisetta",]$gf_offset_af_lew) # 0.8234259, p = 0.02282
cor.test(offsets_all[offsets_all$bag_site == "YorkRiver",]$lea_offset_yrk,  offsets_all[offsets_all$bag_site == "YorkRiver",]$gf_offset_af_yrk) # 0.836363, p = 0.01901

ggplot(data = offsets_all) + 
  geom_point(data = offsets_all[offsets_all$bag_site == "Lewisetta",], aes(x = lea_offset_lew, y = gf_offset_af_lew, fill = pop), pch = 21, size = 3) +
  geom_smooth(method = "lm", data = offsets_all[offsets_all$bag_site == "Lewisetta",], aes(x = lea_offset_lew, y = gf_offset_af_lew), col = "black") +
  labs(title = "LEA vs GF Offsets Lewisetta", x = "LEA Offset", y = "GF Offset", fill = "Broodstock Group") +
  theme_classic()

ggplot(data = offsets_all) + 
  geom_point(data = offsets_all[offsets_all$bag_site == "YorkRiver",], aes(x = lea_offset_yrk, y = gf_offset_af_yrk, fill = pop), pch = 21, size = 3) +
  geom_smooth(method = "lm", data = offsets_all[offsets_all$bag_site == "YorkRiver",], aes(x = lea_offset_yrk, y = gf_offset_af_yrk), col = "black") +
  labs(title = "LEA vs GF Offsets York River", x = "LEA Offset", y = "GF Offset", fill = "Broodstock Group") +
  theme_classic()

# corr gf/rda


# corr lea/rda

```

