---
title: "CviMVP_oyster_GF"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GF Genomic Offsets

This script is a gradientForest pipeline for calculating genomic offset. I am using gradientForest, which build on randomForest to use predictions of genome-wide allele frequency variation along multivariate environmental gradients to quantify the disruption of genotype-environment associations under environmental change. I first train the models using gradientForest on a large seascape dataset, then predict offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment. Offset is calculated from the environment of origin to common garden conditions.

## Setup

```{r}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(dplyr)
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(raster) # for environmental pixels
library(ggplot2) # for plotting
#library(magrittr) # unsure
#library(geosphere) # unsure
#library(gdm) # unsure
#library(foreach) # unsure
library(parallel) # for parallelizing processes
library(doParallel) # for parallelizing processes 
library(gradientForest) # for running gradient forest
#library(fields) # unsure
#library(sf) # for spatial files

# data
# env
seascapeEnv <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv") # keep this but also get list of reduced variables
expEnv <- read.csv("data/EnvDat/exp/envr_summary_stats.csv")[,-1]

# gen
# add once i get it from madeline, for now using exp indivs
geno_mat
```


## Data Prep

