---
title: "CviMVP_oyster_GenDatProcess"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Genetic Data Processing

This script is to process individual and (genetic for seascape) level data for the seascape and experiment by identifying the number and identities of individuals in each dataset and preparing this information and the genomic data into formats needed for downstream analysis.

The seascape data comes from...

The experimental data comes from...


FOR NOW I AM DOING THIS JUST WITH EXPERIMENTAL INDIVIDUALS GENOTYPED, BUT WE WILL NEED TO THINK ABOUT HOW TO DO THIS WHEN WE HAVE NO TRUE EXPERIMENTAL FITNESS DATA ON AN INDIVIDUAL LEVEL (IG WE DO FOR TAGGED INDS LENGTHS? BUT NOT MORT)


## Setup

```{r}
# set wd
setwd("~/Documents/Github/MVP_oyster_GO")

# packages
library(dplyr)
library(ggplot2)
library(tidyverse)
library(LEA)
library(stringr)

# data
geno_mat <- readRDS("data/GenDat/20250606_FULLSNPs_seascape_labeled.rds")
geno_mat_ld <- readRDS("data/GenDat/20250606_THINNEDSNPs_seascape_labeled.rds")
sea_inds <- readRDS("data/IndDat/20250604_seascape_indsmatrix.rds")
```


## Pre-processing

Process data to remove missing data, non-variable genomic sites.

```{r}
# check data
dim(geno_mat) # 746 inds, 153921 snps
geno_mat[1:10,1:10]

dim(geno_mat_ld) # 746 inds, 115185 snps (~40k less than unthinned)

# any missing data?
length(which(geno_mat == 9)) # nope!
length(which(geno_mat_ld == 9)) # nope!

# non-variable genomic sites
nonvarsite <- apply(geno_mat, 2, function(x) all(x == x[1]))
which(nonvarsite == T) # none
nonvarsite_ld <- apply(geno_mat_ld, 2, function(x) all(x == x[1]))
which(nonvarsite_ld == T) # none
```


## Individual Data

```{r}
dim(sea_inds) # 746 inds (presumably in same order), 79 cols with identifying info incl pop
<<<<<<< HEAD
=======

# let's get rid of the empty columns
sea_inds_rm <- sea_inds[, apply(sea_inds, 2, function(x) !any(is.na(x)))] # only 29 columns now
>>>>>>> 71caa11d457877753593e32f3e7f18fc6ec7efff

# let's get rid of the empty columns
sea_inds_rm <- sea_inds[, apply(sea_inds, 2, function(x) !any(is.na(x)))] # only 29 columns now

# rownames of genomic data match clean_ID in individual df, and appear to be in the same order
```


## Genomic Data Formatting

```{r}
# write these matrices as geno files
write.geno(geno_mat, paste0("data/GenDat/geno_seascape_", Sys.Date(), ".geno"))
write.geno(geno_mat_ld, paste0("data/GenDat/geno_seascape_LDthin_", Sys.Date(), ".geno"))

<<<<<<< HEAD
# for LEA need lfmm format data
geno2lfmm("data/GenDat/geno_seascape_2025-08-13.geno")
geno2lfmm("data/GenDat/geno_seascape_LDthin_2025-08-13.geno")

# for RDA and gf geno matrix should work, will also output geno as rds object
write_rds(geno_mat, paste0("data/GenDat/geno_seascape_", Sys.Date(),".rds"))
write_rds(geno_mat_ld, paste0("data/GenDat/geno_seascape_LDthin_", Sys.Date(),".rds"))
=======
# for RDA and gf geno matrix should work, will also output geno as rds object
write.csv(geno_mat, "data/GenDat/geno_file_exp_all.csv")
>>>>>>> 71caa11d457877753593e32f3e7f18fc6ec7efff
```

