---
title: "CviMVP_oyster_RDA.Rmd"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LEA Genomic offsets

This script is a RDA-based pipeline for calculating genomic offset. I am using rdadapt, a package developed for genome scans for selection, adaptive index projection on the landscape, estimation of genomic offsets, identification of co-varying modules of adaptive loci, and/or identification of adaptive groups of individual or populations. I first calculate "adaptive index" across seascape sites, then calculate offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment. Offset is calculated from the environment of origin to common garden conditions.

A vignette detailing the rdadapt pipeline is found [here](https://landscape-genomics.github.io/rdadapt/articles/vignette_tutorial.html).

## Setup

```{r}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(rdadapt) # for offsets
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(vegan) # for env dat processing
library(terra) # for env dat processing
library(reshape2) # for env dat processing
library(raster) # for env dat processing
library(dplyr) # for data wrangling
library(plyr) # for data wrangling
library(stringr) # for data wrangling
library(ggplot2) # for data vis
library(viridis) # for data vis
library(wesanderson) # for data vis
library(rdadapt) # for rda

# data
# env
env_all <- as.data.frame(read.csv("data/EnvDat/env_scaled_2025-08-13.csv")[,-1])
env_sea <- as.data.frame(read.csv("data/EnvDat/seascape/sea_scaled_expanded_reord_2025-08-14.csv")[,-1])
env_exp <- as.data.frame(read.csv("data/EnvDat/exp/exp_scaled_expanded_2025-08-14.csv")[,-1])

# sites
env_sea_site <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv")[,-1]
env_sea_date <- read.csv("data/EnvDat/seascape/SeascapeSamples - date.csv")[,-1]
env_exp_site <- read.csv("data/EnvDat/exp_sitenames.csv")[,-1]
exp_site_latlon <- read.csv("data/EnvDat/exp/exp_site_info.csv")[,-1]

# gen
genoMat <- read.geno("data/GenDat/geno_seascape_2025-08-14.geno")
genoThinMat <- read.geno("data/GenDat/geno_seascape_LDthin_2025-08-14.geno")

# individuals
expInds <- readRDS("data/IndDat/20240922_experimental_indsmatrix.rds")
seascapeInds <- readRDS("data/IndDat/20250604_seascape_indsmatrix.rds")
inds_genomat <- read.csv("data/GenDat/inds_genomat.csv")[,-1]
```


## Data Checks

I will start by double checking that I have my genomic and environmental data in the correct formats.

```{r}
# env data
##########
# seascape
env_sea
env_sea[which(is.na(env_sea$Mean_Annual_Salinity_ppt)),] # no na

# need to find the more complete YRK & LEW data, talk to nicole about this

# check exp data
env_exp

# check individual data
seascapeInds

# keep just relevant data
seascapeInds_red <- seascapeInds[,c("clean_ID","Label_ind","Key_ind","ID_SiteDate","SiteDate_Key","EstuaryRegion_name","Site_Order","Site_Color","Region","Region_color")]
levels(as.factor(seascapeInds$ID_SiteDate))

# pops list
pops_list <- levels(as.factor(seascapeInds$ID_SiteDate))

# get some lat lons
sea_latlon <- merge(env_sea_site[,c("ID_Site", "Latitude_manual", "Longitude_manual")], env_sea_date[,c("ID_Site", "ID_SiteDate")], by = "ID_Site")

# add latlon to env data
env_sea_complete <- merge(sea_latlon, env_sea, by = c("ID_SiteDate"))

# rasterize these data
#expEnv_pop_mut_rm <- expEnv_pop_mut[, !names(expEnv_pop_mut) %in% c("lat.y", "lon.y")]
#expEnv_pop_mut_rn <- expEnv_pop_mut_rm %>% dplyr::rename(lat = lat.x, lon = lon.x)

#ras_env <- rast(data.frame(x = rep(1:12, time = 12) # NEED TO COME BACK HERE
#                                  , y = rep(1:12, each = 12)
#                                  , aggregate(xx.current, by = list(ceiling(coord$x)
#                                                                    , ceiling(coord$y))
#                                              , mean)[, -c(1, 2)])
#                       , type = "xyz"
#                       , crs = crs("epsg:4326"))
##########

# gen data
##########
# check gen dat
dim(genoMat)
dim(genoThinMat)

# assign meaningful rownames - the inds_genomat vector has individual names in the order they appear in the genotype matrix (as determined in GenDatProcess script), which is also the order of the seascape env matrix
rownames(genoMat) <- rownames(genoThinMat) <- inds_genomat
##########
```

## Outlier Detection

While I do not intend to use an outlier dataset to calculate adaptive index/offset, I will still go through the process of outlier detection with RDA. This process is also part of training the RDA model for offsets later.

```{r}
# reduce size of env data set
env_sea_red <- env_sea[,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")]

# run rda
RDA_sea <- rda(genoMat ~ ., as.data.frame(env_sea_red))
plot(RDA_sea)
screeplot(RDA_sea, main = "Eigenvalues of constrained axes")

RDA_sea_thin <- rda(genoThinMat ~ ., as.data.frame(env_sea_red))
plot(RDA_sea_thin)
screeplot(RDA_sea_thin, main = "Eigenvalues of constrained axes")

# run rdadapt with K = 2 for first two axes
rdadapt_sea <- rdadapt(RDA = RDA_sea, K = 2)
rdadapt_sea_thin <- rdadapt(RDA = RDA_sea_thin, K = 2)

# bonferroni correction
# P-values threshold after Bonferroni correction
thres_env <- 0.01 / length(rdadapt_sea$p.values)
thres_env_thin <- 0.01 / length(rdadapt_sea_thin$p.values)

# Identifying the loci that are below the p-value threshold
outliers <- (1:ncol(genoMat))[which(rdadapt_sea$p.values < thres_env)]
length(outliers) # 2863 outliers

outliers_thin <- (1:ncol(genoThinMat))[which(rdadapt_sea_thin$p.values < thres_env_thin)]
length(outliers_thin) # 1203 outliers

# Formatting table for ggplot
locus_scores <- as.data.frame(scores(RDA_sea, choices=c(1:2), display="species", scaling="none"))
TAB_loci <- data.frame(names = row.names(locus_scores), locus_scores)
TAB_loci$type <- "Neutral"
TAB_loci$type[outliers] <- "Outliers"
TAB_loci$type <- factor(TAB_loci$type, levels = c("Neutral", "Outliers"))
TAB_loci <- TAB_loci[order(TAB_loci$type),]
TAB_var <- as.data.frame(scores(RDA_sea, choices=c(1,2), display="bp")) 

locus_scores_thin <- as.data.frame(scores(RDA_sea_thin, choices=c(1:2), display="species", scaling="none"))
TAB_loci_thin <- data.frame(names = row.names(locus_scores_thin), locus_scores_thin)
TAB_loci_thin$type <- "Neutral"
TAB_loci_thin$type[outliers_thin] <- "Outliers"
TAB_loci_thin$type <- factor(TAB_loci_thin$type, levels = c("Neutral", "Outliers"))
TAB_loci_thin <- TAB_loci_thin[order(TAB_loci_thin$type),]
TAB_var_thin <- as.data.frame(scores(RDA_sea_thin, choices=c(1,2), display="bp")) 


# Biplot of RDA loci and variables scores
ggplot() +
  geom_hline(yintercept=0, linetype="dashed", color = gray(.80), linewidth=0.6) +
  geom_vline(xintercept=0, linetype="dashed", color = gray(.80), linewidth=0.6) +
  geom_point(data = TAB_loci, aes(x=RDA1*20, y=RDA2*20, colour = type), size = 1.4, alpha = 0.7) +
  scale_color_manual(values = c("gray90", "#EB8055FF")) +
  geom_segment(data = TAB_var, aes(xend=RDA1, yend=RDA2, x=0, y=0), colour="black", size=0.15, linetype=1, arrow=arrow(length = unit(0.02, "npc"))) +
  geom_text(data = TAB_var, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(TAB_var)), size = 2.5, family = "Times") +
  guides(color=guide_legend(title="Locus type")) +
  labs(x = "RDA 1 (83.4%)", y = "RDA 2 (10.4%)", subtitle = "RDA space") +
  theme_bw(base_size = 11, base_family = "Times") +
  theme(panel.grid = element_blank()
        , plot.background = element_blank()
        , panel.background = element_blank())

# plot on thinned geno dat
ggplot() +
  geom_hline(yintercept=0, linetype="dashed", color = gray(.80), linewidth=0.6) +
  geom_vline(xintercept=0, linetype="dashed", color = gray(.80), linewidth=0.6) +
  geom_point(data = TAB_loci_thin, aes(x=RDA1*20, y=RDA2*20, colour = type), size = 1.4, alpha = 0.7) +
  scale_color_manual(values = c("gray90", "#EB8055FF")) +
  geom_segment(data = TAB_var_thin, aes(xend=RDA1, yend=RDA2, x=0, y=0), colour="black", size=0.15, linetype=1, arrow=arrow(length = unit(0.02, "npc"))) +
  geom_text(data = TAB_var_thin, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(TAB_var_thin)), size = 2.5, family = "Times") +
  guides(color=guide_legend(title="Locus type")) +
  labs(x = "RDA 1 (73.7%)", y = "RDA 2 (14.2%)", subtitle = "RDA space") +
  theme_bw(base_size = 11, base_family = "Times") +
  theme(panel.grid = element_blank()
        , plot.background = element_blank()
        , panel.background = element_blank())
```

## Adaptive Index

Now, I will use rdadapt to calculate adaptive index at the two experimental sites (NOT across the range, though I could potentially do this if I decide to impute the seascape environmental data to have a more complete coastal environment map - either way the offsets for each site will be the same). 

NOTE: For now, I am skipping this as it isn't necessary for calculating offsets.

```{r}
# overall, including lew & yrk - I can come back here if I want to do rasterization/interpolation, but it isn't necessary to calculate offset
# RDA_proj_current <- adaptive_index(RDA = RDA_env
#                                        , K = 2
#                                        , env = expEnv_red # alas, this needs to be a raster - must rasterize my environmental data
#                                        , env_mask = NULL
#                                        , method = "loadings")
```


## Offset

Now, I calculate genomic offset for all of my populations!

```{r}
# common garden info
lew_env <- env_exp[!duplicated(env_exp$site_name) ,c("lew_salinity_quantile_10_scaled", "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled")]
yrk_env <- env_exp[!duplicated(env_exp$site_name) ,c("yrk_salinity_quantile_10_scaled", "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")]
colnames(lew_env) <- colnames(yrk_env) <- c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")

env_exp_red <- env_exp[!duplicated(env_exp$site_name) ,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")]

# Estimate genomic offset
offset_lew <- genomic_offset(RDA = RDA_sea, 
                              K = 2, # will need to reassess where I make this cutoff
                              env_pres = env_exp_red, 
                              env_fut = lew_env, 
                              method = "loadings")

offset_yrk <- genomic_offset(RDA = RDA_sea, 
                              K = 2, 
                              env_pres = env_exp_red, 
                              env_fut = yrk_env, 
                              method = "loadings")

# get the experimental long/lat data
env_exp_latlon <- merge(exp_site_latlon, env_exp, by = c("site_name"))

# reassign the metadata
offsets_rda_df <- data.frame(site_name = unique(env_exp_latlon$site_name), lat = unique(env_exp_latlon$lat), lon = unique(env_exp_latlon$lon), offset_lew = offset_lew$genomic_offset$Global, offset_yrk = offset_yrk$genomic_offset$Global)

write.csv(offsets_rda_df, paste0("results/GO_results/offsets_RDA_expSites_",Sys.Date(),".csv", sep = ""))
```

## Visualization

```{r}
# add the color palette
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME"), 
                     order = 1:8, 
                     cols = c("#3e4989" ,"#31688e", "#35b779", "#6ece58", "#440154", "#482878", "#fde725", "#b5de2b"),
                     shape = c(21,21,21,21,23,23,21,21),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME"))

offsets_plotting <- merge(offsets_rda_df, popsdf, by = c("site_name"))

# try mapping the offsets
library(s2) # mapping
library(rnaturalearth) # mapping
library(rnaturalearthdata) # mapping
library(maps) # mapping
library(ggspatial) # mapping

# map
sf::sf_use_s2(FALSE) # sf package doesn't work

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin =-102 , xmax = -68, ymin = 24, ymax = 47.049130))
world_chesapeake <- sf::st_crop(world, c(xmin =-77.856476 , xmax = -72.762787, ymin = 36.019557, ymax = 39.939163))

# bottom right: 24.353558, -55.373595
# top left: 47.049130, -101.983412

# offset to lew
map_lew <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting, aes(x=lon, y=lat, 
                                       fill = offset_lew), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,3.5))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to Lewisetta (RDA)", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_lew)

popout_lew <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting, aes(x=lon, y=lat, 
                                       fill = offset_lew), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,3.5))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
popout_lew

map_yrk <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting, aes(x=lon, y=lat, 
                                       fill = offset_yrk), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,3.5))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to York (RDA)", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_yrk)

popout_york <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting, aes(x=lon, y=lat, 
                                       fill = offset_yrk), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,3.5))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
popout_york

cor(offsets_plotting$offset_lew, offsets_plotting$offset_yrk)
plot(x = offsets_plotting$offset_lew, y = offsets_plotting$offset_yrk)
abline(offsets_plotting$offset_yrk ~ offsets_plotting$offset_lew)
```

