---
title: "CviMVP_oyster_compareGOexp"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "~/Documents/GitHub/MVP_oyster_GO")
```

# Compare Predicted GO to Experimental Results

This script is for analyzing how well various genomic offset methods predict experimentally obtained fitness. I compare LEA, GF, and RDA offsets to experimental mortality, length, and condition index at the conclusion of the experiment. 

## Setup

```{r}
# setwd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(dplyr)
library(tidyr)
library(stringr)
library(vegan)
library(ggplot2)
library(plotrix)

# data
# offset predictions
lea_go <- read.csv("results/GO_results/offsets_LEA_expsites_2025-11-04.csv")[,-1]
rda_go <- read.csv("results/GO_results/offsets_RDA_expsites_2025-10-09.csv")[,-1]
gf_go <- read.csv("results/GO_results/offsets_GF_expSites_2025-08-27.csv")[,-1]

# fitness metrics
mort_bags <- read.csv("results/exp_results/mortality_long_bags_2025-06-06.csv")[,-1]
mort <- read.csv("results/exp_results/mortality_long_popsite_2025-06-06.csv")[,-1]
len <- read.csv("results/exp_results/lengths_bags_event_2025-06-03.csv")[,-1]
```

## Data Prep

Prepare data from offsets & experiment to be directly compared.

```{r}
# inspect data
lea_go
gf_go
rda_go
mort
len

# trim mort
mort_popsite <- mort[!(mort$event == 0), c("bag_site","pop","mean_surv","sd","event")]
colnames(mort_popsite) <- c("bag_site","pop","mean_surv","sd_surv","event")

# also get fitness rankings at each time
# mortality
ord_mort_popsite <- data.frame()
for (i in 1:4) {
  ordt_lew <- mort_popsite[mort_popsite$event == i & mort_popsite$bag_site == "Lewisetta",]
  ordt_lew <- ordt_lew %>% arrange(desc(mean_surv))
  ordt_lew$order_mort <- 1:length(ordt_lew$pop)
  ordt_yrk <- mort_popsite[mort_popsite$event == i & mort_popsite$bag_site == "YorkRiver",]
  ordt_yrk <- ordt_yrk %>% arrange(desc(mean_surv))
  ordt_yrk$order_mort <- 1:length(ordt_yrk$pop)
  ord_mort_popsite <- rbind(ord_mort_popsite, ordt_lew, ordt_yrk)
}
ord_mort_popsite

# do the same with length
# first length needs a pop label and site label
len_bags <- as.data.frame(cbind(len$bag, str_split_fixed(len$bag, "-", 3)[,2:3]))
colnames(len_bags) <- c("bag","pop","bagnum")
len_bags$bagnum <- as.numeric(len_bags$bagnum)
len_bags$site <- ifelse(len_bags$bagnum > 3, "YorkRiver", "Lewisetta")
len_all <- merge(len_bags[,c("bag","pop","site")], len, by = c("bag"))

len_t1 <- len_all[len_all$monitoring_event == 1,]
t1_lens <- aggregate(length ~ pop + site, data = len_t1, FUN = mean)
t1_sd <- aggregate(length ~ pop + site, data = len_t1, FUN = sd)
colnames(t1_sd) <- c("pop", "site", "len_sd")
t1_lens_popsite <- merge(t1_lens, t1_sd, by = c("pop", "site"))
t1_lens_popsite$monitoring_event <- 1
len_t2 <- len_all[len_all$monitoring_event == 2,]
t2_lens <- aggregate(length ~ pop + site, data = len_t2, FUN = mean)
t2_sd <- aggregate(length ~ pop + site, data = len_t2, FUN = sd)
colnames(t2_sd) <- c("pop", "site", "len_sd")
t2_lens_popsite <- merge(t2_lens, t2_sd, by = c("pop", "site"))
t2_lens_popsite$monitoring_event <- 2
len_t3 <- len_all[len_all$monitoring_event == 3,]
t3_lens <- aggregate(length ~ pop + site, data = len_t3, FUN = mean)
t3_sd <- aggregate(length ~ pop + site, data = len_t3, FUN = sd)
colnames(t3_sd) <- c("pop", "site", "len_sd")
t3_lens_popsite <- merge(t3_lens, t3_sd, by = c("pop", "site"))
t3_lens_popsite$monitoring_event <- 3
len_t4 <- len_all[len_all$monitoring_event == 4,]
t4_lens <- aggregate(length ~ pop + site, data = len_t4, FUN = mean)
t4_sd <- aggregate(length ~ pop + site, data = len_t4, FUN = sd)
colnames(t4_sd) <- c("pop", "site", "len_sd")
t4_lens_popsite <- merge(t4_lens, t4_sd, by = c("pop", "site"))
t4_lens_popsite$monitoring_event <- 4

lens_popsite <- rbind(t1_lens_popsite, t2_lens_popsite, t3_lens_popsite, t4_lens_popsite)
colnames(lens_popsite) <- c("pop","bag_site","mean_len","sd_len","event")

# and orders
ord_len_popsite <- data.frame()
for (i in 1:4) {
  ordt_lew <- lens_popsite[lens_popsite$event == i & lens_popsite$bag_site == "Lewisetta",]
  ordt_lew <- ordt_lew %>% arrange(desc(mean_len))
  ordt_lew$order_len <- 1:length(ordt_lew$pop)
  ordt_yrk <- lens_popsite[lens_popsite$event == i & lens_popsite$bag_site == "YorkRiver",]
  ordt_yrk <- ordt_yrk %>% arrange(desc(mean_len))
  ordt_yrk$order_len <- 1:length(ordt_yrk$pop)
  ord_len_popsite <- rbind(ord_len_popsite, ordt_lew, ordt_yrk)
}
ord_len_popsite

# put length and mortality together
ord_fitness_popsite <- merge(ord_mort_popsite, ord_len_popsite, by = c("pop","bag_site","event"))

# need a site name column
ord_fitness_popsite$site_name <- ord_fitness_popsite$pop
map = c("TX"="W1_TX", "LA"="W2_LA", "FL"="W3_FL", "LOLA"="S1_LOLA", "DEBY"="S2_DEBY", "JR"="W4_VA", "NH"="W5_NH", "ME"="W6_ME")
ord_fitness_popsite$site_name <- recode(ord_fitness_popsite$site_name,
                                           "TX"="W1_TX", "LA"="W2_LA", "FL"="W3_FL",
                                           "LOLA"="S1_LOLA", "DEBY"="S2_DEBY", "JR"="W4_VA",
                                           "NH"="W5_NH", "ME"="W6_ME")


# do bag level data as well
colnames(mort_bags) <- c("bag","event","surv")
colnames(len) <- c("bag","event","length")
bagfit <- merge(mort_bags, len, by = c("bag","event"))

# pop info
bagfit$pop <- str_split_fixed(bagfit$bag, "-", 3)[,2]
bagfit

# add this to other fitness data
bagfullfit <- merge(ord_fitness_popsite, bagfit, by = c("pop","event"))

# need to align naming conventions to merge
# LEA
# colnames
colnames(lea_go) <- c("site_name","lew_go_lea","yrk_go_lea","lew_gothin_lea","yrk_gothin_lea")

# add order
order_lew_lea <- as.data.frame(lea_go[order(lea_go$lew_go_lea, decreasing = T),]$site_name)
colnames(order_lew_lea) <- c("site_name")
order_lew_lea$order_lew_lea <- 1:length(order_lew_lea$site_name)
order_yrk_lea <- as.data.frame(lea_go[order(lea_go$yrk_go_lea, decreasing = T),]$site_name)
colnames(order_yrk_lea) <- c("site_name")
order_yrk_lea$order_yrk_lea <- 1:length(order_yrk_lea$site_name)
orders_lea <- merge(order_lew_lea, order_yrk_lea, by = c("site_name"))
lea_go1 <- merge(lea_go, orders_lea, by = c("site_name"))

# GF
# colnames are fine
# add order
order_lew_af <- as.data.frame(gf_go[order(gf_go$offset_gf_af_lew, decreasing = T),]$site_name)
colnames(order_lew_af) <- c("site_name")
order_lew_af$order_lew_af_gf <- 1:length(order_lew_af$site_name)
order_yrk_af <- as.data.frame(lea_go[order(gf_go$offset_gf_af_yrk, decreasing = T),]$site_name)
colnames(order_yrk_af) <- c("site_name")
order_yrk_af$order_yrk_af_gf <- 1:length(order_yrk_af$site_name)
orders_gf <- merge(order_lew_af, order_yrk_af, by = c("site_name"))
gf_go1 <- merge(gf_go, orders_gf, by = c("site_name"))

# RDA
# colnames
colnames(rda_go) <- c("site_name","lat","lon","offset_lew_rda","offset_yrk_rda")

# add order
order_lew_rda <- as.data.frame(rda_go[order(rda_go$offset_lew, decreasing = T),]$site_name)
colnames(order_lew_rda) <- c("site_name")
order_lew_rda$order_lew_rda <- 1:length(order_lew_af$site_name)
order_yrk_rda <- as.data.frame(rda_go[order(rda_go$offset_yrk, decreasing = T),]$site_name)
colnames(order_yrk_rda) <- c("site_name")
order_yrk_rda$order_yrk_rda <- 1:length(order_yrk_rda$site_name)
orders_rda <- merge(order_lew_rda, order_yrk_rda, by = c("site_name"))
rda_go1 <- merge(rda_go, orders_rda, by = c("site_name"))

# put together full go & fitness df
go <- merge(rda_go1, gf_go1, by = "site_name")
go1 <- merge(go, lea_go1, by = "site_name")
fullfit <- merge(bagfullfit, go1, by = "site_name")
```


## Compare Offsets to Fitness

Now that offsets and fitness are in a comparable format, look at how well the two match.

### Performance Metric Calculation

Here I calculate the R^2 and pearson's correlation between fitness and GO for each method at each site. I do this with fitness data both at the bag level and summarized at the population level (since offset is population level).

```{r}
# correlations between go/fit at each time point
#################################################
# function for correlations
cor_fit <- function(site, go_method) {
  cor_df <- data.frame(corr_mort = NA, r2_mort = NA, corr_len = NA, r2_len = NA, corrpop_mort = NA, r2pop_mort = NA, corrpop_len = NA, r2pop_len = NA, method = NA, bag_site = NA, event = NA)
  go_metric <- ifelse(go_method == "lea" & site == "Lewisetta", "lew_go_lea", ifelse(go_method == "gf" & site == "Lewisetta", "offset_gf_af_lew", ifelse(go_method == "rda" & site == "Lewisetta", "offset_lew_rda", ifelse(go_method == "lea" & site == "YorkRiver", "yrk_go_lea", ifelse(go_method == "gf" & site == "YorkRiver", "offset_gf_af_yrk", ifelse(go_method == "rda" & site == "YorkRiver", "offset_yrk_rda", NULL))))))

  for (i in 1:4) {
    cor_df[i,]$corr_mort <- cor(fullfit[fullfit$bag_site == (site) & fullfit$event == i, c(go_metric)], fullfit[fullfit$bag_site == (site) & fullfit$event == i,]$surv)
    cor_df[i,]$r2_mort <- summary(lm(fullfit[fullfit$bag_site == (site) & fullfit$event == i,]$surv ~ fullfit[fullfit$bag_site == (site) & fullfit$event == i, c(go_metric)]))$r.squared
    
    cor_df[i,]$corrpop_mort <- cor(fullfit[fullfit$bag_site == (site) & fullfit$event == i, c(go_metric)], fullfit[fullfit$bag_site == (site) & fullfit$event == i,]$mean_surv)
    cor_df[i,]$r2pop_mort <- summary(lm(fullfit[fullfit$bag_site == (site) & fullfit$event == i,]$mean_surv ~ fullfit[fullfit$bag_site == (site) & fullfit$event == i, c(go_metric)]))$r.squared
    
    cor_df[i,]$corr_len <- cor(fullfit[fullfit$bag_site == (site) & fullfit$event == i, c(go_metric)], fullfit[fullfit$bag_site == (site) & fullfit$event == i,]$length)
    cor_df[i,]$r2_len <- summary(lm(fullfit[fullfit$bag_site == (site) & fullfit$event == i,]$length ~ fullfit[fullfit$bag_site == (site) & fullfit$event == i, c(go_metric)]))$r.squared
    
    cor_df[i,]$corrpop_len <- cor(fullfit[fullfit$bag_site == (site) & fullfit$event == i, c(go_metric)], fullfit[fullfit$bag_site == (site) & fullfit$event == i,]$mean_len)
    cor_df[i,]$r2pop_len <- summary(lm(fullfit[fullfit$bag_site == (site) & fullfit$event == i,]$mean_len ~ fullfit[fullfit$bag_site == (site) & fullfit$event == i, c(go_metric)]))$r.squared
    
    cor_df[i,]$method <- as.character(go_method)
    cor_df[i,]$bag_site <- as.character(site)
    cor_df[i,]$event <- i
  }
  return(cor_df)
}

# correlations
cor_lea_lew <- cor_fit("Lewisetta","lea")
cor_lea_yrk <- cor_fit("YorkRiver","lea")
cor_gf_af_lew <- cor_fit("Lewisetta","gf")
cor_gf_af_yrk <- cor_fit("YorkRiver","gf")
cor_rda_lew <- cor_fit("Lewisetta","rda")
cor_rda_yrk <- cor_fit("YorkRiver","rda")

corrs_df <- rbind(cor_lea_lew, cor_lea_yrk, cor_gf_af_lew, cor_gf_af_yrk, cor_rda_lew, cor_rda_yrk)
rownames(corrs_df) <- NULL

# save this
write.csv(corrs_df, paste0("results/GO_results/go_fit_corrs_",Sys.Date(),".csv"), row.names = F)
#################################################
```


### Plot GO vs mortality against each other

I'll start by plotting LEA results:

```{r}
# plot lea vs mort at each site/time point
goFitLEA_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = lew_go_lea, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = lew_go_lea, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = lew_go_lea, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = lew_go_lea, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = lew_go_lea, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = lew_go_lea, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Mortality at Lewisetta", x = "LEA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_lew_plotbag

goFitLEA_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = yrk_go_lea, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = yrk_go_lea, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = yrk_go_lea, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = yrk_go_lea, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = yrk_go_lea, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = yrk_go_lea, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Mortality at York River", x = "LEA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_plot

# plot go lea vs mort at both sites just final time point
goFitLEA_plotfinalT <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = surv), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = surv), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (LEA) vs Mortality at Final Time Point", x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotfinalT
```


Next, I'll plot gradient forest results:

```{r}
# plot lea vs mort at each site/time point
goFitGF_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (GFaf) vs Mortality at Lewisetta", x = "GF Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_lew_plotbag

goFitGF_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (GFaf) vs Mortality at York River", x = "GF Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_yrk_plot

# plot go gf vs mort at both sites just final time point
goFitGF_plotfinalT <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (GFaf) vs Mortality at Final Time Point", x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGF_plotfinalT
```


Finally, I'll plot rda results:

```{r}
# plot lea vs mort at each site/time point
goFitRDA_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lew_rda, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lew_rda, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lew_rda, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lew_rda, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lew_rda, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lew_rda, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Mortality at Lewisetta", x = "RDA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_lew_plotbag

goFitRDA_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_yrk_rda, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_yrk_rda, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_yrk_rda, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_yrk_rda, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_yrk_rda, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_yrk_rda, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Mortality at York River", x = "RDA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_yrk_plot

# plot go gf vs mort at both sites just final time point
goFitRDA_plotfinalT <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = surv), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = surv), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (RDA) vs Mortality at Final Time Point", x = "RDA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitRDA_plotfinalT
```


### Plot GO vs length against each other

```{r}
# plot lea vs len at each site/time point
goFitLEA_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = lew_go_lea, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = lew_go_lea, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = lew_go_lea, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = lew_go_lea, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = lew_go_lea, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = lew_go_lea, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Size at Lewisetta", x = "LEA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_lew_len

goFitLEA_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = yrk_go_lea, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = yrk_go_lea, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = yrk_go_lea, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = yrk_go_lea, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = yrk_go_lea, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = yrk_go_lea, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Size at York River", x = "LEA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_len

# plot go lea vs len at both sites just final time point
goFitLEA_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = length), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = length), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (LEA) vs Mortality at Final Time Point", x = "LEA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotfinalTlen

# plot gf vs len at each site/time point
goFitGF_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (GF) vs Size at Lewisetta", x = "GF Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_lew_len

goFitGF_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (GF) vs Size at York River", x = "GF Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGF_yrk_len

# plot gf vs len at both sites just final time point
goFitGF_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (GF) vs Size at Final Time Point", x = "GF Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGF_plotfinalTlen

# plot rda vs len at each site/time point
goFitRDA_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lew_rda, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lew_rda, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lew_rda, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lew_rda, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lew_rda, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lew_rda, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Size at Lewisetta", x = "RDA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_lew_len

goFitRDA_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_yrk_rda, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_yrk_rda, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_yrk_rda, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_yrk_rda, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_yrk_rda, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_yrk_rda, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Size at York River", x = "RDA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_yrk_len

# plot rda vs len at both sites just final time point
goFitRDA_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = length), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = length), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (RDA) vs Size at Final Time Point", x = "RDA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitRDA_plotfinalTlen
```


### Environmental Distance

Does environmental distance between home sites and common garden conditions predict mortality better than genomic offset does? 

```{r}
# read in environmental distance matrix (comes from "EnvDatProcess")
env_dist_mat <- read.csv("data/EnvDat/exp/distance_matrix_2025-09-01.csv")[,-c(1:2)]
rownames(env_dist_mat) <- colnames(env_dist_mat)
env_dist <- read.csv("data/EnvDat/exp/distance_matrix_longer_2025-09-01.csv")[-1]

# check env dist and fitness data
env_dist
mort_bags

# keep just comparisons to experimental sites in dist df
env_dist_exp <- env_dist[env_dist$site1 %in% c("YorkRiver", "Lewisetta"),]
env_dist_exprm <- env_dist_exp[!(env_dist_exp$site2 %in% c("YorkRiver", "Lewisetta")), c("site2", "site1", "dist")]
colnames(env_dist_exprm) <- c("site_name","bag_site","euc_dist")

# add popsdf
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME", "LARMIX", "SEEDMIX"), 
                     order_pop = 1:10, 
                     cols = c("#3e4989" ,"#31688e", "#26828e", "#440154", "#482878", "#1f9e89", "#35b779", "#6ece58", "#fde725", "#b5de2b"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME", "H1_LARMIX", "H2_SEEDMIX"))
fit_popsdf <- merge(fullfit, popsdf, by = c("pop","site_name"))

# merge
distFit <- merge(fit_popsdf, env_dist_exprm, by = c("site_name", "bag_site"))

# compare fitness to environmental distance
# mortality across all times
cor.test(distFit$mean_surv, distFit$euc_dist) # -0.2299484 pval 0.0003686
cor.test(distFit[distFit$bag_site == "Lewisetta",]$mean_surv, distFit[distFit$bag_site == "Lewisetta",]$euc_dist) # -0.2588923 pval 0.004643
cor.test(distFit[distFit$bag_site == "YorkRiver",]$mean_surv, distFit[distFit$bag_site == "YorkRiver",]$euc_dist) # 0.09547117 pval 0.3038

cor_dist_lew <- data.frame(corr_mort = NA, r2_mort = NA, corr_len = NA, r2_len = NA, corrpop_mort = NA, r2pop_mort = NA, corrpop_len = NA, r2pop_len = NA, method = NA, bag_site = NA, event = NA)
for (i in 1:4) {
  cor_dist_lew[i,]$corr_mort <- cor(distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$euc_dist, distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$surv)
  cor_dist_lew[i,]$r2_mort <- summary(lm(distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$surv ~ distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$euc_dist))$r.squared
  
  cor_dist_lew[i,]$corrpop_mort <- cor(distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$euc_dist, distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$mean_surv)
  cor_dist_lew[i,]$r2pop_mort <- summary(lm(distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$mean_surv ~ distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$euc_dist))$r.squared
  
  cor_dist_lew[i,]$corr_len <- cor(distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$euc_dist, distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$length)
  cor_dist_lew[i,]$r2_len <- summary(lm(distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$length ~ distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$euc_dist))$r.squared
  
  cor_dist_lew[i,]$corrpop_len <- cor(distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$euc_dist, distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$mean_len)
  cor_dist_lew[i,]$r2pop_len <- summary(lm(distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$mean_len ~ distFit[distFit$bag_site == "Lewisetta" & distFit$event == i,]$euc_dist))$r.squared
  
  cor_dist_lew[i,]$method <- "Euclidean"
  cor_dist_lew[i,]$bag_site <- "Lewisetta"
  cor_dist_lew[i,]$event <- i
}

cor_dist_yrk <- data.frame(corr_mort = NA, r2_mort = NA, corr_len = NA, r2_len = NA, corrpop_mort = NA, r2pop_mort = NA, corrpop_len = NA, r2pop_len = NA, method = NA, bag_site = NA, event = NA)
for (i in 1:4) {
  cor_dist_yrk[i,]$corr_mort <- cor(distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$euc_dist, distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$surv)
  cor_dist_yrk[i,]$r2_mort <- summary(lm(distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$surv ~ distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$euc_dist))$r.squared
  
  cor_dist_yrk[i,]$corrpop_mort <- cor(distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$euc_dist, distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$mean_surv)
  cor_dist_yrk[i,]$r2pop_mort <- summary(lm(distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$mean_surv ~ distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$euc_dist))$r.squared
  
  cor_dist_yrk[i,]$corr_len <- cor(distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$euc_dist, distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$length)
  cor_dist_yrk[i,]$r2_len <- summary(lm(distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$length ~ distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$euc_dist))$r.squared
  
  cor_dist_yrk[i,]$corrpop_len <- cor(distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$euc_dist, distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$mean_len)
  cor_dist_yrk[i,]$r2pop_len <- summary(lm(distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$mean_len ~ distFit[distFit$bag_site == "YorkRiver" & distFit$event == i,]$euc_dist))$r.squared
  
  cor_dist_yrk[i,]$method <- "Euclidean"
  cor_dist_yrk[i,]$bag_site <- "YorkRiver"
  cor_dist_yrk[i,]$event <- i
}

# add these correlations to corrs_df
corrs_df_dist <- rbind(corrs_df, cor_dist_lew, cor_dist_yrk)
write.csv(corrs_df_dist, paste0("results/GO_results/go_fit_corrs_dist_",Sys.Date(),".csv"), row.names = F)

# visualize mort
# lewisetta
envdist_mort_plot_lew <- ggplot(data = distFit[distFit$bag_site == "Lewisetta",]) + 
  geom_point(data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 1,], aes(x = euc_dist, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 1,], aes(x = euc_dist, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 2,], aes(x = euc_dist, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 2,], aes(x = euc_dist, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 3,], aes(x = euc_dist, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 3,], aes(x = euc_dist, y = surv), col = "slateblue", se = T) +
  geom_point(data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 4,], aes(x = euc_dist, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 4,], aes(x = euc_dist, y = surv), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Mortality at Lewisetta", x = "Euclidean Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  ylim(0,1)+
  theme_classic()
envdist_mort_plot_lew

# york river
envdist_mort_plot_yrk <- ggplot(data = distFit[distFit$bag_site == "YorkRiver",]) + 
  geom_point(data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 1,], aes(x = euc_dist, y = surv, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 1,], aes(x = euc_dist, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 2,], aes(x = euc_dist, y = surv, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 2,], aes(x = euc_dist, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 3,], aes(x = euc_dist, y = surv, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 3,], aes(x = euc_dist, y = surv), col = "slateblue", se = T) +
  geom_point(data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 4,], aes(x = euc_dist, y = surv, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 4,], aes(x = euc_dist, y = surv), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Mortality at York River", x = "Euclidean Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  ylim(0,1)+
  theme_classic()
envdist_mort_plot_yrk

# compare
envdist_mort_plotT4 <- ggplot(data = distFit) +
  geom_point(data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 4,], aes(x = euc_dist, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 4,], aes(x = euc_dist, y = surv), col = "mediumspringgreen", se = T) +
  geom_point(data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 4,], aes(x = euc_dist, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 4,], aes(x = euc_dist, y = surv), col = "dodgerblue", se = T) +
  labs(title = "Environmental Distance vs Mortality at Final Time Point", x = "Euclidean Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  ylim(0,1)+
  theme_classic()
envdist_mort_plotT4

# visualize length
# lewisetta
envdist_len_plot_lew <- ggplot(data = distFit[distFit$bag_site == "Lewisetta",]) + 
  geom_point(data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 1,], aes(x = euc_dist, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 1,], aes(x = euc_dist, y = length), col = "cadetblue3", se = T) +
  geom_point(data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 2,], aes(x = euc_dist, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 2,], aes(x = euc_dist, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 3,], aes(x = euc_dist, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 3,], aes(x = euc_dist, y = length), col = "slateblue", se = T) +
  geom_point(data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 4,], aes(x = euc_dist, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 4,], aes(x = euc_dist, y = length), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Length at Lewisetta", x = "Euclidean Environmental Distance", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
envdist_len_plot_lew

# york river
envdist_len_plot_yrk <- ggplot(data = distFit[distFit$bag_site == "YorkRiver",]) + 
  geom_point(data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 1,], aes(x = euc_dist, y = length, fill = "Fall'23"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 1,], aes(x = euc_dist, y = length), col = "cadetblue3", se = T) +
  geom_point(data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 2,], aes(x = euc_dist, y = length, fill = "Spring'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 2,], aes(x = euc_dist, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 3,], aes(x = euc_dist, y = length, fill = "Fall'24"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 3,], aes(x = euc_dist, y = length), col = "slateblue", se = T) +
  geom_point(data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 4,], aes(x = euc_dist, y = length, fill = "Spring'25"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 4,], aes(x = euc_dist, y = length), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Length at York River", x = "Euclidean Environmental Distance", y = "Length (mm)") +
  scale_fill_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
envdist_len_plot_yrk

# compare
envdist_len_plotT4 <- ggplot(data = distFit) +
  geom_point(data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 4,], aes(x = euc_dist, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "Lewisetta" & distFit$event == 4,], aes(x = euc_dist, y = length), col = "mediumspringgreen", se = T) +
  geom_point(data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 4,], aes(x = euc_dist, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = distFit[distFit$bag_site == "YorkRiver" & distFit$event == 4,], aes(x = euc_dist, y = length), col = "dodgerblue", se = T) +
  labs(title = "Environmental Distance vs Length at Final Time Point", x = "Euclidean Environmental Distance", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
envdist_len_plotT4
```


## Compare Offset Methods

Now that I have predictions from different methods, I will look at how offset predictions made by these methods compare to each other and how different their ability is to predict experimental fitness.


### Compare Offsets to Each Other

```{r Offset Comparison}
# corr gf/lea
cor.test(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$lew_go_lea,  fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$offset_gf_af_lew) # -0.08873151, p = 0.6068
cor.test(fullfit[fullfit$bag_site == "YorkRiver",]$yrk_go_lea,  fullfit[fullfit$bag_site == "YorkRiver",]$offset_gf_af_yrk) # 0.4729701, p = 2.788e-09

ggplot(data = fullfit) + 
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = lew_go_lea, y = offset_gf_af_lew, fill = pop), pch = 21, size = 3) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = lew_go_lea, y = offset_gf_af_lew), col = "black") +
  labs(title = "LEA vs GF Offsets Lewisetta", x = "LEA Offset", y = "GF Offset", fill = "Broodstock Group") +
  theme_classic()

ggplot(data = fullfit) + 
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = yrk_go_lea, y = offset_gf_af_yrk, fill = pop), pch = 21, size = 3) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = yrk_go_lea, y = offset_gf_af_yrk), col = "black") +
  labs(title = "LEA vs GF Offsets York River", x = "LEA Offset", y = "GF Offset", fill = "Broodstock Group") +
  theme_classic()

# corr gf/rda
cor.test(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$offset_lew_rda,  fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$offset_gf_af_lew) # 0.03513667, p = 0.8388
cor.test(fullfit[fullfit$bag_site == "YorkRiver",]$offset_yrk_rda,  fullfit[fullfit$bag_site == "YorkRiver",]$offset_gf_af_yrk) # 0.2381278, p = 0.004321

# corr lea/rda
cor.test(fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$offset_lew_rda,  fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,]$lew_go_lea) # 0.08615492, p = 0.6173
cor.test(fullfit[fullfit$bag_site == "YorkRiver",]$offset_lew_rda,  fullfit[fullfit$bag_site == "YorkRiver",]$yrk_go_lea) # -0.1736073, p = 0.03881
```

