---
title: "CviMVP_oyster_compareGOexp"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "~/Documents/GitHub/MVP_oyster_GO")
```

# Compare Predicted GO to Experimental Results

This script is for analyzing how well various genomic offset methods predict experimentally obtained fitness. I compare LEA, GF, and RDA offsets to experimental mortality, length, and condition index at the conclusion of the experiment. 

## Setup

```{r}
# setwd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(dplyr)
library(tidyr)
library(stringr)
library(vegan)
library(ggplot2)
library(plotrix)

# data
# offset predictions
lea_go <- read.csv("results/GO_results/offsets_LEA_expsites_2025-11-10.csv")[,-1]
rda_go <- read.csv("results/GO_results/offsets_RDA_expsites_2025-11-10.csv")[,-1]
gf_go <- read.csv("results/GO_results/offsets_GF_expSites_2025-11-19.csv")[,-1]
env_dist <- read.csv("data/EnvDat/exp/distance_matrix_2025-10-30.csv")[28:35,c("X","Lewisetta","YorkRiver")]

# fitness metrics
mort_bags <- read.csv("results/exp_results/mortality_long_bags_2025-06-06.csv")[,-1]
mort <- read.csv("results/exp_results/mortality_long_popsite_2025-06-06.csv")[,-1]
len <- read.csv("results/exp_results/lengths_bags_event_2025-06-03.csv")[,-1]
```

## Data Prep

Prepare data from offsets & experiment to be directly compared.

```{r}
# inspect data
lea_go
gf_go
rda_go
mort
len

# trim mort
mort_popsite <- mort[!(mort$event == 0), c("bag_site","pop","mean_surv","sd","event")]
colnames(mort_popsite) <- c("bag_site","pop","mean_surv","sd_surv","event")

# also get fitness rankings at each time
# mortality
ord_mort_popsite <- data.frame()
for (i in 1:4) {
  ordt_lew <- mort_popsite[mort_popsite$event == i & mort_popsite$bag_site == "Lewisetta",]
  ordt_lew <- ordt_lew %>% arrange(desc(mean_surv))
  ordt_lew$order_mort <- 1:length(ordt_lew$pop)
  ordt_yrk <- mort_popsite[mort_popsite$event == i & mort_popsite$bag_site == "YorkRiver",]
  ordt_yrk <- ordt_yrk %>% arrange(desc(mean_surv))
  ordt_yrk$order_mort <- 1:length(ordt_yrk$pop)
  ord_mort_popsite <- rbind(ord_mort_popsite, ordt_lew, ordt_yrk)
}
ord_mort_popsite

# do the same with length
# first length needs a pop label and site label
len_bags <- as.data.frame(cbind(len$bag, str_split_fixed(len$bag, "-", 3)[,2:3]))
colnames(len_bags) <- c("bag","pop","bagnum")
len_bags$bagnum <- as.numeric(len_bags$bagnum)
len_bags$site <- ifelse(len_bags$bagnum > 3, "YorkRiver", "Lewisetta")
len_all <- merge(len_bags[,c("bag","pop","site")], len, by = c("bag"))
colnames(len_all) <- c("bag","pop","site","event","length")

len_t1 <- len_all[len_all$event == 1,]
t1_lens <- aggregate(length ~ pop + site, data = len_t1, FUN = mean)
t1_sd <- aggregate(length ~ pop + site, data = len_t1, FUN = sd)
colnames(t1_sd) <- c("pop", "site", "len_sd")
t1_lens_popsite <- merge(t1_lens, t1_sd, by = c("pop", "site"))
t1_lens_popsite$event <- 1
len_t2 <- len_all[len_all$event == 2,]
t2_lens <- aggregate(length ~ pop + site, data = len_t2, FUN = mean)
t2_sd <- aggregate(length ~ pop + site, data = len_t2, FUN = sd)
colnames(t2_sd) <- c("pop", "site", "len_sd")
t2_lens_popsite <- merge(t2_lens, t2_sd, by = c("pop", "site"))
t2_lens_popsite$event <- 2
len_t3 <- len_all[len_all$event == 3,]
t3_lens <- aggregate(length ~ pop + site, data = len_t3, FUN = mean)
t3_sd <- aggregate(length ~ pop + site, data = len_t3, FUN = sd)
colnames(t3_sd) <- c("pop", "site", "len_sd")
t3_lens_popsite <- merge(t3_lens, t3_sd, by = c("pop", "site"))
t3_lens_popsite$event <- 3
len_t4 <- len_all[len_all$event == 4,]
t4_lens <- aggregate(length ~ pop + site, data = len_t4, FUN = mean)
t4_sd <- aggregate(length ~ pop + site, data = len_t4, FUN = sd)
colnames(t4_sd) <- c("pop", "site", "len_sd")
t4_lens_popsite <- merge(t4_lens, t4_sd, by = c("pop", "site"))
t4_lens_popsite$event <- 4

lens_popsite <- rbind(t1_lens_popsite, t2_lens_popsite, t3_lens_popsite, t4_lens_popsite)
colnames(lens_popsite) <- c("pop","bag_site","mean_len","sd_len","event")

# and orders
ord_len_popsite <- data.frame()
for (i in 1:4) {
  ordt_lew <- lens_popsite[lens_popsite$event == i & lens_popsite$bag_site == "Lewisetta",]
  ordt_lew <- ordt_lew %>% arrange(desc(mean_len))
  ordt_lew$order_len <- 1:length(ordt_lew$pop)
  ordt_yrk <- lens_popsite[lens_popsite$event == i & lens_popsite$bag_site == "YorkRiver",]
  ordt_yrk <- ordt_yrk %>% arrange(desc(mean_len))
  ordt_yrk$order_len <- 1:length(ordt_yrk$pop)
  ord_len_popsite <- rbind(ord_len_popsite, ordt_lew, ordt_yrk)
}
ord_len_popsite

# put length and mortality together
ord_fitness_popsite_mg <- merge(ord_mort_popsite, ord_len_popsite, by = c("pop","bag_site","event"))

# remove the larmix/seedmix pops
ord_fitness_popsite <- ord_fitness_popsite_mg[!ord_fitness_popsite_mg$pop %in% c("LARMIX","SEEDMIX"),]

# need a site name column
ord_fitness_popsite$site_name <- ord_fitness_popsite$pop
ord_fitness_popsite$site_name <- dplyr::recode(ord_fitness_popsite$site_name,
                                        "TX" = "W1_TX", 
                                        "LA" = "W2_LA", 
                                        "FL" = "W3_FL",
                                        "LOLA" = "S1_LOLA", 
                                        "DEBY" = "S2_DEBY",
                                        "JR" = "W4_VA",
                                        "NH" = "W5_NH", 
                                        "ME" = "W6_ME")

# do bag level data as well
colnames(mort_bags) <- c("bag","event","surv")
colnames(len) <- c("bag","event","length")
bagfit <- merge(mort_bags, len, by = c("bag","event"))

# pop info
bagfit$pop <- str_split_fixed(bagfit$bag, "-", 3)[,2]
bagfit

# site info
bagfit$bag_site <- ifelse(str_split_fixed(bagfit$bag, "-", 3)[,3] %in% c(1,2,3), "Lewisetta","YorkRiver")

# add this to other fitness data
bagfullfit <- merge(ord_fitness_popsite, bagfit, by = c("pop","event","bag_site"))

# need to align naming conventions to merge
# LEA
# colnames
colnames(lea_go) <- c("site_name", "lat", "lon", "lew_go_lea","yrk_go_lea","lew_gothin_lea","yrk_gothin_lea")

# add order
order_lew_lea <- as.data.frame(lea_go[order(lea_go$lew_go_lea, decreasing = T),]$site_name)
colnames(order_lew_lea) <- c("site_name")
order_lew_lea$order_lew_lea <- 1:length(order_lew_lea$site_name)
order_yrk_lea <- as.data.frame(lea_go[order(lea_go$yrk_go_lea, decreasing = T),]$site_name)
colnames(order_yrk_lea) <- c("site_name")
order_yrk_lea$order_yrk_lea <- 1:length(order_yrk_lea$site_name)
orders_lea <- merge(order_lew_lea, order_yrk_lea, by = c("site_name"))
lea_go1 <- merge(lea_go, orders_lea, by = c("site_name"))

# GF
# add order
order_lew_af <- as.data.frame(gf_go[order(gf_go$offset_gf_af_lew, decreasing = T),]$site_name)
colnames(order_lew_af) <- c("site_name")
order_lew_af$order_lew_af_gf <- 1:length(order_lew_af$site_name)
order_yrk_af <- as.data.frame(lea_go[order(gf_go$offset_gf_af_yrk, decreasing = T),]$site_name)
colnames(order_yrk_af) <- c("site_name")
order_yrk_af$order_yrk_af_gf <- 1:length(order_yrk_af$site_name)
orders_gf_af <- merge(order_lew_af, order_yrk_af, by = c("site_name"))
gf_go1 <- merge(gf_go, orders_gf_af, by = c("site_name"))

order_lew_geno <- as.data.frame(gf_go[order(gf_go$offset_gf_geno_lew, decreasing = T),]$site_name)
colnames(order_lew_geno) <- c("site_name")
order_lew_geno$order_lew_geno_gf <- 1:length(order_lew_geno$site_name)
order_yrk_geno <- as.data.frame(lea_go[order(gf_go$offset_gf_geno_yrk, decreasing = T),]$site_name)
colnames(order_yrk_geno) <- c("site_name")
order_yrk_geno$order_yrk_geno_gf <- 1:length(order_yrk_geno$site_name)
orders_gf_geno <- merge(order_lew_geno, order_yrk_geno, by = c("site_name"))
gf_go2 <- merge(gf_go1, orders_gf_geno, by = c("site_name"))

# RDA
# colnames
colnames(rda_go) <- c("site_name","lat","lon","offset_lew_rda","offset_yrk_rda")

# add order
order_lew_rda <- as.data.frame(rda_go[order(rda_go$offset_lew, decreasing = T),]$site_name)
colnames(order_lew_rda) <- c("site_name")
order_lew_rda$order_lew_rda <- 1:length(order_lew_rda$site_name)
order_yrk_rda <- as.data.frame(rda_go[order(rda_go$offset_yrk, decreasing = T),]$site_name)
colnames(order_yrk_rda) <- c("site_name")
order_yrk_rda$order_yrk_rda <- 1:length(order_yrk_rda$site_name)
orders_rda <- merge(order_lew_rda, order_yrk_rda, by = c("site_name"))
rda_go1 <- merge(rda_go, orders_rda, by = c("site_name"))
colnames(rda_go1) <- c("site_name","lat","lon","offset_lew_rda","offset_yrk_rda","order_lew_rda","order_yrk_rda")

# environmental distance
colnames(env_dist) <- c("site_name", "env_dist_lew", "env_dist_yrk")
# add order
order_lew_envdist <- as.data.frame(env_dist[order(env_dist$env_dist_lew, decreasing = T),]$site_name)
colnames(order_lew_envdist) <- "site_name"
order_lew_envdist$order_lew_envdist <- 1:length(order_lew_envdist$site_name)
order_yrk_envdist <- as.data.frame(env_dist[order(env_dist$env_dist_yrk, decreasing = T),]$site_name)
colnames(order_yrk_envdist) <- "site_name"
order_yrk_envdist$order_yrk_envdist <- 1:length(order_yrk_envdist$site_name)
orders_envdist <- merge(order_lew_envdist, order_yrk_envdist, by = "site_name")
env_dist1 <- merge(env_dist, orders_envdist, by = "site_name")

# put together full go & fitness df
go <- merge(rda_go1, gf_go2, by = c("site_name","lat","lon")) 
go1 <- merge(go, lea_go1, by = c("site_name", "lat", "lon"))
go2 <- merge(go1, env_dist1, by = "site_name")
fullfit <- merge(bagfullfit, go2, by = "site_name")
```


## Compare Offsets to Fitness

Now that offsets and fitness are in a comparable format, look at how well the two match.

### Performance Metric Calculation

Here I calculate the R^2 and pearson's correlation between fitness and GO for each method at each site. I do this with fitness data both at the bag level and summarized at the population level (since offset is population level).

```{r}
# correlations between go/fit at each time point
#################################################
# function for correlations
cor_fit <- function(site, go_method, df) {
  cor_df <- data.frame(pearson_surv = NA, pearson_surv_pval = NA, spearman_surv = NA, spearman_surv_pval = NA, kendall_surv = NA, kendall_surv_pval = NA, r2_surv = NA, r2_surv_rankings = NA, pearson_len = NA, pearson_len_pval = NA, spearman_len = NA, spearman_len_pval = NA, kendall_len = NA, kendall_len_pval = NA, r2_len = NA, r2_len_rankings = NA, method = NA, bag_site = NA, event = NA)
  
  # what metric should we use based on method input?
  lookup_metric <- list(
  Lewisetta = list(lea = "lew_gothin_lea", gf_af = "offset_gf_af_lew",
                   gf_geno = "offset_gf_geno_lew", rda = "offset_lew_rda", 
                   env_dist = "env_dist_lew"),
  YorkRiver = list(lea = "yrk_gothin_lea", gf_af = "offset_gf_af_yrk", 
                   gf_geno = "offset_gf_geno_yrk", rda = "offset_yrk_rda", 
                   env_dist = "env_dist_yrk"))

  lookup_rank <- list(Lewisetta = list(lea = "order_lew_lea", gf_af = "offset_gf_af_lew",
                   gf_geno = "order_lew_geno_gf", rda = "order_lew_rda", 
                   env_dist = "order_lew_envdist"),
  YorkRiver = list(lea = "order_yrk_lea", gf_af = "offset_gf_af_yrk", 
                   gf_geno = "order_yrk_geno_gf", rda = "order_yrk_rda", 
                   env_dist = "order_yrk_envdist"))
  go_metric <- lookup_metric[[site]][[go_method]]
  go_metric_rank <- lookup_rank[[site]][[go_method]]

  for (i in 1:4) {
    fitdat <- df[df$bag_site == (site) & df$event == i,]
    
    # survival
    cor_df[i,]$pearson_surv <- cor.test(fitdat[,go_metric], fitdat$surv, method = "pearson")$estimate
    cor_df[i,]$pearson_surv_pval <- cor.test(fitdat[,go_metric], fitdat$surv, method = "pearson", exact = T)$p.value
    cor_df[i,]$spearman_surv <- cor.test(fitdat[,go_metric], fitdat$surv, method = "spearman")$estimate
    cor_df[i,]$spearman_surv_pval <- cor.test(fitdat[,go_metric], fitdat$surv, method = "spearman", exact = T)$p.value
    cor_df[i,]$kendall_surv <- cor.test(as.numeric(fitdat[,go_metric_rank]), as.numeric(fitdat$order_mort), method = "kendall")$estimate
    cor_df[i,]$kendall_surv_pval <- cor.test(fitdat[,go_metric], fitdat$surv, method = "kendall", exact = T)$p.value
    cor_df[i,]$r2_surv <- summary(lm(fitdat$surv ~ fitdat[,go_metric]))$r.squared
    cor_df[i,]$r2_surv_rankings <- summary(lm(fitdat$order_mort ~ fitdat[,go_metric_rank]))$r.squared
    
    # length
    cor_df[i,]$pearson_len <- cor.test(fitdat[,go_metric], fitdat$length, method = "pearson")$estimate
    cor_df[i,]$pearson_len_pval <- cor.test(fitdat[,go_metric], fitdat$length, method = "pearson", exact = T)$p.value
    cor_df[i,]$spearman_len <- cor.test(fitdat[,go_metric], fitdat$length, method = "spearman")$estimate
    cor_df[i,]$spearman_len_pval <- cor.test(fitdat[,go_metric], fitdat$length, method = "spearman", exact = T)$p.value
    cor_df[i,]$kendall_len <- cor.test(as.numeric(fitdat[,go_metric_rank]), as.numeric(fitdat$order_len), method = "kendall")$estimate
    cor_df[i,]$kendall_len_pval <- cor.test(fitdat[,go_metric], fitdat$length, method = "kendall", exact = T)$p.value
    cor_df[i,]$r2_len <- summary(lm(fitdat$length ~ fitdat[,go_metric]))$r.squared
    cor_df[i,]$r2_len_rankings <- summary(lm(fitdat$order_len ~ fitdat[,go_metric_rank]))$r.squared
    
    # metadata
    cor_df[i,]$method <- as.character(go_method)
    cor_df[i,]$bag_site <- as.character(site)
    cor_df[i,]$event <- i
  }
  return(cor_df)
}

# correlations
(cor_lea_lew <- cor_fit("Lewisetta","lea", fullfit))
(cor_lea_yrk <- cor_fit("YorkRiver","lea", fullfit))
(cor_gf_af_lew <- cor_fit("Lewisetta","gf_af", fullfit))
(cor_gf_af_yrk <- cor_fit("YorkRiver","gf_af", fullfit))
(cor_gf_geno_lew <- cor_fit("Lewisetta","gf_geno", fullfit))
(cor_gf_geno_yrk <- cor_fit("YorkRiver","gf_geno", fullfit))
(cor_rda_lew <- cor_fit("Lewisetta","rda", fullfit))
(cor_rda_yrk <- cor_fit("YorkRiver","rda", fullfit))
(cor_envdist_lew <- cor_fit("Lewisetta", "env_dist", fullfit))
(cor_envdist_yrk <- cor_fit("YorkRiver", "env_dist", fullfit))

corrs_df <- rbind(cor_lea_lew, cor_lea_yrk, cor_gf_af_lew, cor_gf_af_yrk, cor_gf_geno_lew, cor_gf_geno_yrk, cor_rda_lew, cor_rda_yrk, cor_envdist_lew, cor_envdist_yrk)
rownames(corrs_df) <- NULL
corrs_df$signif_surv <- ifelse(corrs_df$spearman_surv_pval < .05, T, F)
corrs_df$signif_len <- ifelse(corrs_df$spearman_len_pval < .05, T, F)

# save this
write.csv(corrs_df, paste0("results/GO_results/go_fit_corrs_",Sys.Date(),".csv"), row.names = F)

# # what about without selection lines?
# (cor_lea_lew_ns <- cor_fit("Lewisetta","lea", fullfit[!fullfit$site_name %in% c("S1_LOLA", "S2_DEBY"),]))
# (cor_lea_yrk_ns <- cor_fit("YorkRiver","lea", fullfit[!fullfit$site_name %in% c("S1_LOLA", "S2_DEBY"),]))
# (cor_gf_af_lew_ns <- cor_fit("Lewisetta","gf", fullfit[!fullfit$site_name %in% c("S1_LOLA", "S2_DEBY"),]))
# (cor_gf_af_yrk_ns <- cor_fit("YorkRiver","gf", fullfit[!fullfit$site_name %in% c("S1_LOLA", "S2_DEBY"),]))
# (cor_rda_lew_ns <- cor_fit("Lewisetta","rda", fullfit[!fullfit$site_name %in% c("S1_LOLA", "S2_DEBY"),]))
# (cor_rda_yrk_ns <- cor_fit("YorkRiver","rda", fullfit[!fullfit$site_name %in% c("S1_LOLA", "S2_DEBY"),]))
# (cor_envdist_lew_ns <- cor_fit("Lewisetta", "env_dist", fullfit[!fullfit$site_name %in% c("S1_LOLA", "S2_DEBY"),]))
# (cor_envdist_yrk_ns <- cor_fit("YorkRiver", "env_dist", fullfit[!fullfit$site_name %in% c("S1_LOLA", "S2_DEBY"),]))
# 
# corrs_df_ns <- rbind(cor_lea_lew_ns, cor_lea_yrk_ns, cor_gf_af_lew_ns, cor_gf_af_yrk_ns, cor_rda_lew_ns, cor_rda_yrk_ns, cor_envdist_lew_ns, cor_envdist_yrk_ns)
# rownames(corrs_df_ns) <- NULL
# 
# cor(corrs_df$spearman_mort, corrs_df_ns$spearman_mort) # .96667 cor sel/no sel
# cor(corrs_df$spearman_len, corrs_df_ns$spearman_len) # .92388 cor sel/no sel
#################################################
```


### Plot GO vs mortality against each other

I'll start by plotting LEA results:

```{r}
# plot lea vs mort at each site/time point
goFitLEA_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = lew_go_lea, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = lew_go_lea, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = lew_go_lea, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = lew_go_lea, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = lew_go_lea, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = lew_go_lea, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Survival at Lewisetta", x = "LEA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_lew_plotbag

goFitLEA_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = yrk_go_lea, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = yrk_go_lea, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = yrk_go_lea, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = yrk_go_lea, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = yrk_go_lea, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = yrk_go_lea, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Survival at York River", x = "LEA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_plot

# plot go lea vs mort at both sites just final time point
goFitLEA_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = surv), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = surv), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (LEA) vs Survival at Final Time Point", x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotfinalTsurv
```

Next, I'll plot gradient forest results, starting with the allele frequency model.

```{r}
# plot gf af vs mort at each site/time point
goFitGFaf_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (GFaf) vs Survival at Lewisetta", x = "GF Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFaf_lew_plotbag

goFitGFaf_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (GFaf) vs Survival at York River", x = "GF Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFaf_yrk_plot

# plot gf vs mort at both sites just final time point
goFitGFaf_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (GFaf) vs Survival at Final Time Point", x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGFaf_plotfinalTsurv
```

Then the genotype model results:

```{r}
# plot gf geno vs mort at each site/time point
goFitGFgeno_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (GFgeno) vs Survival at Lewisetta", x = "GF Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFgeno_lew_plotbag

goFitGFgeno_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (GFgeno) vs Survival at York River", x = "GF Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFgeno_yrk_plot

# plot gf vs mort at both sites just final time point
goFitGFgeno_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = surv), col = "mediumspringgreen", linetype = "dashed", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = surv), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (GFgeno) vs Survival at Final Time Point", x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGFgeno_plotfinalTsurv
```


Then I'll plot rda results:

```{r}
# plot rda vs mort at each site/time point
goFitRDA_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lew_rda, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lew_rda, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lew_rda, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lew_rda, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lew_rda, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lew_rda, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Survival at Lewisetta", x = "RDA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_lew_plotbag

goFitRDA_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_yrk_rda, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_yrk_rda, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_yrk_rda, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_yrk_rda, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_yrk_rda, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_yrk_rda, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Survival at York River", x = "RDA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_yrk_plot

# plot rda vs mort at both sites just final time point
goFitRDA_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = surv), col = "mediumspringgreen", linetype = "dashed", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = surv), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (RDA) vs Survival at Final Time Point", x = "RDA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitRDA_plotfinalTsurv
```

And finally, environmental distance:

```{r}
# plot env dist vs mort at each site/time point
goFitEnvdist_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = surv), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Survival at Lewisetta", x = "Environmental Distance", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvdist_lew_plotbag

goFitEnvdist_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = surv), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Survival at York River", x = "Environmental Distance", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvdist_yrk_plot

# plot env dist vs mort at both sites just final time point
goFitEnvdist_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = surv), col = "mediumspringgreen", linetype = "dashed", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = surv), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Environmental Distance vs Survival at Final Time Point", x = "Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitEnvdist_plotfinalTsurv
```


### Plot GO vs length against each other

LEA vs length:

```{r}
# plot lea vs len at each site/time point
goFitLEA_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = lew_go_lea, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = lew_go_lea, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = lew_go_lea, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = lew_go_lea, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = lew_go_lea, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = lew_go_lea, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Size at Lewisetta", x = "LEA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_lew_len

goFitLEA_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = yrk_go_lea, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = yrk_go_lea, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = yrk_go_lea, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = yrk_go_lea, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = yrk_go_lea, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = yrk_go_lea, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Size at York River", x = "LEA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_len

# plot go lea vs len at both sites just final time point
goFitLEA_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = lew_go_lea, y = length), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = yrk_go_lea, y = length), col = "dodgerblue", se = T) +
  labs(title = "Predicted GO (LEA) vs Size at Final Time Point", x = "LEA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotfinalTlen
```


GF AF GO vs length: 

```{r}
# plot gf vs len at each site/time point
goFitGFaf_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (GF) vs Size at Lewisetta", x = "GF Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFaf_lew_len

goFitGFaf_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (GF) vs Size at York River", x = "GF Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFaf_yrk_len

# plot gf vs len at both sites just final time point
goFitGFaf_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (GF) vs Size at Final Time Point", x = "GF Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGFaf_plotfinalTlen
```


GF geno GO vs length: 

```{r}
# plot gf vs len at each site/time point
goFitGFgeno_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (GF) vs Size at Lewisetta", x = "GF Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFgeno_lew_len

goFitGFgeno_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (GF) vs Size at York River", x = "GF Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFgeno_yrk_len

# plot gf vs len at both sites just final time point
goFitGFgeno_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (GF) vs Size at Final Time Point", x = "GF Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGFgeno_plotfinalTlen
```


RDA GO vs length:

```{r}
# plot rda vs len at each site/time point
goFitRDA_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lew_rda, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lew_rda, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lew_rda, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lew_rda, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lew_rda, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lew_rda, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Size at Lewisetta", x = "RDA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_lew_len

goFitRDA_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_yrk_rda, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_yrk_rda, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_yrk_rda, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_yrk_rda, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_yrk_rda, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_yrk_rda, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Size at York River", x = "RDA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_yrk_len

# plot rda vs len at both sites just final time point
goFitRDA_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lew_rda, y = length), col = "mediumspringgreen", linetype = "dashed", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_yrk_rda, y = length), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (RDA) vs Size at Final Time Point", x = "RDA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitRDA_plotfinalTlen
```

Environmental distance vs length:

```{r}
# plot env dist vs length at each site/time point
goFitEnvdist_len_lew_plot <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Size at Lewisetta", x = "Environmental Distance", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvdist_len_lew_plot

goFitEnvdist_len_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Size at York River", x = "Environmental Distance", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvdist_len_yrk_plot

# plot go gf vs len at both sites just final time point
goFitEnvdist_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length), col = "mediumspringgreen", linetype = "dashed", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length), col = "dodgerblue", se = T) +
  labs(title = "Environmental Distance vs Size at Final Time Point", x = "Environmental Distance", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitEnvdist_plotfinalTlen
```


Export all of these plots to pdf:

```{r}
pdf(file = "results/figures/GO_comparisons.pdf", 
    width = 9, height = 6)
goFitLEA_plotfinalTsurv
goFitGF_plotfinalTsurv
goFitRDA_plotfinalTsurv
goFitEnvdist_plotfinalTsurv
goFitLEA_plotfinalTlen
goFitGF_plotfinalTlen
goFitRDA_plotfinalTlen
goFitEnvdist_plotfinalTlen
dev.off()
```

## Compare Offset Methods

Now that I have predictions from different methods, I will look at how offset predictions made by these methods compare to each other and how different their ability is to predict experimental fitness.


### Compare Offsets to Each Other

```{r Offset Comparison}
# create a correlation matrix for offsets
go_forcorrs <- fullfit %>% distinct(site_name, bag_site, event, .keep_all = TRUE) %>% filter(event == 4) %>% dplyr::select(site_name, pop, bag_site, offset_lew_rda, offset_yrk_rda, order_lew_rda, order_yrk_rda, offset_gf_af_lew, offset_gf_af_yrk, order_lew_af_gf, order_yrk_af_gf, lew_gothin_lea, yrk_gothin_lea, order_lew_lea, order_yrk_lea, env_dist_lew, env_dist_yrk, order_lew_envdist, order_yrk_envdist)

# check correlations
cor_go_raw <- cor(go_forcorrs[,c("offset_lew_rda", "lew_gothin_lea", "offset_gf_af_lew", "env_dist_lew", "offset_yrk_rda", "yrk_gothin_lea", "offset_gf_af_yrk", "env_dist_yrk")])
cor_go_raw_pmat <- cor_pmat(go_forcorrs[,c("offset_lew_rda", "lew_gothin_lea", "offset_gf_af_lew", "env_dist_lew", "offset_yrk_rda", "yrk_gothin_lea", "offset_gf_af_yrk", "env_dist_yrk")])
cor_go_ord <- cor(go_forcorrs[,c("order_lew_rda", "order_lew_lea", "order_lew_af_gf",  "order_lew_envdist", "order_yrk_rda", "order_yrk_lea", "order_yrk_af_gf", "order_yrk_envdist")])
cor_go_ord_pmat <- cor_pmat(go_forcorrs[,c("order_lew_rda", "order_lew_lea", "order_lew_af_gf",  "order_lew_envdist", "order_yrk_rda", "order_yrk_lea", "order_yrk_af_gf", "order_yrk_envdist")])

# plot these correlations
library(ggcorrplot)
# raw go
corrs_go_raw_all <- ggcorrplot(cor_go_raw, 
                               lab = T,
                               colors = c("#6D9EC1", "white", "#E46726")) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Genomic Offset Correlations",
       x = NULL,
       y = NULL)
corrs_go_raw_all

corrs_go_raw_nons <- ggcorrplot(cor_go_raw,
                                lab = T, 
                                p.mat = cor_go_raw_pmat, 
                                insig = "blank",
                                colors = c("#6D9EC1", "white", "#E46726")) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Genomic Offset Correlations (Significant)",
       x = NULL,
       y = NULL)
corrs_go_raw_nons

# go orders
corrs_go_ord_all <- ggcorrplot(cor_go_ord, 
                               lab = T,
                               colors = c("#6D9EC1", "white", "#E46726")) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Genomic Offset Correlations (Rank)",
       x = NULL,
       y = NULL)
corrs_go_ord_all

corrs_go_ord_nons <- ggcorrplot(cor_go_ord, 
                                lab = T, 
                                p.mat = cor_go_ord_pmat, 
                                insig = "blank",
                                colors = c("#6D9EC1", "white", "#E46726")) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Genomic Offset Correlations (Rank, Significant)",
       x = NULL,
       y = NULL)
corrs_go_ord_nons
```

