---
title: "CviMVP_oyster_GenDatProcess"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Genetic Data Processing

This script is to process individual level genetic data for the seascape populations by identifying the number and identities of individuals in each dataset and preparing this information and the genomic data into formats needed for downstream analysis.

The seascape data comes from Madeline's seascape study.

## Setup

```{r}
# set wd
setwd("~/Documents/Github/MVP_oyster_GO")

# packages
library(dplyr)
library(ggplot2)
library(tidyverse)
library(LEA)
library(stringr)

# data
geno_mat_ld <- readRDS("data/GenDat/20250606_THINNEDSNPs_seascape_labeled.rds")
sea_inds <- readRDS("data/IndDat/20250604_seascape_indsmatrix.rds")
env_sea_sites <- read.csv("data/EnvDat/seascape/seascape_sitenames.csv")[,-1]
```


## Pre-processing

Process data to remove missing data, non-variable genomic sites.

```{r}
# check data
dim(geno_mat_ld) # 746 inds, 115185 snps (~40k less than unthinned)

# any missing data?
length(which(geno_mat_ld == 9)) # nope!

# non-variable genomic sites
nonvarsite_ld <- apply(geno_mat_ld, 2, function(x) all(x == x[1]))
which(nonvarsite_ld == T) # none
```


## Individual Data

```{r}
dim(sea_inds) # 746 inds (presumably in same order), 79 cols with identifying info incl pop

# let's get rid of the empty columns
sea_inds_rm <- sea_inds[, apply(sea_inds, 2, function(x) !any(is.na(x)))] # only 29 columns now

# rownames of genomic data match clean_ID in individual df, and appear to be in the same order

# i need to get rid of inds not in my dataset (the ones with no environmental data)
seaRows_keep <- which((sea_inds_rm$ID_SiteDate %in% env_sea_sites))
geno_mat_ld_keep <- geno_mat_ld[seaRows_keep,]

# also be sure to put this data in the correct order to match environmental df
# first let's figure out which population everyone is from
genomat_order <- data.frame(clean_ID = rownames(geno_mat_ld_keep), order_mat = 1:length(rownames(geno_mat_ld_keep)))
sea_inds_order <- sea_inds_rm[seaRows_keep, c("clean_ID","ID_SiteDate")]
sea_inds_order$order_inds <- 1:length(sea_inds_order$clean_ID)
which(!(sea_inds_order$clean_ID %in% genomat_order$clean_ID))
ordered_sea_gen <- merge(genomat_order, sea_inds_order, by = "clean_ID")
ordered_sea_gen
```


## Genomic Data Formatting

```{r}
# write these matrices as geno files
write.geno(geno_mat_keep, paste0("data/GenDat/geno_seascape_", Sys.Date(), ".geno"))
write.geno(geno_mat_ld_keep, paste0("data/GenDat/geno_seascape_LDthin_", Sys.Date(), ".geno"))

# also write out information about the individuals in each row (same inds in ld and not ld mats)
inds_genomat <- rownames(geno_mat_ld_keep)
write.csv(inds_genomat, "data/GenDat/inds_genomat.csv")

# for LEA need lfmm format data
geno2lfmm("data/GenDat/geno_seascape_2026-01-27.geno")
geno2lfmm("data/GenDat/geno_seascape_LDthin_2026-01-27.geno")

# for RDA and gf geno matrix should work
write_rds(geno_mat_keep, paste0("data/GenDat/geno_seascape_rm_", Sys.Date(), ".RDS"))
write_rds(geno_mat_ld_keep, paste0("data/GenDat/geno_seascape_LDthin_rm_", Sys.Date(), ".RDS"))
```

