---
title: "CviMVP_LEA_offset"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LEA Genomic offsets

This script is a LEA pipeline for calculating genomic offset. I am using LEA3, which allows for multivariate environmental datasets. I first train the models using LFMM of a large seascape dataset, then calculate offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment. Offset is calculated from the environment of origin to common garden conditions.

## Setup

```{r}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(LEA) # for offsets
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(ggplot2) # for plotting
library(vegan) # for env dat processing
library(dplyr) # for data wrangling
library(stringr) # for data wrangling

# data
# env
seascapeEnv <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv") # keep this but also get list of reduced variables
expEnv <- read.csv("data/EnvDat/exp/envr_summary_stats.csv")[,-1]

# gen
# add once i get it from madeline
geno_mat
```


## Data Prep

```{r}
# check data
seascapeEnv
expEnv
geno_mat

# Genetic data prep
# need to convert genotype matrix to lfmm format
geno2lfmm("INSERT_PATH&FILENAME.csv", force = T)

# read in lfmm object
geno_lfmm <- read.lfmm("INSERT_PATH&FILENAME.lfmm")

# any missing data?
sum(geno_lfmm == 9)

# naive imputation, remove once i get fully processed genetic data
geno_lfmm_imp <- apply(geno_lfmm, 2, function(x) replace(x, x == 9, as.numeric(names(which.max(table(x))))))
sum((geno_lfmm_imp == 9)) 

# remove non-variable genomic sites
geno_lfmm_imp_df <- as.data.frame(geno_lfmm_imp)
geno_lfmm_imp_trim <- Filter(var, geno_lfmm_imp_df)
geno_lfmm_imp_mat <- as.matrix(geno_lfmm_imp_trim)

# rewrite lfmm files
write.lfmm(geno_lfmm_imp_mat, "data/GenDat/geno_lfmm_imp_mat.lfmm")

# environmental data prep
# add this once I have environmental data
```


## Seascape Training

Now that I have prepped and cleaned my data, I will use it to train LFMM models later used to predict genomic offset to some other (in this case experimental) climate. Importantly, I do not need to trim or impute my genomic data, or determine ancestry - Madeline has already done this. The data has been LD thinned, trimmed for MAF, and ancestry analysis has recovered K=2.
Madeline also compiled the environmental data, which I further processed in the script "CviMVP_oyster_EnvDatProcess."

May decide to train both genotype-based models and allele frequency based models - for now genotype models only.

```{r}
# read in the new lfmm file
geno_lfmm_mod <- read.lfmm("data/GenDat/geno_lfmm_imp_mat.lfmm")

# check the data
geno_lfmm_mod
seascape_env # these will be .env files, read in with data
exp_env

# fit model
mod_lfmm <- lfmm2(input = geno_lfmm_mod,
                  env = seascape_env, 
                  K = 2) # using number of MLLs - should I keep K the same for

# perform test
pv_lfmm <- lfmm2.test(object = mod_lfmm,
                      input = geno_lfmm_mod,
                      env = exp_env, 
                      full = TRUE)

# pvalues
pv <- pv_lfmm$pvalues

# qvalues
qv <- qvalue(pv, fdr.level = 0.01)

# candidates
candidates <- which(qv$significant)
candidates_pvals <- (pv)[candidates]

# using FDR to filter out false positives
length(which(qv$qvalues < 0.01)) # how many

# identify which SNPs have FDR less than 1%
fdr_cands <- colnames(geno_lfmm_mod)[which(qv$qvalues < 0.01)]

# df for candidates
cand_df <- data.frame(index = colnames(geno_lfmm_mod), pos = 1:length(pv),
                      pvals = pv, qvals = qv$qvalues)

# manhattan plot - unnecessary but just a quick visualization
plot(cand_df$pos, -log10(cand_df$pvals), 
     pch = 19, col = "darkslategray4",
     main = "Candidate SNPs",
     xlab = "Position",
     ylab = "-log10(Pvals)")
points(cand_df$pos[cand_df$qvals < 0.01], 
       -log10(cand_df$pvals[cand_df$qvals < 0.01]),
       pch = 19,
       col = "hotpink")
```


## Offset Prediction

```{r}
# check future env data
exp_env # this will be env file for experimental environmental dat

# calculate genetic gap
gap_cands_full <- genetic.gap(input = geno_lfmm_mod, 
                         env = seascape_env, 
                         pred.env = exp_env, 
                         scale = F, 
                         K = 2)

# THESE PARTS NEED TO BE UPDATED FROM EELGRASS SCRIPT VERSION
# check how well each genetic gap measure correlates with euclidean environmental distance
Delta <- current_MLL_env - future_MLL_env
dist_env <- sqrt(rowSums(Delta^2))
plot(dist_env, gap_cands_MLL$distance,
     main = "Genetic gap vs Environmental Distance - MLL", 
     xlab ="Euclidean distance",  ylab ="sqrt(genetic gap)", 
     cex = .6, col = "darkslategray4", pch = 19)

# Extract genomic offset from genetic gap object
offsets_full <- gap_cands_full$offset

# summary stats on offsets
summary(offsets_full)

# offset dfs
offset_full_df <- cbind(indiv_full_merge, offsets_full)
```
