---
title: "CviMVP_oyster_GF"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GF Genomic Offsets

This script is a gradientForest pipeline for calculating genomic offset. I am using gradientForest, which build on randomForest to use predictions of genome-wide allele frequency variation along multivariate environmental gradients to quantify the disruption of genotype-environment associations under environmental change. I first train the models using gradientForest on a large seascape dataset, then predict offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment. Offset is calculated from the environment of origin to common garden conditions.

## Setup

```{r Setup}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(LEA)
library(dplyr) # for data cleaning
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(raster) # for environmental pixels
library(ggplot2) # for plotting
library(parallel) # for parallelizing processes
library(doParallel) # for parallelizing processes 
library(gradientForest) # for running gradient forest

# data
# env
seascapeEnv <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv") # keep this but also get list of reduced variables
expEnv <- read.csv("data/EnvDat/exp/exp_scaled_expanded_04222025.csv")[,-1]
expEnv_pop <- read.csv("data/EnvDat/exp/expquants_scaled_04222025.csv")[,-1]

# gen
geno_dat <- as.data.frame(read.lfmm("data/GenDat/geno_file_exp_all.lfmm"))

# individuals
seascapeInds <- readRDS("data/IndDat/20240922_experimental_indsmatrix.rds")
```


## Data Prep

Inspect the environmental and genomic data to ensure it's ready for gradientForest. Note that environmental data was pre-processed in the script "CviMVP_oyster_EnvDatPRocess.Rmd" and genomic data was pre-processed in the script "CviMVP_oyster_GenDatProcess.Rmd".

### Individual Data Prep

```{r}
head(seascapeInds)

# need ind level dat
indsExp <- readRDS("data/IndDat/20240922_experimental_indsmatrix.rds")

# keep just relevant data
indsExp_red <- indsExp[,c("merge_ID_inds","Label_ind","Key_ind","ID_SiteDate","SiteDate_Key")]
levels(as.factor(indsExp_red$ID_SiteDate))

# need to add scaled env data to this
exp_scaled$site_name
indsExp_red$ID_SiteDate

# the pop (ID_SiteDate/site_name) data doesn't match up
map <- list("TX-CapBay" = "W1-TX", "LA-SisLake" = "W2-LA", "FL-KingPlan" = "W3-FL", "LOLA" = "S1-LOLA", "DEBY" = "S2-DEBY", "VA-DeepWatSh" = "W4-VA", "NH-GrtBay" = "W5-NH", "ME-HogIs" = "W6-ME")
indsExp_red_mut <- indsExp_red %>% mutate(ID_SiteDate = ifelse(ID_SiteDate %in% names(map), map[ID_SiteDate], ID_SiteDate))
colnames(indsExp_red_mut) <- c("merge_ID_inds", "Label_ind", "Key_ind", "site_name", "SiteDate_Key")

# also need to remap expenvpop 
expEnv_pop
map <- list("LEW" = "LEW", "YRK" = "YRK", "W1_TX" = "W1-TX", "W2_LA" = "W2-LA", "W3_FL" = "W3-FL", "S1_LOLA" = "S1-LOLA", "S2_DEBY" = "S2-DEBY", "W4_VA" = "W4-VA", "W5_NH" = "W5-NH", "W6_ME" = "W6-ME")
expEnv_pop_mut <- expEnv_pop %>% mutate(site_name = ifelse(site_name %in% names(map), map[site_name], site_name))

which(!(exp_scaled$site_name %in% expEnv_pop_mut$site_name))
which(!(expEnv_pop_mut$site_name %in% exp_scaled$site_name)) # substitution worked!

expEnv_pop_mut$site_name <- as.character(expEnv_pop_mut$site_name)
```

### Genomic Data Prep

```{r Genomic Data Prep}
# check geno data
dim(geno_dat)

# rotate df so rows are now columns (we want individuals as columns, alleles as rows)
geno_dat_t <- as.matrix(t(geno_dat))

# any missing data?
sum((geno_dat_t == 9)) # returns NA - good

# get vector of populations
seascapePops <- seascapeInds$ID_SiteDate

# add pop data to genomic data
colnames(geno_dat_t) <- seascapePops

# function to calculate allele frequency @ single locus
calc_freq <- function(x){
  a <- sum(x, na.rm=TRUE)
  b <- (2*length(na.omit(x)))
  a/b
}

# function extending this to calculate allele frequencies across whole pop
calcfreq <- function(a, pop){
  tapply(a, pop, calc_freq)
}

# allele freqs
freqs <- apply(t(geno_dat_t), 2, calcfreq, seascapePops)
str(freqs)
```


### Environmental Data Prep

```{r Environmental Data Prep}
# check env data
#dim(seascapeEnv) # for now I only use expEnv object because I don't have any seascape genomic data
dim(expEnv)
dim(expEnv_pop)
dim(indsExp_red_mut)

# rename site names pop df
expEnv_pop$site_name <- c("LEW", "S1-LOLA", "S2-DEBY", "W1-TX", "W2-LA", "W3-FL", "W4-VA", "W5-NH","W6-ME","YRK")

# merge
exp_scaled_expanded <- merge(as.data.frame(indsExp_red_mut), as.data.frame(expEnv_pop[,c("site_name", "sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled"),]), by = c("site_name"))

# reduced variable set
expEnv_red <- exp_scaled_expanded[,c("site_name","sal_quant10_scaled","sal_quant90_scaled","temp_quant10_scaled","temp_quant90_scaled")]
expEnv_pop_red <- expEnv_pop_mut[,c("site_name","sal_quant10_scaled","sal_quant90_scaled","temp_quant10_scaled","temp_quant90_scaled")]
```


## Train Models

To train the models, I will be using the seascape genomic dataset that we processed in the previous chunk, as well as the current environmental data files produced using "CviMVP_oyster_EnvDatProcess.Rmd." NOTE FOR NOW TRAINED ON EXP INDS

```{r}
# start by defining a maximum number of splits
maxLevel <- log2(0.368*nrow(expEnv_red)/2)

# run gradient forest
# allele freq models
start_time <- Sys.time() # time start
gf_af <- gradientForest(cbind(expEnv_pop_red[2:9,c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")], freqs), predictor.vars = colnames(expEnv_pop_red[, c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")]), response.vars=colnames(freqs), ntree=500, maxLevel=maxLevel, trace=T, corr.threshold=0.50) # tons of warnings here, "response has five or fewer unique values. Are you sure you want to do regression?"
end_time <- Sys.time() # time end
(end_time - start_time)

# geno models
start_time <- Sys.time() # time start
gf_geno <- gradientForest(cbind(expEnv_red[,c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")], geno_dat), predictor.vars = colnames(expEnv_red[, c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")]), response.vars=colnames(geno_dat), ntree=500, maxLevel=maxLevel, trace=T, corr.threshold=0.50)
end_time <- Sys.time() # time end
(end_time - start_time)

# plot bar graphs depicting importance of each spatial and climate variable
plot(gf_af, plot.type = "O")
plot(gf_geno, plot.type = "O")

# plots of turnover functions for individual loci - now having issues with ylim
plot(gf_af, plot.type = "C", imp.vars = gf_af$imp.vars, show.overall = T, legend = T, leg.posn = "topleft", leg.nspecies = 5, ylim = c(0,0.5)) # ylim error
plot(gf_geno, plot.type = "C", imp.vars = gf_geno$imp.vars, show.overall = T, legend = T, leg.posn = "topleft", leg.nspecies = 5, ylim = c(0,0.5))
```



