---
title: "CviMVP_oyster_RDA.Rmd"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LEA Genomic offsets

This script is a RDA-based pipeline for calculating genomic offset. I am using rdadapt, a package developed for genome scans for selection, adaptive index projection on the landscape, estimation of genomic offsets, identification of co-varying modules of adaptive loci, and/or identification of adaptive groups of individual or populations. I first calculate "adaptive index" across seascape sites, then calculate offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment. Offset is calculated from the environment of origin to common garden conditions.

A vignette detailing the rdadapt pipeline is found [here](https://landscape-genomics.github.io/rdadapt/articles/vignette_tutorial.html).

## Setup

```{r}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(rdadapt) # for offsets
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(vegan) # for env dat processing
library(terra) # for env dat processing
library(reshape2) # for env dat processing
library(raster) # for env dat processing
library(dplyr) # for data wrangling
library(plyr) # for data wrangling
library(stringr) # for data wrangling
library(ggplot2) # for data vis
library(viridis) # for data vis
library(wesanderson) # for data vis

# data
# env
seascapeEnv <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv") # keep this but also get list of reduced variables
expEnv <- read.csv("data/EnvDat/exp/exp_scaled_expanded_04222025.csv")[,-1]
expEnv_pop <- read.csv("data/EnvDat/exp/expquants_scaled_04222025.csv")[,-1]
# expanded env data
complete_env <- read.csv("data/EnvDat/exp/expsitesquants_scaled_04222025.csv")[,-1]

# gen
geno_dat <- as.data.frame(read.lfmm("data/GenDat/geno_file_exp_all.lfmm"))
geno_exp <- as.data.frame(readRDS("data/GenDat/20250203_experimentalmatrix.RDS"))

# individuals
seascapeInds <- readRDS("data/IndDat/20240922_experimental_indsmatrix.rds") # these are the experimental individuals for now, need the updated seascape stuff from madeline
```


## Data Checks

I will start by double checking that I have my genomic and environmental data in the correct formats.

```{r}
# env data
##########
# seascape
seascapeEnv
seascapeEnv[which(is.na(seascapeEnv$Mean_Annual_Salinity_ppt)),]

# we have NAs! remove those
seascape_nona <- seascapeEnv[!(is.na(seascapeEnv$Mean_Annual_Salinity_ppt)),]

# need to find the more complete YRK & LEW data, talk to nicole about this
# once I have that and the seascape quantile data, I will make predictions - for now use just the experimental pops for training
# also need the seascape genomic data, this will be goal for next week to update all three GO pipelines for seascape genomic data

# check exp data
expEnv
expEnv_pop # expEnv and expEnv_pop have different values for the same pops, need to fix this! come back here

# lets just use what I'm using in gradient forest right now...
indsExp <- as.data.frame(readRDS("data/IndDat/20240922_experimental_indsmatrix.rds")) # for some reason this is a list?

ldply(indsExp, data.frame)

# keep just relevant data
indsExp_red <- indsExp[,c("merge_ID_inds","Label_ind","Key_ind","ID_SiteDate","SiteDate_Key")]
levels(as.factor(indsExp_red$ID_SiteDate))

# need to add scaled env data to this
expEnv$site_name
indsExp_red$ID_SiteDate

# the pop (ID_SiteDate/site_name) data doesn't match up
map <- list("TX-CapBay" = "W1-TX", "LA-SisLake" = "W2-LA", "FL-KingPlan" = "W3-FL", "LOLA" = "S1-LOLA", "DEBY" = "S2-DEBY", "VA-DeepWatSh" = "W4-VA", "NH-GrtBay" = "W5-NH", "ME-HogIs" = "W6-ME")
indsExp_red_mut <- indsExp_red %>% mutate(ID_SiteDate = ifelse(ID_SiteDate %in% names(map), map[ID_SiteDate], ID_SiteDate))
indsExp_red_mut <- as.data.frame(indsExp_red_mut)
colnames(indsExp_red_mut) <- c("merge_ID_inds", "Label_ind", "Key_ind", "site_name", "SiteDate_Key")

# pops list
pops_list <- unique(indsExp_red_mut$site_name)

# also need to remap expenvpop 
expEnv_pop
map <- list("LEW" = "LEW", "YRK" = "YRK", "W1_TX" = "W1-TX", "W2_LA" = "W2-LA", "W3_FL" = "W3-FL", "S1_LOLA" = "S1-LOLA", "S2_DEBY" = "S2-DEBY", "W4_VA" = "W4-VA", "W5_NH" = "W5-NH", "W6_ME" = "W6-ME")
expEnv_pop_mut <- expEnv_pop %>% mutate(site_name = ifelse(site_name %in% names(map), map[site_name], site_name))

which(!(expEnv$site_name %in% expEnv_pop_mut$site_name))
which(!(expEnv_pop_mut$site_name %in% expEnv$site_name)) # substitution worked!

expEnv_pop_mut$site_name <- as.character(expEnv_pop_mut$site_name)

# rasterize these data
expEnv_pop_mut_rm <- expEnv_pop_mut[, !names(expEnv_pop_mut) %in% c("lat.y", "lon.y")]
expEnv_pop_mut_rn <- expEnv_pop_mut_rm %>% dplyr::rename(lat = lat.x, lon = lon.x)

#ras_env <- rast(data.frame(x = rep(1:12, time = 12) # NEED TO COME BACK HERE
#                                  , y = rep(1:12, each = 12)
#                                  , aggregate(xx.current, by = list(ceiling(coord$x)
#                                                                    , ceiling(coord$y))
#                                              , mean)[, -c(1, 2)])
#                       , type = "xyz"
#                       , crs = crs("epsg:4326"))
##########

# gen data
##########
dim(geno_dat) # 154 rows (inds), 32847 cols (loci)
dim(geno_expfilt) # 186 rows, 275671 - this is not the correct data
dim(indsExpSeq) # 154 rows (inds), 195026 cols (loci) - this one is snp chip data, likely unfiltered

# use indiv data to check genomic data
head(seascapeInds)

# put seascapeInds in order of genomic data row names
# to do
##########
```

## Outlier Detection

While I do not intend to use an outlier dataset to calculate adaptive index/offset, I will still go through the process of outlier detection with RDA. This process is also part of training the RDA model for offsets later.

```{r}
expEnv_red <- complete_env[,c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")]

RDA_env <- rda(geno_dat ~ ., as.data.frame(expEnv_red))
plot(RDA_env)
screeplot(RDA_env, main = "Eigenvalues of constrained axes") # probably just RDA 1, or keep all 4

# run rdadapt with K = 4 for all 4 axes, since can't run on just one (K=1)
rdadapt_env <- rdadapt(RDA = RDA_env, K = 4)

# bonferroni correction
# P-values threshold after Bonferroni correction
thres_env <- 0.01 / length(rdadapt_env$p.values)

# Identifying the loci that are below the p-value threshold
outliers <- (1:ncol(Y))[which(rdadapt_env$p.values < thres_env)]
length(outliers) # 118 outliers

# Formatting table for ggplot
locus_scores <- as.data.frame(scores(RDA_env, choices=c(1:2), display="species", scaling="none"))
TAB_loci <- data.frame(names = row.names(locus_scores), locus_scores)
TAB_loci$type <- "Neutral"
TAB_loci$type[outliers] <- "Outliers"
TAB_loci$type <- factor(TAB_loci$type, levels = c("Neutral", "Outliers"))
TAB_loci <- TAB_loci[order(TAB_loci$type),]
TAB_var <- as.data.frame(scores(RDA_env, choices=c(1,2), display="bp")) 

# Biplot of RDA loci and variables scores
ggplot() +
  geom_hline(yintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_vline(xintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_point(data = TAB_loci, aes(x=RDA1*20, y=RDA2*20, colour = type), size = 1.4, alpha = 0.7) +
  scale_color_manual(values = c("gray90", "#EB8055FF")) +
  geom_segment(data = TAB_var, aes(xend=RDA1, yend=RDA2, x=0, y=0), colour="black", size=0.15, linetype=1, arrow=arrow(length = unit(0.02, "npc"))) +
  geom_text(data = TAB_var, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(TAB_var)), size = 2.5, family = "Times") +
  guides(color=guide_legend(title="Locus type")) +
  labs(x = "RDA 1 (XX%)", y = "RDA 2 (XX%)", subtitle = "RDA space") +
  theme_bw(base_size = 11, base_family = "Times") +
  theme(panel.grid = element_blank()
        , plot.background = element_blank()
        , panel.background = element_blank())
```

## Adaptive Index

Now, I will use rdadapt to calculate adaptive index at the two seascape sites (NOT across the range, though I could potentially do this if I decide to impute the seascape environmental data to have a more complete coastal environment map - either way the offsets for each site will be the same). 

```{r}
# overall, including lew & yrk - I can come back here if I want to do rasterization/interpolation, but it isn't necessary to calculate offset
# RDA_proj_current <- adaptive_index(RDA = RDA_env
#                                        , K = 2
#                                        , env = expEnv_red # alas, this needs to be a raster - must rasterize my environmental data
#                                        , env_mask = NULL
#                                        , method = "loadings")
```


## Offset

Now, I calculate genomic offset for all of my populations!

```{r}
# common garden info
lew_env <- complete_env[,c("LEW_sal_quant10_scaled","LEW_sal_quant90_scaled","LEW_temp_quant10_scaled","LEW_temp_quant90_scaled")]
yrk_env <- complete_env[,c("YRK_sal_quant10_scaled","YRK_sal_quant90_scaled","YRK_temp_quant10_scaled","YRK_temp_quant90_scaled")]
colnames(lew_env) <- colnames(yrk_env) <- c("sal_quant10_scaled","sal_quant90_scaled","temp_quant10_scaled","temp_quant90_scaled")

# Estimate genomic offset
offset_lew <- genomic_offset(RDA = RDA_env, 
                              K = 2, # will need to reasses where I make this cutoff
                              env_pres = expEnv_red, 
                              env_fut = lew_env, 
                              method = "loadings")

offset_yrk <- genomic_offset(RDA = RDA_env, 
                              K = 2, 
                              env_pres = expEnv_red, 
                              env_fut = yrk_env, 
                              method = "loadings")

# reassign the metadata
offsets_rda_df <- data.frame(site_name = complete_env$site_name, lat = complete_env$lat.x, lon = complete_env$lon.x, offset_lew = offset_lew$genomic_offset$Global, offset_yrk = offset_yrk$genomic_offset$Global)

write.csv(offsets_rda_df, paste0("results/GO_results/offsets_RDA_expSites_",Sys.Date(),".csv", sep = ""))
```

## Visualization

```{r}
# add the color palette
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME"), 
                     order = 1:8, 
                     cols = c("#3e4989" ,"#31688e", "#35b779", "#6ece58", "#440154", "#482878", "#fde725", "#b5de2b"),
                     shape = c(21,21,21,21,23,23,21,21),
                     site_name = c("W1-TX", "W2-LA", "W3-FL", "W4-VA", "S1-LOLA", "S2-DEBY", "W5-NH", "W6-ME"))

offsets_plotting <- merge(offsets_rda_df, popsdf, by = c("site_name"))

# try mapping the offsets
library(s2) # mapping
library(rnaturalearth) # mapping
library(rnaturalearthdata) # mapping
library(maps) # mapping
library(ggspatial) # mapping

# map
sf::sf_use_s2(FALSE) # sf package doesn't work

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin =-102 , xmax = -68, ymin = 24, ymax = 47.049130))

# bottom right: 24.353558, -55.373595
# top left: 47.049130, -101.983412

# offset to lew
map_lew <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting, aes(x=lon, y=lat, 
                                       fill = offset_lew), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,2.25))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to Lewisetta (RDA)", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_lew)

map_yrk <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting, aes(x=lon, y=lat, 
                                       fill = offset_yrk), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,2.25))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to York (RDA)", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_yrk)
```

