---
title: "CviMVP_oyster_GF"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GF Genomic Offsets

This script is a gradientForest pipeline for calculating genomic offset. I am using gradientForest, which build on randomForest to use predictions of genome-wide allele frequency variation along multivariate environmental gradients to quantify the disruption of genotype-environment associations under environmental change. I first train the models using gradientForest on a large seascape dataset, then predict offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment based in the Chesapeake Bay. Offset is calculated from the environment of origin to common garden conditions.

## Setup

```{r Setup}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(LEA)
library(dplyr) # for data cleaning
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(raster) # for environmental pixels
library(ggplot2) # for plotting
library(parallel) # for parallelizing processes
library(doParallel) # for parallelizing processes 
library(gradientForest) # for running gradient forest

# data
# env
seascapeEnv <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv") # keep this but also get list of reduced variables
expEnv <- read.csv("data/EnvDat/exp/exp_scaled_expanded_04222025.csv")[,-1]
expEnv_pop <- read.csv("data/EnvDat/exp/expquants_scaled_04222025.csv")[,-1]

# gen
geno_dat <- as.data.frame(read.lfmm("data/GenDat/geno_file_exp_all.lfmm"))

# individuals
seascapeInds <- readRDS("data/IndDat/20240922_experimental_indsmatrix.rds")
```


## Data Prep

Inspect the environmental and genomic data to ensure it's ready for gradientForest. Note that environmental data was pre-processed in the script "CviMVP_oyster_EnvDatPRocess.Rmd" and genomic data was pre-processed in the script "CviMVP_oyster_GenDatProcess.Rmd".

### Individual Data Prep

```{r}
head(seascapeInds)

# need ind level dat
indsExp <- readRDS("data/IndDat/20240922_experimental_indsmatrix.rds")

# keep just relevant data
indsExp_red <- indsExp[,c("merge_ID_inds","Label_ind","Key_ind","ID_SiteDate","SiteDate_Key")]
levels(as.factor(indsExp_red$ID_SiteDate))

# need to add scaled env data to this
exp_scaled$site_name
indsExp_red$ID_SiteDate

# the pop (ID_SiteDate/site_name) data doesn't match up
map <- list("TX-CapBay" = "W1-TX", "LA-SisLake" = "W2-LA", "FL-KingPlan" = "W3-FL", "LOLA" = "S1-LOLA", "DEBY" = "S2-DEBY", "VA-DeepWatSh" = "W4-VA", "NH-GrtBay" = "W5-NH", "ME-HogIs" = "W6-ME")
indsExp_red_mut <- indsExp_red %>% mutate(ID_SiteDate = ifelse(ID_SiteDate %in% names(map), map[ID_SiteDate], ID_SiteDate))
colnames(indsExp_red_mut) <- c("merge_ID_inds", "Label_ind", "Key_ind", "site_name", "SiteDate_Key")

# pops list
pops_list <- unique(indsExp_red_mut$site_name)

# also need to remap expenvpop 
expEnv_pop
map <- list("LEW" = "LEW", "YRK" = "YRK", "W1_TX" = "W1-TX", "W2_LA" = "W2-LA", "W3_FL" = "W3-FL", "S1_LOLA" = "S1-LOLA", "S2_DEBY" = "S2-DEBY", "W4_VA" = "W4-VA", "W5_NH" = "W5-NH", "W6_ME" = "W6-ME")
expEnv_pop_mut <- expEnv_pop %>% mutate(site_name = ifelse(site_name %in% names(map), map[site_name], site_name))

which(!(exp_scaled$site_name %in% expEnv_pop_mut$site_name))
which(!(expEnv_pop_mut$site_name %in% exp_scaled$site_name)) # substitution worked!

expEnv_pop_mut$site_name <- as.character(expEnv_pop_mut$site_name)
```

### Genomic Data Prep

```{r Genomic Data Prep}
# check geno data
dim(geno_dat)

# rotate df so rows are now columns (we want individuals as columns, alleles as rows)
geno_dat_t <- as.matrix(t(geno_dat))

# any missing data?
sum((geno_dat_t == 9)) # returns NA - good

# get vector of populations
seascapePops <- seascapeInds$ID_SiteDate

# add pop data to genomic data
colnames(geno_dat_t) <- seascapePops

# function to calculate allele frequency @ single locus
calc_freq <- function(x){
  a <- sum(x, na.rm=TRUE)
  b <- (2*length(na.omit(x)))
  a/b
}

# function extending this to calculate allele frequencies across whole pop
calcfreq <- function(a, pop){
  tapply(a, pop, calc_freq)
}

# allele freqs
freqs <- apply(t(geno_dat_t), 2, calcfreq, seascapePops)
str(freqs)
```


### Environmental Data Prep

```{r Environmental Data Prep}
# check env data
#dim(seascapeEnv) # for now I only use expEnv object because I don't have any seascape genomic data
dim(expEnv)
dim(expEnv_pop)
dim(indsExp_red_mut)
indsExp_red_mut$site_name<- as.character(indsExp_red_mut$site_name)

# rename site names pop df
expEnv_pop$site_name <- c("LEW", "S1-LOLA", "S2-DEBY", "W1-TX", "W2-LA", "W3-FL", "W4-VA", "W5-NH","W6-ME","YRK")

# merge
exp_scaled_expanded <- merge(as.data.frame(indsExp_red_mut), as.data.frame(expEnv_pop[,c("site_name", "sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled"),]), by = c("site_name"))

# reduced variable set
expEnv_red <- exp_scaled_expanded[,c("site_name","sal_quant10_scaled","sal_quant90_scaled","temp_quant10_scaled","temp_quant90_scaled")]
expEnv_pop_red <- expEnv_pop_mut[,c("site_name","sal_quant10_scaled","sal_quant90_scaled","temp_quant10_scaled","temp_quant90_scaled")]
```


## Train Models

To train the models, I will be using the seascape genomic dataset that we processed in the previous chunk, as well as the current environmental data files produced using "CviMVP_oyster_EnvDatProcess.Rmd." NOTE FOR NOW TRAINED ON EXP INDS

```{r GO Model Training}
# start by defining a maximum number of splits
maxLevel <- log2(0.368*nrow(expEnv_red)/2)

# run gradient forest
# allele freq models
start_time <- Sys.time() # time start
gf_af <- gradientForest(cbind(expEnv_pop_red[2:9,c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")], freqs), predictor.vars = colnames(expEnv_pop_red[, c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")]), response.vars=colnames(freqs), ntree=500, maxLevel=maxLevel, trace=T, corr.threshold=0.50) # tons of warnings here, "response has five or fewer unique values. Are you sure you want to do regression?"
end_time <- Sys.time() # time end
(end_time - start_time) # about 2 mins

# geno models
# crashes on personal laptop, run on cluster
# start_time <- Sys.time() # time start
# gf_geno <- gradientForest(cbind(expEnv_red[,c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")], geno_dat), predictor.vars = colnames(expEnv_red[, c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")]), response.vars=colnames(geno_dat), ntree=500, maxLevel=maxLevel, trace=T, corr.threshold=0.50)
# end_time <- Sys.time() # time end
# (end_time - start_time)

# plot bar graphs depicting importance of each spatial and climate variable
plot(gf_af)
# plot(gf_geno, plot.type = "O")

# plots of turnover functions for individual loci - now having issues with ylim
plot(gf_af, plot.type = "C", imp.vars = gf_af$imp.vars, show.overall = T, legend = T, leg.posn = "topleft", leg.nspecies = 5, ylim = c(0,0.5)) # ylim error
# plot(gf_geno, plot.type = "C", imp.vars = gf_geno$imp.vars, show.overall = T, legend = T, leg.posn = "topleft", leg.nspecies = 5, ylim = c(0,0.5))
```

It looks like for the allele frequency based model, temp_quant90 is the most important climate variable, followed by either sal_quant1/temp_quant10 (accuracy importance/R2 weighted importance), then temp_quant10/sal_quant90, and finally sal_quant90/sal_quant10. 

For the genotype based model, ...

For both training runs, it's important to note that I had a significant number of warning messages - specifically:

"Warning in randomForest.default(x = X, y = spec_vec, maxLevel = maxLevel, : The response has five or fewer unique values. Are you sure you want to do regression?"

I think this is coming from non-variable genomic sites, which I could trim in future iterations of this model. Brandon has encountered this issue before and seems to think that the warning is internal to extendedForests (pckg for implementing random forests), and refers to number of unique AFs for a given locus.


### Predict Offset

Now that we have trained our gradient forest models, we need to use these to predict performance in the mesocosm treatments. This requires no new genomic data, just new environmental data that corresponds to each of the treatments and to the sites these populations originated from. Importantly, this new environmental data set must use the same column names as the data set used for training. We will ensure data compatibility below.

```{r Predict Offset Prep}
# check everything again
gf_af
gf_geno

# reverse order of pops list to match previous data
pops_list1 <- rev(pops_list) # VA is in the wrong spot, hard code instead
pops_list1 <- c("S1-LOLA", "S2-DEBY", "W1-TX", "W2-LA", "W3-FL", "W4-VA", "W5-NH", "W6-ME")

# process the env data for the gardens to match up with each pop/ind entry
# AF
af_lewenv <- expEnv_pop_red[expEnv_pop_red$site_name %in% c("LEW"), c("site_name", "sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")]
colnames(af_lewenv) <- c("garden", "sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")
pops_list_lew <- data.frame(site_name = pops_list1)
pops_list_lew$garden <- rep("LEW", 8)
af_lewenv_comp <- merge(pops_list_lew, af_lewenv, by = c("garden"))

af_yrkenv <- expEnv_pop_red[expEnv_pop_red$site_name %in% c("YRK"), c("site_name", "sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")]
colnames(af_yrkenv) <- c("garden", "sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled")
pops_list_yrk <- data.frame(site_name = pops_list1)
pops_list_yrk$garden <- rep("YRK", 8)
af_yrkenv_comp <- merge(pops_list_yrk, af_yrkenv, by = c("garden"))

# COME BACK TO DO GENO ONCE THIS MODEL CAN BE RUN ON THE CLUSTER AGAIN
```


Now I can predict offset using these different methods (genotype and allele frequency based).

```{r Predict Offset}
# prediction to home site
pred_af_home <- predict(gf_af, expEnv_pop_red[2:9, 2:5])
#pred_geno_home <- predict(gf_geno, curr_pop_env_gf_red)

# prediction to common gardens
pred_af_lew <- predict(gf_af, af_lewenv_comp[,3:6])
pred_af_yrk <- predict(gf_af, af_yrkenv_comp[,3:6])

#pred_geno_lew <- predict(gf_geno, geno_lewenv_comp[,3:6])
#pred_geno_yrk <- predict(gf_geno, geno_yrkenv_comp[,3:6])

# We now compute the euclidean distance between current and new cumulative importance
nb_ind_af <- nrow(pred_af_lew)
#nb_ind_geno <- nrow(pred_geno_lew)

# calculate offsets
euclidean <- function(a, b) sqrt(sum((a - b)^2)) # Euclidean distance for genetic offset

# af
# lew
df_af_lew <- data.frame() # df
for (i in 1:nb_ind_af){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_af_home[i,], pred_af_lew[i,]))
  df_af_lew <- rbind(df_af_lew, ec_dist)
}
af_offset_lew <- cbind(pops_list1, df_af_lew) # associate metadata with offset calculation, will need to ensure that this list is in the correct order
colnames(af_offset_lew) <- c("site_name", "offset_gf_af_lew")

# york
df_af_yrk <- data.frame() # df
for (i in 1:nb_ind_af){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_af_home[i,], pred_af_yrk[i,]))
  df_af_yrk <- rbind(df_af_yrk, ec_dist)
}
af_offset_yrk <- cbind(pops_list1, df_af_yrk) # associate metadata with offset calculation, will need to ensure that this list is in the correct order
colnames(af_offset_yrk) <- c("site_name", "offset_gf_af_yrk")

# merge
af_offsets <- merge(af_offset_lew, af_offset_yrk, by = c("site_name"))

# geno
# lew
df_geno_lew <- data.frame() # df
for (i in 1:nb_ind_af){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno[i,], pred_geno_garden[i,]))
  df_geno_lew <- rbind(df_geno_lew, ec_dist)
}
geno_offset_lew <- cbind(as.data.frame(pops_list), df_geno_lew) # associate metadata with offset calculation

# york
df_geno_yrk <- data.frame() # df
for (i in 1:nb_ind_af){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno[i,], pred_geno_garden[i,]))
  df_geno_yrk <- rbind(df_geno_yrk, ec_dist)
}
geno_offset_yrk <- cbind(as.data.frame(pops_list), df_geno_yrk) # associate metadata with offset calculation

# merge
geno_offsets <- merge(geno_offset_lew, geno_offset_yrk, by = c("site_name"))
offsets_mvp <- merge(af_offsets, geno_offsets, by = c("site_name"))

# write these offsets to results
write.csv(offsets_mvp, paste0("results/GO_results/offsetsMVP_GF_", Sys.Date(), ".csv"))
```

