---
title: "CviMVP_oyster_compareGOexp"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "~/Documents/GitHub/MVP_oyster_GO")
```

# Compare Predicted GO to Experimental Results

This script is for analyzing how well various genomic offset methods predict experimentally obtained fitness. I compare LEA, GF, and RDA offsets to experimental mortality, length, and condition index at the conclusion of the experiment. 

## Setup

```{r}
# setwd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(dplyr)
library(tidyr)
library(stringr)
library(vegan)
library(ggplot2)
library(plotrix)

# data
# offset predictions
lea_go <- read.csv("results/GO_results/offsets_LEA_expSites_05272025.csv")[,-1]
#rda_go <- 
gf_go <- read.csv("results/GO_results/offsetsMVP_2025-06-03.csv")[,-1]

# fitness metrics
mort <- read.csv("results/exp_results/mortality_bags_052925.csv")[,-1]
#len <- read.csv("data/ExpDat/")
```

## Data Prep

Prepare data from offsets & experiment to be directly compared.

```{r}
# inspect data
lea_go
gf_go
mort

# we want to get avg mort per pop per site
mort_popsite <- mort %>% group_by(bag_site,pop) %>%
  summarise(mean_t1 = mean(as.numeric(t1)), 
            sd_t1 = sd(as.numeric(t1)),
            mean_t2 = mean(as.numeric(t2)), 
            sd_t2 = sd(as.numeric(t2)),
            mean_t3 = mean(as.numeric(t3)), 
            sd_t3 = sd(as.numeric(t3)),
            mean_t4 = mean(as.numeric(t4)),
            sd_t4 = sd(as.numeric(t4)),
            .groups = 'drop') %>% as.data.frame()

# also get fitness rankings at each time
# t1
ordt1_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t1")])
ordt1_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t1")])
mort_popsite$ordt1 <- c(ordt1_lew, ordt1_york)
# t2
ordt2_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t2")])
ordt2_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t2")])
mort_popsite$ordt2 <- c(ordt2_lew, ordt2_york)
# t3
ordt3_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t3")])
ordt3_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t3")])
mort_popsite$ordt3 <- c(ordt3_lew, ordt3_york)
# t4
ordt4_lew <- order(mort_popsite[mort_popsite$bag_site == "Lewisetta", c("mean_t4")])
ordt4_york <- order(mort_popsite[mort_popsite$bag_site == "York River", c("mean_t4")])
mort_popsite$ordt4 <- c(ordt4_lew, ordt4_york)

# need to align naming conventions to merge
# LEA
colnames(lea_go) <- c("pop", "offset_lew", "order_lew", "offset_yrk", "order_yrk")
lea_go$pop <- str_split_fixed(lea_go$pop, "-",2)[,2]
#mort_popsite[mort_popsite$pop == "JR",]$pop <- "VA"

# GF
# add order
gf_go$order_lew <- order(gf_go$offset_af_lew)
gf_go$order_yrk <- order(gf_go$offset_af_yrk)
gf_go$pop <- str_split_fixed(gf_go$site_name,"-",2)[,2]

# merge gos & fitness dat
goFitLEA <- merge(mort_popsite, lea_go, by = "pop")
goFitGF <- merge(mort_popsite, gf_go, by = "pop")
```


## Compare Offsets to Fitness

Now that offsets and fitness are in a comparable format, look at how well the two match.

### LEA Offset

```{r}
# VA ISN'T SHOWING UP HERE????
# ALSO THE REASON CORRS CHANGED SO MUCH BETWEEN LAST ANALYSIS AND CURRENT ANALYSIS MAY BE BECAUSE OF DIFFERENT BAGS BEING TRIMMED FROM ANALYSIS, WHICH LIKELY CHANGED THE RAW SURVIVAL NUMBERS FOR SOME POPS

# correlations between go/lew mort at each time point
# lea
(cor_lea1_lew <- cor(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t1)) # -0.3502308
(cor_lea2_lew <- cor(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t2)) # -0.1718701
(cor_lea3_lew <- cor(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t3)) # -0.1407716
(cor_lea4_lew <- cor(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t4)) # -0.1650885

# correlations between go/york mort at each time point
# lea
(cor_lea1_yrk <- cor(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t1)) # -0.00143748
(cor_lea2_yrk <- cor(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t2)) # 0.01642408
cor_lea3_yrk <- cor(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, (goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t3)) # -0.1755285
(cor_lea4_yrk <- cor(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t4)) # -0.1882313

# these corrs aren't great, try kendall's tau
cor(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t1, method = "kendall") # 0.04761905
cor(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t1, method = "kendall") # -0.2380952

cor(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t2, method = "kendall") # 0.04761905
cor(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t2, method = "kendall") # 0.04761905

cor(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t3, method = "kendall") # -0.04761905
cor(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t3, method = "kendall") # 0.04761905

cor(goFitLEA[goFitLEA$bag_site == "YorkRiver",]$offset_yrk, goFitLEA[goFitLEA$bag_site == "YorkRiver",]$mean_t4, method = "kendall") # 0.04761905
cor(goFitLEA[goFitLEA$bag_site == "Lewisetta",]$offset_lew, goFitLEA[goFitLEA$bag_site == "Lewisetta",]$mean_t4, method = "kendall") # -0.04761905

# put this all into a df with method/time point/fitness metric/site/corr
lea_corr_df <- data.frame(method = rep("LEA", 8), metric = rep("surv", 8), monitoring_event = rep(c(1:4),2), bag_site = c(rep("Lewisetta",4 ), rep("YorkRiver", 4)), pearson = c(cor_lea1_lew, cor_lea2_lew, cor_lea3_lew, cor_lea4_lew, cor_lea1_yrk, cor_lea2_yrk, cor_lea3_yrk, cor_lea4_yrk))
```


Plot these data sets against each other

```{r}
# plot lea vs mort at each site/time point
goFitLEA_lew_plot <- ggplot(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",]) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t1, y = offset_lew, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t1, y = offset_lew), col = "cornflowerblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t2, y = offset_lew, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t2, y = offset_lew), col = "slateblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t3, y = offset_lew, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$bag_site == "Lewisetta",], aes(x = mean_t3, y = offset_lew), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at Lewisetta", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24"), values = c("Fall'23"="cornflowerblue","Spring'24"="slateblue","Fall'24"="purple4"))+
  theme_classic()
goFitLEA_lew_plot
ggsave("results/lea_lew_compare.png")

goFitLEA_yrk_plot <- ggplot(data = goFitLEA[goFitLEA$site == "York River",]) +
  geom_point(data = goFitLEA[goFitLEA$site == "York River",], aes(x = mean_t1, y = offset_yrk, col = "Fall'23")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$site == "York River",], aes(x = mean_t1, y = offset_yrk), col = "cornflowerblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$site == "York River",], aes(x = mean_t2, y = offset_yrk, col = "Spring'24")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$site == "York River",], aes(x = mean_t2, y = offset_yrk), col = "slateblue", se = F) +
  geom_point(data = goFitLEA[goFitLEA$site == "York River",], aes(x = mean_t3, y = offset_yrk, col = "Fall'24")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$site == "York River",], aes(x = mean_t3, y = offset_yrk), col = "purple4", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at York River", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24"), values = c("Fall'23"="cornflowerblue","Spring'24"="slateblue","Fall'24"="purple4"))+
  theme_classic()
goFitLEA_yrk_plot
ggsave("results/lea_york_compare.png")

# plot go lea vs mort at both sites just final time point
goFitLEA_plotT3 <- ggplot(data = goFitLEA) +
  geom_point(data = goFitLEA[goFitLEA$site == "Lewisetta",], aes(x = mean_t3, y = offset_lew, col = "Lewisetta")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$site == "Lewisetta",], aes(x = mean_t3, y = offset_lew), col = "mediumspringgreen", se = F) +
  geom_point(data = goFitLEA[goFitLEA$site == "York River",], aes(x = mean_t3, y = offset_lew, col = "York River")) +
  geom_smooth(method = "lm", data = goFitLEA[goFitLEA$site == "York River",], aes(x = mean_t3, y = offset_lew), col = "dodgerblue", se = F) +
  labs(title = "Predicted GO (LEA) vs Mortality at Final Time Point", x = "Survival", y = "LEA Offset") +
  scale_colour_manual(name = "Experimental Site", breaks = c("Lewisetta", "York River"), values = c("Lewisetta"="mediumspringgreen","York River"="dodgerblue"))+
  theme_classic()
goFitLEA_plotT3
```


### GF Offset

```{r}
# correlations between go/lew mort at each time point
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t1) # -0.08865028
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t2) # 0.1247669
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t3) # 0.08480605
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$offset_af_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t4) # 0.06622808

# correlations between go/yrk mort at each time point
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t1) # 0.3556833
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t2) # 0.3466489
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t3) # 0.1280315
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$offset_af_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t4) # 0.1363403

# really bad correlations, may need to double check offset calculations
# pearson of orders - LEW
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t1) # -0.2063336
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t2) # -0.0903556
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t3) # -0.0654962
cor(goFitGF[goFitGF$bag_site == "Lewisetta",]$order_lew, goFitGF[goFitGF$bag_site == "Lewisetta",]$mean_t4) # -0.1185905
# pearson of orders - YRK
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t1) # 0.1376695
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t2) # 0.1340658
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t3) # -0.1269526
cor(goFitGF[goFitGF$bag_site == "YorkRiver",]$order_yrk, goFitGF[goFitGF$bag_site == "YorkRiver",]$mean_t4) # -0.1199833
```


## Compare Offset Methods

Now that I have predictions from different methods, I will look at how offset predictions made by these methods compare to each other and how different their ability is to predict experimental fitness.


### Compare Offsets to Each Other

```{r Offset Compilation}
# make a df with all of the offsets in them
offsets_lea_exp <- goFitLEA[,c("pop", "bag_site","offset_lew", "offset_yrk", "order_lew", "order_yrk")]
colnames(offsets_lea_exp) <- c("pop", "bag_site","lea_offset_lew", "lea_offset_yrk", "lea_order_lew", "lea_order_yrk")
offsets_gf_exp <- goFitGF[,c("pop", "bag_site","offset_af_lew", "offset_af_yrk", "order_lew", "order_yrk")]
colnames(offsets_gf_exp) <- c("pop", "bag_site","gf_offset_af_lew", "gf_offset_af_yrk", "gf_order_lew", "gf_order_yrk")
offsets_all <- merge(offsets_lea_exp, offsets_gf_exp, by = c("pop", "bag_site"))
```


```{r Offset Comparison}
# corr gf/lea
cor(offsets_all[offsets_all$bag_site == "Lewisetta",]$lea_offset_lew,  offsets_all[offsets_all$bag_site == "Lewisetta",]$gf_offset_af_lew) # 0.8234259
cor(offsets_all[offsets_all$bag_site == "YorkRiver",]$lea_offset_yrk,  offsets_all[offsets_all$bag_site == "YorkRiver",]$gf_offset_af_yrk) # 0.836363

ggplot(data = offsets_all) + 
  geom_point(data = offsets_all[offsets_all$bag_site == "Lewisetta",], aes(x = lea_offset_lew, y = gf_offset_af_lew, fill = pop), pch = 21, size = 3) +
  geom_smooth(method = "lm", data = offsets_all[offsets_all$bag_site == "Lewisetta",], aes(x = lea_offset_lew, y = gf_offset_af_lew), col = "black") +
  labs(title = "LEA vs GF Offsets Lewisetta", x = "LEA Offset", y = "GF Offset", fill = "Broodstock Group") +
  theme_classic()

ggplot(data = offsets_all) + 
  geom_point(data = offsets_all[offsets_all$bag_site == "YorkRiver",], aes(x = lea_offset_yrk, y = gf_offset_af_yrk, fill = pop), pch = 21, size = 3) +
  geom_smooth(method = "lm", data = offsets_all[offsets_all$bag_site == "YorkRiver",], aes(x = lea_offset_yrk, y = gf_offset_af_yrk), col = "black") +
  labs(title = "LEA vs GF Offsets York River", x = "LEA Offset", y = "GF Offset", fill = "Broodstock Group") +
  theme_classic()

# corr gf/rda


# corr lea/rda

```

