---
title: "CviMVP_LEA_offset"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LEA Genomic offsets

This script is a LEA pipeline for calculating genomic offset. I am using LEA3, which allows for multivariate environmental datasets. I first train the models using LFMM of a large seascape dataset, then calculate offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment. Offset is calculated from the environment of origin to common garden conditions.

## Setup

```{r}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(LEA) # for offsets
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(ggplot2) # for plotting
library(vegan) # for env dat processing
library(dplyr) # for data wrangling
library(stringr) # for data wrangling

# data
# env
env_all <- read.csv("data/EnvDat/env_scaled_2025-09-26.csv")[,-1]
env_rast <- read_rds("data/EnvDat/seascape/raster_interpolated_scaled_env.RDS")
env_sea <- .
env_sea_int <- .
env_exp <- .

# sites
sea_site_latlon <- read.csv("data/EnvDat/seascape/seascape_sitenames.csv")[,-1]
exp_site_latlon <- read.csv("data/EnvDat/exp/exp_site_info.csv")[,-1]

# gen
genoMat <- read.geno("data/GenDat/geno_seascape_2025-08-14.geno")
genoLfmm <- read.lfmm("data/GenDat/geno_seascape_2025-08-14.lfmm")
genoThinMat <- read.geno("data/GenDat/geno_seascape_LDthin_2025-08-14.geno")
genoThinLfmm <- read.lfmm("data/GenDat/geno_seascape_LDthin_2025-08-14.lfmm")

# inds
inds_genomat <- read.csv("data/GenDat/inds_genomat.csv")[,-1]
indsSea <- readRDS("data/IndDat/20250604_seascape_indsmatrix.rds")
```


## Data Prep

### Environmental Data Prep

```{r}
# just renaming columns & rows to be more informative
colnames(env_sea) <- colnames(env_sea_expand) <- c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")

colnames(env_exp) <- colnames(env_exp_expand) <- c("sal_quant10_scaled", "sal_quant90_scaled", "temp_quant10_scaled", "temp_quant90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "lew_sal_quant10_scaled", "lew_sal_quant90_scaled", "lew_temp_quant10_scaled", "lew_temp_quant90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled", "yrk_sal_quant10_scaled", "yrk_sal_quant90_scaled", "yrk_temp_quant10_scaled", "yrk_temp_quant90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")

rownames(env_sea) <- unique(env_sea_csv$site_name)
rownames(env_exp) <- env_exp_sites
rownames(env_sea_expand) <- indsSea[indsSea$ID_SiteDate %in% env_sea_sites, "ID_SiteDate"]
rownames(env_exp_expand) <- indsExp[!(indsExp$ID_SiteDate %in% c("DEBY", "LOLA")), "ID_SiteDate"]
```


### Genetic Data Prep

```{r}
# name the rows of genomic data
rownames(genoMat) <- rownames(genoLfmm) <- rownames(genoThinMat) <- rownames(genoThinLfmm) <- inds_genomat

# make sure genotype data is in the same order as individual & environmental data
#env_sea_csv$clean_ID %in% inds_genomat
#env_sea_csv$clean_ID == inds_genomat # yes, everything is in the correct order
```


## Seascape Training

Now that I have prepped and cleaned my data, I will use it to train LFMM models later used to predict genomic offset to some other (in this case experimental) climate. Importantly, I do not need to trim or impute my genomic data, or determine ancestry - Madeline has already done this. The data has been LD thinned, trimmed for MAF, and ancestry analysis has recovered K=2.
Madeline also compiled the environmental data, which I further processed in the script "CviMVP_oyster_EnvDatProcess."

May decide to train both genotype-based models and allele frequency based models - for now genotype models only.

```{r}
# check the data
#genoLfmm[1:10, 1:10]
#genoThinLfmm[1:10, 1:10]

# fit model
mod_lfmm <- lfmm2(input = genoLfmm,
                  env = env_sea_expand, 
                  K = 2, # this was number from Madeline's snmf analysis
                  effect.sizes = T) 
mod_lfmmthin <- lfmm2(input = genoThinLfmm,
                      env = env_sea_expand,
                      K = 2,
                      effect.sizes = T)

# perform test
pv_lfmm <- lfmm2.test(object = mod_lfmm,
                      input = genoLfmm,
                      env = env_sea_expand, 
                      full = TRUE)
pv_lfmmthin <- lfmm2.test(object = mod_lfmmthin,
                      input = genoThinLfmm,
                      env = env_sea_expand, 
                      full = TRUE)

# pvalues
pv <- pv_lfmm$pvalues
pvthin <- pv_lfmmthin$pvalues

# qvalues
library(qvalue)
qv <- qvalue(pv, fdr.level = 0.01)
qvthin <- qvalue(pvthin, fdr.level = 0.01)

# candidates
candidates <- which(qv$significant)
candidates_pvals <- (pv)[candidates]

candidatesthin <- which(qvthin$significant)
candidates_pvalsthin <- (pvthin)[candidates]

# using FDR to filter out false positives
length(which(qv$qvalues < 0.001)) # 12455
length(which(qvthin$qvalues < 0.001)) # 5430

# identify which SNPs have FDR less than 1%
fdr_cands <- colnames(genoLfmm)[which(qv$qvalues < 0.001)]
fdr_candsthin <- colnames(genoThinLfmm)[which(qv$qvalues < 0.001)]

# df for candidates
cand_df <- data.frame(index = colnames(genoLfmm), pos = 1:length(pv),
                      pvals = pv, qvals = qv$qvalues)
cand_df_thin <- data.frame(index = colnames(genoThinLfmm), pos = 1:length(pvthin),
                      pvals = pvthin, qvals= qvthin$qvalues)

# manhattan plot - unnecessary but just a quick visualization
plot(cand_df$pos, -log10(cand_df$pvals), 
     pch = 19, col = "darkslategray4",
     main = "Candidate SNPs",
     xlab = "Position",
     ylab = "-log10(Pvals)")
points(cand_df$pos[cand_df$qvals < 0.001], 
       -log10(cand_df$pvals[cand_df$qvals < 0.001]),
       pch = 19,
       col = "hotpink")

plot(cand_df_thin$pos, -log10(cand_df_thin$pvals), 
     pch = 19, col = "darkslategray4",
     main = "Candidate SNPs (Thinned)",
     xlab = "Position",
     ylab = "-log10(Pvals)")
points(cand_df_thin$pos[cand_df_thin$qvals < 0.001], 
       -log10(cand_df_thin$pvals[cand_df_thin$qvals < 0.001]),
       pch = 19,
       col = "hotpink")
```


## Offset Prediction

```{r}
# check env data
colnames(env_exp_csv)
colnames(env_exp_expand)
colnames(env_exp_csv)
colnames(env_exp_expand) <- c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "lew_salinity_quantile_10_scaled",  "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled", "yrk_salinity_quantile_10_scaled",  "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")

# I CANNOT CALCULATE OFFSET FOR MY EXPERIMENTAL SITES WITH THE BASE FUNCTION - I NEED TO FOLLOW THE VIGNETTE FROM SSMPG https://github.com/bcm-uga/SSMPG2025/blob/main/Tutorials/Tutorial%20visualization/Spatial_Prediction_Genomic_Offset.Rmd

# first, I need to get environmental effect sizes from the trained models
B <- mod_lfmm@B
Bthin <- mod_lfmmthin@B

# not using candidate loci for now
# calculate "genetic gap" for all the experimental sites
gg_lew = NULL
for (i in 1:nrow(env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")])){
gg_lew[i] = mean(((env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")] - env_exp_expand[,c("lew_salinity_quantile_10_scaled",  "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled")])[i,]  %*% t(B))^2, na.rm = TRUE)
}

gg_yrk = NULL
for (i in 1:nrow(env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")])){
gg_yrk[i] = mean(((env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")] - env_exp_expand[,c("yrk_salinity_quantile_10_scaled",  "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")])[i,]  %*% t(B))^2, na.rm = TRUE)
}

gg_lewthin = NULL
for (i in 1:nrow(env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")])){
gg_lewthin[i] = mean(((env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")] - env_exp_expand[,c("lew_salinity_quantile_10_scaled",  "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled")])[i,]  %*% t(Bthin))^2, na.rm = TRUE)
}

gg_yrkthin = NULL
for (i in 1:nrow(env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")])){
gg_yrkthin[i] = mean(((env_exp_expand[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")] - env_exp_expand[,c("yrk_salinity_quantile_10_scaled",  "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")])[i,]  %*% t(Bthin))^2, na.rm = TRUE)
}

# associate predictions with site metadata
gap_df <- env_exp_csv[,c("site_name","merge_ID_inds","Label_ind")]
gap_df$lew_go <- gg_lew
gap_df$yrk_go <- gg_yrk
gap_df$lew_gothin <- gg_lewthin
gap_df$yrk_gothin <- gg_yrkthin

# reduce to single pop level prediction here, no need to have indiv level unless we end up with individual level disease data for these pops
gap_df_red <- gap_df[!duplicated(gap_df$site_name), c("site_name", "lew_go", "yrk_go", "lew_gothin", "yrk_gothin")]

# output results
write.csv(gap_df_red, paste0("results/GO_results/offsets_LEA_",Sys.Date(),".csv", sep = ""))
```

## Offset Visualization

```{r}
# need to add latlongs for each site
offsets_plotting <- merge(exp_site_latlon, gap_df_red, by =  c("site_name"))

# add the color palette
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME", "LARMIX", "SEEDMIX"), 
                     order = 1:10, 
                     cols = c("#3e4989" ,"#31688e", "#26828e", "#1f9e89", "#440154", "#482878", "#35b779", "#6ece58", "#fde725", "#b5de2b"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME", "H1_LARMIX", "H2_SEEDMIX"))
offsets_plotting_full <- merge(offsets_plotting, popsdf, by = c("site_name"))

# save popdf object
write.csv(popsdf, "data/IndDat/popsdf.csv")

# try mapping the offsets
library(s2) # mapping
library(rnaturalearth) # mapping
library(rnaturalearthdata) # mapping
library(maps) # mapping
library(ggspatial) # mapping

# map
sf::sf_use_s2(FALSE) # sf package doesn't work

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin =-102 , xmax = -68, ymin = 24, ymax = 47.049130))

# bottom right: 24.353558, -55.373595
# top left: 47.049130, -101.983412

# offset to lew
map_lew <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting_full, aes(x=lon, y=lat, 
                                       fill = lew_go), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,1.2))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to Lewisetta (LEA)", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_lew)

world_chesapeake <- sf::st_crop(world, c(xmin =-77.856476 , xmax = -72.762787, ymin = 36.019557, ymax = 39.939163))

# top left: 39.939163, -77.856476
# bottom right: 36.019557, -72.762787

popout_lew <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting_full, aes(x=lon, y=lat, 
                                       fill = lew_go), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,1.2))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
popout_lew

# offset to york
map_york <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting_full, aes(x=lon, y=lat, 
                                       fill = yrk_go), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,1.2))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to York River (LEA)", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_york)

popout_york <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = offsets_plotting_full, aes(x=lon, y=lat, 
                                       fill = yrk_go), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,1.2))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
popout_york

cor(offsets_plotting_full$offset_lew, offsets_plotting_full$offset_yrk)
plot(x = offsets_plotting_full$offset_lew, y = offsets_plotting_full$offset_yrk)
```


