---
title: "CviMVP_oyster_compareGOexp"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "~/Documents/GitHub/MVP_oyster_GO")
```

# Compare Predicted GO to Experimental Results

This script is for analyzing how well various genomic offset methods predict experimentally obtained fitness. I compare LEA, GF, and RDA offsets to experimental mortality, length, and condition index at the conclusion of the experiment. 

## Setup

```{r}
# setwd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(dplyr)
library(tidyr)
library(stringr)
library(vegan)
library(ggplot2)
library(plotrix)

# data
# offset predictions
lea_go <- read.csv("results/GO_results/offsets_LEA_expsites_2025-12-08.csv")[,-1]
rda_go <- read.csv("results/GO_results/offsets_RDA_expsites_2025-12-08.csv")[,-1]
gf_go <- read.csv("results/GO_results/offsets_GF_expSites_2025-11-24.csv")[,-1]
env_dist <- read.csv("data/EnvDat/exp/distance_matrix_2025-12-05.csv")[28:35,c("X","Lewisetta","YorkRiver")]
env_dist_nb <- read.csv("data/EnvDat/exp/distance_matrix_abiotic_2025-12-05.csv")[28:35,c("X","Lewisetta","YorkRiver")]

# fitness metrics
mort_bags <- read.csv("results/exp_results/mortality_long_bags_2025-06-06.csv")[,-1]
mort <- read.csv("results/exp_results/mortality_long_popsite_2025-06-06.csv")[,-1]
len <- read.csv("results/exp_results/lengths_bags_event_2025-06-03.csv")[,-1]
```

## Data Prep

Prepare data from offsets & experiment to be directly compared.

```{r}
# inspect data
lea_go
gf_go
rda_go
mort
len

# trim mort
mort_popsite <- mort[!(mort$event == 0), c("bag_site","pop","mean_surv","sd","event")]
colnames(mort_popsite) <- c("bag_site","pop","mean_surv","sd_surv","event")

# also get fitness rankings at each time
# mortality
ord_mort_popsite <- data.frame()
for (i in 1:4) {
  ordt_lew <- mort_popsite[mort_popsite$event == i & mort_popsite$bag_site == "Lewisetta",]
  ordt_lew <- ordt_lew %>% arrange(desc(mean_surv))
  ordt_lew$order_mort <- 1:length(ordt_lew$pop)
  ordt_yrk <- mort_popsite[mort_popsite$event == i & mort_popsite$bag_site == "YorkRiver",]
  ordt_yrk <- ordt_yrk %>% arrange(desc(mean_surv))
  ordt_yrk$order_mort <- 1:length(ordt_yrk$pop)
  ord_mort_popsite <- rbind(ord_mort_popsite, ordt_lew, ordt_yrk)
}
ord_mort_popsite

# do the same with length
# first length needs a pop label and site label
len_bags <- as.data.frame(cbind(len$bag, str_split_fixed(len$bag, "-", 3)[,2:3]))
colnames(len_bags) <- c("bag","pop","bagnum")
len_bags$bagnum <- as.numeric(len_bags$bagnum)
len_bags$site <- ifelse(len_bags$bagnum > 3, "YorkRiver", "Lewisetta")
len_all <- merge(len_bags[,c("bag","pop","site")], len, by = c("bag"))
colnames(len_all) <- c("bag","pop","site","event","length")

len_t1 <- len_all[len_all$event == 1,]
t1_lens <- aggregate(length ~ pop + site, data = len_t1, FUN = mean)
t1_sd <- aggregate(length ~ pop + site, data = len_t1, FUN = sd)
colnames(t1_sd) <- c("pop", "site", "len_sd")
t1_lens_popsite <- merge(t1_lens, t1_sd, by = c("pop", "site"))
t1_lens_popsite$event <- 1
len_t2 <- len_all[len_all$event == 2,]
t2_lens <- aggregate(length ~ pop + site, data = len_t2, FUN = mean)
t2_sd <- aggregate(length ~ pop + site, data = len_t2, FUN = sd)
colnames(t2_sd) <- c("pop", "site", "len_sd")
t2_lens_popsite <- merge(t2_lens, t2_sd, by = c("pop", "site"))
t2_lens_popsite$event <- 2
len_t3 <- len_all[len_all$event == 3,]
t3_lens <- aggregate(length ~ pop + site, data = len_t3, FUN = mean)
t3_sd <- aggregate(length ~ pop + site, data = len_t3, FUN = sd)
colnames(t3_sd) <- c("pop", "site", "len_sd")
t3_lens_popsite <- merge(t3_lens, t3_sd, by = c("pop", "site"))
t3_lens_popsite$event <- 3
len_t4 <- len_all[len_all$event == 4,]
t4_lens <- aggregate(length ~ pop + site, data = len_t4, FUN = mean)
t4_sd <- aggregate(length ~ pop + site, data = len_t4, FUN = sd)
colnames(t4_sd) <- c("pop", "site", "len_sd")
t4_lens_popsite <- merge(t4_lens, t4_sd, by = c("pop", "site"))
t4_lens_popsite$event <- 4

lens_popsite <- rbind(t1_lens_popsite, t2_lens_popsite, t3_lens_popsite, t4_lens_popsite)
colnames(lens_popsite) <- c("pop","bag_site","mean_len","sd_len","event")

# and orders
ord_len_popsite <- data.frame()
for (i in 1:4) {
  ordt_lew <- lens_popsite[lens_popsite$event == i & lens_popsite$bag_site == "Lewisetta",]
  ordt_lew <- ordt_lew %>% arrange(desc(mean_len))
  ordt_lew$order_len <- 1:length(ordt_lew$pop)
  ordt_yrk <- lens_popsite[lens_popsite$event == i & lens_popsite$bag_site == "YorkRiver",]
  ordt_yrk <- ordt_yrk %>% arrange(desc(mean_len))
  ordt_yrk$order_len <- 1:length(ordt_yrk$pop)
  ord_len_popsite <- rbind(ord_len_popsite, ordt_lew, ordt_yrk)
}
ord_len_popsite

# put length and mortality together
ord_fitness_popsite_mg <- merge(ord_mort_popsite, ord_len_popsite, by = c("pop","bag_site","event"))

# remove the larmix/seedmix pops
ord_fitness_popsite <- ord_fitness_popsite_mg[!ord_fitness_popsite_mg$pop %in% c("LARMIX","SEEDMIX"),]

# need a site name column
ord_fitness_popsite$site_name <- ord_fitness_popsite$pop
ord_fitness_popsite$site_name <- dplyr::recode(ord_fitness_popsite$site_name,
                                        "TX" = "W1_TX", 
                                        "LA" = "W2_LA", 
                                        "FL" = "W3_FL",
                                        "LOLA" = "S1_LOLA", 
                                        "DEBY" = "S2_DEBY",
                                        "JR" = "W4_VA",
                                        "NH" = "W5_NH", 
                                        "ME" = "W6_ME")

# do bag level data as well
colnames(mort_bags) <- c("bag","event","surv")
colnames(len) <- c("bag","event","length")
bagfit <- merge(mort_bags, len, by = c("bag","event"))

# pop info
bagfit$pop <- str_split_fixed(bagfit$bag, "-", 3)[,2]
bagfit

# site info
bagfit$bag_site <- ifelse(str_split_fixed(bagfit$bag, "-", 3)[,3] %in% c(1,2,3), "Lewisetta","YorkRiver")

# add this to other fitness data
bagfullfit <- merge(ord_fitness_popsite, bagfit, by = c("pop","event","bag_site"))

# need to align naming conventions to merge
# LEA
# colnames
colnames(lea_go) <- c("site_name", "lat", "lon", "offset_lea_go_lew","offset_lea_go_yrk","offset_lea_go_lew_nb","offset_lea_go_yrk_nb")

# add order
order_lew_lea <- as.data.frame(lea_go[order(lea_go$offset_lea_go_lew, decreasing = T),]$site_name)
colnames(order_lew_lea) <- c("site_name")
order_lew_lea$order_lew_lea <- 1:length(order_lew_lea$site_name)

order_lew_lea_nb <- as.data.frame(lea_go[order(lea_go$offset_lea_go_lew_nb, decreasing = T),]$site_name)
colnames(order_lew_lea_nb) <- c("site_name")
order_lew_lea_nb$order_lew_lea_nb <- 1:length(order_lew_lea_nb$site_name)

order_yrk_lea <- as.data.frame(lea_go[order(lea_go$offset_lea_go_yrk, decreasing = T),]$site_name)
colnames(order_yrk_lea) <- c("site_name")
order_yrk_lea$order_yrk_lea <- 1:length(order_yrk_lea$site_name)

order_yrk_lea_nb <- as.data.frame(lea_go[order(lea_go$offset_lea_go_yrk_nb, decreasing = T),]$site_name)
colnames(order_yrk_lea_nb) <- c("site_name")
order_yrk_lea_nb$order_yrk_lea_nb <- 1:length(order_yrk_lea_nb$site_name)

orders_lea_b <- merge(order_lew_lea, order_yrk_lea, by = c("site_name"))
orders_lea_nb <- merge(order_lew_lea_nb, order_yrk_lea_nb, by = ("site_name"))
orders_lea <- merge(orders_lea_b, orders_lea_nb, by = c("site_name"))
lea_go1 <- merge(lea_go, orders_lea, by = c("site_name"))

# GF
# add order
order_lew_af <- as.data.frame(gf_go[order(gf_go$offset_gf_af_lew, decreasing = T),]$site_name)
colnames(order_lew_af) <- c("site_name")
order_lew_af$order_lew_af_gf <- 1:length(order_lew_af$site_name)

order_lew_af_nb <- as.data.frame(gf_go[order(gf_go$offset_gf_af_lew_nb, decreasing = T),]$site_name)
colnames(order_lew_af_nb) <- c("site_name")
order_lew_af_nb$order_lew_af_gf_nb <- 1:length(order_lew_af_nb$site_name)

order_yrk_af <- as.data.frame(lea_go[order(gf_go$offset_gf_af_yrk, decreasing = T),]$site_name)
colnames(order_yrk_af) <- c("site_name")
order_yrk_af$order_yrk_af_gf <- 1:length(order_yrk_af$site_name)

order_yrk_af_nb <- as.data.frame(lea_go[order(gf_go$offset_gf_af_yrk_nb, decreasing = T),]$site_name)
colnames(order_yrk_af_nb) <- c("site_name")
order_yrk_af_nb$order_yrk_af_gf_nb <- 1:length(order_yrk_af_nb$site_name)

orders_gf_af_b <- merge(order_lew_af, order_yrk_af, by = c("site_name"))
orders_gf_af_nb <- merge(order_lew_af_nb, order_yrk_af_nb, by = c("site_name"))
orders_gf_af <- merge(orders_gf_af_b, orders_gf_af_nb, by = c("site_name"))
gf_go1 <- merge(gf_go, orders_gf_af, by = c("site_name"))

order_lew_geno <- as.data.frame(gf_go[order(gf_go$offset_gf_geno_lew, decreasing = T),]$site_name)
colnames(order_lew_geno) <- c("site_name")
order_lew_geno$order_lew_geno_gf <- 1:length(order_lew_geno$site_name)

order_lew_geno_nb <- as.data.frame(gf_go[order(gf_go$offset_gf_geno_lew_nb, decreasing = T),]$site_name)
colnames(order_lew_geno_nb) <- c("site_name")
order_lew_geno_nb$order_lew_geno_gf_nb <- 1:length(order_lew_geno_nb$site_name)

order_yrk_geno <- as.data.frame(lea_go[order(gf_go$offset_gf_geno_yrk, decreasing = T),]$site_name)
colnames(order_yrk_geno) <- c("site_name")
order_yrk_geno$order_yrk_geno_gf <- 1:length(order_yrk_geno$site_name)

order_yrk_geno_nb <- as.data.frame(lea_go[order(gf_go$offset_gf_geno_yrk_nb, decreasing = T),]$site_name)
colnames(order_yrk_geno_nb) <- c("site_name")
order_yrk_geno_nb$order_yrk_geno_gf_nb <- 1:length(order_yrk_geno_nb$site_name)

orders_gf_geno_b <- merge(order_lew_geno, order_yrk_geno, by = c("site_name"))
orders_gf_geno_nb <- merge(order_lew_geno_nb, order_yrk_geno_nb, by = c("site_name"))
orders_gf_geno <- merge(orders_gf_geno_b, orders_gf_geno_nb, by = c("site_name"))
gf_go2 <- merge(gf_go1, orders_gf_geno, by = c("site_name"))

# RDA
# colnames
colnames(rda_go) <- c("site_name","lat","lon","offset_rda_lew","offset_rda_yrk","offset_rda_lew_nb","offset_rda_yrk_nb")

# add order
order_lew_rda <- as.data.frame(rda_go[order(rda_go$offset_rda_lew, decreasing = T),]$site_name)
colnames(order_lew_rda) <- c("site_name")
order_lew_rda$order_lew_rda <- 1:length(order_lew_rda$site_name)

order_lew_rda_nb <- as.data.frame(rda_go[order(rda_go$offset_rda_lew_nb, decreasing = T),]$site_name)
colnames(order_lew_rda_nb) <- c("site_name")
order_lew_rda_nb$order_lew_rda_nb <- 1:length(order_lew_rda_nb$site_name)

order_yrk_rda <- as.data.frame(rda_go[order(rda_go$offset_rda_yrk, decreasing = T),]$site_name)
colnames(order_yrk_rda) <- c("site_name")
order_yrk_rda$order_yrk_rda <- 1:length(order_yrk_rda$site_name)

order_yrk_rda_nb <- as.data.frame(rda_go[order(rda_go$offset_rda_yrk_nb, decreasing = T),]$site_name)
colnames(order_yrk_rda_nb) <- c("site_name")
order_yrk_rda_nb$order_yrk_rda_nb <- 1:length(order_yrk_rda_nb$site_name)

orders_rda_b <- merge(order_lew_rda, order_yrk_rda, by = c("site_name"))
orders_rda_nb <- merge(order_lew_rda_nb, order_yrk_rda_nb, by = c("site_name"))
orders_rda <- merge(orders_rda_b, orders_rda_nb, by = c("site_name"))
rda_go1 <- merge(rda_go, orders_rda, by = c("site_name"))

# environmental distance
colnames(env_dist) <- c("site_name", "env_dist_lew", "env_dist_yrk")
colnames(env_dist_nb) <- c("site_name", "env_dist_lew_nb", "env_dist_yrk_nb")
env_dist1 <- merge(env_dist, env_dist_nb, by = "site_name")
# add order
order_lew_envdist <- as.data.frame(env_dist1[order(env_dist1$env_dist_lew, decreasing = T),]$site_name)
colnames(order_lew_envdist) <- "site_name"
order_lew_envdist$order_lew_envdist <- 1:length(order_lew_envdist$site_name)

order_lew_envdist_nb <- as.data.frame(env_dist1[order(env_dist1$env_dist_lew_nb, decreasing = T),]$site_name)
colnames(order_lew_envdist_nb) <- "site_name"
order_lew_envdist_nb$order_lew_envdist_nb <- 1:length(order_lew_envdist_nb$site_name)

order_yrk_envdist <- as.data.frame(env_dist1[order(env_dist1$env_dist_yrk, decreasing = T),]$site_name)
colnames(order_yrk_envdist) <- "site_name"
order_yrk_envdist$order_yrk_envdist <- 1:length(order_yrk_envdist$site_name)

order_yrk_envdist_nb <- as.data.frame(env_dist1[order(env_dist1$env_dist_yrk_nb, decreasing = T),]$site_name)
colnames(order_yrk_envdist_nb) <- "site_name"
order_yrk_envdist_nb$order_yrk_envdist_nb <- 1:length(order_yrk_envdist_nb$site_name)

orders_envdist_b <- merge(order_lew_envdist, order_yrk_envdist, by = "site_name")
orders_envdist_nb <- merge(order_lew_envdist_nb, order_yrk_envdist_nb, by = "site_name")
orders_envdist <- merge(orders_envdist_b, orders_envdist_nb, by = "site_name")
env_dist2 <- merge(env_dist1, orders_envdist, by = "site_name")

# put together full go & fitness df
go <- merge(rda_go1, gf_go2, by = c("site_name","lat","lon")) 
go1 <- merge(go, lea_go1, by = c("site_name", "lat", "lon"))
go2 <- merge(go1, env_dist2, by = "site_name")
fullfit <- merge(bagfullfit, go2, by = "site_name")

# make sure everything is coded correctly
fullfit$order_lew_envdist <- as.numeric(fullfit$order_lew_envdist)
fullfit$order_yrk_envdist <- as.numeric(fullfit$order_yrk_envdist)
fullfit$order_lew_envdist_nb <- as.numeric(fullfit$order_lew_envdist_nb)
fullfit$order_yrk_envdist_nb <- as.numeric(fullfit$order_yrk_envdist_nb)
fullfit$env_dist_lew <- as.numeric(fullfit$env_dist_lew)
fullfit$env_dist_yrk <- as.numeric(fullfit$env_dist_yrk)
fullfit$env_dist_lew_nb <- as.numeric(fullfit$env_dist_lew_nb)
fullfit$env_dist_yrk_nb <- as.numeric(fullfit$env_dist_yrk_nb)
```


## Compare Offsets to Fitness

Now that offsets and fitness are in a comparable format, look at how well the two match.

### Performance Metric Calculation

Here I calculate the R^2 and pearson's correlation between fitness and GO for each method at each site. I do this with fitness data both at the bag level and summarized at the population level (since offset is population level).

```{r}
# correlations between go/fit at each time point
#################################################
# function for correlations
cor_fit <- function(site, go_method, bio, sel, df) {
  cor_df <- data.frame(method = NA, biotic = NA, sel = NA, bag_site = NA, event = NA, pearson_surv = NA, pearson_surv_pval = NA, pearson_surv_rank = NA, pearson_surv_rank_pval = NA, spearman_surv = NA, spearman_surv_pval = NA, spearman_surv_rank = NA, spearman_surv_rank_pval = NA, kendall_surv = NA, kendall_surv_pval = NA, kendall_surv_rank = NA, kendall_surv_rank_pval = NA, r2_surv = NA, r2_surv_rank = NA, pearson_len = NA, pearson_len_pval = NA, pearson_len_rank = NA, pearson_len_rank_pval = NA, spearman_len = NA, spearman_len_pval = NA, spearman_len_rank = NA, spearman_len_rank_pval = NA, kendall_len = NA, kendall_len_pval = NA, kendall_len_rank = NA, kendall_len_rank_pval = NA, r2_len = NA, r2_len_rank = NA)
  
  # what metric should we use based on inputs
  lookup_metric <- list(
  Lewisetta = list(
    b = list(lea = "offset_lea_go_lew", gf_af = "offset_gf_af_lew",
                   gf_geno = "offset_gf_geno_lew", rda = "offset_rda_lew", 
                   env_dist = "env_dist_lew"),
    nb = list(lea = "offset_lea_go_lew_nb", gf_af = "offset_gf_af_lew_nb",
                   gf_geno = "offset_gf_geno_lew_nb", rda = "offset_rda_lew_nb", 
                   env_dist = "env_dist_lew_nb")),
  YorkRiver = list(
    b = list(lea = "offset_lea_go_yrk", gf_af = "offset_gf_af_yrk",
                   gf_geno = "offset_gf_geno_yrk", rda = "offset_rda_yrk", 
                   env_dist = "env_dist_yrk"),
    nb = list(lea = "offset_lea_go_yrk_nb", gf_af = "offset_gf_af_yrk_nb",
                   gf_geno = "offset_gf_geno_yrk_nb", rda = "offset_rda_yrk_nb", 
                   env_dist = "env_dist_yrk_nb")))
  
  lookup_rank <- list(
  Lewisetta = list(
    b = list(lea = "order_lew_lea", gf_af = "order_lew_af_gf",
                   gf_geno = "order_lew_geno_gf", rda = "order_lew_rda", 
                   env_dist = "order_lew_envdist"),
    nb = list(lea = "order_lew_lea_nb", gf_af = "order_lew_af_gf_nb",
                   gf_geno = "order_lew_geno_gf_nb", rda = "order_lew_rda_nb", 
                   env_dist = "order_lew_envdist_nb")),
  YorkRiver = list(
    b = list(lea = "order_yrk_lea", gf_af = "order_yrk_af_gf",
                   gf_geno = "order_yrk_geno_gf", rda = "order_yrk_rda", 
                   env_dist = "order_yrk_envdist"),
    nb = list(lea = "order_yrk_lea_nb", gf_af = "order_yrk_af_gf_nb",
                   gf_geno = "order_yrk_geno_gf_nb", rda = "order_yrk_rda_nb", 
                   env_dist = "order_yrk_envdist_nb")))
  
  go_metric <- lookup_metric[[site]][[bio]][[go_method]]
  go_rank <- lookup_rank[[site]][[bio]][[go_method]]

  for (i in 1:4) {
    if (sel == "sel") {
      fitdat <- df[df$bag_site == (site) & df$event == i,]
    } else {
      fitdat <- df[df$bag_site == (site) & df$event == i & df$site_name %in% c("W1_TX","W2_LA","W3_FL","W4_VA","W5_NH","W6_ME"),]
    }
    
    # Convert to numeric to handle any character/factor columns
    metric_values <- na.omit(as.numeric(fitdat[,go_metric]))
    rank_values <- na.omit(as.numeric(fitdat[,go_rank]))
    surv_values <- na.omit(as.numeric(fitdat$surv))
    surv_ranks <- na.omit(as.numeric(fitdat$order_mort))
    len_values <- na.omit(as.numeric(fitdat$length))
    len_ranks <- na.omit(as.numeric(fitdat$order_len))
    
    # metadata
    cor_df[i,]$method <- as.character(go_method)
    cor_df[i,]$bag_site <- as.character(site)
    cor_df[i,]$event <- i
    cor_df[i,]$biotic <- as.character(bio)
    cor_df[i,]$sel <- as.character(sel)    
    
    # survival
    cor_df[i,]$pearson_surv <- cor.test(metric_values, surv_values, method = "pearson")$estimate
    cor_df[i,]$pearson_surv_pval <- cor.test(metric_values, surv_values, method = "pearson")$p.value
    cor_df[i,]$pearson_surv_rank <- cor.test(rank_values, surv_ranks, method = "pearson")$estimate
    cor_df[i,]$pearson_surv_rank_pval <- cor.test(rank_values, surv_ranks, method = "pearson")$p.value
    cor_df[i,]$spearman_surv <- cor.test(metric_values, surv_values, method = "spearman")$estimate
    cor_df[i,]$spearman_surv_pval <- cor.test(metric_values, surv_values, method = "spearman")$p.value
    cor_df[i,]$spearman_surv_rank <- cor.test(rank_values, surv_ranks, method = "spearman")$estimate
    cor_df[i,]$spearman_surv_rank_pval <- cor.test(rank_values, surv_ranks, method = "spearman")$p.value
    cor_df[i,]$kendall_surv <- cor.test(metric_values, surv_values, method = "kendall")$estimate
    cor_df[i,]$kendall_surv_pval <- cor.test(metric_values, surv_values, method = "kendall")$p.value
    cor_df[i,]$kendall_surv_rank <- cor.test(rank_values, surv_ranks, method = "kendall")$estimate
    cor_df[i,]$kendall_surv_rank_pval <- cor.test(rank_values, surv_ranks, method = "kendall")$p.value
    cor_df[i,]$r2_surv <- summary(lm(surv_values ~ metric_values))$r.squared
    cor_df[i,]$r2_surv_rank <- summary(lm(surv_ranks ~ rank_values))$r.squared

    # length
    cor_df[i,]$pearson_len <- cor.test(metric_values, len_values, method = "pearson")$estimate
    cor_df[i,]$pearson_len_pval <- cor.test(metric_values, len_values, method = "pearson")$p.value
    cor_df[i,]$pearson_len_rank <- cor.test(rank_values, len_ranks, method = "pearson")$estimate
    cor_df[i,]$pearson_len_rank_pval <- cor.test(rank_values, len_ranks, method = "pearson")$p.value
    cor_df[i,]$spearman_len <- cor.test(metric_values, len_values, method = "spearman")$estimate
    cor_df[i,]$spearman_len_pval <- cor.test(metric_values, len_values, method = "spearman")$p.value
    cor_df[i,]$spearman_len_rank <- cor.test(rank_values, len_ranks, method = "spearman")$estimate
    cor_df[i,]$spearman_len_rank_pval <- cor.test(rank_values, len_ranks, method = "spearman")$p.value
    cor_df[i,]$kendall_len <- cor.test(metric_values, len_values, method = "kendall")$estimate
    cor_df[i,]$kendall_len_pval <- cor.test(metric_values, len_values, method = "kendall")$p.value
    cor_df[i,]$kendall_len_rank <- cor.test(rank_values, len_ranks, method = "kendall")$estimate
    cor_df[i,]$kendall_len_rank_pval <- cor.test(rank_values, len_ranks, method = "kendall")$p.value
    cor_df[i,]$r2_len <- summary(lm(len_values ~ metric_values))$r.squared
    cor_df[i,]$r2_len_rank <- summary(lm(len_ranks ~ rank_values))$r.squared
  }
  return(cor_df)
}

# correlations
(cor_lea_lew <- cor_fit("Lewisetta","lea","b", "sel", fullfit))
(cor_lea_yrk <- cor_fit("YorkRiver","lea","b", "sel", fullfit))
(cor_gf_af_lew <- cor_fit("Lewisetta","gf_af","b", "sel", fullfit))
(cor_gf_af_yrk <- cor_fit("YorkRiver","gf_af","b", "sel", fullfit))
(cor_gf_geno_lew <- cor_fit("Lewisetta","gf_geno","b", "sel", fullfit))
(cor_gf_geno_yrk <- cor_fit("YorkRiver","gf_geno","b", "sel", fullfit))
(cor_rda_lew <- cor_fit("Lewisetta","rda","b", "sel", fullfit))
(cor_rda_yrk <- cor_fit("YorkRiver","rda","b", "sel", fullfit))
(cor_envdist_lew <- cor_fit("Lewisetta","env_dist", "b", "sel", fullfit))
(cor_envdist_yrk <- cor_fit("YorkRiver","env_dist", "b", "sel", fullfit))

(cor_lea_lew_nb <- cor_fit("Lewisetta","lea","nb", "sel", fullfit))
(cor_lea_yrk_nb <- cor_fit("YorkRiver","lea","nb", "sel", fullfit))
(cor_gf_af_lew_nb <- cor_fit("Lewisetta","gf_af","nb", "sel", fullfit))
(cor_gf_af_yrk_nb <- cor_fit("YorkRiver","gf_af","nb", "sel", fullfit))
(cor_gf_geno_lew_nb <- cor_fit("Lewisetta","gf_geno","nb", "sel", fullfit))
(cor_gf_geno_yrk_nb <- cor_fit("YorkRiver","gf_geno","nb", "sel", fullfit))
(cor_rda_lew_nb <- cor_fit("Lewisetta","rda","nb", "sel", fullfit))
(cor_rda_yrk_nb <- cor_fit("YorkRiver","rda","nb", "sel", fullfit))
(cor_envdist_lew_nb <- cor_fit("Lewisetta","env_dist", "nb", "sel", fullfit))
(cor_envdist_yrk_nb <- cor_fit("YorkRiver","env_dist", "nb", "sel", fullfit))

corrs_df <- rbind(cor_lea_lew, cor_lea_yrk, cor_gf_af_lew, cor_gf_af_yrk, cor_gf_geno_lew, cor_gf_geno_yrk, cor_rda_lew, cor_rda_yrk, cor_envdist_lew, cor_envdist_yrk, cor_lea_lew_nb, cor_lea_yrk_nb, cor_gf_af_lew_nb, cor_gf_af_yrk_nb, cor_gf_geno_lew_nb, cor_gf_geno_yrk_nb, cor_rda_lew_nb, cor_rda_yrk_nb, cor_envdist_lew_nb, cor_envdist_yrk_nb)
rownames(corrs_df) <- NULL
corrs_df$signif_surv <- ifelse(corrs_df$kendall_surv_pval < .05, T, F)
corrs_df$signif_len <- ifelse(corrs_df$kendall_len_pval < .05, T, F)

# save this
write.csv(corrs_df, paste0("results/GO_results/go_fit_corrs_",Sys.Date(),".csv"), row.names = F)

# another version without pearson or spearman corrs
corrs_df_kend <- corrs_df %>% dplyr::select(-contains(c("pearson", "spearman")))

# and for only final time point
corrs_df_kend_t4 <- corrs_df_kend[corrs_df_kend$event == 4,]

write.csv(corrs_df_kend_t4, paste0("results/GO_results/go_fit_corrs_final_",Sys.Date(),".csv"), row.names = F)
#################################################
```


### Plot GO vs mortality against each other

I'll start by plotting LEA results:

```{r}
# plot lea vs mort at each site/time point
goFitLEA_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lea_go_lew, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lea_go_lew, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lea_go_lew, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lea_go_lew, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lea_go_lew, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lea_go_lew, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Survival at Lewisetta", x = "LEA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_lew_plotbag

goFitLEA_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_lea_go_yrk, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_lea_go_yrk, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_lea_go_yrk, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_lea_go_yrk, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_lea_go_yrk, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_lea_go_yrk, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Survival at York River", x = "LEA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_plot

# plot go lea vs mort at both sites just final time point
goFitLEA_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = surv), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = surv), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (LEA) vs Survival at Final Time Point", x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotfinalTsurv
```

Next, I'll plot gradient forest results, starting with the allele frequency model.

```{r}
# plot gf af vs mort at each site/time point
goFitGFaf_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (GFaf) vs Survival at Lewisetta", x = "GF Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFaf_lew_plotbag

goFitGFaf_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (GFaf) vs Survival at York River", x = "GF Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFaf_yrk_plot

# plot gf vs mort at both sites just final time point
goFitGFaf_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = surv), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = surv), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (GFaf) vs Survival at Final Time Point", x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGFaf_plotfinalTsurv
```

Then the genotype model results:

```{r}
# plot gf geno vs mort at each site/time point
goFitGFgeno_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (GFgeno) vs Survival at Lewisetta", x = "GF Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFgeno_lew_plotbag

goFitGFgeno_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (GFgeno) vs Survival at York River", x = "GF Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFgeno_yrk_plot

# plot gf vs mort at both sites just final time point
goFitGFgeno_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = surv), col = "mediumspringgreen", linetype = "dashed", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = surv), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (GFgeno) vs Survival at Final Time Point", x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGFgeno_plotfinalTsurv
```


Then I'll plot rda results:

```{r}
# plot rda vs mort at each site/time point
goFitRDA_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_rda_lew, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_rda_lew, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_rda_lew, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_rda_lew, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_rda_lew, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_rda_lew, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Survival at Lewisetta", x = "RDA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_lew_plotbag

goFitRDA_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_rda_yrk, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_rda_yrk, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_rda_yrk, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_rda_yrk, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_rda_yrk, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_rda_yrk, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = surv), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Survival at York River", x = "RDA Offset", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_yrk_plot

# plot rda vs mort at both sites just final time point
goFitRDA_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = surv), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = surv), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (RDA) vs Survival at Final Time Point", x = "RDA Offset", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitRDA_plotfinalTsurv
```

And finally, environmental distance:

```{r}
# plot env dist vs mort at each site/time point
goFitEnvdist_lew_plotbag <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = surv), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Survival at Lewisetta", x = "Environmental Distance", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvdist_lew_plotbag

goFitEnvdist_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = surv, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = surv), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = surv, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = surv), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = surv, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = surv), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = surv, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = surv), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Survival at York River", x = "Environmental Distance", y = "Survival") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvdist_yrk_plot

# plot env dist vs mort at both sites just final time point
goFitEnvdist_plotfinalTsurv <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = surv, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = surv), col = "mediumspringgreen", linetype = "dashed", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = surv, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = surv), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Environmental Distance vs Survival at Final Time Point", x = "Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitEnvdist_plotfinalTsurv
```


### Plot GO vs length against each other

LEA vs length:

```{r}
# plot lea vs len at each site/time point
goFitLEA_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lea_go_lew, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_lea_go_lew, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lea_go_lew, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_lea_go_lew, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lea_go_lew, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_lea_go_lew, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Size at Lewisetta", x = "LEA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_lew_len

goFitLEA_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_lea_go_yrk, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_lea_go_yrk, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_lea_go_yrk, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_lea_go_yrk, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_lea_go_yrk, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_lea_go_yrk, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (LEA) vs Size at York River", x = "LEA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitLEA_yrk_len

# plot go lea vs len at both sites just final time point
goFitLEA_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_lea_go_lew, y = length), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_lea_go_yrk, y = length), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (LEA) vs Size at Final Time Point", x = "LEA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitLEA_plotfinalTlen
```


GF AF GO vs length: 

```{r}
# plot gf vs len at each site/time point
goFitGFaf_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_af_lew, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_af_lew, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_af_lew, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (GF) vs Size at Lewisetta", x = "GF Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFaf_lew_len

goFitGFaf_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_af_yrk, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_af_yrk, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_af_yrk, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (GF) vs Size at York River", x = "GF Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFaf_yrk_len

# plot gf vs len at both sites just final time point
goFitGFaf_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_af_lew, y = length), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_af_yrk, y = length), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (GF) vs Size at Final Time Point", x = "GF Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGFaf_plotfinalTlen
```


GF geno GO vs length: 

```{r}
# plot gf vs len at each site/time point
goFitGFgeno_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_gf_geno_lew, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_gf_geno_lew, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_gf_geno_lew, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (GF) vs Size at Lewisetta", x = "GF Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFgeno_lew_len

goFitGFgeno_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_gf_geno_yrk, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_gf_geno_yrk, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_gf_geno_yrk, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (GF) vs Size at York River", x = "GF Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitGFgeno_yrk_len

# plot gf vs len at both sites just final time point
goFitGFgeno_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_gf_geno_lew, y = length), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_gf_geno_yrk, y = length), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (GF) vs Size at Final Time Point", x = "GF Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitGFgeno_plotfinalTlen
```


RDA GO vs length:

```{r}
# plot rda vs len at each site/time point
goFitRDA_lew_len <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_rda_lew, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = offset_rda_lew, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_rda_lew, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = offset_rda_lew, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_rda_lew, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = offset_rda_lew, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Size at Lewisetta", x = "RDA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_lew_len

goFitRDA_yrk_len <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_rda_yrk, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = offset_rda_yrk, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_rda_yrk, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = offset_rda_yrk, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_rda_yrk, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = offset_rda_yrk, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length), col = "purple4", se = T) +
  labs(title = "Predicted GO (RDA) vs Size at York River", x = "RDA Offset", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitRDA_yrk_len

# plot rda vs len at both sites just final time point
goFitRDA_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = offset_rda_lew, y = length), col = "mediumspringgreen", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = offset_rda_yrk, y = length), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Predicted GO (RDA) vs Size at Final Time Point", x = "RDA Offset", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitRDA_plotfinalTlen
```

Environmental distance vs length:

```{r}
# plot env dist vs length at each site/time point
goFitEnvdist_len_lew_plot <- ggplot(data = fullfit[fullfit$bag_site == "Lewisetta",]) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 1,], aes(x = env_dist_lew, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 2,], aes(x = env_dist_lew, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 3,], aes(x = env_dist_lew, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Size at Lewisetta", x = "Environmental Distance", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvdist_len_lew_plot

goFitEnvdist_len_yrk_plot <-  ggplot(data = fullfit[fullfit$bag_site == "YorkRiver",]) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = length, col = "Fall'23")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 1,], aes(x = env_dist_yrk, y = length), col = "cadetblue3", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = length, col = "Spring'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 2,], aes(x = env_dist_yrk, y = length), col = "cornflowerblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = length, col = "Fall'24")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 3,], aes(x = env_dist_yrk, y = length), col = "slateblue", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length, col = "Spring'25")) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length), col = "purple4", se = T) +
  labs(title = "Environmental Distance vs Size at York River", x = "Environmental Distance", y = "Length (mm)") +
  scale_colour_manual(name = "Monitoring Events", breaks = c("Fall'23", "Spring'24", "Fall'24","Spring'25"), values = c("Fall'23"="cadetblue3","Spring'24"="cornflowerblue","Fall'24"="slateblue","Spring'25"="purple4"))+
  theme_classic()
goFitEnvdist_len_yrk_plot

# plot go gf vs len at both sites just final time point
goFitEnvdist_plotfinalTlen <- ggplot(data = fullfit) +
  geom_point(data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length, fill = "Lewisetta"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,], aes(x = env_dist_lew, y = length), col = "mediumspringgreen", linetype = "dashed", se = T) +
  geom_point(data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length, fill = "YorkRiver"), shape = 21, size = 4) +
  geom_smooth(method = "lm", data = fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,], aes(x = env_dist_yrk, y = length), col = "dodgerblue", linetype = "dashed", se = T) +
  labs(title = "Environmental Distance vs Size at Final Time Point", x = "Environmental Distance", y = "Length (mm)") +
  scale_fill_manual(name = "Experimental Site", breaks = c("Lewisetta", "YorkRiver"), values = c("Lewisetta"="mediumspringgreen", "YorkRiver"="dodgerblue"))+
  theme_classic()
goFitEnvdist_plotfinalTlen
```


Export all of these plots to pdf:

```{r}
# this version lets me export directly as an editable powerpoint with vector graphics
library(officer) # for powerpoint export/editing
library(rvg) # for converting r graphics to editable vector graphics

doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = goFitLEA_plotfinalTsurv), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = goFitGFaf_plotfinalTsurv), location = ph_location_type(type = "body")) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = goFitRDA_plotfinalTsurv), location = ph_location_type(type = "body")) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = goFitEnvdist_plotfinalTsurv), location = ph_location_type(type = "body")) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = goFitLEA_plotfinalTlen), location = ph_location_type(type = "body")) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = goFitGFaf_plotfinalTlen), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = goFitRDA_plotfinalTlen), location = ph_location_type(type = "body")) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = goFitEnvdist_plotfinalTlen), location = ph_location_type(type = "body"))
print(doc, target = "results/figures/GO_compare.pptx")
```

## Compare Offset Methods

Now that I have predictions from different methods, I will look at how offset predictions made by these methods compare to each other and how different their ability is to predict experimental fitness.


### Compare Offsets to Each Other

```{r Offset Comparison}
# create a correlation matrix for offsets
go_forcorrs <- fullfit %>% distinct(site_name, bag_site, event, .keep_all = TRUE) %>% filter(event == 4) %>% dplyr::select(site_name, pop, bag_site, offset_rda_lew, offset_rda_yrk, order_lew_rda, order_yrk_rda, offset_gf_af_lew, offset_gf_af_yrk, order_lew_af_gf, order_yrk_af_gf, offset_gf_geno_lew, offset_gf_geno_yrk, order_lew_geno_gf, order_yrk_geno_gf, offset_lea_go_lew, offset_lea_go_yrk, order_lew_lea, order_yrk_lea, env_dist_lew, env_dist_yrk, order_lew_envdist, order_yrk_envdist)

# check correlations
library(ggcorrplot)
cor_go_raw <- cor(go_forcorrs[,c("offset_rda_lew", "offset_gf_af_lew", "offset_gf_geno_lew", "offset_lea_go_lew", "env_dist_lew", "offset_rda_yrk", "offset_gf_af_yrk", "offset_gf_geno_yrk", "offset_lea_go_yrk", "env_dist_yrk")])
cor_go_raw_pmat <- cor_pmat(go_forcorrs[,c("offset_rda_lew", "offset_gf_af_lew", "offset_gf_geno_lew", "offset_lea_go_lew", "env_dist_lew", "offset_rda_yrk", "offset_gf_af_yrk", "offset_gf_geno_yrk", "offset_lea_go_yrk", "env_dist_yrk")])
cor_go_ord <- cor(go_forcorrs[,c("order_lew_rda", "order_lew_af_gf", "order_lew_geno_gf", "order_lew_lea", "order_lew_envdist", "order_yrk_rda", "order_yrk_af_gf", "order_yrk_geno_gf", "order_yrk_lea", "order_yrk_envdist")])
cor_go_ord_pmat <- cor_pmat(go_forcorrs[,c("order_lew_rda", "order_lew_lea", "order_lew_af_gf", "order_lew_geno_gf", "order_lew_envdist", "order_yrk_rda", "order_yrk_lea", "order_yrk_af_gf", "order_yrk_geno_gf", "order_yrk_envdist")])

# plot these correlations
# raw go
corrs_go_raw_all <- ggcorrplot(cor_go_raw, 
                               p.mat = cor_go_raw_pmat,
                               insig = "blank",
                               lab = T,
                               colors = c("#6D9EC1", "white", "#E46726")) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Genomic Offset Correlations",
       x = NULL,
       y = NULL)
corrs_go_raw_all

# go orders
corrs_go_ord_all <- ggcorrplot(cor_go_ord, 
                               p.mat = cor_go_ord_pmat,
                               insig = "blank",
                               lab = T,
                               colors = c("#6D9EC1", "white", "#E46726")) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Genomic Offset Correlations (Rank)",
       x = NULL,
       y = NULL)
corrs_go_ord_all
```

Look also at the difference between go models with and without biotic variables.

```{r}
# create a correlation matrix for offsets
go_forcorrs_bio <- fullfit %>% distinct(site_name, bag_site, event, .keep_all = TRUE) %>% filter(event == 4) %>% dplyr::select(site_name, pop, bag_site, offset_rda_lew, offset_rda_yrk, order_lew_rda, order_yrk_rda, offset_gf_af_lew, offset_gf_af_yrk, order_lew_af_gf, order_yrk_af_gf, offset_gf_geno_lew, offset_gf_geno_yrk, order_lew_geno_gf, order_yrk_geno_gf, offset_lea_go_lew, offset_lea_go_yrk, order_lew_lea, order_yrk_lea, env_dist_lew, env_dist_yrk, order_lew_envdist, order_yrk_envdist, offset_rda_lew_nb, offset_rda_yrk_nb, order_lew_rda_nb, order_yrk_rda_nb, offset_gf_af_lew_nb, offset_gf_af_yrk_nb, order_lew_af_gf_nb, order_yrk_af_gf_nb, offset_gf_geno_lew_nb, offset_gf_geno_yrk_nb, order_lew_geno_gf_nb, order_yrk_geno_gf_nb, offset_lea_go_lew_nb, offset_lea_go_yrk_nb, order_lew_lea_nb, order_yrk_lea_nb, env_dist_lew_nb, env_dist_yrk_nb, order_lew_envdist_nb, order_yrk_envdist_nb)
  
# check correlations
library(ggcorrplot)
cor_go_raw_bio <- cor(go_forcorrs_bio[,c("offset_rda_lew", "offset_gf_af_lew", "offset_gf_geno_lew", "offset_lea_go_lew", "env_dist_lew", "offset_rda_yrk", "offset_gf_af_yrk", "offset_gf_geno_yrk", "offset_lea_go_yrk", "env_dist_yrk", "offset_rda_lew_nb", "offset_gf_af_lew_nb", "offset_gf_geno_lew_nb", "offset_lea_go_lew_nb", "env_dist_lew_nb", "offset_rda_yrk_nb", "offset_gf_af_yrk_nb", "offset_gf_geno_yrk_nb", "offset_lea_go_yrk_nb", "env_dist_yrk_nb")])
cor_go_raw_pmat_bio <- cor_pmat(go_forcorrs_bio[,c("offset_rda_lew", "offset_gf_af_lew", "offset_gf_geno_lew", "offset_lea_go_lew", "env_dist_lew", "offset_rda_yrk", "offset_gf_af_yrk", "offset_gf_geno_yrk", "offset_lea_go_yrk", "env_dist_yrk", "offset_rda_lew_nb", "offset_gf_af_lew_nb", "offset_gf_geno_lew_nb", "offset_lea_go_lew_nb", "env_dist_lew_nb", "offset_rda_yrk_nb", "offset_gf_af_yrk_nb", "offset_gf_geno_yrk_nb", "offset_lea_go_yrk_nb", "env_dist_yrk_nb")])
cor_go_ord_bio <- cor(go_forcorrs_bio[,c("order_lew_rda", "order_lew_af_gf", "order_lew_geno_gf", "order_lew_lea", "order_lew_envdist", "order_yrk_rda", "order_yrk_af_gf", "order_yrk_geno_gf", "order_yrk_lea", "order_yrk_envdist", "order_lew_rda_nb", "order_lew_af_gf_nb", "order_lew_geno_gf_nb", "order_lew_lea_nb", "order_lew_envdist_nb", "order_yrk_rda_nb", "order_yrk_af_gf_nb", "order_yrk_geno_gf_nb", "order_yrk_lea_nb", "order_yrk_envdist_nb")])
cor_go_ord_pmat_bio <- cor_pmat(go_forcorrs_bio[,c("order_lew_rda", "order_lew_af_gf", "order_lew_geno_gf", "order_lew_lea", "order_lew_envdist", "order_yrk_rda", "order_yrk_af_gf", "order_yrk_geno_gf", "order_yrk_lea", "order_yrk_envdist", "order_lew_rda_nb", "order_lew_af_gf_nb", "order_lew_geno_gf_nb", "order_lew_lea_nb", "order_lew_envdist_nb", "order_yrk_rda_nb", "order_yrk_af_gf_nb", "order_yrk_geno_gf_nb", "order_yrk_lea_nb", "order_yrk_envdist_nb")])

# plot these correlations
# raw go
corrs_go_raw_all_bio <- ggcorrplot(cor_go_raw_bio, 
                               p.mat = cor_go_raw_pmat_bio,
                               insig = "blank",
                               lab = T,
                               colors = c("#6D9EC1", "white", "#E46726")) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Genomic Offset Correlations",
       x = NULL,
       y = NULL)
corrs_go_raw_all_bio

# go orders
corrs_go_ord_all_bio <- ggcorrplot(cor_go_ord_bio, 
                               p.mat = cor_go_ord_pmat_bio,
                               insig = "blank",
                               lab = T,
                               colors = c("#6D9EC1", "white", "#E46726")) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Genomic Offset Correlations (Rank)",
       x = NULL,
       y = NULL)
corrs_go_ord_all_bio

# how different were the offset predictions?
lew_final <- fullfit[fullfit$bag_site == "Lewisetta" & fullfit$event == 4,]
yrk_final <- fullfit[fullfit$bag_site == "YorkRiver" & fullfit$event == 4,]

diff_lea_lew <- mean((lew_final$offset_lea_go_lew_nb - lew_final$offset_lea_go_lew) / lew_final$offset_lea_go_lew)
diff_lea_yrk <- mean((yrk_final$offset_lea_go_yrk_nb - yrk_final$offset_lea_go_yrk) / yrk_final$offset_lea_go_yrk)

diff_gf_lew <- mean((lew_final$offset_gf_geno_lew_nb - lew_final$offset_gf_geno_lew) / lew_final$offset_gf_geno_lew)
diff_gf_yrk <- mean((yrk_final$offset_gf_geno_yrk_nb - yrk_final$offset_gf_geno_yrk) / yrk_final$offset_gf_geno_yrk)

diff_rda_lew <- mean((lew_final$offset_rda_lew_nb - lew_final$offset_rda_lew) / lew_final$offset_rda_lew)
diff_rda_yrk <-  mean((yrk_final$offset_rda_yrk_nb - yrk_final$offset_rda_yrk) / yrk_final$offset_rda_yrk)

diff_envdist_lew <- mean((lew_final$env_dist_lew_nb - lew_final$env_dist_lew) / lew_final$env_dist_lew)
diff_envdist_yrk <- mean((yrk_final$env_dist_yrk_nb - yrk_final$env_dist_yrk) / yrk_final$env_dist_yrk)

offset_diffs_bio <- data.frame(method = c("lea","lea","gf","gf","rda","rda","envdist","envdist"), bag_site = rep(c("Lewisetta","YorkRiver"),4), diff_percent = c(diff_lea_lew, diff_lea_yrk, diff_gf_lew, diff_gf_yrk, diff_rda_lew, diff_rda_yrk, diff_envdist_lew, diff_envdist_yrk))

# how different were the offset performances? 
final_corrs_b <- corrs_df[corrs_df$event == 4 & corrs_df$biotic == "b", c("method","bag_site","kendall_surv","signif_surv","kendall_len","signif_len")]
final_corrs_nb <- corrs_df[corrs_df$event == 4 & corrs_df$biotic == "nb", c("method","bag_site","kendall_surv","signif_surv","kendall_len","signif_len")]
colnames(final_corrs_nb) <- c("method","bag_site","kendall_surv_nb","signif_surv_nb","kendall_len_nb","signif_len_nb")
final_corrs <- merge(final_corrs_b, final_corrs_nb, by = c("method", "bag_site"))

final_corrs$diff_surv <- final_corrs$kendall_surv - final_corrs$kendall_surv_nb
final_corrs$diff_len <- final_corrs$kendall_len - final_corrs$kendall_len_nb

write.csv(final_corrs, "results/GO_results/final_correlation_GOfit.csv", row.names = F)

# plot performance differences
disease_vs_no_performance <- ggplot(data = final_corrs) +
  geom_point(data = final_corrs, aes(x = kendall_surv_nb, y = kendall_surv, fill = method), shape = 21, size = 4) +
  ylim(-1,0.05)+
  xlim(-1,0.05)+
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Genomic Offset Performance with and without \nthe Biotic Landscape", x = "Genomic Offset (Environment Only)", y = "Genomic Offset (Disease & Environment)") +
  theme_classic()
disease_vs_no_performance

disease_vs_no_len <- ggplot(data = final_corrs) +
  geom_point(data = final_corrs, aes(x = kendall_len_nb, y = kendall_len, fill = method), shape = 21, size = 4) +
  ylim(-1,0.05)+
  xlim(-1,0.05)+
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Genomic Offset Performance (Length) with and without \nthe Biotic Landscape", x = "Genomic Offset (Environment Only)", y = "Genomic Offset (Disease & Environment)") +
  theme_classic()
disease_vs_no_len
```


