---
title: "CviMVP_oyster_GenDatProcess"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Genetic Data Processing

This script is to process individual and (genetic for seascape) level data for the seascape and experiment by identifying the number and identities of individuals in each dataset and preparing this information and the genomic data into formats needed for downstream analysis.

The seascape data comes from...

The experimental data comes from...


FOR NOW I AM DOING THIS JUST WITH EXPERIMENTAL INDIVIDUALS GENOTYPED, BUT WE WILL NEED TO THINK ABOUT HOW TO DO THIS WHEN WE HAVE NO TRUE EXPERIMENTAL FITNESS DATA ON AN INDIVIDUAL LEVEL (IG WE DO FOR TAGGED INDS LENGTHS? BUT NOT MORT)


## Setup

```{r}
# set wd
setwd("~/Documents/Github/MVP_oyster_GO")

# packages
library(dplyr)
library(ggplot2)
library(tidyverse)
library(LEA)
library(stringr)

# data
geno_mat <- read.geno("data/GenDat/geno_file_exp_all.geno")
```


## Pre-processing

Process data to remove missing data, non-variable genomic sites.

```{r}
# check data
dim(geno_mat) # 154 inds, 32487 snps
geno_mat[1:10,1:10]

# any missing data?
length(which(geno_mat == 9)) # nope!

# non-variable genomic sites? do later if necessary, this isn't the final iteration of this script anyways
```


## Individual Data

```{r}
geno_mat
```


## Genomic Data Formatting

```{r}

```

