---
title: "CviMVP_oyster_GF"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# GF Genomic Offsets

This script is a gradientForest pipeline for calculating genomic offset. I am using gradientForest, which build on randomForest to use predictions of genome-wide allele frequency variation along multivariate environmental gradients to quantify the disruption of genotype-environment associations under environmental change. I first train the models using gradientForest on a large seascape dataset, then predict offset for a subset of 8 populations from this dataset and 2 selection lines used in a common garden experiment based in the Chesapeake Bay. Offset is calculated from the environment of origin to common garden conditions.

## Setup

```{r Setup}
# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(LEA)
library(dplyr) # for data cleaning
library(plyr) # for data cleaning
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(raster) # for environmental pixels
library(ggplot2) # for plotting
library(parallel) # for parallelizing processes
library(doParallel) # for parallelizing processes 
library(gradientForest) # for running gradient forest
library(sf) # plotting
library(rnaturalearth) # plotting

# data
# env
env_all <- as.data.frame(read.csv("data/EnvDat/env_scaled_2025-10-30.csv")[,-1])
env_sea <- as.data.frame(read.csv("data/EnvDat/seascape/sea_scaled_expanded_reord_2025-10-30.csv")[,-1])
env_exp <- as.data.frame(read.csv("data/EnvDat/exp/exp_scaled_expanded_2025-10-30.csv")[,-1])

# sites
env_sea_site <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv")[,-1]
env_sea_date <- read.csv("data/EnvDat/seascape/SeascapeSamples - date.csv")[,-1]
env_exp_site <- read.csv("data/EnvDat/exp/exp_sitenames.csv")[,-1]
exp_site_latlon <- read.csv("data/EnvDat/exp/exp_site_info.csv")[,-1]

# gen
genoMat <- read.geno("data/GenDat/geno_seascape_2025-08-14.geno")
genoThinMat <- read.geno("data/GenDat/geno_seascape_LDthin_2025-08-14.geno")

# individuals
expInds <- readRDS("data/IndDat/20240922_experimental_indsmatrix.rds")
seascapeInds <- readRDS("data/IndDat/20250604_seascape_indsmatrix.rds")
inds_genomat <- read.csv("data/GenDat/inds_genomat.csv")[,-1]
```


## Data Prep

Inspect the environmental and genomic data to ensure it's ready for gradientForest. Note that environmental data was pre-processed in the script "CviMVP_oyster_EnvDatPRocess.Rmd" and genomic data was pre-processed in the script "CviMVP_oyster_GenDatProcess.Rmd".

### Individual Data Prep

```{r}
# order this data by genotype info, first by trimming inds to only those in df
seascapeInds_reord <- seascapeInds[seascapeInds$clean_ID %in% inds_genomat,]
seascapeInds_reord$clean_ID <- factor(seascapeInds_reord$clean_ID, levels = inds_genomat)
seascapeInds_reord2 <- seascapeInds_reord[order(seascapeInds_reord$clean_ID),]

# double check
(seascapeInds_reord2$clean_ID == inds_genomat) # looks good!

# get a full list of pops
pops_list_full <- seascapeInds_reord2$ID_SiteDate
```

### Genomic Data Prep

```{r Genomic Data Prep}
# check geno data
dim(genoMat)
dim(genoThinMat)

# rotate df so rows are now columns (we want individuals as columns, alleles as rows)
genoMatT <- as.matrix(t(genoMat))
genoThinMatT <- as.matrix(t(genoThinMat))

# any missing data?
sum((genoMatT == 9)) # 0 - good 
sum((genoThinMatT == 9)) # 0 - good

# add pop data to genomic data
rownames(genoMat) <- rownames(genoThinMat) <- pops_list_full
colnames(genoMatT) <- colnames(genoThinMatT) <- pops_list_full

# function to calculate allele frequency @ single locus
calc_freq <- function(x){
  a <- sum(x, na.rm=TRUE)
  b <- (2*length(na.omit(x)))
  a/b
}

# function extending this to calculate allele frequencies across whole pop
calcfreq <- function(a, pop){
  tapply(a, pop, calc_freq)
}

# allele freqs
freqs <- apply((genoMat), 2, calcfreq, pops_list_full)
str(freqs)

freqs_thin <- apply((genoThinMat), 2, calcfreq, pops_list_full)
str(freqs_thin)

# update colnames for gf
colnames(freqs) <- make.names(as.character(1:ncol(genoMat)))
colnames(freqs_thin) <- make.names(as.character(1:ncol(genoThinMat)))

# do the same for the full geno matrix, just in case
colnames(genoMat) <- make.names(as.character(1:ncol(genoMat)))
colnames(genoThinMat) <- make.names(as.character(1:ncol(genoThinMat)))
```


### Environmental Data Prep

```{r Environmental Data Prep}
# check env data
dim(env_sea)
dim(env_exp)
dim(seascapeInds_reord2)

# check seascape env in the same order as the genotype matrix
env_sea$clean_ID == seascapeInds_reord2$clean_ID # in same order

# we also need one just per site
env_sea_pop <- env_sea[!duplicated(env_sea$ID_SiteDate),]

# now experimental data
# no need to order this by any geno data
# should get pop level info 
env_exp_pop <- env_exp[!duplicated(env_exp$site_name),]

# reduced variable set
env_sea_red <- env_sea[,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_Crab_scaled")]
env_sea_pop_red <- env_sea_pop[,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_Crab_scaled")]
env_exp_pop_red <- env_exp_pop[,c("site_name","salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_Crab_scaled")]
env_exp_lew_red <- env_exp_pop[,c("site_name", "lew_salinity_quantile_10_scaled", "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_Crab_scaled")]
env_exp_yrk_red <- env_exp_pop[,c("site_name", "yrk_salinity_quantile_10_scaled", "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_Crab_scaled")]

# change exp colnames to match up
colnames(env_exp_lew_red) <- colnames(env_exp_yrk_red) <- colnames(env_exp_pop_red)
```


## Train Models

To train the models, I will be using the seascape genomic dataset that we processed in the previous chunk, as well as the current environmental data files produced using "CviMVP_oyster_EnvDatProcess.Rmd."

```{r GO Model Training}
# start by defining a maximum number of splits
maxLevel <- log2(0.368*nrow(env_sea_pop_red)/2)

# run gradient forest
# allele freq models
start_time <- Sys.time() # time start
gf_af <- gradientForest(cbind(env_sea_pop_red, freqs_thin), predictor.vars = colnames(env_sea_pop_red), response.vars = (colnames(freqs_thin)), ntree = 500, maxLevel = maxLevel, trace = T, corr.threshold=0.50) # tons of warnings here, "response has five or fewer unique values. Are you sure you want to do regression?" - warning doesn't appear on the cluster
end_time <- Sys.time() # time end
(end_time - start_time) # about 1.5hrs

## COME BACK TO THIS ON CLUSTER, DOESN'T RUN ON PERSONAL COMPUTER
# geno models
# crashes on personal laptop, run on cluster
start_time <- Sys.time() # time start
gf_geno <- gradientForest(cbind(env_sea[,c("salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_Crab_scaled")], genoThinMat), predictor.vars = c("salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_Crab_scaled"), response.vars=colnames(genoThinMat), ntree=500, maxLevel=maxLevel, trace=T, corr.threshold=0.50)
end_time <- Sys.time() # time end
(end_time - start_time) # about 2hrs on cluster

# plot bar graphs depicting importance of each spatial and climate variable
plot(gf_af)
#plot(gf_geno)
saveRDS(gf_af, paste0("results/lg_results/af_geno_",Sys.Date(),".RDS"))
#saveRDS(gf_geno, paste0("results/GO_results/gf_geno_",Sys.Date(),".RDS"))

# plots of turnover functions for individual loci - now having issues with ylim
plot(gf_af, plot.type = "C", imp.vars = gf_af$imp.vars, show.overall = T, legend = T, leg.posn = "topleft", leg.nspecies = 5, ylim = c(0,0.5)) # ylim error, if i provide none I get "need finite ylim" if I provide one I get "formal argument "ylim" matched by multiple actual arguments"
#plot(gf_geno, plot.type = "C", imp.vars = gf_geno$imp.vars, show.overall = T, legend = T, leg.posn = "topleft", leg.nspecies = 5, ylim = c(0,0.5)) # doesn't finish running, had to interrupt
```

It looks like for the allele frequency based model, sal_quant90 is the most important climate variable, followed by either dermo prevalence/temp_quant90 (accuracy importance/R2 weighted importance), then dermo prevalence/sal_quantile10, then temp_quantil10, and finally pea crabs. 

For the genotype based model, temp_quant10 is the most important variable, then sal_quant10, then temp_quant90, and finally sal_quant90 (consistent across accuracy importance and R2 weighted importance) FIX THIS ONCE RUN ON CLUSTER

For both training runs, it's important to note that I had a significant number of warning messages - specifically:

"Warning in randomForest.default(x = X, y = spec_vec, maxLevel = maxLevel, : The response has five or fewer unique values. Are you sure you want to do regression?"

I think this is coming from non-variable genomic sites, which I could trim in future iterations of this model. Brandon has encountered this issue before and seems to think that the warning is internal to extendedForests (pckg for implementing random forests), and refers to number of unique AFs for a given locus. Something I've noticed while rerunning this code on the cluster is that these warnings do NOT appear, so it may alternatively be an issue with computing power.


### Predict Offset

Now that we have trained our gradient forest models, we need to use these to predict performance in the mesocosm treatments. This requires no new genomic data, just new environmental data that corresponds to each of the treatments and to the sites these populations originated from. Importantly, this new environmental data set must use the same column names as the data set used for training. We will ensure data compatibility below.

```{r Predict Offset Prep}
# check everything again
gf_af
gf_geno

# use exp data to predict from/to, seascape just to train the gf
env_exp_pop_red
env_exp_lew_red
env_exp_yrk_red

# pops list experimental groups
pops_list1 <- c("S1_LOLA", "S2_DEBY", "W1_TX", "W2_LA", "W3_FL", "W4_VA", "W5_NH", "W6_ME")
```


Now I can predict offset using these different methods (genotype and allele frequency based).

```{r Predict Offset}
# prediction to home site
pred_af_home <- predict(gf_af, env_exp_pop_red[,-1])
#pred_geno_home <- predict(gf_geno, expEnv_red[,2:5])

# prediction to common gardens
pred_af_lew <- predict(gf_af, env_exp_lew_red[,-1])
pred_af_yrk <- predict(gf_af, env_exp_yrk_red[,-1])

#pred_geno_lew <- predict(gf_geno, geno_lewenv_comp[,3:6])
#pred_geno_yrk <- predict(gf_geno, geno_yrkenv_comp[,3:6])

# We now compute the euclidean distance between current and new cumulative importance
nb_ind_af <- nrow(pred_af_lew)
#nb_ind_geno <- nrow(pred_geno_lew)

# calculate offsets
euclidean <- function(a, b) sqrt(sum((a - b)^2)) # Euclidean distance for genetic offset

# af
# lew
df_af_lew <- data.frame() # df
for (i in 1:nb_ind_af){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_af_home[i,], pred_af_lew[i,]))
  df_af_lew <- rbind(df_af_lew, ec_dist)
}
af_offset_lew <- cbind(pops_list1[3:8], df_af_lew) # associate metadata with offset calculation, will need to ensure that this list is in the correct order
colnames(af_offset_lew) <- c("site_name", "offset_gf_af_lew")

# york
df_af_yrk <- data.frame() # df
for (i in 1:nb_ind_af){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_af_home[i,], pred_af_yrk[i,]))
  df_af_yrk <- rbind(df_af_yrk, ec_dist)
}
af_offset_yrk <- cbind(pops_list1[3:8], df_af_yrk) # associate metadata with offset calculation, will need to ensure that this list is in the correct order
colnames(af_offset_yrk) <- c("site_name", "offset_gf_af_yrk")

# merge
af_offsets <- merge(af_offset_lew, af_offset_yrk, by = c("site_name"))

# geno
# lew
df_geno_lew <- data.frame() # df
for (i in 1:nb_ind_geno){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home[i,], pred_geno_lew[i,]))
  df_geno_lew <- rbind(df_geno_lew, ec_dist)
}
geno_offset_lew <- cbind(as.data.frame(expEnv[c("site_name","Label_ind")]), df_geno_lew) # associate metadata with offset calculation
colnames(geno_offset_lew) <- c("site_name", "Label_ind", "offset_gf_geno_lew")

# york
df_geno_yrk <- data.frame() # df
for (i in 1:nb_ind_geno){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home[i,], pred_geno_yrk[i,]))
  df_geno_yrk <- rbind(df_geno_yrk, ec_dist)
}
geno_offset_yrk <- cbind(as.data.frame(expEnv[c("site_name","Label_ind")]), df_geno_yrk) # associate metadata with offset calculation
colnames(geno_offset_yrk) <- c("site_name", "Label_ind", "offset_gf_geno_yrk")

# merge
geno_offsets <- merge(geno_offset_lew, geno_offset_yrk, by = c("site_name","Label_ind"))

# only one actual prediction for the geno offsets, so reduce the df
geno_offsets_red <- geno_offsets[!duplicated(geno_offsets$site_name),]

offsets_mvp <- merge(geno_offsets_red, af_offsets, by = c("site_name"))

# write these offsets to results
write.csv(offsets_mvp, paste0("results/GO_results/offsets_GF_expSites_", Sys.Date(), ".csv"))
```

### Plotting

I've got my offsets, and now want to plot them on a map! 

```{r}
# prep data for plotting
offsets_mvp
# for now just allele frequency stuff
af_offsets
env_exp_pop_red
exp_site_latlon
popsdf <- read.csv("data/IndDat/popsdf.csv")[,-1]
#colnames(offsets_mvp) <- c("site_name", "Label_ind", "offset_gf_geno_lew", "offset_gf_geno_yrk", "offset_gf_af_lew", "offset_gf_af_yrk")
popsplot <- merge(popsdf, exp_site_latlon, by = "site_name")
popsplot1 <- merge(popsplot, env_exp_pop_red, by = "site_name")
popsplot2 <- merge(popsplot1, af_offsets, by = "site_name")

# map
sf::sf_use_s2(FALSE) # sf package doesn't work

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin =-102 , xmax = -68, ymin = 24, ymax = 47.049130))

# bottom right: 24.353558, -55.373595
# top left: 60.109743, 9.489737

# offset to lew geno
map_lew_geno <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = popsplot2, aes(x=lon, y=lat, 
                                   fill = offset_gf_geno_lew), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.045))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to Lewisetta (GF, Geno)", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_lew_geno)

world_chesapeake <- sf::st_crop(world, c(xmin =-77.856476 , xmax = -72.762787, ymin = 36.019557, ymax = 39.939163))

# top left: 39.939163, -77.856476
# bottom right: 36.019557, -72.762787

popout_lew_geno <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = popsplot2, aes(x=lon, y=lat, 
                                   fill = offset_gf_geno_lew), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.045))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
popout_lew_geno

# offset to lew af
map_lew_af <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = popsplot2, aes(x=lon, y=lat, 
                                   fill = offset_gf_af_lew), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.02))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to Lewisetta (GF, AF)", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_lew_af)

popout_lew_af <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = popsplot2, aes(x=lon, y=lat, 
                                   fill = offset_gf_af_lew), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.02))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
popout_lew_af

# offset to york geno
map_york_geno <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = popsplot2, aes(x=lon, y=lat, 
                                   fill = offset_gf_geno_yrk), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.045))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to York River (GF, Geno)", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_york_geno)

popout_york_geno <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = popsplot2, aes(x=lon, y=lat, 
                                   fill = offset_gf_geno_yrk), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.07))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
popout_york_geno

# offset to york af
map_york_af <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = popsplot2, aes(x=lon, y=lat, 
                                   fill = offset_gf_af_yrk), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.02))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
        labs(title = "Genomic Offset of Oysters \n to York River (GF, AF)", x = "Longitude", y = "Latitude", fill = "Genomic \n Offset")
(map_york_af)

popout_york_af <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = popsplot2, aes(x=lon, y=lat, 
                                   fill = offset_gf_af_yrk), 
             shape = 21, size = 4)+
  scale_fill_gradient(low = "green", high = "red", space = "Lab", limits = c(0,.02))+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
popout_york_af
```