---
title: "CviMVP_oyster_GO_plotting"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# wd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# libraries
library(ggplot2) # for plotting
library(raster) # for environmental pixels
library(sf) # mapping
library(s2) # mapping
library(rnaturalearth) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ggrepel) # mapping
library(officer) # for powerpoint export/editing
library(rvg) # for converting r graphics to editable vector graphics
library(stars)

# data
# metadata
popsdf <- read.csv("data/IndDat/popsdf.csv")[,-1]
latlons <- read.csv("data/IndDat/seascape_latlon.csv")[,-1]

# interpolated
lea_interpolated <- readRDS("results/GO_results/lea_interpolated_go.RDS")
rda_interpolated <- readRDS("results/GO_results/rda_interpolated_go.RDS")
gf_interpolated <- readRDS("results/GO_results/gf_interpolated_go.RDS")
envdist_interpolated <- readRDS("results/GO_results/envdist_interpolated.RDS")

# experimental
lea_go <- read.csv("results/GO_results/offsets_LEA_expsites_2026-04-06.csv")[,-1]
rda_go <- read.csv("results/GO_results/offsets_RDA_expSites_2026-04-07.csv")[,-1]
gf_go <- read.csv("results/GO_results/offsets_GF_expSites_2026-04-23.csv")[,-1]

# seascape
lea_sea <- read.csv("results/GO_results/offsets_LEA_fullseascape_2026-04-06.csv")[,-1]
rda_sea <- read.csv("results/GO_results/offsets_RDA_fullseascape_2026-04-07.csv")[,-1]
gf_sea <- read.csv("results/GO_results/offsets_GF_fullseascape_2026-04-23.csv")[,-1]
envdist <- read.csv("data/EnvDat/exp/distance_matrix_2026-01-30.csv")[,c("X","Lewisetta","YorkRiver")] # this has both the seascape sites and the experimental sites
envdist_abiotic <- read.csv("data/EnvDat/exp/distance_matrix_abiotic_2026-01-30.csv")[,c("X","Lewisetta","YorkRiver")] # this has both the seascape sites and the experimental sites
```


# Genomic Offset Plotting

This script is for high quality plotting and export of genomic offset prediction maps across all methods (LEA, GF, RDA, EnvDist) to an editable powerpoint with vectorized graphics. 


## Data Prep

Start by inspecting data & reformatting as necessary.

```{r}
# check data
lea_go
rda_go
gf_go
lea_sea
rda_sea
gf_sea
envdist

# experimental go have the wrong latlon for deby
lea_go[lea_go$site_name == "S2_DEBY",]$Latitude_manual <- 37.24728
lea_go[lea_go$site_name == "S2_DEBY",]$Longitude_manual <- -76.49937
rda_go[rda_go$site_name == "S2_DEBY",]$lat <- 37.24728
rda_go[rda_go$site_name == "S2_DEBY",]$lon <- -76.49937
gf_go[gf_go$site_name == "S2_DEBY",]$lat <- 37.24728
gf_go[gf_go$site_name == "S2_DEBY",]$lon <- -76.49937

# latlons
latlons$site_name <- latlons$ID_SiteDate

# rename envdist columns
colnames(envdist) <- colnames(envdist_abiotic) <- c("site_name", "dist_lew", "dist_yrk")

# split into seascape vs exp pops for future plotting
envdist_sea <- envdist[1:27,]
envdist_exp <- envdist[28:35,]

envdist_sea_nb <- envdist_abiotic[1:27,]
envdist_exp_nb <- envdist_abiotic[28:35,]

# add latlons
envdist_sea2 <- merge(latlons[,c("site_name","Latitude_manual","Longitude_manual")], envdist_sea, by = c("site_name"))
envdist_exp2 <- merge(popsdf, envdist_exp, by = c("site_name"))

envdist_sea2_nb <- merge(latlons[,c("site_name","Latitude_manual","Longitude_manual")], envdist_sea_nb, by = c("site_name"))
envdist_exp2_nb <- merge(popsdf, envdist_exp_nb, by = c("site_name"))

# rda also needs latlons
rda_sea2 <- merge(latlons[,c("site_name","Latitude_manual","Longitude_manual")], rda_sea, by = c("site_name"))
#########################
```

Prep the plotting range.

```{r}
# set spherical geometry
sf::sf_use_s2(FALSE)

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin =-102, xmax = -65, ymin = 24, ymax = 47.05))
coast <- ne_coastline(scale = "medium", returnclass = "sf")
coast_crop <- sf::st_crop(coast, c(xmin =-102, xmax = -65, ymin = 24, ymax = 47.05))
map_bounds <- data.frame(xmin =-102, xmax = -65, ymin = 24, ymax = 47.05)

# chesapeake extent
ches_ext <- st_bbox(c(xmin =-77.856476 , xmax = -72.762787, ymin = 36.019557, ymax = 39.939163),
                    crs = st_crs(world))
world_chesapeake <- sf::st_crop(world, ches_ext)
ches_bounds <- data.frame(xmin =-77.856476 , xmax = -72.762787, ymin = 36.019557, ymax = 39.939163)

# crop the GO objects as well
lea_interpolated_chesapeake <- lapply(lea_interpolated, function(x) {st_crop(x, ches_ext)})
rda_interpolated_chesapeake <- lapply(rda_interpolated, function(x) {st_crop(x, ches_ext)})
gf_interpolated_chesapeake <- lapply(gf_interpolated, function(x) {st_crop(x, ches_ext)})
envdist_interpolated_chesapeake <- lapply(envdist_interpolated, function(x) {st_crop(x, ches_ext)})

# and the GO dataframes
# exp
lea_go_chesapeake <- lea_go[lea_go$Latitude_manual <= ches_bounds$ymax & lea_go$Latitude_manual >= ches_bounds$ymin & lea_go$Longitude_manual <= ches_bounds$xmax & lea_go$Longitude_manual >= ches_bounds$xmin,]
rda_go_chesapeake <- rda_go[rda_go$lat <= ches_bounds$ymax & rda_go$lat >= ches_bounds$ymin & rda_go$lon <= ches_bounds$xmax & rda_go$lon >= ches_bounds$xmin,]
gf_go_chesapeake <- gf_go[gf_go$lat <= ches_bounds$ymax & gf_go$lat >= ches_bounds$ymin & gf_go$lon <= ches_bounds$xmax & gf_go$lon >= ches_bounds$xmin,]
envdist_exp_chesapeake <- envdist_exp2[envdist_exp2$latitude <= ches_bounds$ymax & envdist_exp2$latitude >= ches_bounds$ymin & envdist_exp2$longitude <= ches_bounds$xmax & envdist_exp2$longitude >= ches_bounds$xmin,]

# seascape
lea_sea_chesapeake <- lea_sea[lea_sea$Latitude_manual <= ches_bounds$ymax & lea_sea$Latitude_manual >= ches_bounds$ymin & lea_sea$Longitude_manual <= ches_bounds$xmax & lea_sea$Longitude_manual >= ches_bounds$xmin,]
rda_sea_chesapeake <- rda_sea2[rda_sea2$Latitude_manual <= ches_bounds$ymax & rda_sea2$Latitude_manual >= ches_bounds$ymin & rda_sea2$Longitude_manual <= ches_bounds$xmax & rda_sea2$Longitude_manual >= ches_bounds$xmin,]
gf_sea_chesapeake <- gf_sea[gf_sea$lat <= ches_bounds$ymax & gf_sea$lat >= ches_bounds$ymin & gf_sea$lon <= ches_bounds$xmax & gf_sea$lon >= ches_bounds$xmin,]
envdist_sea_chesapeake <- envdist_sea2[envdist_sea2$Latitude_manual <= ches_bounds$ymax & envdist_sea2$Latitude_manual >= ches_bounds$ymin & envdist_sea2$Longitude_manual <= ches_bounds$xmax & envdist_sea2$Longitude_manual >= ches_bounds$xmin,]
```

## Plotting

Now plot the genomic offsets. For each method, we make a map of genomic offset to Lewisetta and York River, with large points showing predictions for experimental populations, and small points showing predictions for seascape populations. The shaded region shows genomic offset interpolated over the seascape range (extends 0.25º from coast). 

### LEA

```{r}
map_lew_int_lea <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(lea_interpolated$interpolated_lew_go), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = lea_sea, aes(x = Longitude_manual, y = Latitude_manual, 
                                              fill = lew_go), 
             shape = 21, size = 2) +
  geom_point(data = lea_go[lea_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),],
             aes(x=Longitude_manual, y=Latitude_manual, fill = lew_go), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.06)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.06)) +
  geom_label_repel(data = lea_go[lea_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),],
                  aes(x=Longitude_manual, y=Latitude_manual, label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Predicted ", LEA[offset], " to Lewisetta")), x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")

map_lew_int_lea_inset <- ggplot(data = world_chesapeake) +
  geom_sf(data = st_as_sf(lea_interpolated_chesapeake$interpolated_lew_go), mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.06)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.06)) +
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = lea_sea_chesapeake, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = lew_go), 
             shape = 21, size = 2) +
  geom_point(data = lea_go_chesapeake, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = lew_go), 
             shape = 21, size = 4)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 0.08)) +
  geom_point(data = popsdf[popsdf$site_name == "LEWISETTA",], 
             aes(x = longitude-0.2, y = latitude), fill = "deeppink", shape = 24, size = 4)+
  geom_label_repel(data = lea_go_chesapeake,
                  aes(x = Longitude_manual, y = Latitude_manual, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.4,
                  fill = "white",
                  size = 4)+
  guides(fill = "none")+
  theme_void() +
  xlim(ches_bounds$xmin, ches_bounds$xmax) +
  ylim(ches_bounds$ymin, ches_bounds$ymax) +
  theme(legend.position = "none", 
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
        panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"))

# plot panel
map_lew_lea <- map_lew_int_lea + 
  annotate("rect", xmin = ches_bounds$xmin, xmax = ches_bounds$xmax, 
           ymin = ches_bounds$ymin, ymax = ches_bounds$ymax, 
           fill = NA, color = "black", size = 0.5) +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 47.05, yend = ches_bounds$ymax, color = "black", linetype = "solid") +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 35.6, yend = ches_bounds$ymin, color = "black", linetype = "solid") +
  inset_element(map_lew_int_lea_inset, 
    left = -0.05,
    bottom = 0.5,
    right = 0.5,
    top = 0.95)
map_lew_lea

# yrk
map_yrk_int_lea <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(lea_interpolated$interpolated_yrk_go), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = lea_sea, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = yrk_go), 
             shape = 21, size = 2) +
  geom_point(data = lea_go[lea_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], 
             aes(x = Longitude_manual, y = Latitude_manual, fill = yrk_go), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.06)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.06)) +
  geom_label_repel(data = lea_go[lea_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x = Longitude_manual, y = Latitude_manual, label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = "right")+
  labs(title = expression(paste("(B) Predicted ", LEA[offset], " to York River")), x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")

map_yrk_int_lea_inset <- ggplot(data = world_chesapeake) +
  theme_bw()+
  geom_sf(data = st_as_sf(lea_interpolated_chesapeake$interpolated_yrk_go), mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.06)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.06)) +
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = lea_sea_chesapeake, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = yrk_go), 
             shape = 21, size = 2) +
  geom_point(data = lea_go_chesapeake, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = yrk_go), 
             shape = 21, size = 4)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 0.08)) +
  geom_point(data = popsdf[popsdf$site_name == "YORKRIVER",], 
             aes(x = longitude-0.1, y = latitude+0.1), fill = "deeppink", shape = 24, size = 4)+
  geom_label_repel(data = lea_go_chesapeake, 
                  aes(x = Longitude_manual, y = Latitude_manual, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.75,
                  fill = "white",
                  size = 4)+
guides(fill = "none")+
  theme_void() +
  xlim(ches_bounds$xmin, ches_bounds$xmax) +
  ylim(ches_bounds$ymin, ches_bounds$ymax) +
  theme(legend.position = "none", 
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
        panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"))

# plot panel
map_yrk_lea <- map_yrk_int_lea + 
  annotate("rect", xmin = ches_bounds$xmin, xmax = ches_bounds$xmax, 
           ymin = ches_bounds$ymin, ymax = ches_bounds$ymax, 
           fill = NA, color = "black", size = 0.5) +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 47.05, yend = ches_bounds$ymax, color = "black", linetype = "solid") +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 35.6, yend = ches_bounds$ymin, color = "black", linetype = "solid") +
  inset_element(map_yrk_int_lea_inset, 
    left = -0.05,
    bottom = 0.5,
    right = 0.5,
    top = 0.95)
map_yrk_lea
```

### RDA

```{r}
map_lew_int_rda <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(rda_interpolated$interpolated_lew_go), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = rda_sea2, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = offset_lew_sea), 
             shape = 21, size = 2) +
  geom_point(data = rda_go[rda_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], 
             aes(x = lon, y = lat, fill = offset_lew), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,4)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,4)) +
  geom_label_repel(data = rda_go[rda_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),],
                  aes(x = lon, y = lat, label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(C) Predicted ", RDA[offset], " to Lewisetta")), x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")

map_lew_int_rda_inset <- ggplot(data = world_chesapeake) +
  geom_sf(data = st_as_sf(rda_interpolated_chesapeake$interpolated_lew_go), mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,4)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,4)) +
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = rda_sea_chesapeake, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = offset_lew_sea), 
             shape = 21, size = 2) +
  geom_point(data = rda_go_chesapeake, 
             aes(x = lon, y = lat, fill = offset_lew), 
             shape = 21, size = 4)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 3.5)) +
  geom_point(data = popsdf[popsdf$site_name == "LEWISETTA",], 
             aes(x = longitude-0.2, y = latitude), fill = "deeppink", shape = 24, size = 4)+
  geom_label_repel(data = rda_go_chesapeake, 
                  aes(x = lon, y = lat, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.75,
                  fill = "white",
                  size = 4)+
  guides(fill = "none")+
  theme_void() +
  xlim(ches_bounds$xmin, ches_bounds$xmax) +
  ylim(ches_bounds$ymin, ches_bounds$ymax) +
  theme(legend.position = "none", 
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
        panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"))

# plot panel
map_lew_rda <- map_lew_int_rda + 
  annotate("rect", xmin = ches_bounds$xmin, xmax = ches_bounds$xmax, 
           ymin = ches_bounds$ymin, ymax = ches_bounds$ymax, 
           fill = NA, color = "black", size = 0.5) +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 47.05, yend = ches_bounds$ymax, color = "black", linetype = "solid") +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 35.6, yend = ches_bounds$ymin, color = "black", linetype = "solid") +
  inset_element(map_lew_int_rda_inset, 
    left = -0.05,
    bottom = 0.5,
    right = 0.5,
    top = 0.95)
map_lew_rda

# yrk
map_yrk_int_rda <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(rda_interpolated$interpolated_yrk_go), mapping = aes(col = var1.pred, fill = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = rda_sea2, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = offset_yrk_sea), 
             shape = 21, size = 2) +
  geom_point(data = rda_go[rda_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], 
             aes(x = lon, y = lat, fill = offset_yrk), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,4)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,4)) +
  geom_label_repel(data = rda_go[rda_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),],
                  aes(x = lon, y = lat, label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = "right")+
  labs(title = expression(paste("(D) Predicted ", RDA[offset], " to York River")), x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")

map_yrk_int_rda_inset <- ggplot(data = world_chesapeake) +
  geom_sf(data = st_as_sf(rda_interpolated_chesapeake$interpolated_yrk_go), mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,4)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,4)) +
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = rda_sea_chesapeake, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = offset_yrk_sea), 
             shape = 21, size = 2) +
  geom_point(data = rda_go_chesapeake, 
             aes(x = lon, y = lat, fill = offset_yrk), 
             shape = 21, size = 4)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 3.5)) +
  geom_point(data = popsdf[popsdf$site_name == "YORKRIVER",], 
             aes(x = longitude-0.1, y = latitude+0.1), fill = "deeppink", shape = 24, size = 4)+
  geom_label_repel(data = rda_go_chesapeake, 
                  aes(x = lon, y = lat, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.7,
                  fill = "white",
                  size = 4)+
  guides(fill = "none")+
  theme_void() +
  xlim(ches_bounds$xmin, ches_bounds$xmax) +
  ylim(ches_bounds$ymin, ches_bounds$ymax) +
  theme(legend.position = "none", 
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
        panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"))

# put together plot panel
map_yrk_rda <- map_yrk_int_rda + 
  annotate("rect", xmin = ches_bounds$xmin, xmax = ches_bounds$xmax, 
           ymin = ches_bounds$ymin, ymax = ches_bounds$ymax, 
           fill = NA, color = "black", size = 0.5) +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 47.05, yend = ches_bounds$ymax, color = "black", linetype = "solid") +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 35.6, yend = ches_bounds$ymin, color = "black", linetype = "solid") +
  inset_element(map_yrk_int_rda_inset, 
    left = -0.05,
    bottom = 0.5,
    right = 0.5,
    top = 0.95)
map_yrk_rda
```

### GF

```{r}
# lew
map_lew_int_gf_geno <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(gf_interpolated$interpolated_lew_go_geno), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = gf_sea, 
             aes(x = lon, y = lat, fill = offset_gf_geno_lew), 
             shape = 21, size = 2) +
  geom_point(data = gf_go[gf_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], 
             aes(x = lon, y = lat, fill = offset_gf_geno_lew), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  geom_label_repel(data = gf_go[gf_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], 
                  aes(x = lon, y = lat, label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(E) Predicted ", GF[offset], " to Lewisetta")), x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")

map_lew_int_gf_geno_inset <- ggplot(data = world_chesapeake) +
  geom_sf(data = st_as_sf(gf_interpolated_chesapeake$interpolated_lew_go_geno), mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = gf_sea_chesapeake, aes(x = lon, y = lat, fill = offset_gf_geno_lew), 
             shape = 21, size = 2) +
  geom_point(data = gf_go_chesapeake, 
             aes(x = lon, y = lat, fill = offset_gf_geno_lew), 
             shape = 21, size = 4)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 0.04)) +
  geom_point(data = popsdf[popsdf$site_name == "LEWISETTA",], 
             aes(x = longitude-0.2, y = latitude), fill = "deeppink", shape = 24, size = 4)+
  geom_label_repel(data = gf_go_chesapeake, 
                  aes(x = lon, y = lat, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.4,
                  fill = "white",
                  size = 4)+
  guides(fill = "none")+
  theme_void() +
  xlim(ches_bounds$xmin, ches_bounds$xmax) +
  ylim(ches_bounds$ymin, ches_bounds$ymax) +
  theme(legend.position = "none", 
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
        panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"))

# put together plot panel
map_lew_gf <- map_lew_int_gf_geno + 
  annotate("rect", xmin = ches_bounds$xmin, xmax = ches_bounds$xmax, 
           ymin = ches_bounds$ymin, ymax = ches_bounds$ymax, 
           fill = NA, color = "black", size = 0.5) +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 47.05, yend = ches_bounds$ymax, color = "black", linetype = "solid") +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 35.6, yend = ches_bounds$ymin, color = "black", linetype = "solid") +
  inset_element(map_lew_int_gf_geno_inset, 
    left = -0.05,
    bottom = 0.5,
    right = 0.5,
    top = 0.95)
map_lew_gf

# yrk 
map_yrk_int_gf_geno <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(gf_interpolated$interpolated_yrk_go_geno), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = gf_sea, 
             aes(x = lon, y = lat, fill = offset_gf_geno_yrk), 
             shape = 21, size = 2) +
  geom_point(data = gf_go[gf_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], 
             aes(x = lon, y = lat, fill = offset_gf_geno_yrk), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  geom_label_repel(data = gf_go[gf_go$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x = lon, y = lat, label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(F) Predicted ", GF[offset], " to York River")), x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")

map_yrk_int_gf_geno_inset <- ggplot(data = world_chesapeake) +
  geom_sf(data = st_as_sf(gf_interpolated_chesapeake$interpolated_yrk_go_geno), mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,0.04)) +
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = gf_sea_chesapeake, 
             aes(x = lon, y = lat, fill = offset_gf_geno_yrk), 
             shape = 21, size = 2) +
  geom_point(data = gf_go_chesapeake, 
             aes(x = lon, y = lat, fill = offset_gf_geno_yrk), 
             shape = 21, size = 4)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 0.04)) +
  geom_point(data = popsdf[popsdf$site_name == "YORKRIVER",], 
             aes(x = longitude-0.1, y = latitude+0.1), fill = "deeppink", shape = 24, size = 4)+
  geom_label_repel(data = gf_go_chesapeake, 
                   aes(x = lon, y = lat, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.4,
                  fill = "white",
                  size = 4)+
  guides(fill = "none")+
  theme_void() +
  xlim(ches_bounds$xmin, ches_bounds$xmax) +
  ylim(ches_bounds$ymin, ches_bounds$ymax) +
  theme(legend.position = "none", 
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
        panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"))

# put together plot panel
map_yrk_gf <- map_yrk_int_gf_geno + 
  annotate("rect", xmin = ches_bounds$xmin, xmax = ches_bounds$xmax, 
           ymin = ches_bounds$ymin, ymax = ches_bounds$ymax, 
           fill = NA, color = "black", size = 0.5) +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 47.05, yend = ches_bounds$ymax, color = "black", linetype = "solid") +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 35.6, yend = ches_bounds$ymin, color = "black", linetype = "solid") +
  inset_element(map_yrk_int_gf_geno_inset, 
    left = -0.05,
    bottom = 0.5,
    right = 0.5,
    top = 0.95)
map_yrk_gf
```

### Environmental Distance

```{r}
# lew
map_lew_int_envdist <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(envdist_interpolated$interpolated_lew_envdist), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = envdist_sea2, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = dist_lew), 
             shape = 21, size = 2) +
  geom_point(data = envdist_exp2[envdist_exp2$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], 
             aes(x = longitude, y = latitude, fill = dist_lew), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  geom_label_repel(data = envdist_exp2[envdist_exp2$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], 
                  aes(x = longitude, y = latitude, label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(G) Environmental Distance to \nLewisetta", x = "Longitude", y = "Latitude", col = "Environmental \nDistance", fill = "Environmental \nDistance")

map_lew_int_envdist_inset <- ggplot(data = world_chesapeake) +
  geom_sf(data = st_as_sf(envdist_interpolated_chesapeake$interpolated_lew_envdist), mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = envdist_sea_chesapeake, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = dist_lew), 
             shape = 21, size = 2) +
  geom_point(data = envdist_exp_chesapeake, 
             aes(x = longitude, y = latitude, fill = dist_lew), 
             shape = 21, size = 4)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 6)) +
  geom_point(data = popsdf[popsdf$site_name == "LEWISETTA",], 
             aes(x = longitude-0.2, y = latitude), fill = "deeppink", 
             shape = 24, size = 4)+
  geom_label_repel(data = envdist_exp_chesapeake, 
                  aes(x = longitude, y = latitude, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.75,
                  fill = "white")+
  guides(fill = "none")+
  theme_void() +
  xlim(ches_bounds$xmin, ches_bounds$xmax) +
  ylim(ches_bounds$ymin, ches_bounds$ymax) +
  theme(legend.position = "none", 
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
        panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"))

# put together plot panel
map_lew_envdist <- map_lew_int_envdist + 
  annotate("rect", xmin = ches_bounds$xmin, xmax = ches_bounds$xmax, 
           ymin = ches_bounds$ymin, ymax = ches_bounds$ymax, 
           fill = NA, color = "black", size = 0.5) +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 47.05, yend = ches_bounds$ymax, color = "black", linetype = "solid") +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 35.6, yend = ches_bounds$ymin, color = "black", linetype = "solid") +
  inset_element(map_lew_int_envdist_inset, 
    left = -0.05,
    bottom = 0.5,
    right = 0.5,
    top = 0.95)
map_lew_envdist

# yrk
map_yrk_int_envdist <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(envdist_interpolated$interpolated_yrk_envdist), mapping = aes(col = var1.pred, fill = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = envdist_sea2, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = dist_yrk), 
             shape = 21, size = 2) +
  geom_point(data = envdist_exp2[envdist_exp2$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], 
             aes(x = longitude, y = latitude, fill = dist_yrk), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  geom_label_repel(data = envdist_exp2[envdist_exp2$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x = longitude, y = latitude, label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = "right")+
  labs(title = "(H) Environmental Distance to \nYork River", x = "Longitude", y = "Latitude", col = "Environmental \nDistance", fill = "Environmental \nDistance")

map_yrk_int_envdist_inset <- ggplot(data = world_chesapeake) +
  geom_sf(data = st_as_sf(envdist_interpolated_chesapeake$interpolated_yrk_envdist), mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  geom_sf(data = world_chesapeake, fill = 'antiquewhite1') +
  geom_point(data = envdist_sea_chesapeake, 
             aes(x = Longitude_manual, y = Latitude_manual, fill = dist_yrk), 
             shape = 21, size = 2) +
  geom_point(data = envdist_exp_chesapeake, 
             aes(x=longitude, y=latitude, fill = dist_yrk), 
             shape = 21, size = 4)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 6)) +
  geom_point(data = popsdf[popsdf$site_name == "YORKRIVER",], 
             aes(x = longitude-0.1, y = latitude+0.1), fill = "deeppink", shape = 24, size = 4)+
  geom_label_repel(data = envdist_exp_chesapeake, 
                  aes(x = longitude, y = latitude, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.7,
                  fill = "white",
                  size = 4)+
  guides(fill = "none")+
  theme_void() +
  xlim(ches_bounds$xmin, ches_bounds$xmax) +
  ylim(ches_bounds$ymin, ches_bounds$ymax) +
  theme(legend.position = "none", 
        panel.border = element_rect(colour = "black", fill = NA, size = 1),
        panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"))

# put together plot panel
map_yrk_envdist <- map_yrk_int_envdist + 
  annotate("rect", xmin = ches_bounds$xmin, xmax = ches_bounds$xmax, 
           ymin = ches_bounds$ymin, ymax = ches_bounds$ymax, 
           fill = NA, color = "black", size = 0.5) +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 47.05, yend = ches_bounds$ymax, color = "black", linetype = "solid") +
  annotate("segment", x = -87.4, xend = ches_bounds$xmin, y = 35.6, yend = ches_bounds$ymin, color = "black", linetype = "solid") +
  inset_element(map_yrk_int_envdist_inset, 
    left = -0.05,
    bottom = 0.5,
    right = 0.5,
    top = 0.95)
map_yrk_envdist
```


## Export

```{r}
# export to ppt as editable vector graphic
# doc <- read_pptx() %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_lew_int_lea), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_lew_int_lea_inset), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_yrk_int_lea), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_yrk_int_lea_inset), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_lew_int_rda), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_lew_int_rda_inset), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_yrk_int_rda), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_yrk_int_rda_inset), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_lew_int_gf_geno), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_lew_int_gf_geno_inset), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_yrk_int_gf_geno), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_yrk_int_gf_geno_inset), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_lew_int_envdist), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_lew_int_envdist_inset), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_yrk_int_envdist), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme") %>% 
#   ph_with(value = dml(ggobj = map_yrk_int_envdist_inset), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
#   add_slide(layout = "Title and Content", master = "Office Theme")

doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = map_lew_lea), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = map_yrk_lea), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = map_lew_rda), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = map_yrk_rda), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = map_lew_gf), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = map_yrk_gf), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = map_lew_envdist), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = map_yrk_envdist), location = ph_location(left = 1, top = 1, width = 6.5, height = 4))

print(doc, target = "results/figures/GO_maps.pptx")
```

Also format and export the raw genomic offsets for both experimental and seasecape sites:

```{r}
# experimental sites
lea_go
rda_go
gf_go
envdist_exp2
envdist_exp2_nb

colnames(lea_go) <- c("site_name","Latitude_manual","Longitude_manual","LEA_lew","LEA_yrk","LEA_lew_abiotic","LEA_yrk_abiotic")
colnames(rda_go) <- c("site_name","Latitude_manual","Longitude_manual","RDA_lew","RDA_yrk","RDA_lew_abiotic","RDA_yrk_abiotic")
colnames(gf_go) <- c("site_name","Latitude_manual","Longitude_manual","GF_geno_lew","GF_geno_yrk","GF_geno_lew_abiotic","GF_geno_yrk_abiotic")
colnames(envdist_exp2) <- c("site_name","pop","order","Latitude_manual","Longitude_manual","cols","shape","EnvDist_lew","EnvDist_yrk")
colnames(envdist_exp2_nb) <- c("site_name","pop","order","Latitude_manual","Longitude_manual","cols","shape","EnvDist_lew_abiotic","EnvDist_yrk_abiotic")

exp_offsets1 <- merge(lea_go, rda_go, by = c("site_name","Latitude_manual","Longitude_manual"))
exp_offsets2 <- merge(exp_offsets1, gf_go, by = c("site_name","Latitude_manual","Longitude_manual"))
exp_offsets3 <- merge(exp_offsets2, envdist_exp2[,c("site_name","EnvDist_lew","EnvDist_yrk")], by = c("site_name"))
exp_offsets <- merge(exp_offsets3, envdist_exp2_nb[,c("site_name","EnvDist_lew_abiotic","EnvDist_yrk_abiotic")], by = c("site_name"))

# now seascape
lea_sea
rda_sea2
gf_sea
envdist_sea2
envdist_sea2_nb

colnames(lea_sea) <- c("site_name","Latitude_manual","Longitude_manual","LEA_lew","LEA_yrk")
colnames(rda_sea2) <- c("site_name","Latitude_manual","Longitude_manual","RDA_lew","RDA_yrk")
colnames(gf_sea) <- c("site_name","Latitude_manual","Longitude_manual","GF_geno_lew","GF_geno_yrk")
colnames(envdist_sea2) <- c("site_name","Latitude_manual","Longitude_manual","EnvDist_lew","EnvDist_yrk")

sea_offsets1 <- merge(lea_sea, rda_sea2, by = c("site_name","Latitude_manual","Longitude_manual"))
sea_offsets2 <- merge(sea_offsets1, gf_sea, by = c("site_name","Latitude_manual","Longitude_manual"))
sea_offsets <- merge(sea_offsets2, envdist_sea2[,c("site_name","EnvDist_lew","EnvDist_yrk")], by = c("site_name"))

write.csv(exp_offsets, paste0("results/GO_results/Offsets_expSites_", Sys.Date(), ".csv"), row.names = F)
write.csv(sea_offsets, paste0("results/GO_results/Offsets_fullSeascape_", Sys.Date(), ".csv"), row.names = F)
```
