---
title: "CviMVP_oyster_GonadIndex"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/projects/lotterhos/MVP_oyster_GO")

# load libraries
library(tidyverse) # for data wrangling
library(stringr) # for splitting strings
library(lubridate) # for dates
library(ggplot2) # for plotting
library(plotrix) # for std error
```

# Calculating Gonadal Index of Oysters Sampled at Conclusion of Field Experiment

This script is to analyze the gonad index of oysters at the end of the MVP Field Experiment from First I need to read in and merge together relevant data. I will then clean this data up to prepare it for analysis by site and population.

## Setup

```{r}
# read in data
bags <- read.csv("data/ExpDat/MVP23-FieldBags - bags.csv")
gonad <- read.csv("data/ExpDat/MVP23-FieldBags - finalTissue.csv")
```

## Data Prep

```{r}
# link gonad to bags
gonad_bags <- merge(gonad, bags, by = "bags_key", keep = T)

# get pop and site data
gonad_bags$pop <- str_split_fixed(gonad_bags$bags_label, "-", 3)[,2]

# trim to relevant data
gonad_trim <- gonad_bags %>% filter(!(pop %in% c("SEEDMIX", "LARMIX"))) %>% select(bag_site, pop, bags_label, finalTissue_VialLabel, finalTissue_condition, finalTissue_gonadIndex)
gonad_trim$condition <- as.factor(str_split_fixed(gonad_trim$finalTissue_condition, " - ", 2)[,1])
gonad_trim$gonad <- as.factor(str_split_fixed(gonad_trim$finalTissue_gonadIndex, " - ", 2)[,1])

# rename populations
map=c("TX"="W1_TX","LA"="W2_LA","FL"="W3_FL","LOLA"="S1_LOLA","DEBY"="S2_DEBY","JR"="W4_VA","NH"="W5_NH","ME"="W6_ME")
gonad_trim_rename <- gonad_trim %>% mutate(pop = ifelse(pop %in% names(map), map[pop], pop)) %>% rename(site_name = pop)

# remove any NAs
gonad_trim_rename_nona <- na.omit(gonad_trim_rename)

# output for downstream analysis
write.csv(gonad_trim_rename_nona, paste0("results/exp_results/gonad_condition_", Sys.Date(), ".csv"))
```

## Population & Site Trends

Now that I have prepped my data, I want to look at population/site level trends

```{r}
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "LOLA", "DEBY", "NH", "ME"), 
                     order = 1:8, 
                     cols = c("#365C8DFF", "#277F8EFF", "#1FA187FF", "#4AC16DFF", "#440154FF", "#46337EFF", "#9FDA3AFF", "#FDE725FF"),
                     shape = c(21,21,21,21,23,23,21,21),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME"),
                     site_name_plotting = c("W1-TX", "W2-LA", "W3-FL", "W4-VA", "S1-LOLA", "S2-DEBY", "W5-NH", "W6-ME"))
gonad_df_full <- merge(gonad_trim_rename_nona, popsdf, by = "site_name")

# gonad index by site/pop
gonad_plot_pop <- ggplot(gonad_df_full, aes(x = site_name_plotting, y = as.numeric(gonad), fill = fct_reorder(site_name, order))) + 
  geom_boxplot(position=position_dodge(1.0)) +  
  facet_grid(.~ bag_site,
             switch = "x") +
  scale_fill_manual(name = "Broodstock Group", label = popsdf$site_name_plotting, values = popsdf$cols)+
#  geom_text(aes(label = letters, y = max_surv), vjust = -0.5, position = position_dodge(1)) +
  labs(title = "Final Gonad Condition at Lewisetta and York River", x = "Experimental Site", y = "Gonad Index") +
  theme_classic()+ 
  theme(strip.background = element_blank(), 
        strip.text.x = element_text(), 
        strip.placement = "outside",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.spacing.x = unit(3, "lines"))
gonad_plot_pop

# condition by site/pop
cond_plot_pop <- ggplot(gonad_df_full, aes(x = site_name_plotting, y = as.numeric(condition), fill = fct_reorder(site_name, order))) + 
  geom_boxplot(position=position_dodge(1.0)) +  
  facet_grid(.~ bag_site,
             switch = "x") +
  scale_fill_manual(name = "Broodstock Group", label = popsdf$site_name_plotting, values = popsdf$cols)+
#  geom_text(aes(label = letters, y = max_surv), vjust = -0.5, position = position_dodge(1)) +
  labs(title = "Final Tissue Condition at Lewisetta and York River", x = "Experimental Site", y = "Condition") +
  theme_classic()+ 
  theme(strip.background = element_blank(), 
        strip.text.x = element_text(), 
        strip.placement = "outside",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.spacing.x = unit(3, "lines"))
cond_plot_pop
```
