---
title: "MVP_oyster_Fitness_Modeling"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/projects/lotterhos/MVP_oyster_GO")

# load libraries
library(tidyverse) # for data wrangling
library(dplyr) # for data wrangling
library(stringr) # for splitting strings
library(lubridate) # for dates
library(ggplot2) # for plotting
library(plotrix) # for std error
library(MASS) # for ordinal regression
library(lme4) # for lmm
library(performance) # for lmm
library(lmerTest) # for lmm
library(car) # for avPlots(), Anova()
library(emmeans) # for significance letters
library(multcomp) # for significance letters
library(multcompView) # for significance letters
library(officer) # for export to ppt
library(rvg) # for converting to vector graphics

# data
surv_long <- read.csv("results/exp_results/mortality_long_bags_2026-01-15.csv")[,-1]
bags <- read.csv("data/ExpDat/MVP23-FieldBags - bags.csv")
lengths <- read.csv("results/exp_results/lengths_long_2025-06-03.csv")[,-1]
condition <- read.csv("results/exp_results/gonad_condition_2026-03-30.csv")[,-1]
```

This script is for modeling differences in survival, length, tissue condition, and gonad index, amongst bags/populations/sites/time points in the MVP Field Experiment. 

I intend to use linear (mixed) models to examine how survival and length vary by group and how genetics and environmental history drive these differences in performance. 

## Modeling differences in survival

First I'll process survival data:

```{r}
# split bags to have more detailed bag data
bags[,5:6] <- str_split_fixed(bags$bags_label, "-", 3)[,2:3]
colnames(bags) <- c("bags_key","bag_site","bag","SpawnTrt_Key","pop","bagnum")

# combine data
surv_for_mod <- merge(surv_long, bags[,c("bag","bag_site")], by = "bag")
surv_for_mod$pop <- str_split_fixed(surv_for_mod$bag, "-", 3)[,2]
surv_mod_ord <- surv_for_mod[,c("bag_site","pop","bag","event","surv")]

# rename populations
map=c("TX"="W1_TX","LA"="W2_LA","FL"="W3_FL","LOLA"="S1_LOLA","DEBY"="S2_DEBY","JR"="W4_VA","NH"="W5_NH","ME"="W6_ME")
surv_mod_ord_rename <- surv_mod_ord %>% mutate(pop = ifelse(pop %in% names(map), map[pop], pop))

# remove any NAs
surv_mod_ord_rename_nona <- na.omit(surv_mod_ord_rename)

# we should also remove seedmix and larmix for now
surv_mod_ord_rm <- surv_mod_ord_rename_nona %>% filter(!pop %in% c("LARMIX","SEEDMIX"))

colnames(surv_mod_ord_rm) <- c("bag_site","pop","bags_label","t","surv")
```


Now I can use LMMs to look at differences in survival by group.

```{r}
# analysis of fixed effects
anova(lm(surv ~ pop + bag_site + t, data = surv_mod_ord_rm)) # all significant

# ensure correct coding of explanatory variables
surv_mod_ord_rm$t <- as.factor(surv_mod_ord_rm$t)

# start with full model, then subtract each interaction, then subtract two of the interaction
surv_lmm <- lmer(surv ~ bag_site * pop * t + (1|bags_label), data = surv_mod_ord_rm)
plot(surv_lmm)
Anova(surv_lmm, type = "III")
drop1(surv_lmm) # keep all
summary(surv_lmm)

# one more model on final time point only
surv_lm <- lm(surv ~ bag_site * pop, data = surv_mod_ord_rm[surv_mod_ord_rm$t == 4,]) # no random effect here because there is only one observation per bag label
plot(surv_lm)
Anova(surv_lm, type = "III") # make sure to always use this method if two categorical predictors and  interaction and unbalanced design ; type III may have to set the contrast, check biostats notes
TukeyHSD(aov(surv_lm))

# plot survival at last time point
surv_t4 <- surv_mod_ord_rm[surv_mod_ord_rm$t == 4,]

surv_t4_plot_site <- ggplot(surv_t4, aes(x = pop, y = surv, fill = bag_site)) + 
  geom_boxplot(position=position_dodge(0.8)) +
  ggtitle("Survival at Final Time Point") +
  theme_classic()
surv_t4_plot_site

surv_t4_plot_pop <- ggplot(surv_t4, aes(x = bag_site, y = surv, fill = pop)) + 
  geom_boxplot(position=position_dodge(1.0)) +
  theme_classic()
surv_t4_plot_pop

#extract significant letters from TK test and store
# add a different pop column that's split
surv_t4$oldpop <- str_split_fixed(surv_t4$pop,"_",2)[,2]

# now redo model
surv_lm2 <- lm(surv ~ bag_site * oldpop, data = surv_t4)
plot(surv_lm2)
Anova(surv_lm2, type = "III") # make sure to always use this method if two categorical predictors and interaction and unbalanced design ; type III may have to set the contrast, check biostats notes
library(multcompView)
tuk_surv_t4 <- TukeyHSD(aov(surv_lm2), conf.level = 0.95)
cld_surv_t4 <- multcompLetters4(aov(surv_lm2), tuk_surv_t4)

# can paste pops with sites to line up with letters output
surv_t4$popsite <- paste0(surv_t4$bag_site,":",str_split_fixed(surv_t4$pop,"_",2)[,2])

# multcomp letters to get letters
letters_surv_t4_df <- data.frame(letters = as.data.frame.list(cld_surv_t4$`bag_site:oldpop`), popsite = names((cld_surv_t4$`bag_site:oldpop`)$Letters))
letters_surv_t4_df_red <- letters_surv_t4_df[,c("letters.Letters","popsite")]
colnames(letters_surv_t4_df_red) <- c("letters","popsite")

# save these outputs too
sink(paste0("results/exp_results/survival_model_results_", Sys.Date(), ".txt"))
Anova(surv_lmm, type = "III")
summary(surv_lmm)
Anova(surv_lm, type = "III")
summary(surv_lm)
Anova(surv_lm2, type = "III")
summary(surv_lm2)
letters_surv_t4_df_red
sink()

# merge
surv_final_time_rename_mg <- merge(surv_t4, letters_surv_t4_df_red, by = "popsite")
(surv_final_time_rename_mg)
surv_final_time_rename_mg$site_name <- surv_final_time_rename_mg$pop

# pop color scheme
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "LOLA", "DEBY", "NH", "ME"), 
                     order = 1:8, 
                     cols = c("#365C8DFF", "#277F8EFF", "#1FA187FF", "#4AC16DFF", "#440154FF", "#46337EFF", "#9FDA3AFF", "#FDE725FF"),
                     shape = c(21,21,21,21,23,23,21,21),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME"),
                     site_name_plotting = c("W1-TX", "W2-LA", "W3-FL", "W4-VA", "S1-LOLA", "S2-DEBY", "W5-NH", "W6-ME"))

surv_final_time_plotting <- merge(surv_final_time_rename_mg, popsdf, by = "site_name")
surv_final_time_plotting$bag_site <- as.factor(surv_final_time_plotting$bag_site)
surv_final_time_plotting$site_name <- as.factor(surv_final_time_plotting$site_name)
surv_final_time_plotting$letters <- as.factor(surv_final_time_plotting$letters)
surv_final_time_plotting$t <- as.factor(surv_final_time_plotting$t)
surv_final_time_plotting$pop.x <- NULL
names(surv_final_time_plotting)[names(surv_final_time_plotting) == "pop.y"] <- "pop"

# y-axis location for letters
letters_loc <- surv_final_time_plotting %>% group_by(popsite) %>% summarise(across(surv, max))
colnames(letters_loc) <- c("popsite", "max_surv")
surv_final_time_plotting_letters <- merge(surv_final_time_plotting, letters_loc, by = "popsite")

# plot by pop
surv_t4_plot_pop <- ggplot(surv_final_time_plotting_letters, aes(x = site_name_plotting, y = surv, fill = fct_reorder(site_name, order))) + 
  geom_boxplot(position=position_dodge(1.0)) +  
  facet_grid(.~ bag_site,
             switch = "x") +
  scale_fill_manual(name = "Broodstock Group", label = popsdf$site_name_plotting, values = popsdf$cols)+
  geom_text(aes(label = letters, y = max_surv), vjust = -0.5, position = position_dodge(1)) +
  labs(title = "Final Survival at Lewisetta and York River", x = "Experimental Site", y = "Survival") +
  theme_classic()+ 
  theme(strip.background = element_blank(), 
        strip.text.x = element_text(), 
        strip.placement = "outside",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.spacing.x = unit(3, "lines"))
surv_t4_plot_pop
```


## Modeling differences in length

Now I'll try modeling differences in length. I'll start by reading in and processing the length data.

```{r}
# trim data
head(lengths)
lengths_mod <- lengths[,c("bag_site","pop","event","bags_label","tags_label","length","width")]
colnames(lengths_mod) <- c("bag_site","pop","t","bags_label","tags_label","length","width")

# rename pops
map=c("TX"="W1_TX","LA"="W2_LA","FL"="W3_FL","LOLA"="S1_LOLA","DEBY"="S2_DEBY",
      "JR"="W4_VA","NH"="W5_NH","ME"="W6_ME")
lengths_mod_rename <- lengths_mod %>% mutate(pop = ifelse(pop %in% names(map), map[pop], pop))

# we should also remove seedmix and larmix
lengths_mod_rm <- lengths_mod_rename %>% filter(!pop %in% c("LARMIX","SEEDMIX"))

# check data
summary(lengths_mod_rm) # we have NAs, and a massive outlier
lengths_mod_rm_clean <- lengths_mod_rm[which(lengths_mod_rm$width < 200),]
lengths_mod_rm_clean0 <- lengths_mod_rm_clean[which(lengths_mod_rm_clean$length < 200),]

# some individuals died/went missing and have "0" in length/width - rm these
lengths_mod_rm_nodead <- lengths_mod_rm_clean0[!lengths_mod_rm_clean0$length == 0 | !(complete.cases(lengths_mod_rm_clean0)),]
sum(!(complete.cases(lengths_mod_rm_nodead)))
lengths_mod_rm_nodead[which(is.na(lengths_mod_rm_nodead$width)),]
lengths_mod_narm_nodead <- lengths_mod_rm_nodead[!(is.na(lengths_mod_rm_nodead$width)),]

# ensure correct coding of variables
lengths_mod_narm_nodead$bag_site <- as.factor(lengths_mod_narm_nodead$bag_site)
lengths_mod_narm_nodead$pop <- as.factor(lengths_mod_narm_nodead$pop)
lengths_mod_narm_nodead$t <- as.factor(lengths_mod_narm_nodead$t)
```

I will now use LMMs to look at differences in length by grouping.

```{r}
# analysis of fixed effects
length_lm <- lm(length ~ pop * bag_site * t, data = lengths_mod_narm_nodead)
Anova(length_lm, type = "III") # all significant
boxplot(residuals(length_lm) ~ lengths_mod_narm_nodead$bags_label)

# with interactions
# using REML
length_lmm <- lmer(length ~ bag_site * pop * t + (1|bags_label), data = lengths_mod_narm_nodead)
Anova(length_lmm, type = "III") # all signif

# start with full model, then subtract each interaction, then subtract two of the interaction
drop1(length_lmm) # keep all

# one more model on final time point only
length_lmm_final <- lmer(length ~ bag_site * pop + (1|bags_label), data = lengths_mod_narm_nodead[lengths_mod_narm_nodead$t == 4,])
plot(length_lmm_final)
Anova(length_lmm_final, type = "III", test.statistic = "F") # all signif
drop1(length_lmm_final)

# can paste pops with sites to do a traditional Tukey
lengths_final_time <- lengths_mod_narm_nodead[lengths_mod_narm_nodead$t == 4,]
lengths_final_time$popsite <- paste0(lengths_final_time$pop, "_", lengths_final_time$bag_site)

# rename pops
lengths_final_time_rename <- lengths_final_time
(aov_len_t4 <- aov(lm(length ~ popsite, data = lengths_final_time_rename)))
plot(aov_len_t4) # looks good!
(tukey_len_t4 <- TukeyHSD(aov_len_t4, conf.level = 0.95))

# Now use multcompLetters
library(multcompView)
anova_len_t4 <- (Anova(lm(length ~ popsite, data = lengths_final_time_rename)))
letters_len_t4 <- multcompLetters4(aov_len_t4, tukey_len_t4)
letters_len_t4_df <- data.frame(letters = as.data.frame.list(letters_len_t4$popsite)$Letters, popsite = rownames(as.data.frame.list(letters_len_t4$popsite)))

lengths_final_time_rename_mg <- merge(lengths_final_time_rename, letters_len_t4_df, by = "popsite")
(lengths_final_time_rename_mg)
lengths_final_time_rename_mg$site_name <- lengths_final_time_rename_mg$pop
lengths_final_time_rename_mg$pop <- str_split_fixed(lengths_final_time_rename_mg$pop, "_", 2)[,2]

# need sd for letters
summary_len_t4 <- lengths_final_time_rename_mg %>%
  dplyr::summarise(mean_len = mean(length), sd = sd(length)) %>%
  dplyr::arrange()

# save all this 
sink(paste0("results/exp_results/length_model_results_", Sys.Date(), ".txt"))
Anova(length_lmm, type = "III")
Anova(length_lmm_final, type = "III")
anova_len_t4
letters_len_t4
sink()

# plot
len_t4_plot_site <- ggplot(lengths_final_time_rename_mg, aes(x = pop, y = length, fill = bag_site)) + 
  geom_boxplot(position=position_dodge(1)) +
  ggtitle("Length at Final Time Point") +
  geom_text(aes(label = letters, y = 127), vjust = -0.5, position=position_dodge(1)) +
  theme_classic()
len_t4_plot_site

len_final_time_plotting <- merge(lengths_final_time_rename_mg, popsdf, by = c("pop","site_name"))
len_final_time_plotting$bag_site <- as.factor(len_final_time_plotting$bag_site)
len_final_time_plotting$site_name <- as.factor(len_final_time_plotting$site_name)
len_final_time_plotting$letters <- as.factor(len_final_time_plotting$letters)
len_final_time_plotting$t <- as.factor(len_final_time_plotting$t)
len_final_time_plotting$pop.x <- NULL
names(len_final_time_plotting)[names(len_final_time_plotting) == "pop.y"] <- "pop"

# y-axis location for letters
letters_loc_len <- len_final_time_plotting %>% group_by(popsite) %>% dplyr::summarise(across(length, max))
colnames(letters_loc_len) <- c("popsite", "max_len")
len_final_time_plotting_letters <- merge(len_final_time_plotting, letters_loc_len, by = "popsite")

# plot by pop
len_t4_plot_pop <- ggplot(len_final_time_plotting_letters, aes(x = site_name_plotting, y = length, fill = fct_reorder(site_name, order))) + 
  geom_boxplot(position=position_dodge(1.0)) +  
  facet_grid(.~ bag_site,
             switch = "x") +
  scale_fill_manual(name = "Broodstock Group", label = popsdf$site_name_plotting, values = popsdf$cols)+
  geom_text(aes(label = letters, y = max_len + 2), vjust = -0.5, position = position_dodge(1)) +
  labs(title = "Final Length at Lewisetta and York River", x = "Experimental Site", y = "Length (mm)") +
  theme_classic()+ 
  theme(strip.background = element_blank(), 
        strip.text.x = element_text(), 
        strip.placement = "outside",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.spacing.x = unit(3, "lines"))
len_t4_plot_pop
```


## Modeling differences in Condition

Now I'll try modeling differences in condition I'll start by reading in and processing the condition data

```{r}
# trim data
head(condition)
cond_mod <- condition[,c("bag_site","site_name","bags_label","finalTissue_VialLabel","gonad","condition")]
colnames(cond_mod) <- c("bag_site","pop","bags_label","individual","gonad","condition")

# check data
summary(cond_mod) # we have NAs, and a massive outlier
table(cond_mod[cond_mod$bag_site == "Lewisetta",]$pop) # note NH & ME have ~1/2 n of other pops
table(cond_mod[cond_mod$bag_site == "YorkRiver",]$pop) # note NH & ME have ~1/4 n of other pops

# ensure correct coding of variables
cond_mod$bag_site <- as.factor(cond_mod$bag_site)
cond_mod$pop <- as.factor(cond_mod$pop)
cond_mod$condition <- as.ordered(cond_mod$condition)
cond_mod$gonad <- as.ordered(cond_mod$gonad)
```

I will now use ordinal logistic regression to look at differences in tissue condition by grouping.

```{r}
# quick distribution checks
hist(as.numeric(cond_mod[cond_mod$bag_site == "Lewisetta",]$condition), main = "Condition Lewisetta")
hist(as.numeric(cond_mod[cond_mod$bag_site == "YorkRiver",]$condition), main = "Condition York River")
hist(as.numeric(cond_mod$condition), main = "Condition Distribution")

# categories
levels(as.ordered(cond_mod$condition))

# here I will use ordinal logistic regression
cond_mod_olr <- polr(condition ~ pop * bag_site, data = cond_mod, Hess = TRUE)
summary(cond_mod_olr)
Anova(cond_mod_olr, type = "II", test.statistic = "Chisq") # all significant

# also paste pop/site
cond_mod$pop_site <- paste0(str_split_fixed(cond_mod$pop, "_", 2)[,2], "_", cond_mod$bag_site)
cond_mod_olr_popsite <- polr(condition ~ pop_site, data = cond_mod, Hess = TRUE)
summary(cond_mod_olr_popsite)

# significance letters
cond_emm <- emmeans(cond_mod_olr_popsite, ~ pop_site)
cond_letters <- cld(cond_emm, Letters = letters)
letters_df <- data.frame(letters = cond_letters$.group, pop_site = cond_letters$pop_site)
letters_df$pop <- str_split_fixed(letters_df$pop_site, "_", 2)[,1]
letters_df$bag_site <- str_split_fixed(letters_df$pop_site, "_", 2)[,2]
letters_df$site_name <- letters_df$pop
letters_df_rename <- letters_df %>% mutate(site_name = ifelse(site_name %in% names(map), map[site_name], site_name))

# also change letters to match a being best and f worst
letters_df_rename2 <- as.character(chartr("abcdef", "fedcba", letters_df_rename$letters))
letters_df_rename3 <- trimws(letters_df_rename2)
letters_df_rename$letters_rev <- letters_df_rename3

# save everything
sink(paste0("results/exp_results/condition_model_results_", Sys.Date(), ".txt"))
summary(cond_mod_olr)
Anova(cond_mod_olr, type = "II", test.statistic = "Chisq") # all significant
letters_df_rename
sink()

# full plotting df
cond_mod_plotting <- merge(cond_mod[,-2], letters_df_rename, by = c("bag_site","pop_site"))
cond_mod_plotting_pops <- merge(cond_mod_plotting, popsdf, by = c("site_name", "pop"))

# y-axis location for letters
letters_loc_cond <- cond_mod_plotting_pops %>% group_by(pop_site) %>% dplyr::summarise(across(condition, max))
colnames(letters_loc_cond) <- c("pop_site", "max_cond")
cond_plotting_letters <- merge(cond_mod_plotting_pops, letters_loc_cond, by = "pop_site")

# plot by pop - something is going wrong here
cond_plot_pop <- ggplot(cond_plotting_letters, aes(x = site_name_plotting, y = as.numeric(condition), fill = fct_reorder(site_name, order))) + 
  geom_boxplot(position=position_dodge(1.0)) +  
  facet_grid(.~ bag_site,
             switch = "x") +
  scale_fill_manual(name = "Broodstock Group", label = popsdf$site_name_plotting, values = popsdf$cols)+
  geom_text(aes(label = letters_rev, y = as.numeric(max_cond) + 0.1), vjust = -0.5, position = position_dodge(1)) +
  labs(title = "Final Condition at Lewisetta and York River", x = "Experimental Site", y = "Tissue Condition (1-5)") +
  theme_classic()+ 
  theme(strip.background = element_blank(), 
        strip.text.x = element_text(), 
        strip.placement = "outside",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.spacing.x = unit(3, "lines"))
cond_plot_pop
```

Now for gonad index

```{r}
# quick distributions checks
hist(as.numeric(cond_mod[cond_mod$bag_site == "Lewisetta",]$gonad), main = "Gonad Lewisetta")
hist(as.numeric(cond_mod[cond_mod$bag_site == "YorkRiver",]$gonad), main = "Gonad York River")
hist(as.numeric(cond_mod$gonad), main = "Gonad Distribution")

# categories
levels(as.ordered(cond_mod$gonad))

# here I will use ordinal logistic regression
gonad_mod_olr <- polr(gonad ~ pop * bag_site, data = cond_mod, Hess = TRUE)
summary(gonad_mod_olr)
Anova(gonad_mod_olr, type = "II", test.statistic = "Chisq") # all significant

# also pop/site
gonad_mod_olr_popsite <- polr(gonad ~ pop_site, data = cond_mod, Hess = TRUE)
summary(gonad_mod_olr_popsite)

# significance letters
gonad_emm <- emmeans(gonad_mod_olr_popsite, ~ pop_site)
gonad_letters <- cld(gonad_emm, Letters = letters) # only gives groupings for three of these groups
letters_df <- data.frame(letters = gonad_letters$.group, pop_site = gonad_letters$pop_site)
letters_df$pop <- str_split_fixed(letters_df$pop_site, "_", 2)[,1]
letters_df$bag_site <- str_split_fixed(letters_df$pop_site, "_", 2)[,2]
letters_df$site_name <- letters_df$pop
letters_df_rename <- letters_df %>% mutate(site_name = ifelse(site_name %in% names(map), map[site_name], site_name))

# also change letters to match a being best and f worst
letters_df_rename2 <- as.character(chartr("abcdef", "fedcba", letters_df_rename$letters))
letters_df_rename3 <- trimws(letters_df_rename2)
letters_df_rename$letters_rev <- letters_df_rename3

# save everything
sink(paste0("results/exp_results/gonad_model_results_", Sys.Date(), ".txt"))
summary(gonad_mod_olr)
Anova(gonad_mod_olr, type = "II", test.statistic = "Chisq") # all significant
letters_df_rename
sink()

# full plotting df
gonad_mod_plotting <- merge(cond_mod[,-2], letters_df_rename, by = c("bag_site","pop_site"))
gonad_mod_plotting_pops <- merge(gonad_mod_plotting, popsdf, by = c("site_name", "pop"))

# y-axis location for letters
letters_loc_gonad <- gonad_mod_plotting_pops %>% group_by(pop_site) %>% dplyr::summarise(across(gonad, max))
colnames(letters_loc_gonad) <- c("pop_site", "max_gonad")
gonad_plotting_letters <- merge(gonad_mod_plotting_pops, letters_loc_gonad, by = "pop_site")

# plot by pop
gonad_plot_pop <- ggplot(gonad_plotting_letters, aes(x = site_name_plotting, y = as.numeric(gonad), fill = fct_reorder(site_name, order))) + 
  geom_boxplot(position=position_dodge(1.0)) +  
  facet_grid(.~ bag_site,
             switch = "x") +
  scale_fill_manual(name = "Broodstock Group", label = popsdf$site_name_plotting, values = popsdf$cols)+
  geom_text(aes(label = letters_rev, y = as.numeric(max_gonad) + 0.1), vjust = -0.5, position = position_dodge(1)) +
  labs(title = "Final Gonad Index at Lewisetta and York River", x = "Experimental Site", y = "Gonad Index (1-5)") +
  theme_classic() + 
  theme(strip.background = element_blank(), 
        strip.text.x = element_text(), 
        strip.placement = "outside",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.spacing.x = unit(3, "lines"))
gonad_plot_pop
```


## Export Figures

```{r}
doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_t4_plot_pop), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = len_t4_plot_pop), location = ph_location_type(type = "body"))
print(doc, target = "results/figures/Final_Experimental_Fitness.pptx")
```
