---
title: "len_surv_modeling_ooters"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/projects/lotterhos/MVP_oyster_GO")

# load libraries
library(tidyverse) # for data wrangling
library(stringr) # for splitting strings
library(lubridate) # for dates
library(ggplot2) # for plotting
library(plotrix) # for std error
library(lme4) # for lmm
library(performance) # for lmm
library(lmerTest) # for lmm
library(car) # for avPlots(), Anova()

# data
surv_long <- read.csv("results/exp_results/mortality_long_bags_052925.csv")[,-1]
bags <- read.csv("data/ExpDat/MVP23-FieldBags - bags.csv")
lengths <- read.csv("results/exp_results/lengths_long_2025-06-03.csv")[,-1]
```

This script is for modeling differences in survival and length amongst bags/populations/sites/time points in the MVP Field Experiment. 

I intend to use linear (mixed) models to examine how survival and length vary by group and how genetics and environmental history drive these differences in performance. 

This script will also be split at some point to separate some of the modeling.

## Modeling differences in survival

First I'll process survival data:

```{r}
# split bags to have more detailed bag data
bags[,5:6] <- str_split_fixed(bags$bags_label, "-", 3)[,2:3]
colnames(bags) <- c("bags_key","bag_site","bag","SpawnTrt_Key","pop","bagnum")

# combine data
surv_for_mod <- merge(surv_long, bags[,c("bag","bag_site")], by = "bag")
surv_for_mod$pop <- str_split_fixed(surv_for_mod$bag, "-", 3)[,2]
surv_mod_ord <- surv_for_mod[,c("bag_site","pop","bag","event","surv")]

# rename populations
map=c("TX"="W1_TX","LA"="W2_LA","FL"="W3_FL","LOLA"="S1_LOLA","DEBY"="S2_DEBY","JR"="W4_VA","NH"="W5_NH","ME"="W6_ME")
surv_mod_ord_rename <- surv_mod_ord %>% mutate(pop = ifelse(pop %in% names(map), map[pop], pop))

# remove any NAs
surv_mod_ord_rename_nona <- na.omit(surv_mod_ord_rename)

# we should also remove seedmix and larmix for now
surv_mod_ord_rm <- surv_mod_ord_rename_nona %>% filter(!pop %in% c("LARMIX","SEEDMIX"))

colnames(surv_mod_ord_rm) <- c("bag_site","pop","bags_label","t","surv")
```


Now I can use LMMs to look at differences in survival by group.

```{r}
# analysis of fixed effects
anova(lm(surv ~ pop + bag_site + t, data = surv_mod_ord_rm)) # all significant

# ensure correct coding of explanatory variables
surv_mod_ord_rm$t <- as.factor(surv_mod_ord_rm$t)

# start with full model, then subtract each interaction, then subtract two of the interaction
surv_lmm <- lmer(surv ~ bag_site * pop * t + (1|bags_label), data = surv_mod_ord_rm)
plot(surv_lmm)
Anova(surv_lmm, type = "III")
drop1(surv_lmm) # keep all
summary(surv_lmm)

# one more model on final time point only
surv_lm <- lm(surv ~ bag_site * pop, data = surv_mod_ord_rm[surv_mod_ord_rm$t == 4,]) # no random effect here because there is only one observation per bag label
plot(surv_lm)
Anova(surv_lm, type = "III") # make sure to always use this method if two categorical predictors and  interaction and unbalanced design ; type III may have to set the contrast, check biostats notes
TukeyHSD(aov(surv_lm))

# plot survival at last time point
surv_t4 <- surv_mod_ord_rm[surv_mod_ord_rm$t == 4,]

surv_t4_plot_site <- ggplot(surv_t4, aes(x = pop, y = surv, fill = bag_site)) + 
  geom_boxplot(position=position_dodge(0.8)) +
  ggtitle("Survival at Final Time Point") +
  theme_classic()
surv_t4_plot_site

surv_t4_plot_pop <- ggplot(surv_t4, aes(x = bag_site, y = surv, fill = pop)) + 
  geom_boxplot(position=position_dodge(1.0)) +
  theme_classic()
surv_t4_plot_pop

#extract significant letters from TK test and store
# add a different pop column that's split
surv_t4$oldpop <- str_split_fixed(surv_t4$pop,"_",2)[,2]

# now redo model
surv_lm2 <- lm(surv ~ bag_site * oldpop, data = surv_t4)
plot(surv_lm2)
Anova(surv_lm2, type = "III") # make sure to always use this method if two categorical predictors and interaction and unbalanced design ; type III may have to set the contrast, check biostats notes
library(multcompView)
tuk_surv_t4 <- TukeyHSD(aov(surv_lm2), conf.level = 0.95)
cld_surv_t4 <- multcompLetters4(aov(surv_lm2), tuk_surv_t4)

# can paste pops with sites to line up with letters output
surv_t4$popsite <- paste0(surv_t4$bag_site,":",str_split_fixed(surv_t4$pop,"_",2)[,2])

# multcomp letters to get letters
letters_surv_t4_df <- data.frame(letters = as.data.frame.list(cld_surv_t4$`bag_site:oldpop`), popsite = names((cld_surv_t4$`bag_site:oldpop`)$Letters))
letters_surv_t4_df_red <- letters_surv_t4_df[,c("letters.Letters","popsite")]
colnames(letters_surv_t4_df_red) <- c("letters","popsite")

# save these outputs too
sink(paste0("results/exp_results/survival_model_results_", Sys.Date(), ".txt"))
Anova(surv_lmm, type = "III")
summary(surv_lmm)
Anova(surv_lm, type = "III")
summary(surv_lm)
Anova(surv_lm2, type = "III")
summary(surv_lm2)
letters_surv_t4_df_red
sink()

# merge
surv_final_time_rename_mg <- merge(surv_t4, letters_surv_t4_df_red, by = "popsite")
(surv_final_time_rename_mg)
surv_final_time_rename_mg$site_name <- surv_final_time_rename_mg$pop

# pop color scheme
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "LOLA", "DEBY", "NH", "ME"), 
                     order = 1:8, 
                     cols = c("#365C8DFF", "#277F8EFF", "#1FA187FF", "#4AC16DFF", "#440154FF", "#46337EFF", "#9FDA3AFF", "#FDE725FF"),
                     shape = c(21,21,21,21,23,23,21,21),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME"))

surv_final_time_plotting <- merge(surv_final_time_rename_mg, popsdf, by = "site_name")

# plot by pop
surv_t4_plot_pop <- ggplot(surv_final_time_plotting, aes(x = bag_site, y = surv, fill = fct_reorder(site_name, order))) + 
  geom_boxplot(position=position_dodge(1.0)) +  
  scale_fill_manual(name = "Broodstock Group", label = popsdf$site_name, values = popsdf$cols)+
  geom_text(aes(label = letters, y = 1), vjust = -0.5, position=position_dodge(1)) +
  labs(title = "Final Survival at Lewisetta and York River", x = "Experimental Site", y = "Survival") +
  theme_classic()
surv_t4_plot_pop
```


## Modeling differences in length

Now I'll try modeling differences in length. I'll start by reading in and processing the length data.

```{r}
# trim data
head(lengths)
lengths_mod <- lengths[,c("bag_site","pop","event","bags_label","tags_label","length","width")]
colnames(lengths_mod) <- c("bag_site","pop","t","bags_label","tags_label","length","width")

# rename pops
map=c("TX"="W1_TX","LA"="W2_LA","FL"="W3_FL","LOLA"="S1_LOLA","DEBY"="S2_DEBY",
      "JR"="W4_VA","NH"="W5_NH","ME"="W6_ME")
lengths_mod_rename <- lengths_mod %>% mutate(pop = ifelse(pop %in% names(map), map[pop], pop))

# we should also remove seedmix and larmix
lengths_mod_rm <- lengths_mod_rename %>% filter(!pop %in% c("LARMIX","SEEDMIX"))

# check data
summary(lengths_mod_rm) # we have NAs, and a massive outlier
lengths_mod_rm_clean <- lengths_mod_rm[which(lengths_mod_rm$width < 200),]
lengths_mod_rm_clean0 <- lengths_mod_rm_clean[which(lengths_mod_rm_clean$length < 200),]

# some individuals died/went missing and have "0" in length/width - rm these
lengths_mod_rm_nodead <- lengths_mod_rm_clean0[!lengths_mod_rm_clean0$length == 0 | !(complete.cases(lengths_mod_rm_clean0)),]
sum(!(complete.cases(lengths_mod_rm_nodead)))
lengths_mod_rm_nodead[which(is.na(lengths_mod_rm_nodead$width)),]
lengths_mod_narm_nodead <- lengths_mod_rm_nodead[!(is.na(lengths_mod_rm_nodead$width)),]

# ensure correct coding of variables
lengths_mod_narm_nodead$bag_site <- as.factor(lengths_mod_narm_nodead$bag_site)
lengths_mod_narm_nodead$pop <- as.factor(lengths_mod_narm_nodead$pop)
lengths_mod_narm_nodead$t <- as.factor(lengths_mod_narm_nodead$t)
```

I will now use LMMs to look at differences in length by grouping.

```{r}
# analysis of fixed effects
length_lm <- lm(length ~ pop * bag_site * t, data = lengths_mod_narm_nodead)
Anova(length_lm, type = "III") # all significant
boxplot(residuals(length_lm) ~ lengths_mod_narm_nodead$bags_label)

# with interactions
# using REML
length_lmm <- lmer(length ~ bag_site * pop * t + (1|bags_label), data = lengths_mod_narm_nodead)
Anova(length_lmm, type = "III") # all signif

# start with full model, then subtract each interaction, then subtract two of the interaction
drop1(length_lmm) # keep all

# one more model on final time point only
length_lmm_final <- lmer(length ~ bag_site * pop + (1|bags_label), data = lengths_mod_narm_nodead[lengths_mod_narm_nodead$t == 4,])
plot(length_lmm_final)
Anova(length_lmm_final, type = "III") # all signif
drop1(length_lmm_final)

# can paste pops with sites to do a traditional Tukey
lengths_final_time <- lengths_mod_narm_nodead[lengths_mod_narm_nodead$t == 4,]
lengths_final_time$popsite <- paste0(lengths_final_time$pop, "_", lengths_final_time$bag_site)

# rename pops
lengths_final_time_rename <- lengths_final_time
(aov_len_t4 <- aov(lm(length ~ popsite, data = lengths_final_time_rename)))
plot(aov_len_t4) # looks good!
(tukey_len_t4 <- TukeyHSD(aov_len_t4, conf.level = 0.95))

# Now use multcompLetters
library(multcompView)
anova_len_t4 <- (Anova(lm(length ~ popsite, data = lengths_final_time_rename)))
letters_len_t4 <- multcompLetters4(aov_len_t4, tukey_len_t4)
letters_len_t4_df <- data.frame(letters = as.data.frame.list(letters_len_t4$popsite)$Letters, popsite = rownames(as.data.frame.list(letters_len_t4$popsite)))

lengths_final_time_rename_mg <- merge(lengths_final_time_rename, letters_len_t4_df, by = "popsite")
(lengths_final_time_rename_mg)

# need sd for letters
summary_len_t4 <- lengths_final_time_rename_mg %>%
  summarise(mean_len = mean(length), sd = sd(length)) %>%
  arrange()

# save all this 
sink(paste0("results/exp_results/length_model_results_", Sys.Date(), ".txt"))
Anova(length_lmm, type = "III")
Anova(length_lmm_final, type = "III")
anova_len_t4
letters_len_t4
sink()

# plot
len_t4_plot_site <- ggplot(lengths_final_time_rename_mg, aes(x = pop, y = length, fill = bag_site)) + 
  geom_boxplot(position=position_dodge(1)) +
  ggtitle("Length at Final Time Point") +
  geom_text(aes(label = letters, y = 127), vjust = -0.5, position=position_dodge(1)) +
  theme_classic()
len_t4_plot_site

len_final_time_plotting <- merge(lengths_final_time_rename_mg, pop_colors_rm, by = "pop")

# plot by pop
len_t4_plot_pop <- ggplot(len_final_time_plotting, aes(x = bag_site, y = length, fill = fct_reorder(pop, order))) + 
  geom_boxplot(position=position_dodge(1.0)) +  
  scale_fill_manual(name = "Broodstock Group", label = pop_colors_rm$pop, values = pop_colors_rm$hex)+
  geom_text(aes(label = letters, y = 127), vjust = -0.5, position=position_dodge(1)) +
  labs(title = "Length at Final Time Point", x = "Experimental Site", y = "Survival") +
  theme_classic()
len_t4_plot_pop

#ggsave("results/figures/length_t4.pdf", plot = len_t4_plot_pop, width = 10, height = 6, dpi = 300)
```
