---
title: "Len_surv_GE_modeling"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# packages
library()

# data
surv_long <- read.csv("results/exp_results/mortality_long_bags_052925.csv")[,-1]
bags <- read.csv("data/ExpDat/MVP23-FieldBags - bags.csv")
```

This script is for modeling differences in survival and length amongst bags/populations/sites/time points in the MVP Field Experiment. 

I intend to use linear (mixed) models to examine how survival and length vary by group and how genetics and environmental history drive these differences in performance. 

This script will also be split at some point to separate some of the modeling.

## Modeling differences in survival

First I'll process survival data:

```{r}
# split bags to have more detailed bag data
bags[,5:6] <- str_split_fixed(bags$bags_label, "-", 3)[,2:3]
colnames(bags) <- c("bags_key","bag_site","bag","SpawnTrt_Key","pop","bagnum")

# combine data
surv_for_mod <- merge(surv_long, bags[,c("bag","bag_site")], by = "bag")
surv_for_mod$pop <- str_split_fixed(surv_for_mod$bag, "-", 3)[,2]
surv_mod_ord <- surv_for_mod[,c("bag_site","pop","bag","event","surv")]

# rename populations
map=c("TX"="W1-TX","LA"="W2-LA","FL"="W3-FL","LOLA"="S1-LOLA","DEBY"="S2-DEBY","JR"="W4-VA","NH"="W5-NH","ME"="W6-ME")
surv_mod_ord_rename <- surv_mod_ord %>% mutate(pop = ifelse(pop %in% names(map), map[pop], pop))

# remove any NAs
surv_mod_ord_rename_nona <- na.omit(surv_mod_ord_rename)

# we should also remove seedmix and larmix for now
surv_mod_ord_rm <- surv_mod_ord_rename_nona %>% filter(!pop %in% c("LARMIX","SEEDMIX"))

colnames(surv_mod_ord_rm) <- c("bag_site","pop","bags_label","t","surv")
```
## Modeling drivers of differences in survival

For the H2F paper, we want to know how environment and genetics are impacting differential survival. I'll start by putting together the complete dataframe.

```{r}
# start by putting together a complete dataframe
# genetic data
gen_parents <- read.csv("data/ExpDat/parent_genetic_effects.csv")
colnames(gen_parents) <- c("pop","ho","exp_hs","ar")
surv_mod_gen <- merge(surv_mod_ord_rm, gen_parents, by = "pop")

# environmental data
env_parents <- read.csv("data/EnvDat/exp/envr_summary_stats.csv")[,-1]
colnames(env_parents) <- c("pop","mean_temp","max_temp","min_temp","sd_temp","mean_sal","max_sal","min_sal","sd_sal")
surv_mod_gen_env <- merge(surv_mod_gen, env_parents, by = c("pop"))
```

Now look at how survival changes with genetics/environment at each time point. These are multiple regressions, so remove interactions here - they are not interpretable in the same way.

```{r}
# full model time point 1
# start with lew
surv_GE1.lew <- lm(surv ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "Lewisetta",])

surv_GE1.lew
#plot(surv_GE1.lew) 
summary(surv_GE1.lew) # none signif
drop1(surv_GE1.lew) # get rid of ho

surv_GE1.0.lew <- lm(surv ~ ar + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE1.0.lew) 
summary(surv_GE1.0.lew) # none signif
drop1(surv_GE1.0.lew) # get rid of ar

surv_GE1.1.lew <- lm(surv ~ max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE1.1.lew)
summary(surv_GE1.1.lew) # none signif
drop1(surv_GE1.1.lew) # get rid of min_sal

surv_GE1.2.lew <- lm(surv ~ max_temp + min_temp + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE1.2.lew)
summary(surv_GE1.2.lew) # min temp, max_sal signif
drop1(surv_GE1.2.lew) # get rid of max_temp

surv_GE1.3.lew <- lm(surv ~ min_temp + max_sal , data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE1.3.lew)
summary(surv_GE1.3.lew) # all signif
drop1(surv_GE1.3.lew) # this is final model

# now we go to york
surv_GE1.york <- lm(surv ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE1.york)
summary(surv_GE1.york) # min_sal signif
drop1(surv_GE1.york) # drop ar

surv_GE1.0.york <- lm(surv ~ ho + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE1.0.york)
summary(surv_GE1.0.york) # min_sal signif
drop1(surv_GE1.0.york) # drop max_temp

surv_GE1.1.york <- lm(surv ~ ho + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE1.1.york)
summary(surv_GE1.1.york) # min_sal, min_temp signif
drop1(surv_GE1.1.york) # drop max_sal

surv_GE1.2.york <- lm(surv ~ ho + min_temp + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE1.2.york)
summary(surv_GE1.2.york) # all signif
drop1(surv_GE1.2.york) # final model

# lew t2
surv_GE2.lew <- lm(surv ~ ho + ar + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 2 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE2.lew)
summary(surv_GE2.lew) # none signif
drop1(surv_GE2.lew) # get rid of min_sal

surv_GE2.0.lew <- lm(surv ~ ar + ho + min_temp + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 2 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE2.0.lew)
summary(surv_GE2.0.lew) # none signif
drop1(surv_GE2.0.lew) # get rid of ar

surv_GE2.1.lew <- lm(surv ~ ho + min_temp + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 2 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE2.1.lew)
summary(surv_GE2.1.lew) # min_temp, max_sal signif
drop1(surv_GE2.1.lew) # final model

# york t2
surv_GE2.york <- lm(surv ~ ho + ar + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 2 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE2.york) # qq still ugly
summary(surv_GE2.york) # min_temp, min_sal signif
drop1(surv_GE2.york) # get rid of ho

surv_GE2.0.york <- lm(surv ~ ar + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 2 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE2.0.york) 
summary(surv_GE2.0.york) # min_temp, max_sal, min_sal signif
drop1(surv_GE2.0.york) # final model

# lew t3
surv_GE3.lew <- lm(surv ~ ho + ar + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 3 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE3.lew)
summary(surv_GE3.lew) # min_temp, max_sal signif
drop1(surv_GE3.lew) # get rid of min_sal

surv_GE3.0.lew <- lm(surv ~ ho + ar + min_temp + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 3 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE3.0.lew)
summary(surv_GE3.0.lew) # ar, min_temp, max_sal signif
drop1(surv_GE3.0.lew) # get rid of ho

surv_GE3.1.lew <- lm(surv ~ ar + min_temp + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 3 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE3.1.lew)
summary(surv_GE3.1.lew) # ar, min_temp, max_sal signif
drop1(surv_GE3.1.lew) # final model

# york t3
surv_GE3.york <- lm(surv ~ ar + ho + min_temp + max_temp + min_sal + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 3 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE3.york)
summary(surv_GE3.york) # ar, max_temp, min_sal, intercept signif
drop1(surv_GE3.york) # get rid of min_temp

surv_GE3.0.york <- lm(surv ~ ar + ho + max_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 3 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE3.0.york)
summary(surv_GE3.0.york) # ar, ho, min_sal, max_temp, max_sal, intercept signif
drop1(surv_GE3.0.york) # final model

# lew t4
surv_GE4.lew <- lm(surv ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 4 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE4.lew)
summary(surv_GE4.lew) # max_temp, min_temp, max_sal signif
drop1(surv_GE4.lew) # get rid of ar

surv_GE4.0.lew <- lm(surv ~ ho + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 4 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE4.0.lew)
summary(surv_GE4.0.lew) # max_temp, min_temp, max_sal signif
drop1(surv_GE4.0.lew) # this is final model

# york t4
surv_GE4.york <- lm(surv ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = c[surv_mod_gen_env$t == 4 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE4.york)
summary(surv_GE4.york) # intercept, ho, ar, max_temp, min_temp, min_sal signif
drop1(surv_GE4.york) # this is final model

sink(paste0("results/exp_results/survival_GE_model_results_",Sys.Date(),".txt"))
summary(surv_GE1.3.lew)
summary(surv_GE1.2.york)
summary(surv_GE2.1.lew)
summary(surv_GE2.0.york)
summary(surv_GE3.1.lew)
summary(surv_GE3.0.york)
summary(surv_GE4.0.lew)
summary(surv_GE4.york)
sink()
```


## Modeling drivers of differences in length


```{r}
library(car)

# start by putting together a complete dataframe
# genetic data
len_mod_gen <- merge(lengths_mod_rm_nodead, gen_parents, by = "pop")
len_mod_gen_env <- merge(len_mod_gen, env_parents, by = c("pop"))
(len_mod_gen_env)

# fixed effects
(anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "Lewisetta",]))) # all signif
(anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "Lewisetta",]))) # ho, min_temp, max_sal, min_sal signif
(anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "Lewisetta",]))) # ho, max_temp, max_sal, min_sal signif
(anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "Lewisetta",]))) # ho, max_temp, max_sal, min_sal signif

anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "YorkRiver",])) # all but ho & ar signif
anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "YorkRiver",])) # ar, min_temp, max_sal, min_sal signif
anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "YorkRiver",])) # ho, min_temp, max_sal, min_sal signif
anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "YorkRiver",])) # ho, ar, max_sal, min_sal min_temp signif

# full model time point 1 - lew
len_GE1.lew <- lmerTest::lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "Lewisetta",]) 
len_GE1.lew
plot(len_GE1.lew) 
Anova(len_GE1.lew, type = "III") # max_temp, min_temp, min_sal signif
summary(len_GE1.lew) # gives just min_temp & min_sal
step(len_GE1.lew) # trying different model selection method

len_GE1.0.lew <- lmerTest::lmer(length ~ max_temp + min_temp + max_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "Lewisetta",]) 
plot(len_GE1.0.lew) 
Anova(len_GE1.0.lew, type = "III") # all signif
summary(len_GE1.0.lew)
step(len_GE1.0.lew) # final model

# york t1
len_GE1.york <- lmerTest::lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "YorkRiver",]) 
len_GE1.york
plot(len_GE1.york) 
Anova(len_GE1.york, type = "III") # max_temp, min_temp, min_sal signif
summary(len_GE1.york) # safe signif
step(len_GE1.york) # get rid of ho and ar

len_GE1.0.york <- lmerTest::lmer(length ~ max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "YorkRiver",]) 
plot(len_GE1.0.york) 
Anova(len_GE1.0.york, type = "III") # all signif
summary(len_GE1.0.york)
step(len_GE1.0.york) # final model

# full model time point 2
# lew t2
len_GE2.lew <- lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE2.lew
plot(len_GE2.lew)
Anova(len_GE2.lew, type = "III") # min sal
step(len_GE2.lew) # drop min_temp, max_sal, ar, max_temp

len_GE2.0.lew <- lmer(length ~ ho + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE2.0.lew
plot(len_GE2.0.lew)
Anova(len_GE2.0.lew, type = "III") # all signif
step(len_GE2.0.lew) # final model

# york t2
len_GE2.york <- lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE2.york
plot(len_GE2.york)
Anova(len_GE2.york, type = "III") # max_temp, min_temp, min_sal signif
step(len_GE2.york) # drop max_sal, ar

len_GE2.0.york <- lmer(length ~ ho + max_temp + min_temp + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE2.0.york
plot(len_GE2.0.york)
Anova(len_GE2.0.york, type = "III") # all signif except intercept
step(len_GE2.0.york) # final model

# full model time point 3
# lew t3
len_GE3.lew <- lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE3.lew
plot(len_GE3.lew)
Anova(len_GE3.lew, type = "III") # intercept, min_temp, max_sal signif
step(len_GE3.lew) # drop everything except min_temp, max_sal

len_GE3.0.lew <- lmer(length ~ min_temp + max_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE3.0.lew
plot(len_GE3.0.lew)
Anova(len_GE3.0.lew, type = "III") # all signif
step(len_GE3.0.lew) # final model

# york t3
len_GE3.york <- lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE3.york
plot(len_GE3.york)
Anova(len_GE3.york, type = "III") # intercept, max_temp, max_sal, min_sal signif
step(len_GE3.york) # drop ho, min_temp

len_GE3.0.york <- lmer(length ~ ar + max_temp + min_sal + max_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE3.0.york
plot(len_GE3.0.york)
Anova(len_GE3.0.york, type = "III") # all signif except ar
step(len_GE3.0.york) # final model

# lew t4
len_GE4.lew <- lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE4.lew
plot(len_GE4.lew)
Anova(len_GE4.lew, type = "III") # intercept, min_temp, max_sal signif
step(len_GE4.lew) # drop ho, max_temp, ar

len_GE4.0.lew <- lmer(length ~ min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE4.0.lew
plot(len_GE4.0.lew)
Anova(len_GE4.0.lew, type = "III") # intercept, min_temp, max_sal, min_sal signif
step(len_GE4.0.lew) # this is final model

# york t4
len_GE4.york <- lmer(length ~ ar + ho + min_temp + max_temp + min_sal + max_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE4.york
plot(len_GE4.york)
Anova(len_GE4.york, type = "III") # all signif except ho
step(len_GE4.york) # drop ho, min_temp

len_GE4.0.york <- lmer(length ~ ar + max_temp + min_sal + max_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE4.0.york
plot(len_GE4.0.york)
Anova(len_GE4.0.york, type = "III") # all signif
step(len_GE4.0.york) # final model

sink(paste0("results/exp_results/len_GE_model_results_", Sys.Date(), ".txt"))
Anova(len_GE1.0.lew, type = "III")
Anova(len_GE1.0.york, type = "III")
Anova(len_GE2.0.lew, type = "III")
Anova(len_GE2.0.york, type = "III")
Anova(len_GE3.0.lew, type = "III")
Anova(len_GE3.0.york, type = "III")
Anova(len_GE4.lew, type = "III")
Anova(len_GE4.0.york, type = "III")
sink()
```