---
title: "Oyster_len_calc"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/projects/lotterhos/MVP_oyster_GO")

# load libraries
library(tidyverse) # for data wrangling
library(stringr) # for splitting strings
library(lubridate) # for dates
library(ggplot2) # for plotting
library(plotrix) # for std error
```

## Calculating Length/growth of Oysters in Field

This script is to analyze length of oysters in the MVP Field Experiment from 2023-2025. First I need to read in and merge together relevant data. I will then clean this data up to prepare it for analysis by time point, site, population, and more.

```{r}
bags <- read.csv("data/expDat/MVP23-FieldBags - bags.csv")
expsites <- read.csv("data/expDat/MVP23-FieldBags - sites.csv")
spawntrt <- read.csv("data/expDat/MVP23-FieldBags - spawn_trt2.csv")
tags <- read.csv("data/expDat/MVP23-FieldBags - tags.csv")
tagleng <- read.csv("data/expDat/MVP23-FieldBags - tagsLength_new.csv")

# split bags to have more detailed bag data
bags[,5:6] <- str_split_fixed(bags$bags_label, "-", 3)[,2:3]
colnames(bags) <- c("bags_key","bag_site","bags_label","SpawnTrt_Key","pop","bagnum")

# and remove practice bags
bags_rm <- bags[-(61:62),]

# also put together tagleng data
tagsbags <- merge(tags, bags_rm, by = c("bags_key"), all = T)
tagsbags1 <- merge(tagsbags, tagleng, by = c("tags_key"), all = T)

# trim dataset
tagsbags2 <- tagsbags1[,c("tags_label","tags_timestamp","bag_site","bags_label",
                          "pop","bagnum","length","width","tagsLength_timestamp")]

# split timestamps to get dates
# split timestamp column into two
tagsbags2[,c("tagsLength_timestamp")]
datetime_split_tag <- str_split_fixed(tagsbags2[,c("tagsLength_timestamp")], " ", 2) # split timestamp
colnames(datetime_split_tag) <- c("tagsLength_date", "tagsLength_time") # name columns
tagsbags3 <- cbind(tagsbags2, datetime_split_tag) # add to full dataset

# use dates to split by monitoring event
# dates are 11/7/23 - 11/15/23, 4/29/24 - 5/7/24, 10/21/24-11/12/24, 5/20/25-5/26/25
tagsbags3$tagsLength_date <- mdy(tagsbags3$tagsLength_date)

# remove practice
tagsbags4 <- tagsbags3[!(tagsbags3$tagsLength_date == "2024-03-13"),]
tagsbags5 <- tagsbags4[!(is.na(tagsbags4$tagsLength_date)),]

tagsbags5$event <- ifelse(as_date(tagsbags5$tagsLength_date) > as_date("2023-11-16"), ifelse(as_date(tagsbags5$tagsLength_date) > as_date("2024-05-09"), ifelse(as_date(tagsbags5$tagsLength_date) > "2024-11-12", 4, 3), 2), 1)

tagsbags6 <- tagsbags5[!(tagsbags5$pop == ""),]
tagsbags6 <- tagsbags6[!(tagsbags6$pop == "LEW"),]
tagsbags6 <- tagsbags6[!(tagsbags6$pop == "YORK"),]

write.csv(tagsbags6, paste0("results/exp_results/lengths_long_", Sys.Date(), ".csv"))
```

Now that I have prepared the data, I will look at some larger data trends. 

```{r}
###########
# calc bag level averages for each event
event1_bags <- tapply(tagsbags6[tagsbags6$event == 1,]$length, 
                      tagsbags6[tagsbags6$event == 1,]$bags_label,
                      mean)
event2_bags <- tapply(tagsbags6[tagsbags6$event == 2,]$length,
                      tagsbags6[tagsbags6$event == 2,]$bags_label,
                      mean)
event3_bags <- tapply(tagsbags6[tagsbags6$event == 3,]$length,
                      tagsbags6[tagsbags6$event == 3,]$bags_label,
                      mean)
event4_bags <- tapply(tagsbags6[tagsbags6$event == 4,]$length,
                      tagsbags6[tagsbags6$event == 4,]$bags_label,
                      mean)

# get sites and pops
popsite <- as.data.frame(cbind(rownames(event1_bags), str_split_fixed(rownames(event1_bags), "-", 3)[,2:3]))
colnames(popsite) <- c("bag","pop","bagnum")
popsite$site <- ifelse(popsite$bagnum < 4, "Lewisetta", "YorkRiver")

# put this in data frame
event1_bags_df <- data.frame(bag = rownames(event1_bags), 
                             monitoring_event = 1, 
                             length = event1_bags)
event1_bags_df2 <- merge(event1_bags_df, popsite, by = "bag")
rownames(event1_bags_df) <- NULL
event2_bags_df <- data.frame(bag = rownames(event2_bags), 
                             monitoring_event = 2, 
                             length = event2_bags)
event2_bags_df2 <- merge(event2_bags_df, popsite, by = "bag")
rownames(event2_bags_df) <- NULL
event3_bags_df <- data.frame(bag = rownames(event3_bags), 
                             monitoring_event = 3, 
                             length = event3_bags)
event3_bags_df2 <- merge(event3_bags_df, popsite, by = "bag")
rownames(event3_bags_df) <- NULL
event4_bags_df <- data.frame(bag = rownames(event4_bags), 
                             monitoring_event = 4, 
                             length = event4_bags)
event4_bags_df2 <- merge(event4_bags_df, popsite, by = "bag")
rownames(event4_bags_df) <- NULL

# put it together
lengths_bags_df <- rbind(event1_bags_df, event2_bags_df, event3_bags_df, event4_bags_df)
lengths_bags_df

# save it
write.csv(lengths_bags_df, paste0("results/exp_results/lengths_bags_event_", Sys.Date(), ".csv"))
```


```{r}
# calc pop level averages for each event
event1_york <- tapply(tagsbags6[tagsbags6$event == 1 & tagsbags6$bag_site == "YorkRiver",]$length, 
                      tagsbags6[tagsbags6$event == 1 & tagsbags6$bag_site == "YorkRiver",]$pop, 
                      mean)
event1_lew <- tapply(tagsbags6[tagsbags6$event == 1 & tagsbags6$bag_site == "Lewisetta",]$length, 
                     tagsbags6[tagsbags6$event == 1 & tagsbags6$bag_site == "Lewisetta",]$pop, 
                     mean)
event1_tot <- tapply(tagsbags6[tagsbags6$event == 1,]$length, 
                     tagsbags6[tagsbags6$event == 1,]$pop, 
                     mean)
event2_york <- tapply(tagsbags6[tagsbags6$event == 2 & tagsbags6$bag_site == "YorkRiver",]$length, 
                      tagsbags6[tagsbags6$event == 2 & tagsbags6$bag_site == "YorkRiver",]$pop, 
                      mean)
event2_lew <- tapply(tagsbags6[tagsbags6$event == 2 & tagsbags6$bag_site == "Lewisetta",]$length, 
                     tagsbags6[tagsbags6$event == 2 & tagsbags6$bag_site == "Lewisetta",]$pop, 
                     mean)
event2_tot <- tapply(tagsbags6[tagsbags6$event == 2,]$length, 
                     tagsbags6[tagsbags6$event == 2,]$pop, 
                     mean)
event3_york <- tapply(tagsbags6[tagsbags6$event == 3 & tagsbags6$bag_site == "YorkRiver",]$length, 
                      tagsbags6[tagsbags6$event == 3 & tagsbags6$bag_site == "YorkRiver",]$pop, 
                      mean)
event3_lew <- tapply(tagsbags6[tagsbags6$event == 3 & tagsbags6$bag_site == "Lewisetta",]$length, 
                     tagsbags6[tagsbags6$event == 3 & tagsbags6$bag_site == "Lewisetta",]$pop, 
                     mean)
event3_tot <- tapply(tagsbags6[tagsbags6$event == 3,]$length, 
                     tagsbags6[tagsbags6$event == 3,]$pop, 
                     mean)
event4_york <- tapply(tagsbags6[tagsbags6$event == 4 & tagsbags6$bag_site == "YorkRiver",]$length, 
                      tagsbags6[tagsbags6$event == 4 & tagsbags6$bag_site == "YorkRiver",]$pop, 
                      mean)
event4_lew <- tapply(tagsbags6[tagsbags6$event == 4 & tagsbags6$bag_site == "Lewisetta",]$length, 
                     tagsbags6[tagsbags6$event == 4 & tagsbags6$bag_site == "Lewisetta",]$pop, 
                     mean)
event4_tot <- tapply(tagsbags6[tagsbags6$event == 4,]$length, 
                     tagsbags6[tagsbags6$event == 4,]$pop, 
                     mean)

# put it all into a df
df <- data.frame(site = "Lewisetta", pop = rownames(event1_lew), 
                 monitoring_event = 1, length = event1_lew, t1 = event1_lew)
df1 <- data.frame(site = "Lewisetta", pop = rownames(event1_lew),
                  monitoring_event = 2, length = event2_lew, t2 = event2_lew)
df3 <- data.frame(site = "Lewisetta", pop = rownames(event1_lew),
                  monitoring_event = 3, length = event3_lew, t3 = event3_lew)
df4 <- data.frame(site = "Lewisetta", pop = rownames(event1_lew),
                  monitoring_event = 4, length = event4_lew, t4 = event4_lew)
df5 <- data.frame(site = "YorkRiver", pop = rownames(event1_york), 
                  monitoring_event = 1, length = event1_york, t1 = event1_york)
df6 <- data.frame(site = "YorkRiver", pop = rownames(event1_york),
                  monitoring_event = 2, length = event2_york, t2 = event2_york)
df7 <- data.frame(site = "YorkRiver", pop = rownames(event1_york),
                  monitoring_event = 3, length = event3_york, t3 = event3_york)
df8 <- data.frame(site = "YorkRiver", pop = rownames(event1_york),
                  monitoring_event = 4, length = event4_york, t4 = event4_york)
df9 <- data.frame(site = "Overall", pop = rownames(event1_tot), 
                  monitoring_event = 1, length = event1_tot, t1 = event1_tot)
df10 <- data.frame(site = "Overall", pop = rownames(event1_lew),
                  monitoring_event = 2, length = event2_tot, t2 = event2_tot)
df11 <- data.frame(site = "Overall", pop = rownames(event1_lew),
                  monitoring_event = 3, length = event3_tot, t3 = event3_tot)
df12 <- data.frame(site = "Overall", pop = rownames(event1_lew),
                  monitoring_event = 4, length = event4_tot, t4 = event4_tot)

lengths <- rbind(df[,-5], df1[,-5], df3[,-5], df4[,-5], df5[,-5], df6[,-5], df7[,-5], df8[,-5], df9[,-5], df10[,-5], df11[,-5], df12[,-5])

len_summary <- lengths %>% 
  #  filter(monitoring_event == 1, bag_site == "Lewisetta") %>%
  arrange(-length, .by_group = T) %>%
  as.data.frame()

lew4 <- len_summary[len_summary$monitoring_event == 4 & len_summary$site == "Lewisetta",]
lew3 <- len_summary[len_summary$monitoring_event == 3 & len_summary$site == "Lewisetta",]
lew2 <- len_summary[len_summary$monitoring_event == 2 & len_summary$site == "Lewisetta",]
lew1 <- len_summary[len_summary$monitoring_event == 1 & len_summary$site == "Lewisetta",]
york4 <- len_summary[len_summary$monitoring_event == 4 & len_summary$site == "YorkRiver",]
york3 <- len_summary[len_summary$monitoring_event == 3 & len_summary$site == "YorkRiver",]
york2 <- len_summary[len_summary$monitoring_event == 2 & len_summary$site == "YorkRiver",]
york1 <- len_summary[len_summary$monitoring_event == 1 & len_summary$site == "YorkRiver",]

sink(paste0("results/exp_results/len_sites_", Sys.Date(), ".txt"))
york1
york2
york3
york4
lew1
lew2
lew3
lew4
sink()
```

Now visualize some of these trends.

```{r}
# color scheme
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "LOLA", "DEBY", "NH", "ME"), 
                     order = 1:8, 
                     cols = c("#365C8DFF", "#277F8EFF", "#1FA187FF", "#4AC16DFF", "#440154FF", "#46337EFF", "#9FDA3AFF", "#FDE725FF"),
                     shape = c(21,21,21,21,23,23,21,21),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME"))

# merge with popsdf
lengths_mg <- merge(lengths, popsdf, by = c("pop"))

# plot
lew_len <- ggplot(data = lengths_mg[lengths_mg$site == "Lewisetta",]) + 
  geom_line(aes(x = monitoring_event, y = length, color = fct_reorder(pop, order))) + 
  geom_point(aes(x = monitoring_event, y = length, fill = fct_reorder(pop, order), 
                 shape = fct_reorder(pop, order)),
             size = 4) + 
  scale_color_manual(name = "Population", 
                     labels = setNames(popsdf$site_name, popsdf$pop),
                     values = setNames(popsdf$cols, popsdf$pop)) +
  scale_shape_manual(name = "Population", 
                     labels = setNames(popsdf$site_name, popsdf$pop),
                     values = setNames(popsdf$shape, popsdf$pop)) +
  scale_fill_manual(name = "Population", 
                    labels = setNames(popsdf$site_name, popsdf$pop),
                    values = setNames(popsdf$cols, popsdf$pop)) +
  ylim(20, 100) + 
  labs(title = "Lewisetta: Oyster Shell Length thru Time", 
       x = "Monitoring Event", 
       y = "Shell Length (mm)") +
  scale_x_continuous("Monitoring Event", breaks = c(1, 2, 3, 4),
                     labels = c("Fall '23", "Spring '24", "Fall '24", "Spring '25")) +
  theme_classic() + 
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.justification = "top")

yrk_len <- ggplot(data = lengths_mg[lengths_mg$site == "YorkRiver",]) + 
  geom_line(aes(x = monitoring_event, y = length, color = fct_reorder(pop, order))) + 
  geom_point(aes(x = monitoring_event, y = length, fill = fct_reorder(pop, order), 
                 shape = fct_reorder(pop, order)),
             size = 4) + 
  scale_color_manual(name = "Population", 
                     labels = setNames(popsdf$site_name, popsdf$pop),
                     values = setNames(popsdf$cols, popsdf$pop)) +
  scale_shape_manual(name = "Population", 
                     labels = setNames(popsdf$site_name, popsdf$pop),
                     values = setNames(popsdf$shape, popsdf$pop)) +
  scale_fill_manual(name = "Population", 
                    labels = setNames(popsdf$site_name, popsdf$pop),
                    values = setNames(popsdf$cols, popsdf$pop)) +
  ylim(20, 100) + 
  labs(title = "York River: Oyster Shell Length thru Time", 
       x = "Monitoring Event", 
       y = "Shell Length (mm)") +
  scale_x_continuous("Monitoring Event", breaks = c(1, 2, 3, 4),
                     labels = c("Fall '23", "Spring '24", "Fall '24", "Spring '25")) +
  theme_classic() + 
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.justification = "top")

all_len <- ggplot(data = lengths_mg[lengths_mg$site == "Overall",]) + 
  geom_line(aes(x = monitoring_event, y = length, color = fct_reorder(pop, order))) + 
  geom_point(aes(x = monitoring_event, y = length, fill = fct_reorder(pop, order), 
                 shape = fct_reorder(pop, order)),
             size = 4) + 
  scale_color_manual(name = "Population", 
                     labels = setNames(popsdf$site_name, popsdf$pop),
                     values = setNames(popsdf$cols, popsdf$pop)) +
  scale_shape_manual(name = "Population", 
                     labels = setNames(popsdf$site_name, popsdf$pop),
                     values = setNames(popsdf$shape, popsdf$pop)) +
  scale_fill_manual(name = "Population", 
                    labels = setNames(popsdf$site_name, popsdf$pop),
                    values = setNames(popsdf$cols, popsdf$pop)) +
  ylim(20, 100) + 
  labs(title = "Oyster Shell Length thru Time", 
       x = "Monitoring Event", 
       y = "Shell Length (mm)") +
  scale_x_continuous("Monitoring Event", breaks = c(1, 2, 3, 4),
                     labels = c("Fall '23", "Spring '24", "Fall '24", "Spring '25")) +
  theme_classic() + 
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.justification = "top")

all_len
lew_len
yrk_len
```


Now we'll look more in depth at length at just the final time point. 

```{r}
lengths_bags_df$pop <- str_split_fixed(lengths_bags_df$bag, "-", 3)[,2]
lengths_t4 <- merge(lengths_bags_df[lengths_bags_df$monitoring_event == 4,], popsdf, by = c("pop"))
lengths_t4$bagnum <- str_split_fixed(lengths_t4$bag, "-", 3)[,3]
lengths_t4$site <- ifelse(lengths_t4$bagnum < 4, "Lewisetta", "York")

# plot by pop
len_t4_plot_pop <- ggplot(data = lengths_t4, aes(x = site, y = length, fill = fct_reorder(pop, order))) + 
  geom_boxplot(position=position_dodge(0.9)) +
    scale_fill_manual(name = "Population", label = popsdf$label, values = popsdf$cols)+
#  geom_text(aes(label = letters, y = 1), vjust = -0.5, position=position_dodge(1)) +
  labs(title = "Length at Final Time Point", x = "Experimental Site", y = "Length (mm)") +
  theme_classic()
len_t4_plot_pop

# plot by site
len_t4_plot_site <- ggplot(data = lengths_t4, aes(x = site, y = length, fill = site)) + 
  geom_boxplot(position=position_dodge(0.9)) +
  scale_fill_manual(name = "Site", label = c("Lewisetta", "York"), values = c("mediumspringgreen", "dodgerblue")) +
#  geom_text(aes(label = letters, y = 1), vjust = -0.5, position=position_dodge(1)) +
  labs(title = "Length at Final Time Point", x = "Experimental Site", y = "Length (mm)") +
  theme_classic()
len_t4_plot_site
```
