---
title: "Oyster_len_calc"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/MVP_Oyster_Analysis_Camille/MVP_Oysters_Analysis_Camille")

# load libraries
library(tidyverse) # for data wrangling
library(stringr) # for splitting strings
library(lubridate) # for dates
library(ggplot2) # for plotting
library(plotrix) # for std error
```

## Calculating Length/growth of Oysters in Field

This script is to analyze length of oysters in the MVP Field Experiment from 2023-2025. First I need to read in and merge together relevant data. I will then clean this data up to prepare it for analysis by time point, site, population, and more.

```{r}
bags <- read.csv("data/ooters_prelim/MVP23-FieldBags - bags.csv")
expsites <- read.csv("data/ooters_prelim/MVP23-FieldBags - sites.csv")
spawntrt <- read.csv("data/ooters_prelim/MVP23-FieldBags - spawn_trt2.csv")
tags <- read.csv("data/ooters_prelim/MVP23-FieldBags - tags.csv")
tagleng <- read.csv("data/ooters_prelim/MVP23-FieldBags - tagsLength.csv")

# split bags to have more detailed bag data
bags[,5:6] <- str_split_fixed(bags$bags_label, "-", 3)[,2:3]
colnames(bags) <- c("bags_key","bag_site","bags_label","SpawnTrt_Key","pop","bagnum")

# and remove practice bags
bags_rm <- bags[-(61:62),]

# also put together tagleng data
tagsbags <- merge(tags, bags_rm, by = c("bags_key"), all = T)
tagsbags1 <- merge(tagsbags, tagleng, by = c("tags_key"), all = T)

# trim dataset
tagsbags2 <- tagsbags1[,c("tags_label","tags_timestamp","bag_site","bags_label",
                          "pop","bagnum","length","width","tagsLength_timestamp")]

# split timestamps to get dates, dates 4/29-5/7 for most recent monitoring
# split timestamp column into two
tagsbags2[,c("tagsLength_timestamp")]
datetime_split_tag <- str_split_fixed(tagsbags2[,c("tagsLength_timestamp")], " ", 2) # split timestamp
colnames(datetime_split_tag) <- c("tagsLength_date", "tagsLength_time") # name columns
tagsbags3 <- cbind(tagsbags2, datetime_split_tag) # add to full dataset

# use dates to split by monitoring event
# dates are 11/7/23 - 11/15/23, 4/29/24 - 5/7/24
tagsbags3$tagsLength_date <- mdy(tagsbags3$tagsLength_date)

# remove practice
tagsbags4 <- tagsbags3[!(tagsbags3$tagsLength_date == "2024-03-13"),]
tagsbags5 <- tagsbags4[!(is.na(tagsbags4$tagsLength_date)),]

tagsbags5$monitoring_event <- ifelse(as_date(tagsbags5$tagsLength_date) > as_date("2023-11-16"),
                                     ifelse(as_date(tagsbags5$tagsLength_date) > as_date("2024-05-09"), 3, 2),
                                     1)

tagsbags6 <- tagsbags5[!(tagsbags5$pop == ""),]
tagsbags6 <- tagsbags6[!(tagsbags6$pop == "LEW"),]
tagsbags6 <- tagsbags6[!(tagsbags6$pop == "YORK"),]

write.csv(tagsbags6, "results/lengths_long.csv")
```

Now that I have prepared the data, I will look at some larger data trends. 

```{r}
###########
# calc bag level averages for each event
event1_bags <- tapply(tagsbags6[tagsbags6$monitoring_event == 1,]$length, 
                      tagsbags6[tagsbags6$monitoring_event == 1,]$bags_label,
                      mean)
event2_bags <- tapply(tagsbags6[tagsbags6$monitoring_event == 2,]$length,
                      tagsbags6[tagsbags6$monitoring_event == 2,]$bags_label,
                      mean)
event3_bags <- tapply(tagsbags6[tagsbags6$monitoring_event == 3,]$length,
                      tagsbags6[tagsbags6$monitoring_event == 3,]$bags_label,
                      mean)

# put this in data frame
event1_bags_df <- data.frame(bag = rownames(event1_bags), 
                             monitoring_event = 1, 
                             length = event1_bags)
rownames(event1_bags_df) <- NULL
event2_bags_df <- data.frame(bag = rownames(event2_bags), 
                             monitoring_event = 2, 
                             length = event2_bags)
rownames(event2_bags_df) <- NULL
event3_bags_df <- data.frame(bag = rownames(event3_bags), 
                             monitoring_event = 3, 
                             length = event3_bags)
rownames(event3_bags_df) <- NULL

# put it together
lengths_bags_df <- rbind(event1_bags_df, event2_bags_df, event3_bags_df)

# save it
write.csv(lengths_bags_df, "results/lengths_bags_022025.csv")

# calc pop level averages for each event
event1_york <- tapply(tagsbags6[tagsbags6$monitoring_event == 1 & tagsbags6$bag_site == "YorkRiver",]$length, 
                      tagsbags6[tagsbags6$monitoring_event == 1 & tagsbags6$bag_site == "YorkRiver",]$pop, 
                      mean)
event1_lew <- tapply(tagsbags6[tagsbags6$monitoring_event == 1 & tagsbags6$bag_site == "Lewisetta",]$length, 
                     tagsbags6[tagsbags6$monitoring_event == 1 & tagsbags6$bag_site == "Lewisetta",]$pop, 
                     mean)
event1_tot <- tapply(tagsbags6[tagsbags6$monitoring_event == 1,]$length, 
                     tagsbags6[tagsbags6$monitoring_event == 1,]$pop, 
                     mean)
event2_york <- tapply(tagsbags6[tagsbags6$monitoring_event == 2 & tagsbags6$bag_site == "YorkRiver",]$length, 
                      tagsbags6[tagsbags6$monitoring_event == 2 & tagsbags6$bag_site == "YorkRiver",]$pop, 
                      mean)
event2_lew <- tapply(tagsbags6[tagsbags6$monitoring_event == 2 & tagsbags6$bag_site == "Lewisetta",]$length, 
                     tagsbags6[tagsbags6$monitoring_event == 2 & tagsbags6$bag_site == "Lewisetta",]$pop, 
                     mean)
event2_tot <- tapply(tagsbags6[tagsbags6$monitoring_event == 2,]$length, 
                     tagsbags6[tagsbags6$monitoring_event == 2,]$pop, 
                     mean)
event3_york <- tapply(tagsbags6[tagsbags6$monitoring_event == 3 & tagsbags6$bag_site == "YorkRiver",]$length, 
                      tagsbags6[tagsbags6$monitoring_event == 3 & tagsbags6$bag_site == "YorkRiver",]$pop, 
                      mean)
event3_lew <- tapply(tagsbags6[tagsbags6$monitoring_event == 3 & tagsbags6$bag_site == "Lewisetta",]$length, 
                     tagsbags6[tagsbags6$monitoring_event == 3 & tagsbags6$bag_site == "Lewisetta",]$pop, 
                     mean)
event3_tot <- tapply(tagsbags6[tagsbags6$monitoring_event == 3,]$length, 
                     tagsbags6[tagsbags6$monitoring_event == 3,]$pop, 
                     mean)

# put it all into a df
df <- data.frame(site = "Lewisetta", pop = rownames(event1_lew), 
                 monitoring_event = 1, length = event1_lew)
df1 <- data.frame(site = "Lewisetta", pop = rownames(event1_lew),
                  monitoring_event = 2, length = event2_lew)
df3 <- data.frame(site = "Lewisetta", pop = rownames(event1_lew),
                  monitoring_event = 3, length = event3_lew)
df4 <- data.frame(site = "York", pop = rownames(event1_york), 
                  monitoring_event = 1, length = event1_york)
df5 <- data.frame(site = "York", pop = rownames(event1_york),
                  monitoring_event = 2, length = event2_york)
df6 <- data.frame(site = "York", pop = rownames(event1_york),
                  monitoring_event = 3, length = event3_york)
df7 <- data.frame(site = "Overall", pop = rownames(event1_tot), 
                  monitoring_event = 1, length = event1_tot)
df8 <- data.frame(site = "Overall", pop = rownames(event1_lew),
                  monitoring_event = 2, length = event2_tot)
df9 <- data.frame(site = "Overall", pop = rownames(event1_lew),
                  monitoring_event = 3, length = event3_tot)

lengths <- rbind(df, df1, df3, df4, df5, df6, df7, df8, df9)

len_summary <- lengths %>% 
  #  filter(monitoring_event == 1, bag_site == "Lewisetta") %>%
  arrange(-length, .by_group = T) %>%
  as.data.frame()

lew3 <- len_summary[len_summary$monitoring_event == 3 & len_summary$site == "Lewisetta",]
lew2 <- len_summary[len_summary$monitoring_event == 2 & len_summary$site == "Lewisetta",]
lew1 <- len_summary[len_summary$monitoring_event == 1 & len_summary$site == "Lewisetta",]
york3 <- len_summary[len_summary$monitoring_event == 3 & len_summary$site == "York",]
york2 <- len_summary[len_summary$monitoring_event == 2 & len_summary$site == "York",]
york1 <- len_summary[len_summary$monitoring_event == 1 & len_summary$site == "York",]

sink("results/len_sites_02202025.txt")
york1
york2
york3
lew1
lew2
lew3
sink()
```

Now visualize some of these trends.

```{r}
popsdf <- data.frame(pop = c("TX","LA","FL","JR","DEBY","LOLA","NH","ME","LARMIX","SEEDMIX"), 
                     order = 1:10, 
#                     cols = c("#6f1926","#de324c","#f4895f","#f8e16f","#95cf92","#369acc",
#                              "#9656a2","#cbabd1","black","gray"),
                     cols = c("#332288","#117733","#44AA99","#88CCEE","#DDCC77","#CC6677",
                              "#AA4499","#882255","black","gray"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     label = c("W1-TX","W2-LA","W3-FL","W4-VA","S1-LOLA","S2-DEBY",
                               "W5-NH","W6-ME","H1-LARMIX","H2-SEEDMIX"))

# merge with popsdf
lengths_mg <- merge(lengths, popsdf, by = c("pop"))

# plot
lew_len <- ggplot(data = lengths_mg[lengths_mg$site == "Lewisetta",]) + 
  geom_line(aes(x = monitoring_event, y = length, color = fct_reorder(pop,order))) + 
  scale_color_manual(name = "Population", label = popsdf$label, values = popsdf$cols)+
  geom_point(aes(x = monitoring_event, y = length, fill = fct_reorder(pop,order), 
                 shape = fct_reorder(pop,order)),
             size = 4) + 
  scale_shape_manual(name = "Population", label = popsdf$label, values = popsdf$shape)+
  scale_fill_manual(name = "Population", label = popsdf$label, values = popsdf$cols)+
  ylim(20,80) + 
  labs(color = "Population") +
  ylab("Shell Length (mm)") +
  scale_x_continuous("Monitoring Event", breaks = c(1,2,3),
                     labels = c("Fall '23", "Spring '24", "Fall '24")) +
  theme_classic() + 
  ggtitle("(C) Lewisetta: Oyster Shell Length thru Time") +
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.justification = "top") +
  ggtitle("Lewisetta: Oyster Shell Length thru Time")
lew_len

york_len <- ggplot(data = lengths_mg[lengths_mg$site == "York",]) + 
  geom_line(aes(x = monitoring_event, y = length, color = fct_reorder(pop,order))) + 
  scale_color_manual(name = "Population", label = popsdf$label, values = popsdf$cols)+
  geom_point(aes(x = monitoring_event, y = length, fill = fct_reorder(pop,order), 
                 shape = fct_reorder(pop,order)),
             size = 4) + 
  scale_shape_manual(name = "Population", label = popsdf$label, values = popsdf$shape)+
  scale_fill_manual(name = "Population", label = popsdf$label, values = popsdf$cols)+
  ylim(20,80) + 
  labs(color = "Population") +
  ylab("Shell Length (mm)") +
  scale_x_continuous("Monitoring Event", breaks = c(1,2,3),
                     labels = c("Fall '23", "Spring '24", "Fall '24")) +
  theme_classic() + 
  ggtitle("(C) Lewisetta: Oyster Shell Length thru Time") +
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.justification = "top") +
  ggtitle("York River: Oyster Shell Length thru Time")
york_len

all_len <- ggplot(data = lengths[lengths$site == "Overall",]) + 
  geom_line(aes(x = monitoring_event, y = length, color = fct_reorder(pop,order))) + 
  scale_color_manual(name = "Population", label = popsdf$label, values = popsdf$cols)+
  geom_point(aes(x = monitoring_event, y = length, fill = fct_reorder(pop,order), 
                 shape = fct_reorder(pop,order)),
             size = 4) + 
  scale_shape_manual(name = "Population", label = popsdf$label, values = popsdf$shape)+
  scale_fill_manual(name = "Population", label = popsdf$label, values = popsdf$cols)+
  ylim(20,80) + 
  labs(color = "Population") +
  ylab("Shell Length (mm)") +
  scale_x_continuous("Monitoring Event", breaks = c(1,2,3),
                     labels = c("Fall '23", "Spring '24", "Fall '24")) +
  theme_classic()


all_len
lew_len
york_len
```
