---
title: "len_surv_modeling_ooters"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/projects/lotterhos/MVP_oyster_GO")

# load libraries
library(tidyverse) # for data wrangling
library(stringr) # for splitting strings
library(lubridate) # for dates
library(ggplot2) # for plotting
library(plotrix) # for std error
library(lme4) # for lmm
library(performance) # for lmm
library(lmerTest) # for lmm
library(car) # for avPlots(), Anova()
```

This script is for modeling differences in survival and length amongst bags/populations/sites/time points in the MVP Field Experiment. 

I intend to use linear (mixed) models to examine how survival and length vary by group and how genetics and environmental history drive these differences in performance. 

This script will also be split at some point to separate some of the modeling.

## Modeling differences in survival

First I'll read in and process survival data

```{r}
# read in data
surv_long <- read.csv("results/exp_results/mortality_long_bags_052925.csv")[,-1]
bags <- read.csv("data/ExpDat/MVP23-FieldBags - bags.csv")

# split bags to have more detailed bag data
bags[,5:6] <- str_split_fixed(bags$bags_label, "-", 3)[,2:3]
colnames(bags) <- c("bags_key","bag_site","bag","SpawnTrt_Key","pop","bagnum")

# combine data
surv_for_mod <- merge(surv_long, bags[,c("bag","bag_site")], by = "bag")
surv_for_mod$pop <- str_split_fixed(surv_for_mod$bag, "-", 3)[,2]
surv_mod_ord <- surv_for_mod[,c("bag_site","pop","bag","event","surv")]

# rename populations
map=c("TX"="W1-TX","LA"="W2-LA","FL"="W3-FL","LOLA"="S1-LOLA","DEBY"="S2-DEBY","JR"="W4-VA","NH"="W5-NH","ME"="W6-ME")
surv_mod_ord_rename <- surv_mod_ord %>% mutate(pop = ifelse(pop %in% names(map), map[pop], pop))

# remove any NAs
surv_mod_ord_rename_nona <- na.omit(surv_mod_ord_rename)

# we should also remove seedmix and larmix for now
surv_mod_ord_rm <- surv_mod_ord_rename_nona %>% filter(!pop %in% c("LARMIX","SEEDMIX"))

colnames(surv_mod_ord_rm) <- c("bag_site","pop","bags_label","t","surv")
```

Now I can use LMMs to look at differences in survival by group.

```{r}
# analysis of fixed effects
anova(lm(surv ~ pop + bag_site + t, data = surv_mod_ord_rm)) # all significant

# ensure correct coding of explanatory variables
surv_mod_ord_rm$t <- as.factor(surv_mod_ord_rm$t)

# start with full model, then subtract each interaction, then subtract two of the interaction
lmm1 <- lmer(surv ~ bag_site * pop * t + (1|bags_label), data = surv_mod_ord_rm)
plot(lmm1)
Anova(lmm1, type = "III")
drop1(lmm1) # keep all
summary(lmm1)

# would be good to create line plot - this is final thing to report, could create supplemental table of results, lots of two way interactions and changes in survival through time in different "populations" and which "populations" have highest survival at different time points changes through time
avPlots(lmm1) # these are added variable plots, not sure if this was the plot type we wanted
# i guess we technically have a line plot of survival through time already, those are what I produced earlier from raw survival proportion data

# one more model on final time point only
lmm2 <- lm(surv ~ bag_site * pop, data = surv_mod_ord_rm[surv_mod_ord_rm$t == 4,]) # no random effect here because there is only one observation per bag label
plot(lmm2)
Anova(lmm2, type = "III") # make sure to always use this method if two categorical predictors and  interaction and unbalanced design ; type III may have to set the contrast, check biostats notes
TukeyHSD(aov(lmm2)) # anova doesn't use Type III SS, so use aov - look into aov vs anova

# plot survival at last time point
surv_t4 <- surv_mod_ord_rm[surv_mod_ord_rm$t == 4,]

surv_t4_plot_site <- ggplot(surv_t4, aes(x = pop, y = surv, fill = bag_site)) + 
  geom_boxplot(position=position_dodge(0.8)) +
  ggtitle("Survival at Final Time Point") +
  theme_classic()
surv_t4_plot_site

surv_t4_plot_pop <- ggplot(surv_t4, aes(x = bag_site, y = surv, fill = pop)) + 
  geom_boxplot(position=position_dodge(1.0)) +
  theme_classic()
surv_t4_plot_pop

#extract significant letters from TK test and store
# try adding a different pop column that's split
surv_t4$oldpop <- str_split_fixed(surv_t4$pop,"-",2)[,2]

# now redo model
lmm2.1 <- lm(surv ~ bag_site * oldpop, data = surv_t4)
plot(lmm2.1)
Anova(lmm2.1, type = "III") # make sure to always use this method if two categorical predictors and interaction and unbalanced design ; type III may have to set the contrast, check biostats notes
library(multcompView)
tuk_surv_t4 <- TukeyHSD(aov(lmm2.1), conf.level = 0.95)
cld_surv_t4 <- multcompLetters4(aov(lmm2.1), tuk_surv_t4)

# can paste pops with sites to line up with letters output
surv_t4$popsite <- paste0(surv_t4$bag_site,":",str_split_fixed(surv_t4$pop,"-",2)[,2])

# multcomp letters to get letters
letters_surv_t4_df <- data.frame(letters = as.data.frame.list(cld_surv_t4$`bag_site:oldpop`), popsite = names((cld_surv_t4$`bag_site:oldpop`)$Letters))
letters_surv_t4_df_red <- letters_surv_t4_df[,c("letters.Letters","popsite")]
colnames(letters_surv_t4_df_red) <- c("letters","popsite")

# save these outputs too
sink(paste0("results/survival_model_results_", Sys.Date(), ".txt"))
Anova(lmm1, type = "III")
summary(lmm1)
Anova(lmm2, type = "III")
summary(lmm2)
Anova(lmm2.1, type = "III")
summary(lmm2.1)
letters_surv_t4_df_red
sink()

# merge
surv_final_time_rename_mg <- merge(surv_t4, letters_surv_t4_df_red, by = "popsite")
(surv_final_time_rename_mg)

# plot
surv_t4_plot_site <- ggplot(surv_final_time_rename_mg, aes(x = pop, y = surv, fill = bag_site)) + 
  geom_boxplot(position=position_dodge(1)) +
  ggtitle("Survival at Final Time Point") +
  geom_text(aes(label = letters, y = 1), vjust = -0.5, position=position_dodge(1)) +
  theme_classic()
surv_t4_plot_site

# pop color scheme
pop_colors <- data.frame(pop = c("W1-TX","W2-LA","W3-FL","W4-VA","S1-LOLA","S2-DEBY","W5-NH","W6-ME","H1-LARMIX","H2-SEEDMIX"), hex = c("#3e4989","#31688e","#26828e","#1f9e89","#440154","#482878","#35b779","#6ece58","#fde725","#b5de2b"), order = 1:10)

surv_final_time_plotting <- merge(surv_final_time_rename_mg, pop_colors, by = "pop")

# plot by pop
surv_t4_plot_pop <- ggplot(surv_final_time_plotting, aes(x = bag_site, y = surv, fill = fct_reorder(pop, order))) + 
  geom_boxplot(position=position_dodge(1.0)) +  
  scale_fill_manual(name = "Broodstock Group", label = pop_colors$pop, values = pop_colors$hex)+
  geom_text(aes(label = letters, y = 1), vjust = -0.5, position=position_dodge(1)) +
  labs(title = "Final Survival at Lewisetta and York River", x = "Experimental Site", y = "Survival") +
  theme_classic()
surv_t4_plot_pop
```


## Modeling differences in length

Now I'll try modeling differences in length. I'll start by reading in and processing the length data.

```{r}
# read in and trim data
lengths <- read.csv("results/exp_results/lengths_long_2025-06-03.csv")[,-1]
head(lengths)
lengths_mod <- lengths[,c("bag_site","pop","event","bags_label","tags_label","length","width")]
colnames(lengths_mod) <- c("bag_site","pop","t","bags_label","tags_label","length","width")

# rename pops
map=c("TX"="W1-TX","LA"="W2-LA","FL"="W3-FL","LOLA"="S1-LOLA","DEBY"="S2-DEBY",
      "JR"="W4-VA","NH"="W5-NH","ME"="W6-ME")
lengths_mod_rename <- lengths_mod %>% mutate(pop = ifelse(pop %in% names(map), map[pop], pop))

# we should also remove seedmix and larmix for now
lengths_mod_rm <- lengths_mod_rename %>% filter(!pop %in% c("LARMIX","SEEDMIX"))

# some individuals died/went missing and have "0" in length/width - rm these
lengths_mod_rm_nodead <- lengths_mod_rm[!lengths_mod_rm$length == 0 | !(complete.cases(lengths_mod_rm)),]
sum(!(complete.cases(lengths_mod_rm_nodead))) # use which to inspect those rows
lengths_mod_rm_nodead[which(!(complete.cases(lengths_mod_rm_nodead))),] # they're all rows containing na
rows_to_rm <- as.numeric(rownames(lengths_mod_rm_nodead[which(!(complete.cases(lengths_mod_rm_nodead))),]))
lengths_mod_rm_nodead[which(is.na(lengths_mod_rm_nodead$width)),]
lengths_mod_narm_nodead <- lengths_mod_rm_nodead[!(is.na(lengths_mod_rm_nodead$width)),]
```

I will now use LMMs to look at differences in length by grouping.

```{r}
# analysis of fixed effects
lengths_mod_narm_nodead$t <- as.factor(lengths_mod_narm_nodead$t)
modtemp0.1 <- lm(length ~ pop * bag_site * t, data = lengths_mod_narm_nodead)
modtemp.1 <- Anova(lm(length ~ pop * bag_site * t, data = lengths_mod_narm_nodead), type = "III") # all significant
boxplot(residuals(modtemp0.1) ~ lengths_mod_narm_nodead$bags_label) # looks good!

# with interactions
# using REML
lmm3 <- lmer(length ~ bag_site * pop * t + (1|bags_label), data = lengths_mod_narm_nodead)
Anova(lmm3, type = "III") # all signif

# start with full model, then subtract each interaction, then subtract two of the interaction
drop1(lmm3) # keep all

# one more model on final time point only
lmm4 <- lmer(length ~ bag_site * pop + (1|bags_label), data = lengths_mod_narm_nodead[lengths_mod_narm_nodead$t == 4,])
plot(lmm4)
Anova(lmm4, type = "III") # all signif
drop1(lmm4)

library(emmeans)
emmeans_lmm4 <- emmeans(lmm4, ~bag_site * pop) # get estimated marginal means
(pairwise_contrasts_lmm4 <- emmeans(lmm4, pairwise ~ bag_site * pop) )

pairs(emmeans_lmm4, simple = "bag_site")
pairs(emmeans_lmm4, simple = "pop")

test(pairs(emmeans_lmm4, by = "bag_site"), by = NULL, adjust = "mvt")
test(pairs(emmeans_lmm4, by = "pop"), by = NULL, adjust = "mvt")

# can paste pops with sites to do a traditional Tukey - this is separate from the rest of the modeling
lengths_final_time <- lengths_mod_narm_nodead[lengths_mod_narm_nodead$t == 4,]
lengths_final_time$popsite <- paste0(lengths_final_time$pop, "_", lengths_final_time$bag_site)

# rename pops
lengths_final_time_rename <- lengths_final_time
lengths_final_time_rename$popsite <- paste0(str_split_fixed(lengths_final_time$pop,"-",2)[,2],"_",lengths_final_time$bag_site)

(aov_len_t4 <- aov(lm(length ~ popsite, data = lengths_final_time_rename)))
plot(aov_len_t4) # looks good!
(tukey_len_t4 <- TukeyHSD(aov_len_t4, conf.level = 0.95))

# multcomp letters to get letters
library(multcompView)
anova_len_t4 <- (Anova(lm(length ~ popsite, data = lengths_final_time_rename)))
letters_len_t4 <- multcompLetters4(aov_len_t4, tukey_len_t4)

letters_len_t4_df <- data.frame(letters = as.data.frame.list(letters_len_t4$popsite)$Letters, popsite = rownames(as.data.frame.list(letters_len_t4$popsite)))

lengths_final_time_rename_mg <- merge(lengths_final_time_rename, letters_len_t4_df, by = "popsite")
(lengths_final_time_rename_mg)

# need sd for letters
summary_len_t4 <- lengths_final_time_rename_mg %>%
  summarise(mean_len = mean(length), sd = sd(length)) %>%
  arrange()

# save all this 
sink(paste0("results/exp_results/length_model_results_", Sys.Date(), ".txt"))
Anova(lmm3, type = "III")
Anova(lmm4, type = "III")
anova_len_t4
letters_len_t4
sink()

# plot
len_t4_plot_site <- ggplot(lengths_final_time_rename_mg, aes(x = pop, y = length, fill = bag_site)) + 
  geom_boxplot(position=position_dodge(1)) +
  ggtitle("Length at Final Time Point") +
  geom_text(aes(label = letters, y = 127), vjust = -0.5, position=position_dodge(1)) +
  theme_classic()
len_t4_plot_site

# pop color scheme
pop_colors <- data.frame(pop = c("W1-TX","W2-LA","W3-FL","W4-VA","S1-LOLA","S2-DEBY","W5-NH","W6-ME","H1-LARMIX","H2-SEEDMIX"), hex = c("#3e4989","#31688e","#26828e","#1f9e89","#440154","#482878","#35b779","#6ece58","#fde725","#b5de2b"), order = 1:10)

len_final_time_plotting <- merge(lengths_final_time_rename_mg, pop_colors, by = "pop")

# plot by pop
len_t4_plot_pop <- ggplot(len_final_time_plotting, aes(x = bag_site, y = length, fill = fct_reorder(pop, order))) + 
  geom_boxplot(position=position_dodge(1.0)) +  
  scale_fill_manual(name = "Broodstock Group", label = pop_colors$pop, values = pop_colors$hex)+
  geom_text(aes(label = letters, y = 127), vjust = -0.5, position=position_dodge(1)) +
  labs(title = "Length at Final Time Point", x = "Experimental Site", y = "Survival") +
  theme_classic()
len_t4_plot_pop

#ggsave("results/figures/length_t4.pdf", plot = len_t4_plot_pop, width = 10, height = 6, dpi = 300)
```


## Modeling drivers of differences in survival

For the H2F paper, we want to know how environment and genetics are impacting differential survival. I'll start by putting together the complete dataframe.

```{r}
# start by putting together a complete dataframe
# genetic data
gen_parents <- read.csv("data/ExpDat/parent_genetic_effects.csv")
colnames(gen_parents) <- c("pop","ho","exp_hs","ar")
surv_mod_gen <- merge(surv_mod_ord_rm, gen_parents, by = "pop")

# environmental data
env_parents <- read.csv("data/EnvDat/exp/envr_summary_stats.csv")[,-1]
colnames(env_parents) <- c("pop","mean_temp","max_temp","min_temp","sd_temp","mean_sal","max_sal","min_sal","sd_sal")
surv_mod_gen_env <- merge(surv_mod_gen, env_parents, by = c("pop"))
```

Now look at how survival changes with genetics/environment at each time point. These are multiple regressions, so remove interactions here - they are not interpretable in the same way.

```{r}
# full model time point 1
# start with lew
surv_GE1.lew <- lm(surv ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "Lewisetta",])

surv_GE1.lew
#plot(surv_GE1.lew) 
summary(surv_GE1.lew) # none signif
drop1(surv_GE1.lew) # get rid of ho

surv_GE1.0.lew <- lm(surv ~ ar + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE1.0.lew) 
summary(surv_GE1.0.lew) # none signif
drop1(surv_GE1.0.lew) # get rid of ar

surv_GE1.1.lew <- lm(surv ~ max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE1.1.lew)
summary(surv_GE1.1.lew) # none signif
drop1(surv_GE1.1.lew) # get rid of min_sal

surv_GE1.2.lew <- lm(surv ~ max_temp + min_temp + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE1.2.lew)
summary(surv_GE1.2.lew) # min temp, max_sal signif
drop1(surv_GE1.2.lew) # get rid of max_temp

surv_GE1.3.lew <- lm(surv ~ min_temp + max_sal , data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE1.3.lew)
summary(surv_GE1.3.lew) # all signif
drop1(surv_GE1.3.lew) # this is final model

# now we go to york
surv_GE1.york <- lm(surv ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE1.york)
summary(surv_GE1.york) # min_sal signif
drop1(surv_GE1.york) # drop ar

surv_GE1.0.york <- lm(surv ~ ho + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE1.0.york)
summary(surv_GE1.0.york) # min_sal signif
drop1(surv_GE1.0.york) # drop max_temp

surv_GE1.1.york <- lm(surv ~ ho + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE1.1.york)
summary(surv_GE1.1.york) # min_sal, min_temp signif
drop1(surv_GE1.1.york) # drop max_sal

surv_GE1.2.york <- lm(surv ~ ho + min_temp + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 1 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE1.2.york)
summary(surv_GE1.2.york) # all signif
drop1(surv_GE1.2.york) # final model

# lew t2
surv_GE2.lew <- lm(surv ~ ho + ar + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 2 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE2.lew)
summary(surv_GE2.lew) # none signif
drop1(surv_GE2.lew) # get rid of min_sal

surv_GE2.0.lew <- lm(surv ~ ar + ho + min_temp + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 2 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE2.0.lew)
summary(surv_GE2.0.lew) # none signif
drop1(surv_GE2.0.lew) # get rid of ar

surv_GE2.1.lew <- lm(surv ~ ho + min_temp + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 2 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE2.1.lew)
summary(surv_GE2.1.lew) # min_temp, max_sal signif
drop1(surv_GE2.1.lew) # final model

# york t2
surv_GE2.york <- lm(surv ~ ho + ar + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 2 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE2.york) # qq still ugly
summary(surv_GE2.york) # min_temp, min_sal signif
drop1(surv_GE2.york) # get rid of ho

surv_GE2.0.york <- lm(surv ~ ar + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 2 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE2.0.york) 
summary(surv_GE2.0.york) # min_temp, max_sal, min_sal signif
drop1(surv_GE2.0.york) # final model

# lew t3
surv_GE3.lew <- lm(surv ~ ho + ar + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 3 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE3.lew)
summary(surv_GE3.lew) # min_temp, max_sal signif
drop1(surv_GE3.lew) # get rid of min_sal

surv_GE3.0.lew <- lm(surv ~ ho + ar + min_temp + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 3 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE3.0.lew)
summary(surv_GE3.0.lew) # ar, min_temp, max_sal signif
drop1(surv_GE3.0.lew) # get rid of ho

surv_GE3.1.lew <- lm(surv ~ ar + min_temp + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 3 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE3.1.lew)
summary(surv_GE3.1.lew) # ar, min_temp, max_sal signif
drop1(surv_GE3.1.lew) # final model

# york t3
surv_GE3.york <- lm(surv ~ ar + ho + min_temp + max_temp + min_sal + max_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 3 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE3.york)
summary(surv_GE3.york) # ar, max_temp, min_sal, intercept signif
drop1(surv_GE3.york) # get rid of min_temp

surv_GE3.0.york <- lm(surv ~ ar + ho + max_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 3 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE3.0.york)
summary(surv_GE3.0.york) # ar, ho, min_sal, max_temp, max_sal, intercept signif
drop1(surv_GE3.0.york) # final model

# lew t4
surv_GE4.lew <- lm(surv ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 4 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE4.lew)
summary(surv_GE4.lew) # max_temp, min_temp, max_sal signif
drop1(surv_GE4.lew) # get rid of ar

surv_GE4.0.lew <- lm(surv ~ ho + max_temp + min_temp + max_sal + min_sal, data = surv_mod_gen_env[surv_mod_gen_env$t == 4 & surv_mod_gen_env$bag_site == "Lewisetta",])
#plot(surv_GE4.0.lew)
summary(surv_GE4.0.lew) # max_temp, min_temp, max_sal signif
drop1(surv_GE4.0.lew) # this is final model

# york t4
surv_GE4.york <- lm(surv ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = c[surv_mod_gen_env$t == 4 & surv_mod_gen_env$bag_site == "YorkRiver",])
#plot(surv_GE4.york)
summary(surv_GE4.york) # intercept, ho, ar, max_temp, min_temp, min_sal signif
drop1(surv_GE4.york) # this is final model

sink(paste0("results/exp_results/survival_GE_model_results_",Sys.Date(),".txt"))
summary(surv_GE1.3.lew)
summary(surv_GE1.2.york)
summary(surv_GE2.1.lew)
summary(surv_GE2.0.york)
summary(surv_GE3.1.lew)
summary(surv_GE3.0.york)
summary(surv_GE4.0.lew)
summary(surv_GE4.york)
sink()
```


## Modeling drivers of differences in length


```{r}
library(car)

# start by putting together a complete dataframe
# genetic data
len_mod_gen <- merge(lengths_mod_rm_nodead, gen_parents, by = "pop")
len_mod_gen_env <- merge(len_mod_gen, env_parents, by = c("pop"))
(len_mod_gen_env)

# fixed effects
(anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "Lewisetta",]))) # all signif
(anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "Lewisetta",]))) # ho, min_temp, max_sal, min_sal signif
(anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "Lewisetta",]))) # ho, max_temp, max_sal, min_sal signif
(anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "Lewisetta",]))) # ho, max_temp, max_sal, min_sal signif

anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "YorkRiver",])) # all but ho & ar signif
anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "YorkRiver",])) # ar, min_temp, max_sal, min_sal signif
anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "YorkRiver",])) # ho, min_temp, max_sal, min_sal signif
anova(lm(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal, data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "YorkRiver",])) # ho, ar, max_sal, min_sal min_temp signif

# full model time point 1 - lew
len_GE1.lew <- lmerTest::lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "Lewisetta",]) 
len_GE1.lew
plot(len_GE1.lew) 
Anova(len_GE1.lew, type = "III") # max_temp, min_temp, min_sal signif
summary(len_GE1.lew) # gives just min_temp & min_sal
step(len_GE1.lew) # trying different model selection method

len_GE1.0.lew <- lmerTest::lmer(length ~ max_temp + min_temp + max_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "Lewisetta",]) 
plot(len_GE1.0.lew) 
Anova(len_GE1.0.lew, type = "III") # all signif
summary(len_GE1.0.lew)
step(len_GE1.0.lew) # final model

# york t1
len_GE1.york <- lmerTest::lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "YorkRiver",]) 
len_GE1.york
plot(len_GE1.york) 
Anova(len_GE1.york, type = "III") # max_temp, min_temp, min_sal signif
summary(len_GE1.york) # safe signif
step(len_GE1.york) # get rid of ho and ar

len_GE1.0.york <- lmerTest::lmer(length ~ max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 1 & len_mod_gen_env$bag_site == "YorkRiver",]) 
plot(len_GE1.0.york) 
Anova(len_GE1.0.york, type = "III") # all signif
summary(len_GE1.0.york)
step(len_GE1.0.york) # final model

# full model time point 2
# lew t2
len_GE2.lew <- lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE2.lew
plot(len_GE2.lew)
Anova(len_GE2.lew, type = "III") # min sal
step(len_GE2.lew) # drop min_temp, max_sal, ar, max_temp

len_GE2.0.lew <- lmer(length ~ ho + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE2.0.lew
plot(len_GE2.0.lew)
Anova(len_GE2.0.lew, type = "III") # all signif
step(len_GE2.0.lew) # final model

# york t2
len_GE2.york <- lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE2.york
plot(len_GE2.york)
Anova(len_GE2.york, type = "III") # max_temp, min_temp, min_sal signif
step(len_GE2.york) # drop max_sal, ar

len_GE2.0.york <- lmer(length ~ ho + max_temp + min_temp + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 2 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE2.0.york
plot(len_GE2.0.york)
Anova(len_GE2.0.york, type = "III") # all signif except intercept
step(len_GE2.0.york) # final model

# full model time point 3
# lew t3
len_GE3.lew <- lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE3.lew
plot(len_GE3.lew)
Anova(len_GE3.lew, type = "III") # intercept, min_temp, max_sal signif
step(len_GE3.lew) # drop everything except min_temp, max_sal

len_GE3.0.lew <- lmer(length ~ min_temp + max_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE3.0.lew
plot(len_GE3.0.lew)
Anova(len_GE3.0.lew, type = "III") # all signif
step(len_GE3.0.lew) # final model

# york t3
len_GE3.york <- lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE3.york
plot(len_GE3.york)
Anova(len_GE3.york, type = "III") # intercept, max_temp, max_sal, min_sal signif
step(len_GE3.york) # drop ho, min_temp

len_GE3.0.york <- lmer(length ~ ar + max_temp + min_sal + max_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 3 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE3.0.york
plot(len_GE3.0.york)
Anova(len_GE3.0.york, type = "III") # all signif except ar
step(len_GE3.0.york) # final model

# lew t4
len_GE4.lew <- lmer(length ~ ho + ar + max_temp + min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE4.lew
plot(len_GE4.lew)
Anova(len_GE4.lew, type = "III") # intercept, min_temp, max_sal signif
step(len_GE4.lew) # drop ho, max_temp, ar

len_GE4.0.lew <- lmer(length ~ min_temp + max_sal + min_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "Lewisetta",])
len_GE4.0.lew
plot(len_GE4.0.lew)
Anova(len_GE4.0.lew, type = "III") # intercept, min_temp, max_sal, min_sal signif
step(len_GE4.0.lew) # this is final model

# york t4
len_GE4.york <- lmer(length ~ ar + ho + min_temp + max_temp + min_sal + max_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE4.york
plot(len_GE4.york)
Anova(len_GE4.york, type = "III") # all signif except ho
step(len_GE4.york) # drop ho, min_temp

len_GE4.0.york <- lmer(length ~ ar + max_temp + min_sal + max_sal + (1|bags_label), data = len_mod_gen_env[len_mod_gen_env$t == 4 & len_mod_gen_env$bag_site == "YorkRiver",])
len_GE4.0.york
plot(len_GE4.0.york)
Anova(len_GE4.0.york, type = "III") # all signif
step(len_GE4.0.york) # final model

sink(paste0("results/exp_results/len_GE_model_results_", Sys.Date(), ".txt"))
Anova(len_GE1.0.lew, type = "III")
Anova(len_GE1.0.york, type = "III")
Anova(len_GE2.0.lew, type = "III")
Anova(len_GE2.0.york, type = "III")
Anova(len_GE3.0.lew, type = "III")
Anova(len_GE3.0.york, type = "III")
Anova(len_GE4.lew, type = "III")
Anova(len_GE4.0.york, type = "III")
sink()
```
