---
title: "Oyster_survival_calc"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/projects/lotterhos/MVP_oyster_GO")

# load libraries
library(tidyverse) # for data wrangling
library(stringr) # for splitting strings
library(lubridate) # for dates
library(ggplot2) # for plotting
library(plotrix) # for std error
```

## Calculating Oyster Mortality in the Field

This script is to calculate mortality of oysters in the MVP Field Experiment from 2023-2025. 

First I need to read in and merge together relevant data. I will then clean this data up to prepare it for analysis by time point, site, population, and more.

```{r}
# read in data
bags <- read.csv("data/ExpDat/MVP23-FieldBags - bags.csv")
surv <- read.csv("data/ExpDat/MVP23-FieldBags - mortality.csv")
expsites <- read.csv("data/ExpDat/MVP23-FieldBags - sites.csv")
spawntrt <- read.csv("data/ExpDat/MVP23-FieldBags - spawn_trt2.csv")

# merging bags and mortality
head(bags)
colnames(bags) <- c("bags_key", "bag_site", "bags_label", "SpawnTrt_Key")
head(surv)
merge1 <- merge(bags, surv, by = c("bags_key"), all = T)
head(merge1)

# add spawntrt data
head(merge1)
head(spawntrt)
merge2 <- merge(merge1, spawntrt, by = c("SpawnTrt_Key"), all = T)
head(merge2)

# trim dataset
merge2_trim <- merge2[,c("bags_key","bag_site","bags_label","SpawnTrt_Key","SpawnTrt_Label","latitudeDecimal","longitudeDecimal","mortality_key","mortality_label","mortality_timestamp","alive_count","dead_count","alive_returned")]

# split timestamps to get dates - I'll use mortality_timestamp for this
# dates are 5/19/2025 and onward for final monitoring event
# quick checks - looks good!
head(merge2_trim[,c("mortality_timestamp")])
length(merge2_trim[,c("mortality_timestamp")])

# split timestamp column into two
datetime_split <- str_split_fixed(merge2_trim[,c("mortality_timestamp")], " ", 2) # split timestamp
colnames(datetime_split) <- c("mortality_date", "mortality_time") # name columns
merge3 <- cbind(merge2_trim, datetime_split) # add to full dataset

# take out practice rows
merge3_rm <- merge3[!merge3$SpawnTrt_Label == "Practice",]

# also put a monitoring event column in here
merge3_rm$mortality_date <- mdy(merge3_rm$mortality_date)
merge3_rm$event <- ifelse(as_date(merge3_rm$mortality_date) > as_date("2023-11-16"), ifelse(as_date(merge3_rm$mortality_date) > as_date("2024-05-09"), ifelse(as_date(merge3_rm$mortality_date) > "2024-11-12", 4, 3), 2), 1)

# need to set a bag order
bag_order <- data.frame(bags_label = levels(as.factor(merge3_rm$bags_label)), order = 1:60)
merge4 <- merge(merge3_rm, bag_order, by = "bags_label")

# check that data makes sense!! alive count should never be greater than alive returned from previous event
# event 2
which(merge4[merge4$event == 2,]$alive_count > merge4[merge4$event == 1,]$alive_returned)
event2 <- merge4[merge4$event == 2,]
event2[which(merge4[merge4$event == 2,]$alive_count > merge4[merge4$event == 1,]$alive_returned),]$bags_label
# bags DEBY-3 and LA-3 had more alive in event 2 than in event 1...

# how much more?
## DEBY-3, 14 more oysters at t2 than t1
merge4[merge4$event == 2 & merge4$bags_label == "MVP-DEBY-3",]$alive_count - merge4[merge4$event == 1 & merge4$bags_label == "MVP-DEBY-3",]$alive_returned
## LA-3, 198 more oysters at t2 than t1
merge4[merge4$event == 2 & merge4$bags_label == "MVP-LA-3",]$alive_count - merge4[merge4$event == 1 & merge4$bags_label == "MVP-LA-3",]$alive_returned

# event 3
which(merge4[merge4$event == 3,]$alive_count > merge4[merge4$event == 2,]$alive_returned)
# all good!

# event 4
# remove na la-1
merge4_narm <- merge4[(!(is.na(merge4$alive_count))),]
# checks
which(merge4_narm[merge4_narm$event == 4,]$alive_count > merge4_narm[merge4_narm$event == 3,]$alive_returned)
merge4_narm[which(merge4_narm[merge4_narm$event == 4,]$alive_count > merge4_narm[merge4_narm$event == 3,]$alive_returned),]$bags_label
# MVP-DEBY-2 has more

# how much more?
## DEBY-2, 27 more
merge4_narm[merge4_narm$event == 4 & merge4_narm$bags_label == "MVP-DEBY-2",]$alive_count - merge4_narm[merge4_narm$event == 3 & merge4_narm$bags_label == "MVP-DEBY-2",]$alive_returned

# remove the funky bags from analysis
# keep for now
#merge4_rm <- merge4_narm[!(merge4_narm$bags_label == "MVP-DEBY-3" & merge4_narm$event == 2),]
#merge4_rm1 <- merge4_rm[!(merge4_rm$bags_label == "MVP-LA-3" & merge4_rm$event == 2),]
#merge4_rm2 <- merge4_rm1[!(merge4_rm1$bags_label == "MVP-DEBY-2" & merge4_rm1$event == 4),]
#merge4_rm <- merge4_narm[!merge4_narm$bags_label == "MVP-DEBY-3",]
#merge4_rm1 <- merge4_rm[!merge4_rm$bags_label == "MVP-LA-3",]
#merge4_rm2 <- merge4_rm1[!merge4_rm1$bags_label == "MVP-DEBY-2",]

# add pop to df
merge4_narm$pop <- str_split_fixed(merge4_narm$bags_label, "-", 3)[,2]
#merge4_rm$pop <- str_split_fixed(merge4_rm2$bags_label, "-", 3)[,2]
```


Data is now ready for analysis. I will calculate survival as a proportion of those surviving from the previous timepoint.


```{r}
# function for survival calculation - this works across rows, it's naive to population/bag id
surv_calc_bag <- function(dat, e) {
  dat_ord = dat[order(dat$order),]
  O = dat_ord[dat_ord$event == e,]$alive_count
  if(e - 1 < 1) {
    D = (dat_ord[dat_ord$event == e,]$alive_count) + (dat_ord[dat_ord$event == e,]$dead_count)
    S = 1
  } else {
    D = dat_ord[dat_ord$event == (e-1),]$alive_returned
    S = surv_calc_bag(dat_ord, (e-1))$surv
  }
  survival = (O/D) * S
  df = data.frame(bag = dat_ord[dat_ord$event == e,]$bags_label, event = e, surv = survival)
  return(df)
}

# calc survival
e1_surv <- surv_calc_bag(merge4_narm, 1)
e2_surv <- surv_calc_bag(merge4_narm, 2)
e3_surv <- surv_calc_bag(merge4_narm, 3)
e4_surv <- surv_calc_bag(merge4_narm, 4)

# long format df
surv_long <- rbind(e1_surv,e2_surv,e3_surv,e4_surv)
surv_long_narm <- surv_long[!(is.na(surv_long$bag)),]

# wide format df
colnames(e1_surv) <- c("bags_label","event","t1")
colnames(e2_surv) <- c("bags_label","event","t2")
colnames(e3_surv) <- c("bags_label","event","t3")
colnames(e4_surv) <- c("bags_label","event","t4")

# final survival df
surv_df <- merge(e1_surv[,c("bags_label","t1")], 
                 e2_surv[,c("bags_label","t2")], 
                 by = c("bags_label"))
surv_df2 <- merge(surv_df, 
                  e3_surv[,c("bags_label","t3")], 
                  by = c("bags_label"))
surv_df3 <- merge(surv_df2, 
                  e4_surv[,c("bags_label","t4")], 
                  by = c("bags_label"))
surv_df4 <- surv_df3[1:length(levels(as.factor(surv_df3$bags_label))),]
surv_df4[surv_df4$bags_label == "MVP-NH-4",]$t4 <- 0
surv_df4$t0 <- 1
surv_df4$pop <- str_split_fixed(surv_df4$bags_label, "-", 3)[,2]
surv_df4$bagnum <- 1:length(levels(as.factor(surv_df4$bags_label)))
surv_df_mg <- merge(surv_df4, bags, by = c("bags_label"))
surv_df_ord <- surv_df_mg[,c("pop","bags_label","bagnum","bag_site","SpawnTrt_Key","t0","t1","t2","t3","t4")]

# can now calculate averages across sites and populations
# overall pop survival
pop_surv <- surv_df_ord %>% group_by(pop) %>% 
  summarise(mean_t0 = mean(as.numeric(t0)),
            sd_t0 = sd(as.numeric(t0)),
            se_t0 = std.error(as.numeric(t0)),
            mean_t1 = mean(as.numeric(t1)), 
            sd_t1 = sd(as.numeric(t1)),
            se_t1 = std.error(as.numeric(t1)),
            mean_t2 = mean(as.numeric(t2)), 
            sd_t2 = sd(as.numeric(t2)),
            se_t2 = std.error(as.numeric(t2)),
            mean_t3 = mean(as.numeric(t3)), 
            sd_t3 = sd(as.numeric(t3)),
            se_t3 = std.error(as.numeric(t3)),
            mean_t4 = mean(as.numeric(t4)), 
            sd_t4 = sd(as.numeric(t4)),
            se_t4 = std.error(as.numeric(t4)),
            .groups = 'drop')
pop_surv

# overall site survival
site_surv <- surv_df_ord %>% group_by(bag_site) %>%
  summarise(mean_t0 = mean(as.numeric(t0)),
            sd_t0 = sd(as.numeric(t0)),
            se_t0 = std.error(as.numeric(t0)),
            mean_t1 = mean(as.numeric(t1)), 
            sd_t1 = sd(as.numeric(t1)),
            se_t1 = std.error(as.numeric(t1)),
            mean_t2 = mean(as.numeric(t2)), 
            sd_t2 = sd(as.numeric(t2)),
            se_t2 = std.error(as.numeric(t2)),
            mean_t3 = mean(as.numeric(t3)), 
            sd_t3 = sd(as.numeric(t3)),
            se_t3 = std.error(as.numeric(t3)),
            mean_t4 = mean(as.numeric(t4)), 
            sd_t4 = sd(as.numeric(t4)),
            se_t4 = std.error(as.numeric(t4)),
            .groups = 'drop')
site_surv

# by site and pop
site_pop_surv <- surv_df_ord %>% group_by(bag_site,pop) %>%
  summarise(mean_t0 = mean(as.numeric(t0)),
            sd_t0 = sd(as.numeric(t0)),
            se_t0 = std.error(as.numeric(t0)),
            mean_t1 = mean(as.numeric(t1)), 
            sd_t1 = sd(as.numeric(t1)),
            se_t1 = std.error(as.numeric(t1)),
            mean_t2 = mean(as.numeric(t2)), 
            sd_t2 = sd(as.numeric(t2)),
            se_t2 = std.error(as.numeric(t2)),
            mean_t3 = mean(as.numeric(t3)), 
            sd_t3 = sd(as.numeric(t3)),
            se_t3 = std.error(as.numeric(t3)),
            mean_t4 = mean(as.numeric(t4)), 
            sd_t4 = sd(as.numeric(t4)),
            se_t4 = std.error(as.numeric(t4)),
            .groups = 'drop')
site_pop_surv

# lets also put this in long format for plotting
e0 <- as.data.frame(site_pop_surv)[,c("bag_site","pop","mean_t0","sd_t0","se_t0")]
e0$event <- 0
e1 <- as.data.frame(site_pop_surv)[,c("bag_site","pop","mean_t1","sd_t1","se_t1")]
e1$event <- 1
e2 <- as.data.frame(site_pop_surv)[,c("bag_site","pop","mean_t2","sd_t2","se_t2")]
e2$event <- 2
e3 <- as.data.frame(site_pop_surv)[,c("bag_site","pop","mean_t3","sd_t3","se_t3")]
e3$event <- 3
e4 <- as.data.frame(site_pop_surv)[,c("bag_site","pop","mean_t4","sd_t4","se_t4")]
e4$event <- 4
colnames(e0) <- colnames(e1) <- colnames(e2) <- colnames(e3) <- colnames(e4) <-  c("bag_site","pop","mean_surv","sd","se","event")
surv_df <- rbind(e0, e1, e2, e3, e4)

# save all these dataframes
write.csv(surv_df, paste0("results/exp_results/mortality_long_popsite_",Sys.Date(),".csv"))
write.csv(surv_long_narm, paste0("results/exp_results/mortality_long_bags_",Sys.Date(),".csv"))
write.csv(surv_df_ord, paste0("results/exp_results/mortality_bags_", Sys.Date(), ".csv"))
```


All of my survival data is calculated, now I want to visualize trends in survival.

```{r}
# pops df for plotting
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "LOLA", "DEBY", "NH", "ME"), 
                     order = 1:8, 
                     cols = c("#365C8DFF", "#277F8EFF", "#1FA187FF", "#4AC16DFF", "#440154FF", "#46337EFF", "#9FDA3AFF", "#FDE725FF"),
                     shape = c(21,21,21,21,23,23,21,21),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME"))
surv_df[surv_df$pop == "JR",]$pop <- "VA"
surv_df_full <- merge(surv_df, popsdf, by = c("pop"))

# lewisetta
lew_surv <- ggplot(data = surv_df_full[surv_df_full$bag_site == "Lewisetta",]) + 
  geom_line(aes(x = event, y = mean_surv, color = fct_reorder(pop, order))) + 
  geom_point(aes(x = event, y = mean_surv, fill = fct_reorder(pop, order), 
                 shape = fct_reorder(pop, order)),
             size = 4) + 
  scale_color_manual(name = "Population", 
                     labels = setNames(popsdf$site_name, popsdf$pop),
                     values = setNames(popsdf$cols, popsdf$pop)) +
  scale_shape_manual(name = "Population", 
                     labels = setNames(popsdf$site_name, popsdf$pop),
                     values = setNames(popsdf$shape, popsdf$pop)) +
  scale_fill_manual(name = "Population", 
                    labels = setNames(popsdf$site_name, popsdf$pop),
                    values = setNames(popsdf$cols, popsdf$pop)) +
  ylim(0,1) + 
  labs(title = "Lewisetta: Oyster Survival thru Time", 
       x = "Monitoring Event", 
       y = "Survival") +
  scale_x_continuous("Monitoring Event", breaks = c(1, 2, 3, 4),
                     labels = c("Fall '23", "Spring '24", "Fall '24", "Spring '25")) +
  theme_classic() + 
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.justification = "top")
lew_surv

# york river
yrk_surv <- ggplot(data = surv_df_full[surv_df_full$bag_site == "YorkRiver",]) + 
  geom_line(aes(x = event, y = mean_surv, color = fct_reorder(pop, order))) + 
  geom_point(aes(x = event, y = mean_surv, fill = fct_reorder(pop, order), 
                 shape = fct_reorder(pop, order)),
             size = 4) + 
  scale_color_manual(name = "Population", 
                     labels = setNames(popsdf$site_name, popsdf$pop),
                     values = setNames(popsdf$cols, popsdf$pop)) +
  scale_shape_manual(name = "Population", 
                     labels = setNames(popsdf$site_name, popsdf$pop),
                     values = setNames(popsdf$shape, popsdf$pop)) +
  scale_fill_manual(name = "Population", 
                    labels = setNames(popsdf$site_name, popsdf$pop),
                    values = setNames(popsdf$cols, popsdf$pop)) +
  ylim(0,1) + 
  labs(title = "York River: Oyster Survival thru Time", 
       x = "Monitoring Event", 
       y = "Survival") +
  scale_x_continuous("Monitoring Event", breaks = c(1, 2, 3, 4),
                     labels = c("Fall '23", "Spring '24", "Fall '24", "Spring '25")) +
  theme_classic() + 
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.justification = "top")
yrk_surv

# export these graphs for publication
library(officer) # for powerpoint export/editing
library(rvg) # for converting r graphics to editable vector graphics

doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = lew_surv), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = yrk_surv), location = ph_location_type(type = "body"))
print(doc, target = "results/figures/surv_thru_time.pptx")
```

