---
title: "Oyster_survival_calc"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/MVP_Oyster_Analysis_Camille/MVP_Oysters_Analysis_Camille")

# load libraries
library(tidyverse) # for data wrangling
library(stringr) # for splitting strings
library(lubridate) # for dates
library(ggplot2) # for plotting
library(plotrix) # for std error
```

## Calculating Oyster Mortality in the Field

This script is to calculate mortality of oysters in the MVP Field Experiment from 2023-2025. 

First I need to read in and merge together relevant data. I will then clean this data up to prepare it for analysis by time point, site, population, and more.

```{r}
# read in data
bags <- read.csv("data/ooters_prelim/MVP23-FieldBags - bags.csv")
mort <- read.csv("data/ooters_prelim/MVP23-FieldBags - mortality.csv")
expsites <- read.csv("data/ooters_prelim/MVP23-FieldBags - sites.csv")
spawntrt <- read.csv("data/ooters_prelim/MVP23-FieldBags - spawn_trt2.csv")

# merging bags and mortality
head(bags)
colnames(bags) <- c("bags_key", "bag_site", "bags_label", "SpawnTrt_Key")
head(mort)
merge1 <- merge(bags, mort, by = c("bags_key"), all = T)
head(merge1)

# add spawntrt data
head(merge1)
head(spawntrt)
merge2 <- merge(merge1, spawntrt, by = c("SpawnTrt_Key"), all = T)
head(merge2)

# trim dataset
merge2_trim <- merge2[,c("bags_key","bag_site","bags_label","SpawnTrt_Key","SpawnTrt_Label",
                         "latitudeDecimal","longitudeDecimal","mortality_key","mortality_label",
                         "mortality_timestamp","alive_count","dead_count","alive_returned")]

# split timestamps to get dates - I'll use mortality_timestamp for this
# dates are 4/29/24 - 5/7/24 for most recent monitoring event
# quick checks - looks good!
head(merge2_trim[,c("mortality_timestamp")])
length(merge2_trim[,c("mortality_timestamp")])

# split timestamp column into two
datetime_split <- str_split_fixed(merge2_trim[,c("mortality_timestamp")], " ", 2) # split timestamp
colnames(datetime_split) <- c("mortality_date", "mortality_time") # name columns
merge3 <- cbind(merge2_trim, datetime_split) # add to full dataset

# take out practice rows
merge3_rm <- merge3[!merge3$SpawnTrt_Label == "Practice",]

# also put a monitoring event column in here
merge3_rm$mortality_date <- mdy(merge3_rm$mortality_date)
merge3_rm$event <- ifelse(as_date(merge3_rm$mortality_date) > as_date("2023-11-16"),
                                     ifelse(as_date(merge3_rm$mortality_date) > as_date("2024-05-09"), 3, 2),
                                     1)

# need to set a bag order
bag_order <- data.frame(bags_label = levels(as.factor(merge3_rm$bags_label)), order = 1:60)
merge4 <- merge(merge3_rm, bag_order, by = "bags_label")

# check that data makes sense!! alive count should never be greater than alive returned from previous event
# event 2
which(merge4[merge4$event == 2,]$alive_count > merge4[merge4$event == 1,]$alive_returned)
event2 <- merge4[merge4$event == 2,]
event2[which(merge4[merge4$event == 2,]$alive_count > merge4[merge4$event == 1,]$alive_returned),]$bags_label
# bags DEBY-3 and LA-3 had more alive in event 2 than in event 1...

# how much more?
## DEBY-3, 14 more oysters at t2 than t1
merge4[merge4$event == 2 & merge4$bags_label == "MVP-DEBY-3",]$alive_count - merge4[merge4$event == 1 & merge4$bags_label == "MVP-DEBY-3",]$alive_returned
## LA-3, 198 more oysters at t2 than t1
merge4[merge4$event == 2 & merge4$bags_label == "MVP-LA-3",]$alive_count - merge4[merge4$event == 1 & merge4$bags_label == "MVP-LA-3",]$alive_returned

# event 3
which(merge4[merge4$event == 3,]$alive_count > merge4[merge4$event == 2,]$alive_returned)
# all good!

# remove the funky bags from analysis
merge4_rm <- merge4[!merge4$bags_label == "MVP-DEBY-3",]
merge4_rm <- merge4_rm[!merge4_rm$bags_label == "MVP-LA-3",]

# add pop to df
merge4_rm$pop <- str_split_fixed(merge4_rm$bags_label, "-", 3)[,2]
```


Data is now ready for analysis. I will calculate survival as a proportion of those surviving from the previous timepoint.


```{r}
# function for survival calculation
surv_calc_bag <- function(dat, e) {
  dat_ord = dat[order(dat$order),]
  O = dat_ord[dat_ord$event == e,]$alive_count
  if(e - 1 < 1) {
    D = (dat_ord[dat_ord$event == e,]$alive_count) + (dat_ord[dat_ord$event == e,]$dead_count)
    S = 1
  } else {
    D = dat_ord[dat_ord$event == (e-1),]$alive_returned
    S = surv_calc_bag(dat_ord, (e-1))$surv
  }
  survival = (O/D) * S
  df = data.frame(bag = dat_ord[dat_ord$event == e,]$bags_label, event = e, surv = survival)
  return(df)
}

# calc survival
e1_surv <- surv_calc_bag(merge4_rm, 1)
e2_surv <- surv_calc_bag(merge4_rm, 2)
e3_surv <- surv_calc_bag(merge4_rm, 3)

# long format df
surv_long <- rbind(e1_surv,e2_surv,e3_surv)

# wide format df
colnames(e1_surv) <- c("bags_label","event","t1")
colnames(e2_surv) <- c("bags_label","event","t2")
colnames(e3_surv) <- c("bags_label","event","t3")

# final survival df
surv_df <- merge(e1_surv[,c("bags_label","t1")], 
                 e2_surv[,c("bags_label","t2")], 
                 by = c("bags_label"))
surv_df2 <- merge(surv_df, 
                  e3_surv[,c("bags_label","t3")], 
                  by = c("bags_label"))
surv_df2$t0 <- 1
surv_df2$pop <- str_split_fixed(surv_df2$bags_label, "-", 3)[,2]
surv_df2$bagnum <- 1:length(levels(as.factor(surv_df2$bags_label)))
surv_df_mg <- merge(surv_df2, bags, by = c("bags_label"))
surv_df_ord <- surv_df_mg[,c("pop","bags_label","bagnum","bag_site","SpawnTrt_Key","t0","t1","t2","t3")]

# can now calculate averages across sites and populations
# overall pop survival
pop_surv <- surv_df_ord %>% group_by(pop) %>% 
  summarise(mean_t0 = mean(as.numeric(t0)),
            sd_t0 = sd(as.numeric(t0)),
            se_t0 = std.error(as.numeric(t0)),
            mean_t1 = mean(as.numeric(t1)), 
            sd_t1 = sd(as.numeric(t1)),
            se_t1 = std.error(as.numeric(t1)),
            mean_t2 = mean(as.numeric(t2)), 
            sd_t2 = sd(as.numeric(t2)),
            se_t2 = std.error(as.numeric(t2)),
            mean_t3 = mean(as.numeric(t3)), 
            sd_t3 = sd(as.numeric(t3)),
            se_t3 = std.error(as.numeric(t3)),
            .groups = 'drop')
pop_surv

# overall site survival
site_surv <- surv_df_ord %>% group_by(bag_site) %>%
  summarise(mean_t0 = mean(as.numeric(t0)),
            sd_t0 = sd(as.numeric(t0)),
            se_t0 = std.error(as.numeric(t0)),
            mean_t1 = mean(as.numeric(t1)), 
            sd_t1 = sd(as.numeric(t1)),
            se_t1 = std.error(as.numeric(t1)),
            mean_t2 = mean(as.numeric(t2)), 
            sd_t2 = sd(as.numeric(t2)),
            se_t2 = std.error(as.numeric(t2)),
            mean_t3 = mean(as.numeric(t3)), 
            sd_t3 = sd(as.numeric(t3)),
            se_t3 = std.error(as.numeric(t3)),
            .groups = 'drop')
site_surv

# by site and pop
site_pop_surv <- surv_df_ord %>% group_by(bag_site,pop) %>%
  summarise(mean_t0 = mean(as.numeric(t0)),
            sd_t0 = sd(as.numeric(t0)),
            se_t0 = std.error(as.numeric(t0)),
            mean_t1 = mean(as.numeric(t1)), 
            sd_t1 = sd(as.numeric(t1)),
            se_t1 = std.error(as.numeric(t1)),
            mean_t2 = mean(as.numeric(t2)), 
            sd_t2 = sd(as.numeric(t2)),
            se_t2 = std.error(as.numeric(t2)),
            mean_t3 = mean(as.numeric(t3)), 
            sd_t3 = sd(as.numeric(t3)),
            se_t3 = std.error(as.numeric(t3)),
            .groups = 'drop')
site_pop_surv

# lets also put this in long format for plotting
e0 <- as.data.frame(site_pop_surv)[,c("bag_site","pop","mean_t0","sd_t0","se_t0")]
e0$event <- 0
e1 <- as.data.frame(site_pop_surv)[,c("bag_site","pop","mean_t1","sd_t1","se_t1")]
e1$event <- 1
e2 <- as.data.frame(site_pop_surv)[,c("bag_site","pop","mean_t2","sd_t2","se_t2")]
e2$event <- 2
e3 <- as.data.frame(site_pop_surv)[,c("bag_site","pop","mean_t3","sd_t3","se_t3")]
e3$event <- 3
colnames(e0) <- colnames(e1) <- colnames(e2) <- colnames(e3) <- c("bag_site","pop","mean_surv","sd","se","event")
mort_df <- rbind(e0, e1, e2, e3)

# save all these dataframes
write.csv(mort_df, "results/mortality_long_popsite_022025.csv")
write.csv(surv_df_ord, "results/mortality_bags_022025.csv")
```

All of my survival data is calculated, now I want to visualize trends in survival.

```{r}
# pops df for plotting
popsdf <- data.frame(pop = c("TX","LA","FL","JR","DEBY","LOLA","NH","ME","LARMIX","SEEDMIX"), 
                     order = 1:10, 
#                     cols = c("#6f1926","#de324c","#f4895f","#f8e16f","#95cf92","#369acc",
#                              "#9656a2","#cbabd1","black","gray"),
                     cols = c("#332288","#117733","#44AA99","#88CCEE","#DDCC77","#CC6677",
                              "#AA4499","#882255","black","gray"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     label = c("W1-TX","W2-LA","W3-FL","W4-VA","S1-LOLA","S2-DEBY",
                               "W5-NH","W6-ME","H1-LARMIX","H2-SEEDMIX"))
mort_df_full <- merge(mort_df, popsdf, by = c("pop"))

# york river
york_surv <- ggplot(data = mort_df_full[mort_df_full$bag_site == "YorkRiver",]) + 
  geom_line(aes(x = event, y = mean_surv, color = fct_reorder(pop,order))) + 
  scale_color_manual(name = "Population", label = popsdf$label, values = popsdf$cols)+
  geom_point(aes(x = event, y = mean_surv, fill = fct_reorder(pop,order), 
                 shape = fct_reorder(pop,order)),
             size = 4) + 
  scale_shape_manual(name = "Population", label = popsdf$label, values = popsdf$shape)+
  scale_fill_manual(name = "Population", label = popsdf$label, values = popsdf$cols)+
  #  geom_errorbar(aes(x = event, ymin = mean_surv-se, ymax = mean_surv+se), width = 0.05) + # error bars look terrible
  ylim(0,1) + 
  labs(color = "Population") +
  ylab("Survival") +
  scale_x_continuous("Monitoring Event", breaks = c(0,1,2,3),
                     labels = c("Spring '23", "Fall '23", "Spring '24", "Fall '24")) +
  theme_classic() + 
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.justification = "top") +
  ggtitle("York River: Oyster Survival thru Time")
york_surv

# lewisetta
lew_surv <- ggplot(data = mort_df_full[mort_df_full$bag_site == "Lewisetta",]) + 
  geom_line(aes(x = event, y = mean_surv, color = fct_reorder(pop,order))) + 
  scale_color_manual(name = "Population", label = popsdf$label, values = popsdf$cols)+
  geom_point(aes(x = event, y = mean_surv, fill = fct_reorder(pop,order), 
                 shape = fct_reorder(pop,order)),
             size = 4) + 
  scale_shape_manual(name = "Population", label = popsdf$label, values = popsdf$shape)+
  scale_fill_manual(name = "Population", label = popsdf$label, values = popsdf$cols)+
  #  geom_errorbar(aes(x = event, ymin = mean_surv-se, ymax = mean_surv+se), width = 0.05) + # error bars look terrible
  ylim(0,1) + 
  labs(color = "Population") +
  ylab("Survival") +
  scale_x_continuous("Monitoring Event", breaks = c(0,1,2,3),
                     labels = c("Spring '23", "Fall '23", "Spring '24", "Fall '24")) +
  theme_classic() + 
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.justification = "top") +
  ggtitle("Lewisetta: Oyster Survival thru Time")
lew_surv
```