---
title: "exp_envdat"
author: "Camille Rumberger, Nicole Mongill0, & Madeline Eppley"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
# set wd
setwd("~/Documents/Github/MVP_oyster_GO")

# packages
library(dplyr) # data cleaning
library(tidyr) # data cleaning
library(lubridate) # dealing with dates
library(readr) # read in data
library(ggplot2) # plotting
library(vroom) # data files

# data
# YORK
# downloaded 10/08/25 from VIMS http://vecos.vims.edu/StationDetail.aspx?param=YRK005.40&program=CMON
yrk_03_22 <- vroom(list.files(path = "data/EnvDat/exp/yrk_raw", pattern = "\\.csv$", full.names = TRUE)[1:20])

yrk_23 <- read.csv("data/EnvDat/exp/yrk_raw/York_2023.csv")
yrk_24 <- read.csv("data/EnvDat/exp/yrk_raw/York_2024.csv")
yrk_25 <- read.csv("data/EnvDat/exp/yrk_raw/York_2025.csv")

# lew data comes from Willy Reas from Coan River
raw_LOLA_env21 <- read.csv("data/EnvDat/exp/lew_raw/Jess_COAN_2021.csv")
raw_LOLA_env22 <- read.csv("data/EnvDat/exp/lew_raw/Jess_COAN_2022.csv")
raw_LOLA_env25 <- read.csv("data/EnvDat/exp/lew_raw/Jess_COAN_2025.csv")
raw_LOLA_env23_24 <- read.csv("data/EnvDat/exp/lew_raw/LEW-raw.csv")
raw_LOLA_correct <- read.csv("data/EnvDat/exp/lola_corrected.csv")

# experimental pops
exppop_sal <- read.csv("data/EnvDat/exp/all_salinity.csv")
exppop_temp <- read.csv("data/EnvDat/exp/all_temperature.csv")
```


## Pre-processing - York River

```{r}
# put together data
raw_yrk_exp <- rbind(yrk_23, yrk_24, yrk_25)

# subset relevant data
yrk_exp0 <- raw_yrk_exp[,c("SAMPLE_DATETIME", "WTEMP", "WTEMP_A", "SALINITY", "SALINITY_A")] 
yrk_exp0$SAMPLE_DATETIME <- as.POSIXct(yrk_exp0$SAMPLE_DATETIME, "%Y-%m-%d %H:%M:%S", tz = "")
yrk_sea0 <- yrk_03_22[,c("SAMPLE_DATETIME", "WTEMP", "WTEMP_A", "SALINITY", "SALINITY_A")]

parsed <- strptime(yrk_sea0$SAMPLE_DATETIME, "%m/%d/%Y %H:%M:%S")
yrk_sea0$SAMPLE_DATETIME <- as.POSIXct(parsed, "%Y/%m/%d %H:%M:%S", tz = "")

yrk_all <- rbind(yrk_sea0, yrk_exp0)

# keep just dates of experiment
yrk_all_exp <- subset(yrk_all,
                      SAMPLE_DATETIME >= as.POSIXct("2023-05-01 00:00:00.00",
                                                    tz = "") &
                      SAMPLE_DATETIME <= as.POSIXct("2025-05-28 00:00:00.00",
                                                    tz = ""))

# temp & sal subset, no error codes
yrk_noerr <- yrk_all[(yrk_all$WTEMP_A %in% c("", NA)), ]
yrk_all_sea <- yrk_noerr[(yrk_noerr$SALINITY_A %in% c("",NA)), ]

yrk_exp_noerr <- yrk_all_exp[(yrk_all_exp$WTEMP_A %in% c("", NA)), ]
yrk_exp <- yrk_exp_noerr[(yrk_exp_noerr$SALINITY_A %in% c("",NA)), ]

# trim + rename cols
yrk_exp <- yrk_exp[,c("SAMPLE_DATETIME", "WTEMP", "SALINITY")]
colnames(yrk_exp) <- c("datetime", "temp", "salinity")

yrk_allyrs <- yrk_all_sea[,c("SAMPLE_DATETIME", "WTEMP", "SALINITY")]
colnames(yrk_allyrs) <- c("datetime", "temp", "salinity")

# remove NAs
yrk_exp <- yrk_exp %>%
    filter(!is.na(datetime))

yrk_all_years <- yrk_allyrs %>%
  filter(!is.na(datetime))

# check
summary(yrk_exp)
summary(yrk_all_years)

# also store variables that we will include in the final data frame
site_name_y <- "YRK"
download_date_y <- "2025-10-08"
source_description_y <- "VECOS"
lat_y <- 37.247284
lon_y <- -76.499369
firstyear_y <- 2023
finalyear_y <- 2025

firstyear_yall <- 2003
finalyear_yall <- 2025
```


## Pre-processing: Lewisetta

This code comes in part from Nicole Mongillo.

```{r}
raw_LOLA_env23_24_prep <- raw_LOLA_env23_24[,c("Date","EST.Time","Temp","Salinity")]
colnames(raw_LOLA_env21) <- colnames(raw_LOLA_env22) <- colnames(raw_LOLA_env23_24) <- colnames(raw_LOLA_env25) <- c("Date","EST.Time","Temp","Salinity")
raw_LOLA_env <- rbind(raw_LOLA_env21, raw_LOLA_env22, raw_LOLA_env23_24_prep, raw_LOLA_env25)

#The data include separate columns for date and time, which I want to merge into one.
raw_LOLA_env$datetime <- paste(raw_LOLA_env$Date, raw_LOLA_env$EST.Time)

#Remove Date and EST.Time columns
raw_LOLA_env <- subset(raw_LOLA_env, select = -c(Date, EST.Time))

#Reorder columns so datetime is first
raw_LOLA_env <- raw_LOLA_env[, c(3,1,2)]

# View how the data are stored. Note the variable names and the format and units that the data are stored in.  
summary(raw_LOLA_env)

#rename columns. "datetime" = date and time of data collection, "temp" = water temperature in degrees C, "salinity" = water salinity in parts per thousand (ppt)
colnames(raw_LOLA_env) <- c("datetime", "temp", "salinity")

#Convert to POSIXct format. Tell R what the current date/time format is so it knows how to convert it. Store it into a column named datetime in the data frame.
raw_LOLA_env$datetime <- as.POSIXct(raw_LOLA_env$datetime, "%m/%d/%y %H:%M:%S", tz = "")

#add site name and create new data frame with full envr data set
raw_LOLA_env$site_name <- "LOLA"
LOLA_env_full <- raw_LOLA_env

#reorder columns with site_name first
LOLA_env_full <- LOLA_env_full[, c(4, 1, 2, 3)]

#separate out salinity and temp data into new frames for other analyses
LOLA_temp_full <- LOLA_env_full[, c(1,2,3)]

LOLA_sal_full <- LOLA_env_full[, c(1,2,4)]

#Store variables that we will include in the final data frame
site_name <- "LEW"
download_date <- as.Date("2025-10-14")
source_description <- "KCB hatchery in Coan River as part of CBNERR water quality monitoring, comes from Willy Reay"
lat_l <- 37.9803
lon_l <- -76.4619
#Lat and long data are from this site(https://www.vims.edu/cbnerr/monitoring/water_quality/). Data came from the KCB hatchery in the Coan as part of the CBNERR water quality monitoring initiative.
firstyear_l <- 2021
finalyear_l <- 2025

# also lat long for york river
lat_y <- 37.247284
lon_y <- -76.499369
```


## Data filtering

Filter any of the variables that have data points outside of normal range. We will use 0-40 as the accepted range for salinity (ppt) and temperature (C) values. Note, in the summer, salinity values can sometimes exceed 40.

```{r}
# qc filtering yrk
filtered_yrk_exp_sal <- yrk_exp %>%
    filter(between(salinity, 0, 40)) 
filtered_yrk_exp_env <- filtered_yrk_exp_sal %>%
    filter(between(temp, 0, 40))

filtered_yrk_sal <- yrk_all_years %>%
    filter(between(salinity, 0, 40)) 
filtered_yrk_env <- filtered_yrk_sal %>%
    filter(between(temp, 0, 40))

# store dat in new obj for ease
yrk <- filtered_yrk_env
yrk_exp <- filtered_yrk_exp_env

# save this
write.csv(yrk, "data/EnvDat/exp/yrk_raw/YRK-raw.csv")

# filter lew
# keep just dates of experiment
LOLA_env_full$datetime <- as.POSIXct(LOLA_env_full$datetime, "%Y-%m-%d %H:%M:%S", tz = "")
LOLA_env_exp <- subset(LOLA_env_full,
                      datetime >= as.POSIXct("2023-05-01 00:00:00.00",
                                                    tz = "") &
                      datetime <= as.POSIXct("2025-05-28 00:00:00.00",
                                                    tz = ""))

# qc filtering lew
filtered_lew_exp_sal <- LOLA_env_exp %>%
    filter(between(salinity, 0, 40)) 
filtered_lew_exp_env <- filtered_lew_exp_sal %>%
    filter(between(temp, 0, 40))

filtered_lew_sal <- LOLA_env_full %>%
    filter(between(salinity, 0, 40)) 
filtered_lew_env <- filtered_lew_sal %>%
    filter(between(temp, 0, 40))

# store dat in new obj for ease
lew_exp <- filtered_lew_exp_env
lew <- filtered_lew_env

# write out
write.csv(lew, "data/EnvDat/exp/lew_raw/lewisetta_coandat.csv")
```

## Plotting Trends

Check trends in environment by visualizing. This code comes in part from M. Eppley

```{r}
salplot_yrk <- ggplot(yrk, aes(x = datetime)) +
    geom_line(aes(y = salinity, color = "Salinity (ppt)")) +
    ylim(0,40) +
    labs(x = "Time", y = "Salinity ppt", title = "Salinity Plot for York River") +
    scale_color_manual(values = c("Salinity (ppt)" = "lightblue3")) +
    theme_minimal()
salplot_yrk

tempplot_yrk <- ggplot(yrk, aes(x = datetime)) +
    geom_line(aes(y = temp, color = "Temperature (C)")) +
    ylim(0, 40) +
    labs(x = "Time", y = "Temperature C", title = "Temperature Plot for York River") +
    scale_color_manual(values = c( "Temperature (C)" = "pink3")) +
    theme_minimal()
tempplot_yrk

salplot_lew <- ggplot(lew, aes(x = datetime)) +
    geom_line(aes(y = salinity, color = "Salinity (ppt)")) +
    ylim(0,40) +
    labs(x = "Time", y = "Salinity ppt", title = "Salinity Plot for Lewisetta") +
    scale_color_manual(values = c("Salinity (ppt)" = "lightblue3")) +
    theme_minimal()
salplot_lew

tempplot_lew <- ggplot(lew, aes(x = datetime)) +
    geom_line(aes(y = temp, color = "Temperature (C)")) +
    ylim(0, 40) +
    labs(x = "Time", y = "Temperature C", title = "Temperature Plot for Lewisetta") +
    scale_color_manual(values = c( "Temperature (C)" = "pink3")) +
    theme_minimal()
tempplot_lew
```

Also plot combined environmental data for just the experimental period

```{r}
# combine lew and yrk
colnames(lew_exp) <- c("site_name", "datetime", "temp_lew", "salinity_lew")
colnames(yrk_exp) <- c("datetime", "temp_yrk", "salinity_yrk")

exp_mg <- merge(lew_exp[,2:4], yrk_exp, by = "datetime", keep = T)

# plot
temp_gardens <- ggplot(exp_mg) +
    geom_line(aes(x = datetime, y = temp_lew, color = "Lewisetta")) +
    geom_line(aes(x = datetime, y = temp_yrk, color = "York River")) +
    ylim(0,35) +
    xlim(as.Date("2023-05-01"), as.Date("2025-05-01")) + 
    labs(x = "Time", y = "Temperature ºC", title = "Temperature in Common Gardens during the Experiment") +
    scale_color_manual(name = "Common Garden", values = c("Lewisetta" = "mediumspringgreen", "York River" = "dodgerblue")) +
    theme_classic()
temp_gardens

sal_gardens <- ggplot(exp_mg) +
    geom_line(aes(x = datetime, y = salinity_lew, color = "Lewisetta")) +
    geom_line(aes(x = datetime, y = salinity_yrk, color = "York River")) +
    ylim(0,30) +
    xlim(as.Date("2023-05-01"), as.Date("2025-05-01")) + 
    labs(x = "Time", y = "Salinity ppt", title = "Temperature in Common Gardens during the Experiment") +
    scale_color_manual(name = "Common Garden", values = c("Lewisetta" = "mediumspringgreen", "York River" = "dodgerblue")) +
    theme_classic()
sal_gardens
```


## Summary statistics

Calculate means, maxes, mins, medians, etc.

```{r}
# this chunk comes modified from environmental processing script written by M. Eppley to retain consistency in variable naming schemes

# YORK RIVER
# calculate mean, maximum, and minimum values for salinity and temperature for each month. 
yrk_env_mo <- yrk %>%
    dplyr::mutate(year = year(datetime), month = month(datetime)) %>%
    dplyr::group_by(year, month, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9),
      length_temp = length(temp))
  
(yrk_env_mo)


#Calculate the mean, maximum, and minimum values for salinity and temperature for each year. 
yrk_env_yr <- yrk %>%
    dplyr::mutate(year = year(datetime)) %>%
    dplyr::group_by(year, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9))

(yrk_env_yr)

#Calculate the mean, maximum, and minimum values for salinity and temperature for each day. 
yrk_env_day <- yrk %>%
    dplyr::mutate(year = year(datetime), month = month(datetime), day = day(datetime)) %>%
    dplyr::group_by(year, month, day, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9),
      length_temp = length(temp))
  
(yrk_env_day)

# check plots
timeplot <- ggplot(yrk_env_mo, aes(x = year)) +
    geom_point(aes(y = month, color = length_salinity), size = 4) +
    labs(x = "Time", y = "Month", title = "Salinity Timeplot for YRK - Gloucester Point, Virginia") +
    ylim(1,12) +
    theme_minimal()
timeplot

# LEWISETTA - use the "corrected" data from "LOLA_corrections.Rmd"
# calculate mean, maximum, and minimum values for salinity and temperature for each month. 
lew_env_mo <- lew %>%
    dplyr::mutate(year = year(datetime), month = month(datetime)) %>%
    dplyr::group_by(year, month, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9),
      length_temp = length(temp))
(lew_env_mo)

#Calculate the mean, maximum, and minimum values for salinity and temperature for each year. 
lew_env_yr <- lew %>%
    dplyr::mutate(year = year(datetime)) %>%
    dplyr::group_by(year, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9),
      length_temp = length(temp))

(lew_env_yr)

#Calculate the mean, maximum, and minimum values for salinity and temperature for each day. - this is for the 2021-2025 data, since I have already done the daily averages for 2014-2021 in the lola_corrections script
lew_env_day <- lew %>%
    dplyr::mutate(year = year(datetime), month = month(datetime), day = day(datetime)) %>%
    dplyr::group_by(year, month, day, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9),
      length_temp = length(temp))
  
(lew_env_day)

# check plots
timeplot <- ggplot(lew_env_mo, aes(x = year)) +
    geom_point(aes(y = month, color = length_salinity), size = 4) +
    labs(x = "Time", y = "Month", title = "Salinity Timeplot for LEW - Virginia") +
    ylim(1,12) +
    theme_minimal()
timeplot


# now for experimental time frame
# York River
yrk_envexp_yr <- yrk_exp %>%
    dplyr::mutate(year = year(datetime)) %>%
    dplyr::group_by(year, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9))

yrk_envexp_day <- yrk_exp %>%
    dplyr::mutate(year = year(datetime), month = month(datetime), day = day(datetime)) %>%
    dplyr::group_by(year, month, day, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9),
      length_temp = length(temp))

yrk_envexp_mo <- yrk_exp %>%
    dplyr::mutate(year = year(datetime), month = month(datetime)) %>%
    dplyr::group_by(year, month, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9),
      length_temp = length(temp))


# Lewisetta
lew_envexp_yr <- lew_exp %>%
    dplyr::mutate(year = year(datetime)) %>%
    dplyr::group_by(year, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9))

lew_envexp_day <- lew_exp %>%
    dplyr::mutate(year = year(datetime), month = month(datetime), day = day(datetime)) %>%
    dplyr::group_by(year, month, day, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9),
      length_temp = length(temp))
  
lew_envexp_mo <- lew_exp %>%
    dplyr::mutate(year = year(datetime), month = month(datetime)) %>%
    dplyr::group_by(year, month, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      salinity_quantile10 = quantile(salinity, 0.1),
      salinity_quantile90 = quantile(salinity, 0.9),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      temp_quantile10 = quantile(temp, 0.1),
      temp_quantile90 = quantile(temp, 0.9),
      length_temp = length(temp))
```


## Ground-truthing Lewisetta

```{r}
# directory for files
lew_sonde <- "data/EnvDat/exp/lew_raw/lew_sonde" 
pot_sonde <- "data/EnvDat/exp/lew_raw/potomac_sonde"

# Get a list of all .txt files in the directory
files_lew <- list.files(path = lew_sonde, pattern = "\\.txt$", full.names = T)
files_pot <- list.files(path = pot_sonde, pattern = "\\.txt$", full.names = T)

# Read each file into a list of data frames
lew_sonde_df <- lapply(files_lew, function(file) {
  read.table(file, sep = " ")
})
pot_sonde_df <- lapply(files_pot, function(file) {
  read.table(file, sep = " ")
})

# Combine all data frames in the list into a single data frame
lew_check <- bind_rows(lew_sonde_df)
pot_check <- bind_rows(pot_sonde_df)

# colnames
colnames(lew_check) <- c("YY", "MM", "DD", "hh", "mm", "WDIR", "WSPD", "GST", "WVHT", "DPD", "APD", "MWD", "PRES", "ATMP", "temp", "DEWP", "VIS", "TIDE")
colnames(pot_check) <- c("YY", "MM", "DD", "hh", "mm", "DEPTH", "OTMP", "COND", "sal", "O2%", "O2PPM", "CLCON", "TURB", "PH", "EH")

# just relevant data
lew_check_red <- lew_check[,c("YY","MM","DD", "hh", "mm","temp")]
pot_check_red <- pot_check[,c("YY","MM","DD", "hh", "mm", "sal")]

# put together datetime
lew_check_red$datetime <- paste0(lew_check_red$YY, "-", lew_check_red$MM, "-", lew_check_red$DD, " ", lew_check_red$hh, ":", lew_check_red$mm, ":00 EST")
pot_check_red$datetime <- paste0(pot_check_red$YY, "-", pot_check_red$MM, "-", pot_check_red$DD, " ", pot_check_red$hh, ":", pot_check_red$mm, ":00 EST")

lew_check_red$datetime <- as.POSIXct(lew_check_red$datetime, "%Y-%m-%d %H:%M:%S", tz = "")
pot_check_red$datetime <- as.POSIXct(pot_check_red$datetime, "%Y-%m-%d %H:%M:%S", tz = "")

# merge these datesets
check_lewisetta <- merge(lew_check_red[,c("datetime","temp")], pot_check_red[,c("datetime","sal")], by = "datetime", all = T)

# check data
check_lewisetta$temp[check_lewisetta$temp == 999] <- NA
check_lewisetta$sal[check_lewisetta$sal == 99] <- NA
summary(check_lewisetta)

# qc filter
filtered_check_lewisetta <- check_lewisetta %>%
    filter(between(sal, 0, 40)) 
filtered_check_lewisetta <- filtered_check_lewisetta %>%
    filter(between(temp, 0, 40))

# vims data is every 15 mins vs more every hour at noaa buoys -- use daily averages to calculate the other summary stats
filtered_check_lewisetta$date <- as.Date(filtered_check_lewisetta$datetime, tz = "EST")
daily_check_lewisetta <- filtered_check_lewisetta %>% 
  dplyr::group_by(date) %>% 
  dplyr::mutate(mean_daily_temp = mean(temp), mean_daily_sal = mean(sal))

# do the same with other lewisetta data
lew_daily <- lew
lew_daily$date <- as.Date(lew$datetime, tz = "EST")
daily_lewisetta <- lew_daily %>% 
  dplyr::group_by(date) %>% 
  dplyr::mutate(mean_daily_temp = mean(temp), mean_daily_sal = mean(salinity))

#filter so each day is represented once
day_check_lewisetta <- daily_check_lewisetta[match(unique(daily_check_lewisetta$date), daily_check_lewisetta$date),]
noaa_day_filtered <- day_check_lewisetta[day_check_lewisetta$date %in% as.Date(daily_lewisetta$datetime, tz = "EST"),]
lola_day_filtered <- daily_lewisetta[as.Date(daily_lewisetta$datetime, tz = "EST") %in% day_check_lewisetta$date,]

# put together full dataset
colnames(noaa_day_filtered) <- c("datetime", "temp_noaa", "sal_noaa", "date", "mean_daily_temp_noaa", "mean_daily_sal_noaa")
lew_mg <- merge(noaa_day_filtered, lola_day_filtered, by = c("datetime", "date"), all = T) 

# plot to look at differences
filter_tempplot <- ggplot(lew_mg, aes(x=datetime, y = mean_daily_temp, color = "VIMS")) +
    geom_line()+
    geom_line(data = lew_mg, aes(x=datetime, y = mean_daily_temp_noaa, color = "NOAA"))+
  scale_color_manual(values=c("black", "red"))+
  labs(x = "Time", y = "Temperature C", title = "Temperature Plot for NOAA and VIMS Data Sets") 
filter_tempplot # basically right on top of each other

filter_salplot <- ggplot(lew_mg, aes(x=datetime, y = mean_daily_sal, color = "VIMS")) +
    geom_line()+
    geom_line(data = lew_mg, aes(x=datetime, y = mean_daily_sal_noaa, color = "NOAA"))+
  scale_color_manual(values=c("black", "blue"))+
  labs(x = "Time", y = "Salinity psu", title = "Salinity Plot for NOAA and VIMS Data Sets") 
filter_salplot # trends follow, but need a large correction

# corrections
lew_mg <- lew_mg %>% 
  mutate(temp_diff = mean_daily_temp-mean_daily_temp_noaa, sal_diff = mean_daily_sal -mean_daily_sal_noaa)

temp_correction <- mean(lew_mg$temp_diff, na.rm = T)
# VIMS data (from Willy Coan) is 0.09287575ºC above NOAA data, apply this correction to past NOAA data (2014-2020)

sal_correction <- mean(lew_mg$sal_diff, na.rm = T)
# VIMS data (from Willy Coan) is -1.192284 below NOAA data, apply this correction to past NOAA data (2014-2020)

# apply correction
day_check_lewisetta$temp_corrected <- day_check_lewisetta$mean_daily_temp + temp_correction
day_check_lewisetta$sal_corrected <- day_check_lewisetta$mean_daily_sal + sal_correction

# write it out
lew_correct <- day_check_lewisetta
lew_correct <- lew_correct[!(is.na(lew_correct$datetime)),]
write.csv(lew_correct, "data/EnvDat/exp/lew_raw/lewisetta_corrected.csv", row.names = F)

# summary of these data
lew_correct_yr <- lew_correct %>%
    dplyr::mutate(year = year(datetime)) %>%
    dplyr::group_by(year, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(sal_corrected),
      max_salinity = max(sal_corrected),
      mean_salinity = mean(sal_corrected),
      salinity_quantile10 = quantile(sal_corrected, 0.1),
      salinity_quantile90 = quantile(sal_corrected, 0.9),
      min_temp = min(temp_corrected),
      max_temp = max(temp_corrected),
      mean_temp = mean(temp_corrected),
      temp_quantile10 = quantile(temp_corrected, 0.1),
      temp_quantile90 = quantile(temp_corrected, 0.9))
```

## Putting it all together

THIS COMES FROM MADELINE'S SCRIPT

We can now calculate a list of variables that we will have collected for all sites. This will allow us to compare sites easily. We will calculate the number of observations from each site, the mean annual, maximum annual, and minimum annual value for all variables. 

Our list of variables includes: 

- Mean_Annual_Temperature_C: average of all available data
- Mean_max_temperature_C: average of maximums for each year
- Mean_min_temperature_C: average of minimums for each year
- temp_quantile10: 10% quantile of all available data
- temp_quantile90: 90% quantile of all available data
- Temperature_st_dev: standard deviation of all available data
- Temperature_n: total number of data points
- Temperature_years: number of years in data set

- Mean_Annual_Salinity_ppt: average of all available data
- Mean_min_Salinity_ppt: average of minimums for each year
- Mean_max_Salinity_ppt: average of maximums for each year
- sal_quantile10: 10% quantile of all available data
- sal_quantile90: 90% quantile of all available data
- Salinity_st_dev: standard deviation of all available data
- Salinity_n: total number of data points
- Salinity_years: number of years in data set

### York River

Start with the temperature variables

```{r}
#Calculate temperature variables. 
Mean_Annual_Temperature_C <- mean(yrk$temp)
Mean_max_temperature_C <- mean(yrk_env_yr$max_temp)
Mean_min_temperature_C <- mean(yrk_env_yr$min_temp)
temp_quantile_10 <- quantile(yrk$temp, 0.1)
temp_quantile_90 <- quantile(yrk$temp, 0.9)
Temperature_st_dev <- sd(yrk$temp)
Temperature_n <- nrow(yrk)
Temperature_years <- nrow(yrk_env_yr)

# Create a data frame to store the temperature results
yrk_temp <- cbind(data.frame(dataset = "sea", site_name = "S2_DEBY"), download_date, source_description, lat = lat_y, lon = lon_y, firstyear = 2003, finalyear = 2025, Mean_Annual_Temperature_C, Mean_max_temperature_C, Mean_min_temperature_C, temp_quantile_10, temp_quantile_90, Temperature_st_dev, Temperature_n, Temperature_years)
print(yrk_temp)

# Write to a unique new CSV file
write.csv(yrk_temp, paste0("data/EnvDat/seascape/yrk_temperature_",Sys.Date(),".csv"))
```

```{r}
#Calculate temperature variables. 
Mean_Annual_Temperature_C <- mean(yrk_exp$temp)
Mean_max_temperature_C <- mean(yrk_envexp_yr$max_temp)
Mean_min_temperature_C <- mean(yrk_envexp_yr$min_temp)
temp_quantile_10 <- quantile(yrk_exp$temp, 0.1)
temp_quantile_90 <- quantile(yrk_exp$temp, 0.9)
Temperature_st_dev <- sd(yrk_exp$temp)
Temperature_n <- nrow(yrk_exp)
Temperature_years <- nrow(yrk_exp)

# Create a data frame to store the temperature results
yrk_exp_temp <- cbind(data.frame(dataset = "exp", site_name = "YRK"), download_date, source_description, lat = lat_y, lon = lon_y, firstyear = 2022, finalyear = 2025, Mean_Annual_Temperature_C, Mean_max_temperature_C, Mean_min_temperature_C, temp_quantile_10, temp_quantile_90, Temperature_st_dev, Temperature_n, Temperature_years)
print(yrk_exp_temp)

# Write to a unique new CSV file
write.csv(yrk_exp_temp, paste0("data/EnvDat/exp/yrk_temperature_",Sys.Date(),".csv"))
```

Now the salinity variables

```{r}
#Calculate the salinity variables
Mean_Annual_Salinity_ppt <- mean(yrk$salinity)
Mean_max_Salinity_ppt <- mean(yrk_env_yr$max_salinity)
Mean_min_Salinity_ppt <- mean(yrk_env_yr$min_salinity)
salinity_quantile_10 <- quantile(yrk$salinity, 0.1)
salinity_quantile_90 <- quantile(yrk$salinity, 0.9)
Salinity_st_dev <- sd(yrk$salinity)
Salinity_n <- nrow(yrk)
Salinity_years <- nrow(yrk_env_yr)

#Create a data frame to store the temperature results
yrk_salinity <- cbind(data.frame(dataset = "sea", site_name = "S2_DEBY"), download_date, source_description, lat = lat_y, lon = lon_y, firstyear = 2003, finalyear = 2025, Mean_Annual_Salinity_ppt, Mean_max_Salinity_ppt, Mean_min_Salinity_ppt, salinity_quantile_10, salinity_quantile_90, Salinity_st_dev, Salinity_n, Salinity_years)
print(yrk_salinity)

# Write all year data to a unique new CSV file
write.csv(yrk_salinity, paste0("data/EnvDat/seascape/yrk_salinity_",Sys.Date(),".csv"))
```

```{r}
#Calculate the salinity variables
Mean_Annual_Salinity_ppt <- mean(yrk_exp$salinity)
Mean_max_Salinity_ppt <- mean(yrk_envexp_yr$max_salinity)
Mean_min_Salinity_ppt <- mean(yrk_envexp_yr$min_salinity)
salinity_quantile_10 <- quantile(yrk_exp$salinity, 0.1)
salinity_quantile_90 <- quantile(yrk_exp$salinity, 0.9)
Salinity_st_dev <- sd(yrk_exp$salinity)
Salinity_n <- nrow(yrk_exp)
Salinity_years <- nrow(yrk_envexp_yr)

#Create a data frame to store the temperature results
yrkexp_salinity <- cbind(data.frame(dataset = "exp", site_name = "YRK"), download_date, source_description, lat = lat_y, lon = lon_y, firstyear = 2022, finalyear = 2025, Mean_Annual_Salinity_ppt, Mean_max_Salinity_ppt, Mean_min_Salinity_ppt, salinity_quantile_10, salinity_quantile_90, Salinity_st_dev, Salinity_n, Salinity_years)
print(yrkexp_salinity)

# Write all year data to a unique new CSV file
write.csv(yrkexp_salinity, paste0("data/EnvDat/exp/exp_yrk_salinity_",Sys.Date(),".csv"))
```

### Lewisetta

Start with the temperature variables

```{r}
#Calculate temperature variables. 
Mean_Annual_Temperature_C <- mean(lew_correct$temp_corrected)
Mean_max_temperature_C <- mean(lew_correct_yr$max_temp)
Mean_min_temperature_C <- mean(lew_correct_yr$min_temp)
temp_quantile_10 <- quantile(lew_correct$temp_corrected, 0.1)
temp_quantile_90 <- quantile(lew_correct$temp_corrected, 0.9)
Temperature_st_dev <- sd(lew_correct$temp_corrected)
Temperature_n <- nrow(lew_correct)
Temperature_years <- nrow(lew_correct_yr)

# Create a data frame to store the temperature results
lew_temp <- cbind(data.frame(dataset = "sea", site_name = "S1_LOLA"), download_date, source_description, lat = lat_l, lon = lon_l, firstyear = 2014, finalyear = 2024, Mean_Annual_Temperature_C, Mean_max_temperature_C, Mean_min_temperature_C, temp_quantile_10, temp_quantile_90, Temperature_st_dev, Temperature_n, Temperature_years)
print(lew_temp)

# Write to a unique new CSV file
write.csv(lew_temp, paste0("data/EnvDat/seascape/lew_temperature_",Sys.Date(),".csv"))
```

```{r}
#Calculate temperature variables. 
Mean_Annual_Temperature_C <- mean(lew_exp$temp)
Mean_max_temperature_C <- mean(lew_envexp_yr$max_temp)
Mean_min_temperature_C <- mean(lew_envexp_yr$min_temp)
temp_quantile_10 <- quantile(lew$temp, 0.1)
temp_quantile_90 <- quantile(lew$temp, 0.9)
Temperature_st_dev <- sd(lew_exp$temp)
Temperature_n <- nrow(lew_exp)
Temperature_years <- nrow(lew_envexp_yr)
high_temp_stress_days <- sum(high_temp_stress_count$high_temp_stress)
frac_high_temp_stress_days <- high_temp_stress_days/nrow(lew_envexp_day)

# Create a data frame to store the temperature results
lew_exp_temp <- cbind(data.frame(dataset = "exp", site_name = "LEW"), download_date, source_description, lat = lat_l, lon = lon_l, firstyear = 2022, finalyear = 2025, Mean_Annual_Temperature_C, Mean_max_temperature_C, Mean_min_temperature_C, temp_quantile_10, temp_quantile_90, Temperature_st_dev, Temperature_n, Temperature_years)
print(lew_exp_temp)

# Write to a unique new CSV file
write.csv(lew_exp_temp, paste0("data/EnvDat/exp/lew_temperature_",Sys.Date(),".csv"))
```

Now the salinity variables

```{r}
#Calculate the salinity variables
Mean_Annual_Salinity_ppt <- mean(lew_correct$sal_corrected, na.rm = T)
Mean_max_Salinity_ppt <- mean(lew_correct_yr$max_salinity, na.rm = T)
Mean_min_Salinity_ppt <- mean(lew_correct_yr$min_salinity, na.rm = T)
salinity_quantile_10 <- quantile(lew_correct$sal_corrected, 0.1, na.rm = T)
salinity_quantile_90 <- quantile(lew_correct$sal_corrected, 0.9, na.rm = T)
Salinity_st_dev <- sd(lew_correct$sal_corrected, na.rm = T)
Salinity_n <- nrow(lew_correct)
Salinity_years <- nrow(lew_correct_yr)

#Create a data frame to store the temperature results
lew_salinity <- cbind(data.frame(dataset = "sea", site_name = "S1_LOLA"), download_date, source_description, lat = lat_l, lon = lon_l, firstyear = 2014, finalyear = 2024, Mean_Annual_Salinity_ppt, Mean_max_Salinity_ppt, Mean_min_Salinity_ppt, salinity_quantile_10, salinity_quantile_90, Salinity_st_dev, Salinity_n, Salinity_years)
print(lew_salinity)

# Write all year data to a unique new CSV file
write.csv(lew_salinity, paste0("data/EnvDat/seascape/lew_salinity_",Sys.Date(),".csv"))
```

```{r}
#Calculate the salinity variables
Mean_Annual_Salinity_ppt <- mean(lew_exp$salinity)
Mean_max_Salinity_ppt <- mean(lew_envexp_yr$max_salinity)
Mean_min_Salinity_ppt <- mean(lew_envexp_yr$min_salinity)
salinity_quantile_10 <- quantile(lew_exp$salinity, 0.1)
salinity_quantile_90 <- quantile(lew_exp$salinity, 0.9)
Salinity_st_dev <- sd(lew_exp$salinity)
Salinity_n <- nrow(lew_exp)
Salinity_years <- nrow(lew_envexp_yr)

#Create a data frame to store the temperature results
lewexp_salinity <- cbind(data.frame(dataset = "exp", site_name = "LEW"), download_date, source_description, lat = lat_l, lon = lon_l, firstyear = 2022, finalyear = 2025, Mean_Annual_Salinity_ppt, Mean_max_Salinity_ppt, Mean_min_Salinity_ppt, salinity_quantile_10, salinity_quantile_90, Salinity_st_dev, Salinity_n, Salinity_years)
print(lewexp_salinity)

# Write all year data to a unique new CSV file
write.csv(lewexp_salinity, paste0("data/EnvDat/exp/exp_lew_salinity_",Sys.Date(),".csv"))
```

Put together the full environmental dataset with all quantiles. 

```{r}
# make datasets compatible
exppop_sal$dataset <- exppop_temp$dataset <- "exp"
exppop_sal$collection_type <- exppop_temp$collection_type <- NULL

# remove exp sites from these datasheets
exppop_sal_mg <- exppop_sal %>% dplyr::filter(!(site_name %in% c("LEW","YRK", "S1-LOLA", "S2-DEBY")))
exppop_temp_mg <- exppop_temp %>% dplyr::filter(!(site_name %in% c("LEW","YRK", "S1-LOLA", "S2-DEBY")))

# final experimental dataset
sal_cols <- colnames(lew_salinity)
all_salinity <- rbind(exppop_sal_mg[,sal_cols], lew_salinity, lewexp_salinity, yrk_salinity, yrkexp_salinity)
temp_cols <- colnames(lew_temp)
all_temp <- rbind(exppop_temp_mg[,temp_cols], lew_temp, lew_exp_temp, yrk_temp, yrk_exp_temp)

write.csv(all_salinity, "data/EnvDat/exp/exp_salinity.csv")
write.csv(all_temp, "data/EnvDat/exp/exp_temp.csv")
```
