---
title: "CviMVP_oyster_EnvDatProcess"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Environmental Data Processing

This script is to process environmental data for both the seascape and experiment by reducing the variable set to retain uncorrelated variables, standardize data, and convert data to the correct formats for downstream analysis. 

The seascape data comes from...

The experimental data comes from...

## Setup

```{r}
# set wd
setwd("~/Documents/Github/MVP_oyster_GO")

# packages
library(vegan)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(psych)
library(LEA)
library(stringr)

# data
seascapeEnv <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv")
expEnv <- read.csv("data/EnvDat/exp/envr_summary_stats.csv")[,-1]
expsal <- read.csv("data/EnvDat/exp/all_salinity.csv")
exptemp <- read.csv("data/EnvDat/exp/all_temperature.csv")
```


## Data Pre-processing

Will start by checking the environmental data structure and processing to retain only  biologically relevant & uncorrelated variables.

```{r}
# check data
seascapeEnv
expEnv
colnames(seascapeEnv)
colnames(expEnv)

# check data correlations within each dataset
# seascape
pairs.panels(seascapeEnv[,c("Mean_Annual_Salinity_ppt","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","Salinity_st_dev","Mean_Annual_Temperature_C","Mean_max_temperature_C","Mean_min_temperature_C","Temperature_st_dev")])
pairs.panels(expEnv[,c("Mean_Annual_Salinity_ppt","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","Salinity_st_dev","Mean_Annual_Temperature_C","Mean_max_temperature_C","Mean_min_temperature_C","Temperature_st_dev")])

# remove some of these correlated variables
# seascape & env keep maxes & mins, remove all else (need the same variables in both sets)
seascapeEnv_trim <- seascapeEnv[,c("ID_Site","Site_name","Site_ID2","Site_Order","Site_Color","Region","Region_color","Latitude_manual","Longitude_manual","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","Mean_max_temperature_C","Mean_min_temperature_C")]
expEnv_trim <- expEnv[,c("site_name","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","Mean_max_temperature_C","Mean_min_temperature_C")]

# THIS IS NEW METHOD, THE ABOVE PROB DOESN'T WORK QUITE RIGHT
# put together temp and sal
expsal_red <- expsal[,c("site_name","lat","lon","Mean_Annual_Salinity_ppt","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","salinity_quantile_10","salinity_quantile_90")]
exptemp_red <- exptemp[,c("site_name","lat","lon","Mean_Annual_Temperature_C","Mean_max_temperature_C","Mean_min_temperature_C","temp_quantile_10","temp_quantile_90")]
expenv_mg <- merge(expsal_red, exptemp_red, by = c("site_name"))

# look at corrs
pairs.panels(expenv_mg[,c("Mean_Annual_Salinity_ppt","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","salinity_quantile_10","salinity_quantile_90","Mean_Annual_Temperature_C","Mean_max_temperature_C","Mean_min_temperature_C","temp_quantile_10","temp_quantile_90")])

# keeping just quantiles
expenv_trim <- expenv_mg[,c("salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90")]
```

Everything is looking good! Now I will standardize data to be scaled and centered for downstream analysis.

```{r}
# quick data inspection
head(seascapeEnv_trim)
head(expEnv_trim)

# scale
std_sea <- scale(seascapeEnv_trim[,10:13])
std_exp <- scale(expEnv_trim[,2:5])
colnames(std_sea) <- colnames(std_exp) <-  c("max_sal_scaled","min_sal_scaled","max_temp_scaled","min_temp_scaled")

# attach scaled information back to env dfs
sea_scaled <- as.data.frame(cbind(seascapeEnv_trim, std_sea))
exp_scaled <- as.data.frame(cbind(expEnv_trim, std_exp))

# keep track of scaling factors
sea_scale_env <- attr(std_sea, "scaled:scale")
sea_center_env <- attr(std_sea, "scaled:center")
exp_scale_env <- attr(std_exp, "scaled:scale")
exp_center_env <- attr(std_exp, "scaled:center")

# put those scaling factors into their own df
scaling <- as.data.frame(rbind(sea_scale_env, sea_center_env, exp_scale_env, exp_center_env))
scaling$factor_type <- rownames(scaling)
rownames(scaling) <- NULL
scaling

# do it again now for the expenv_mg data
# quick data inspection
head(expenv_trim)

# scale
std_expmg <- scale(expenv_trim)
colnames(std_expmg) <-  c("sal_quant10_scaled","sal_quant90_scaled","temp_quant10_scaled","temp_quant90_scaled")

# attach scaled information back to env dfs
expmg_scaled <- as.data.frame(cbind(expenv_mg, std_expmg))

# keep track of scaling factors
expmg_scale_env <- attr(std_expmg, "scaled:scale")
expmg_center_env <- attr(std_expmg, "scaled:center")

# write out this env data
write.csv(expmg_scaled, "data/EnvDat/exp/expquants_scaled_04222025.csv")
```


Finally I want to make sure now that my environmental data will match my genomic data by having one row in my environmental dataset for each individual in my genomic dataset. To do so, I will need to process the individual data and pair this with my environmental data.

```{r}
# experimental individuals - would replace this with the seascape individuals for lining up the environmental data with genetic data, no need to actually have indiv level predictions for experiment just pop level

# need ind level dat
indsExp <- readRDS("data/IndDat/20240922_experimental_indsmatrix.rds")

# keep just relevant data
indsExp_red <- indsExp[,c("merge_ID_inds","Label_ind","Key_ind","ID_SiteDate","SiteDate_Key")]
levels(as.factor(indsExp_red$ID_SiteDate))

# need to add scaled env data to this
exp_scaled$site_name
indsExp_red$ID_SiteDate

# the pop (ID_SiteDate/site_name) data doesn't match up
map <- list("TX-CapBay" = "W1-TX", "LA-SisLake" = "W2-LA", "FL-KingPlan" = "W3-FL", "LOLA" = "S1-LOLA", "DEBY" = "S2-DEBY", "VA-DeepWatSh" = "W4-VA", "NH-GrtBay" = "W5-NH", "ME-HogIs" = "W6-ME")
indsExp_red_mut <- indsExp_red %>% mutate(ID_SiteDate = ifelse(ID_SiteDate %in% names(map), map[ID_SiteDate], ID_SiteDate))
colnames(indsExp_red_mut) <- c("merge_ID_inds", "Label_ind", "Key_ind", "site_name", "SiteDate_Key")

which(!(exp_scaled$site_name %in% indsExp_red_mut$site_name))
which(!(indsExp_red_mut$site_name %in% exp_scaled$site_name)) # substitution worked!

indsExp_red_mut$site_name <- as.character(indsExp_red_mut$site_name)

# how many inds per site?
table(indsExp_red_mut$site_name) # 17-20

# merge
exp_scaled_expanded <- merge(as.data.frame(indsExp_red_mut), as.data.frame(exp_scaled), by = c("site_name"))

# write this out to .csv
write.csv(exp_scaled_expanded, paste0("data/EnvDat/exp/exp_scaled_expanded", Sys.Date(), ".csv"))

# do it for the other exp data
expmg_scaled
indsExp_red_mut

expmg_scaled$site_name <- gsub("_","-",expmg_scaled$site_name)

expmg_scaled_expanded <- merge(as.data.frame(indsExp_red_mut), as.data.frame(expmg_scaled), by = c("site_name"))

# we lost lew and york because they weren't technically lines/pops, I actually need to attach the Lew/york info as separate columns
lewyrk_scaled <- as.data.frame(expmg_scaled[expmg_scaled$site_name %in% c("LEW","YRK"),])
lew_scaled <- as.data.frame(expmg_scaled[expmg_scaled$site_name == "LEW",c("site_name","salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90","sal_quant10_scaled","sal_quant90_scaled","temp_quant10_scaled","temp_quant90_scaled")])
yrk_scaled <- as.data.frame(expmg_scaled[expmg_scaled$site_name == "YRK",c("site_name","salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90","sal_quant10_scaled","sal_quant90_scaled","temp_quant10_scaled","temp_quant90_scaled")])
colnames(lew_scaled) <- paste0("LEW_",colnames(lew_scaled))
colnames(yrk_scaled) <- paste0("YRK_",colnames(yrk_scaled))

# add columns to larger df
expmg_scaled_expanded_mg1 <- cbind(expmg_scaled_expanded, lew_scaled[rep(seq_len(nrow(lew_scaled)), 154),-1])
expmg_scaled_expanded_mg2 <- cbind(expmg_scaled_expanded_mg1, yrk_scaled[rep(seq_len(nrow(yrk_scaled)), 154),-1])

# save this output
write.csv(expmg_scaled_expanded_mg2, "data/EnvDat/exp/expsitesquants_scaled_04222025.csv")
```

## Environmental Distance

In addition to processing this environmental data, I want to calculate environmental distance for downstream analyses. I will also visualize this distance.

```{r}
# data
expenv_dist_df <- expenv_mg

# change the site_names to match
expenv_dist_df$site_name <- sub("_","-", expenv_dist_df$site_name)

# reduce variable set
expenv_dist_red_df <- expenv_dist_df[c("site_name", "lat.x", "lon.x", "salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90")]

# standardize
std_expenv <- scale(expenv_dist_red_df[,c("salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90")])
colnames(std_expenv) <- c("std_salinity_quantile_10", "std_salinity_quantile_90", "std_temp_quantile_10", "std_temp_quantile_90")
expenv_dist_std_df <- cbind(expenv_dist_red_df, std_expenv)

# calculate environmental distance
library(vegan)
env_dist <- round(vegdist(expenv_dist_std_df[,c("std_salinity_quantile_10", "std_salinity_quantile_90", "std_temp_quantile_10", "std_temp_quantile_90")], 
                          method = "euclidian", upper = FALSE, diag = TRUE), 4)
env_dist_df <- as.data.frame(as.matrix(env_dist))
colnames(env_dist_df) <- rownames(env_dist_df) <- expenv_dist_std_df$site_name
env_dist_df

# matrix
env_dist_mat <- cbind(rownames(env_dist_df),env_dist_df)

# export matrix
write.csv(env_dist_mat, paste0("data/EnvDat/exp/distance_matrix_",Sys.Date(),".csv"))

# visualize this distance as heatmap
library(reshape2)
env_dist_longer <- melt(env_dist_mat)
head(env_dist_longer)
colnames(env_dist_longer) <- c("site1","site2","dist")

popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME", "LARMIX", "SEEDMIX"), 
                     order = 2:11, 
                     cols = c("#3e4989" ,"#31688e", "#26828e", "#440154", "#482878", "#1f9e89", "#35b779", "#6ece58", "#fde725", "#b5de2b"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site_name = c("W1-TX", "W2-LA", "W3-FL", "W4-VA", "S1-LOLA", "S2-DEBY", "W5-NH", "W6-ME", "H1-LARMIX", "H2-SEEDMIX"))
popsdf0 <- rbind(data.frame(pop = c("YRK","LEW"), order = 0:1, cols = c("white","black"), shape = c(22,22), site_name = c("YRK","LEW")), popsdf)
popsdf1 <- popsdf0
popsdf2 <- popsdf0
colnames(popsdf1) <- c("pop","order1","cols","shape","site1")
colnames(popsdf2) <- c("pop","order2","cols","shape","site2")
env_dist_long_mg <- merge(popsdf1[,c("site1","order1")], env_dist_longer, by = "site1")
env_dist_long_mg2 <- merge(popsdf2[,c("site2","order2")], env_dist_long_mg, by = "site2")

# keep this too
write.csv(env_dist_long_mg2, paste0("data/EnvDat/exp/distance_matrix_longer_", Sys.Date(), ".csv"))

# visualize
library(viridis)
envdist_plot <- ggplot(env_dist_long_mg2, aes(x = fct_reorder(site1, order1), 
                              y = fct_reorder(site2, order2), 
                              fill = value))+
  theme_classic()+
  geom_tile(color = "black") +
#  geom_text(aes(label = distances), color = "white", size = 4)+
  scale_fill_viridis_c(name = "Euclidean 
Distance", limits = c(0, 5))+
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top") +
  labs(title = "Environmental Distance between Sites", x = "", y = "") +
  theme(plot.title = element_text(size = 18), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top")
envdist_plot

# show just the lew and york comparisons
env_dist_long_expsite <- env_dist_long_mg2[env_dist_long_mg2$site2 %in% c("LEW", "YRK"),]
env_dist_long_expsite_red <- env_dist_long_expsite[!(env_dist_long_expsite$site1 %in% c("LEW","YRK")),]

# plot
envdist_plot_red <- ggplot(env_dist_long_expsite_red, aes(x = fct_reorder(site1, order1), 
                              y = fct_reorder(site2, order2), 
                              fill = value))+
  theme_classic()+
  geom_tile(color = "black") +
#  geom_text(aes(label = distances), color = "white", size = 4)+
  scale_fill_viridis_c(name = "Euclidean 
Distance", limits = c(0, 4))+
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top") +
  labs(title = "Environmental Distance between Sites", x = "", y = "") +
  theme(plot.title = element_text(size = 18), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top")
envdist_plot_red
```


## Formatting for LEA

LEA requires environmental data be formatted as a .env file type.

```{r}
# prep as matrix
#sea_mat <- as.matrix(sea_scaled[,c("max_sal_scaled","min_sal_scaled","max_temp_scaled","min_temp_scaled")])
exp_mat <- as.matrix(exp_scaled_expanded[,c("max_sal_scaled","min_sal_scaled","max_temp_scaled","min_temp_scaled")]) # would only actually need exp_scaled because don't need each individual to have assigned data, just pop level predictions
expquants_mat <- as.matrix(expmg_scaled_expanded_mg2[,c("sal_quant10_scaled","sal_quant90_scaled","temp_quant10_scaled","temp_quant90_scaled")])
lewquants_mat <- as.matrix(expmg_scaled_expanded_mg2[,c("LEW_sal_quant10_scaled","LEW_sal_quant90_scaled","LEW_temp_quant10_scaled","LEW_temp_quant90_scaled")])
yrkquants_mat <- as.matrix(expmg_scaled_expanded_mg2[,c("YRK_sal_quant10_scaled","YRK_sal_quant90_scaled","YRK_temp_quant10_scaled","YRK_temp_quant90_scaled")])

# and convert to env format
#write.env(sea_mat, "data/EnvDat/sea/sea_scaled.env")
#write.env(exp_mat, "data/EnvDat/exp/exp_scaled.env")
write.env(exp_mat, "data/EnvDat/exp/exp_scaled_expanded_04222025.env")
write.env(expquants_mat, "data/EnvDat/exp/expquant_scaled_expanded_04222025.env")
write.env(lewquants_mat, "data/EnvDat/exp/lewquant_scaled_expanded_04222025.env")
write.env(yrkquants_mat, "data/EnvDat/exp/yrkquant_scaled_expanded_04222025.env")
```


## Formatting for RDA & GF

Here, the csv format should work fine so there's no need to go further. I will simply use the csv "data/EnvDat/exp/exp_scaled_expanded_04222025.csv" for the RDA and GF analyses. Could potentially add pcnms here for GF??? - leave that for now.