---
title: "CviMVP_oyster_EnvDatProcess"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Environmental Data Processing

This script is to process environmental data for both the seascape and experiment by reducing the variable set to retain uncorrelated variables, standardize data, and convert data to the correct formats for downstream analysis. 

The seascape data comes from...

The experimental data comes from...

## Setup

```{r}
# set wd
setwd("~/Documents/Github/MVP_oyster_GO")

# packages
library(vegan)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(psych)
library(LEA)
library(stringr)

# data
seascapeEnv <- read.csv("data/EnvDat/seascape/seascape_abiotic_biotic_envr.csv")
expsal <- read.csv("data/EnvDat/exp/all_salinity.csv")
exptemp <- read.csv("data/EnvDat/exp/all_temperature.csv")
expdis <- read.csv("data/EnvDat/exp/sentinel_RFTM_organized_dermo.csv")[,-(6:10)]
peacrab <- read.csv("data/EnvDat/exp/MVP23-FieldBags - finalTissue.csv")
sitedate <- read.csv("data/EnvDat/seascape/SeascapeSamples - date.csv")
```


## Data Pre-processing

First, I need to get the experimental environmental data into a workable format - right now, the disease data & environmental quantiles are all over the place, not in a single df.

Will start by checking the environmental data structure and processing seascape data to retain only  biologically relevant & uncorrelated variables.

```{r}
# check data
seascapeEnv
colnames(seascapeEnv)

# check data correlations within each dataset
# seascape - lots of correlated variables in the quantiles but keep for now, will check vif
pairs.panels(seascapeEnv[,c("salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90","Dermo_Prevalence","MSX_Prevalence","intensity_msx","intensity_dermo","Pea_crab","Boring_sponge")])
pairs.panels(seascapeEnv[,c("salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90","Dermo_Prevalence","MSX_Prevalence","Pea_crab")])

# seascape & env keep maxes & mins, remove all else (need the same variables in both sets)
seascapeEnv_trim <- seascapeEnv[,c("ID_SiteDate","Site_Order","State_Abbrev","Latitude_manual","Longitude_manual","salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90","Dermo_Prevalence","MSX_Prevalence","Pea_crab")]

# match seascape sites with experimental pops
seascapeEnv_trim[,c("ID_SiteDate","Latitude_manual","Longitude_manual")]
```

I need to do some work to put together the experimental data frame

```{r}
# start with the disease data, need to organize averages across experimental period
# this is coming from sentinel population
lewyrk_dis_avgs <- expdis %>% 
  group_by(SITE) %>% 
  dplyr::summarize(Dermo_Prevalence = mean(P_dermo), Dermo_Weighted_Prevalence = mean(WP_dermo, na.rm = T))
lewyrk_dis_avgs <- as.data.frame(lewyrk_dis_avgs)

# get the peacrab data from finalTissue
peacrab_trim <- peacrab[,c("bagID","peaCrab")]

# figure out which sites each bag comes from
peacrab_trim$bagnum <- as.numeric(str_split_fixed(peacrab_trim$bagID, "-", 3)[,3])
peacrab_trim$site_name <- (str_split_fixed(peacrab_trim$bagID, "-", 3)[,2])
peacrab_trim$site <- ifelse(peacrab_trim$bagnum < 4, "Lewisetta", "YorkRiver")

# now calculate numbers of pea crabs
lewyrk_pea_avgs <- peacrab_trim %>% 
  group_by(site) %>% 
  dplyr::summarize(Pea_Crab = sum(peaCrab == T)/(sum(peaCrab == T) + sum(peaCrab == F))) # this is the percent of oysters per pop that had pea crabs
lewyrk_pea_avgs <- as.data.frame(lewyrk_pea_avgs)

# still need msx for these sites
```


```{r}
# move on to other experimental data for now
# put together temp and sal
expsal_red <- expsal[,c("site_name","lat","lon","Mean_Annual_Salinity_ppt","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","salinity_quantile_10","salinity_quantile_90")]
exptemp_red <- exptemp[,c("site_name","lat","lon","Mean_Annual_Temperature_C","Mean_max_temperature_C","Mean_min_temperature_C","temp_quantile_10","temp_quantile_90")]
expenv_mg <- merge(expsal_red, exptemp_red, by = c("site_name", "lat", "lon"))

# experimental site info output
write.csv(expenv_mg[,c("site_name", "lat", "lon")], "data/EnvDat/exp/exp_site_info.csv")

# look at corrs
pairs.panels(expenv_mg[,c("Mean_Annual_Salinity_ppt","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","salinity_quantile_10","salinity_quantile_90","Mean_Annual_Temperature_C","Mean_max_temperature_C","Mean_min_temperature_C","temp_quantile_10","temp_quantile_90")])
pairs.panels(expenv_mg[,c("salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90")])

# keeping just quantiles
expenv_trim <- expenv_mg[,c("site_name","lat","lon","salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90")]

# pull out just lewisetta and york
lewyrk_quants <- expenv_trim[expenv_trim$site_name %in% c("LEW", "YRK"),]
lewyrk_quants$site_name <- ifelse(lewyrk_quants$site_name == "LEW", "Lewisetta", "YorkRiver")

# and add disease data
lewyrk_dis_avgs
lewyrk_pea_avgs
names(lewyrk_dis_avgs)[names(lewyrk_dis_avgs) == "SITE"] <- names(lewyrk_pea_avgs)[names(lewyrk_pea_avgs) == "site"] <- "site_name"
lewyrk1 <- merge(lewyrk_quants, lewyrk_dis_avgs[,1:2], by = c("site_name"))
lewyrkEnv_all <- merge(lewyrk1, lewyrk_pea_avgs, by = c("site_name"))

# the rest of the experimental disease info for experimental pops is in the seascape data; no DEBY or LOLA
seascape2exp <- data.frame(ID_SiteDate = c("CB_TX_2022-6-9", "Sister_Lk_LA_2022-8-9", "KP_FGR_FL_2022-8-23", "WS_CB_VA_2022-10-7", "SR_GB_NH_2022-8-12", "DC_DR_ME_2022-10-25"), site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "W5_NH", "W6_ME"))

# merge
expEnv_all <- merge(seascape2exp, seascapeEnv_trim, by = c("ID_SiteDate"))

# the VA site has no quantile data, but we did have some compiled from H2F, take from there
quantile_forVA <- read.csv("data/EnvDat/exp/quantiles_05232025.csv")[,-1]
VA_only <- quantile_forVA[quantile_forVA$site_name == "W4-VA", c("site_name","salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90")]
expEnv_all[expEnv_all$site_name == "W4_VA", c("salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90")] <- VA_only[,2:5]

pairs.panels(expEnv_all[,c("salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90","Dermo_Prevalence","MSX_Prevalence")]) # don't include pea crab because all are 0
# these are all reasonably uncorrelated, temp quantile 10/90 are 0.62 and temp quantile 90/MSX are -0.63 but otherwise all are low corr (and all are below |0.7|)

# trim to just what we have for the other sites - exclude DEBY & LOLA for now until we can get this disease data
head(seascapeEnv_trim)
head(expEnv_all)
head(lewyrkEnv_all)

# rename lat long in lewyrk2
names(lewyrkEnv_all)[names(lewyrkEnv_all) == "lat.x"] <- "Latitude_manual"
names(lewyrkEnv_all)[names(lewyrkEnv_all) == "lon.x"] <- "Longitude_manual"
names(lewyrkEnv_all)[names(lewyrkEnv_all) == "Pea_Crab"] <- "Pea_crab"

# all have the same column names now, but we don't have MSX for lew or york
```

Everything is looking good! Now I will standardize data to be scaled and centered for downstream analysis. One thing to consider is that I do have individual level disease data, but am standardizing data at the population level and then expanding this scaled data out - I should probably do the expansion first, then scale.

```{r}
# quick data inspection
head(seascapeEnv_trim)
head(expEnv_all)
head(lewyrkEnv_all)

# scale
std_sea <- scale(seascapeEnv_trim[,c("salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90", "Dermo_Prevalence", "Pea_crab")])
std_exp <- scale(expEnv_all[,c("salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90", "Dermo_Prevalence", "Pea_crab")])
std_lewyrk <- scale(lewyrkEnv_all[,c("salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90", "Dermo_Prevalence", "Pea_crab")])
colnames(std_sea) <- colnames(std_exp) <- colnames(std_lewyrk) <- c("salinity_quantile_10_scale", "salinity_quantile_90_scale", "temp_quantile_10_scale", "temp_quantile_90_scale", "Dermo_Prevalence_scale", "Pea_crab_scale")

# attach scaled information back to env dfs - question, should I scale all of these values together in one df, rather than as separate info?
sea_scaled <- as.data.frame(cbind(seascapeEnv_trim, std_sea))
exp_scaled <- as.data.frame(cbind(expEnv_all, std_exp))
lewyrk_scaled <- as.data.frame(cbind(lewyrkEnv_all, std_lewyrk))

# also try scaling all together
seascapeEnv_trim$site_name <- seascapeEnv_trim$ID_SiteDate
allenv1 <- rbind(na.omit(seascapeEnv_trim[!(seascapeEnv_trim$ID_SiteDate %in% seascape2exp$ID_SiteDate), c("site_name","salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90", "Dermo_Prevalence", "Pea_crab")]), expEnv_all[,c("site_name", "salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90", "Dermo_Prevalence", "Pea_crab")], lewyrkEnv_all[,c("site_name", "salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90", "Dermo_Prevalence", "Pea_crab")])

# scale the full env version
std_allenv <- scale(allenv1[,-1])
colnames(std_allenv) <- c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled")
allenv_scaled <- as.data.frame(cbind(allenv1, std_allenv))

# keep track of scaling factors
sea_scale_env <- attr(std_sea, "scaled:scale")
sea_center_env <- attr(std_sea, "scaled:center")
exp_scale_env <- attr(std_exp, "scaled:scale")
exp_center_env <- attr(std_exp, "scaled:center")
lewyrk_scale_env <- attr(std_lewyrk, "scaled:scale")
lewyrk_center_env <- attr(std_lewyrk, "scaled:center")
all_scale_env <- attr(std_allenv, "scaled:scale")
all_center_env <- attr(std_allenv, "scaled:center")

# put those scaling factors into their own df
scaling <- as.data.frame(rbind(sea_scale_env, sea_center_env, exp_scale_env, exp_center_env, lewyrk_scale_env, lewyrk_center_env, all_scale_env, all_center_env))
scaling$factor_type <- rownames(scaling)
rownames(scaling) <- NULL
scaling

# write out this env data
write.csv(allenv_scaled, paste0("data/EnvDat/env_scaled_",Sys.Date(),".csv"))
```


Finally I want to make sure now that my environmental data will match my genomic data by having one row in my environmental dataset for each individual in my genomic dataset. To do so, I will need to process the individual data and pair this with my environmental data.

```{r}
# experimental individuals - would replace this with the seascape individuals for lining up the environmental data with genetic data, no need to actually have indiv level predictions for experiment just pop level

# need ind level dat
indsSea <- readRDS("data/IndDat/20250604_seascape_indsmatrix.rds")
indsExp <- readRDS("data/IndDat/20240922_experimental_indsmatrix.rds")

# keep just relevant data
indsSea_red <- indsSea[,c("clean_ID","Label_ind","Key_ind","ID_SiteDate","ID_Site","SiteDate_Key")]

indsExp_red <- indsExp[,c("merge_ID_inds","Label_ind","Key_ind","ID_SiteDate","SiteDate_Key")]
levels(as.factor(indsExp_red$ID_SiteDate))

# need to add scaled env data to this
allenv_scaled$site_name
indsSea_red$ID_SiteDate
indsExp_red$ID_SiteDate

# the pop (ID_SiteDate/site_name) data doesn't match up for experimental samples
map <- list("TX-CapBay" = "W1_TX", "LA-SisLake" = "W2_LA", "FL-KingPlan" = "W3_FL", "LOLA" = "S1_LOLA", "DEBY" = "S2_DEBY", "VA-DeepWatSh" = "W4_VA", "NH-GrtBay" = "W5_NH", "ME-HogIs" = "W6_ME")
indsExp_red_mut <- indsExp_red %>% mutate(ID_SiteDate = ifelse(ID_SiteDate %in% names(map), map[ID_SiteDate], ID_SiteDate))
indsExp_red_mut <- as.data.frame(indsExp_red_mut)
colnames(indsExp_red_mut) <- c("merge_ID_inds", "Label_ind", "Key_ind", "site_name", "SiteDate_Key")

which(!(exp_scaled$site_name %in% indsExp_red_mut$site_name))
which(!(indsExp_red_mut$site_name %in% exp_scaled$site_name)) # no lola or deby, which is expected

indsExp_red_mut$site_name <- as.character(indsExp_red_mut$site_name)

# how many inds per site?
table(indsSea_red$ID_SiteDate) # oops, we have a laguna madre sample in here, we need to remove this
table(indsExp_red_mut$site_name) # 17-20

# rm laguna madre
indsSea_redrm <- indsSea_red[!(indsSea_red$ID_SiteDate == "LM_TX_2022-7-10"),]

# split the allenv df into respective datasets
sea_env_scale <- allenv_scaled[!(allenv_scaled$site_name %in% c("Lewisetta", "YorkRiver")),] # don't take out the expOrg data here, as it is the data for the seascape sites that I removed before scaling
expOrg_env_scale <- allenv_scaled[allenv_scaled$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "W5_NH", "W6_ME"),]
expSite_env_scale <- allenv_scaled[allenv_scaled$site_name %in% c("Lewisetta", "YorkRiver"),]

# rename the expOrg site_names that are still in the sea_env_scale df
seascape2exp
map2 <- list("W1_TX" = "CB_TX_2022-6-9", "W2_LA" = "Sister_Lk_LA_2022-8-9", "W3_FL" = "KP_FGR_FL_2022-8-23", "W4_VA" = "WS_CB_VA_2022-10-7", "W5_NH" = "SR_GB_NH_2022-8-12", "W6_ME" = "DC_DR_ME_2022-10-25")
sea_env_scale_mut <- sea_env_scale %>% mutate(site_name = ifelse(site_name %in% names(map2), map2[site_name], site_name)) %>% as.data.frame()

# site_name is a list
sea_env_scale_mut$site_name <- as.factor(unlist(sea_env_scale_mut$site_name))
site_names_seascape <- sea_env_scale_mut$site_name
write.csv(site_names_seascape, "data/EnvDat/seascape_sitenames.csv")

# also get the site names for exp data to reattach to env files
site_names_exp <- exp_scaled$site_name
write.csv(site_names_exp, "data/EnvDat/exp_sitenames.csv")

# the experimental sites Lew & York need to be added to the experimental origin data in a single df
expSite_env_scale_yrk <- expSite_env_scale[expSite_env_scale$site_name == "YorkRiver",]
expSite_env_scale_lew <- expSite_env_scale[expSite_env_scale$site_name == "Lewisetta",]
colnames(expSite_env_scale_yrk) <- c("site_name", "yrk_salinity_quantile_10", "yrk_salinity_quantile_90", "yrk_temp_quantile_10", "yrk_temp_quantile_90", "yrk_Dermo_Prevalence", "yrk_Pea_crab", "yrk_salinity_quantile_10_scaled", "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")
colnames(expSite_env_scale_lew) <- c("site_name", "lew_salinity_quantile_10", "lew_salinity_quantile_90", "lew_temp_quantile_10", "lew_temp_quantile_90", "lew_Dermo_Prevalence", "lew_Pea_crab", "lew_salinity_quantile_10_scaled", "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled")

# add this info to the expOrg df
nrow(indsExp_red_mut) # 154 rows
exp_env_scale <- cbind(expOrg_env_scale, (expSite_env_scale_lew[,-1]), (expSite_env_scale_yrk[,-1]))

# add site_name column to seascape inds
indsSea_redrm$site_name <- indsSea_redrm$ID_SiteDate

# merge
sea_scaled_expanded <- merge(as.data.frame(indsSea_redrm), as.data.frame(sea_env_scale_mut), by = c("site_name"))
exp_scaled_expanded <- merge(as.data.frame(indsExp_red_mut), as.data.frame(exp_env_scale), by = c("site_name"))

# we lose a lot of rows when merging seascape, what are we losing?
dim(sea_scaled_expanded)
dim(indsSea_redrm) # 116 rows lost

levels(as.factor(indsSea_redrm[which(!(indsSea_redrm$site_name %in% sea_env_scale_mut$site_name)),]$site_name)) # we lose six seascape sites, namely "DWS_CB_VA_2022-10-7", "EelLake_Canada_2022-9-27", "FC_LIS_CT_2023-10-11", "NPS_PS_NC_2022-10-12", "NR_NB_RI_2022-8-19", and "PrEdIs_CN_A_2022-10-18"

# is it because these had NA's?
View(seascapeEnv_trim[seascapeEnv_trim$site_name %in% c("DWS_CB_VA_2022-10-7", "EelLake_Canada_2022-9-27", "FC_LIS_CT_2023-10-11", "NPS_PS_NC_2022-10-12", "NR_NB_RI_2022-8-19", "PrEdIs_CN_A_2022-10-18"),]) # yes, these sites only have disease data - they should be removed from analysis, so it's ok that they are not in the seascape scaled df

# write these out to .csv
write.csv(exp_env_scale, paste0("data/EnvDat/exp/exp_scaled_", Sys.Date(), ".csv"))
write.csv(as.data.frame(sea_env_scale_mut), paste0("data/EnvDat/exp/sea_scaled_", Sys.Date(), ".csv"))
write.csv(exp_scaled_expanded, paste0("data/EnvDat/exp/exp_scaled_expanded_", Sys.Date(), ".csv"))
write.csv(sea_scaled_expanded, paste0("data/EnvDat/seascape/sea_scaled_expanded_", Sys.Date(), ".csv"))
```

Because I do have individual level disease data, I should probably also integrate that into my environmental seascape data frame. I also want to ensure that this seascape data matches the order of the genetic data

```{r}
# data
sea_scaled_expanded

# remove the pop level disease and scaled quantile data
sea_exp <- sea_scaled_expanded[,c("site_name","clean_ID","Label_ind","Key_ind","ID_SiteDate","SiteDate_Key", "salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90")]

# now add back the individual disease data
# need to get this indiv level data if possible!

# ensure env data matches genomic data
# make sure genotype data is in the same order as individual & environmental data
sea_scaled_expanded$clean_ID %in% inds_genomat
sea_scaled_expanded$clean_ID == inds_genomat # nope, not in the same order - need to reorder the environmental data

# reorder env data
env_idx_reord <- match(inds_genomat, sea_scaled_expanded$clean_ID)
env_sea_ordered <- sea_scaled_expanded[env_idx_reord,]
env_sea_ordered$clean_ID == inds_genomat

# output this file
write.csv(env_sea_ordered, paste0("data/EnvDat/seascape/sea_scaled_expanded_reord_", Sys.Date(), ".csv"))
```


## Environmental Distance

In addition to processing this environmental data, I want to calculate environmental distance for downstream analyses. I will also visualize this distance.

```{r}
# data
expenv_dist_df <- allenv_scaled

# reduce variable set
expenv_dist_red_df <- expenv_dist_df[,c("site_name", "salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90","Dermo_Prevalence","Pea_crab")]

# calculate environmental distance
library(vegan)
envdist_std <- round(decostand(expenv_dist_red_df[,c("salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90","Dermo_Prevalence","Pea_crab")], method = "standardize"), 4)

env_dist <- round(vegdist(envdist_std, method = "euclidian", 
                          upper = FALSE, diag = TRUE), 4)
env_dist_df <- as.data.frame(as.matrix(env_dist))
colnames(env_dist_df) <- rownames(env_dist_df) <- expenv_dist_red_df$site_name
env_dist_df

# matrix
env_dist_mat <- cbind(rownames(env_dist_df),env_dist_df)

# export matrix
write.csv(env_dist_mat, paste0("data/EnvDat/exp/distance_matrix_",Sys.Date(),".csv"))

# visualize this distance as heatmap
library(reshape2)
env_dist_longer <- melt(env_dist_mat)
head(env_dist_longer)
colnames(env_dist_longer) <- c("site1","site2","dist")

popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME", "LARMIX", "SEEDMIX"), 
                     order = 2:11, 
                     cols = c("#3e4989" ,"#31688e", "#26828e", "#440154", "#482878", "#1f9e89", "#35b779", "#6ece58", "#fde725", "#b5de2b"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME", "H1_LARMIX", "H2_SEEDMIX"))
popsdf0 <- rbind(data.frame(pop = c("LEW","YRK"), order = 0:1, cols = c("white","black"), shape = c(22,22), site_name = c("Lewisetta","YorkRiver")), popsdf)
popsdf1 <- popsdf0
popsdf2 <- popsdf0
colnames(popsdf1) <- c("pop","order1","cols","shape","site1")
colnames(popsdf2) <- c("pop","order2","cols","shape","site2")
env_dist_long_mg <- merge(popsdf1[,c("site1","order1")], env_dist_longer, by = "site1")
env_dist_long_mg2 <- merge(popsdf2[,c("site2","order2")], env_dist_long_mg, by = "site2")

# keep this too
write.csv(env_dist_long_mg2, paste0("data/EnvDat/exp/distance_matrix_longer_", Sys.Date(), ".csv"))

# visualize
library(viridis)
envdist_plot <- ggplot(env_dist_long_mg2, aes(x = fct_reorder(site1, order1), 
                              y = fct_reorder(site2, order2), 
                              fill = dist))+
  theme_classic()+
  geom_tile(color = "black") +
#  geom_text(aes(label = distances), color = "white", size = 4)+
  scale_fill_viridis_c(name = "Euclidean 
Distance", limits = c(0, 7))+
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top") +
  labs(title = "Environmental Distance between Sites", x = "", y = "") +
  theme(plot.title = element_text(size = 18), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top")
envdist_plot

# show just the lew and york comparisons
env_dist_long_expsite <- env_dist_long_mg2[env_dist_long_mg2$site2 %in% c("LEW", "YRK"),]
env_dist_long_expsite_red <- env_dist_long_expsite[!(env_dist_long_expsite$site1 %in% c("LEW","YRK")),]

# plot
# envdist_plot_red <- ggplot(env_dist_long_expsite_red, aes(x = fct_reorder(site1, order1), 
#                               y = fct_reorder(site2, order2), 
#                               fill = value))+
#   theme_classic()+
#   geom_tile(color = "black") +
# #  geom_text(aes(label = distances), color = "white", size = 4)+
#   scale_fill_viridis_c(name = "Euclidean 
# Distance", limits = c(0, 4))+
#   theme(plot.title = element_text(size = 20), 
#         legend.title = element_text(size = 14),
#         legend.position = "right", legend.justification = "top") +
#   labs(title = "Environmental Distance between Sites", x = "", y = "") +
#   theme(plot.title = element_text(size = 18), 
#         legend.title = element_text(size = 14),
#         legend.position = "right", legend.justification = "top")
# envdist_plot_red
```


## Formatting for LEA

LEA requires environmental data be formatted as a .env file type.

```{r}
# prep as matrix
sea_mat <- as.matrix(sea_env_scale_mut[,8:13])
# matrix prep
env_sea_ordered_mat <- env_sea_ordered[,c("salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90", "Dermo_Prevalence", "Pea_crab")]
full_exp_mat <- as.matrix(exp_env_scale[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "lew_salinity_quantile_10_scaled",  "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled", "yrk_salinity_quantile_10_scaled",  "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")])
full_exp_expanded_mat <- as.matrix(exp_scaled_expanded[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "lew_salinity_quantile_10_scaled",  "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled", "yrk_salinity_quantile_10_scaled",  "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled")])

# and convert to env format
write.env(sea_mat, paste0("data/EnvDat/seascape/seascape_scaled_", Sys.Date(), ".env"))
write.env(env_sea_ordered_mat, paste0("data/EnvDat/seascape/seascape_scaled_expanded_reord_", Sys.Date(), ".env"))
write.env(full_exp_mat, paste0("data/EnvDat/exp/exp_scaled_", Sys.Date(), ".env"))
write.env(full_exp_expanded_mat, paste0("data/EnvDat/exp/exp_scaled_expanded_", Sys.Date(), ".env"))
```


## Formatting for RDA & GF

Here, the csv format should work fine so there's no need to go further. I will simply use the csv "data/EnvDat/exp/exp_scaled_expanded_04222025.csv" for the RDA and GF analyses. Could potentially add pcnms here for GF??? - leave that for now.