---
title: "CviMVP_oyster_EnvDatProcess"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Environmental Data Processing

This script is to process environmental data for both the seascape and experiment by reducing the variable set to retain uncorrelated variables, standardize data, and convert data to the correct formats for downstream analysis. 

The seascape data comes from...

The experimental data comes from...

## Setup

```{r}
# set wd
setwd("~/Documents/Github/MVP_oyster_GO")

# packages
library(vegan)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(psych)
library(LEA)
library(stringr)

# data
seascapeEnv <- read.csv("data/EnvDat/seascape/SeascapeSamples_site.csv")
expEnv <- read.csv("data/EnvDat/exp/envr_summary_stats.csv")[,-1]
```


## Data Pre-processing

Will start by checking the environmental data structure and processing to retain only  biologically relevant & uncorrelated variables.

```{r}
# check data
seascapeEnv
expEnv
colnames(seascapeEnv)
colnames(expEnv)

# check data correlations within each dataset
# seascape
pairs.panels(seascapeEnv[,c("Mean_Annual_Salinity_ppt","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","Salinity_st_dev","Mean_Annual_Temperature_C","Mean_max_temperature_C","Mean_min_temperature_C","Temperature_st_dev")])
pairs.panels(expEnv[,c("Mean_Annual_Salinity_ppt","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","Salinity_st_dev","Mean_Annual_Temperature_C","Mean_max_temperature_C","Mean_min_temperature_C","Temperature_st_dev")])

# remove some of these correlated variables
# seascape & env keep maxes & mins, remove all else (need the same variables in both sets)
seascapeEnv_trim <- seascapeEnv[,c("ID_Site","Site_name","Site_ID2","Site_Order","Site_Color","Region","Region_color","Latitude_manual","Longitude_manual","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","Mean_max_temperature_C","Mean_min_temperature_C")]
expEnv_trim <- expEnv[,c("site_name","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","Mean_max_temperature_C","Mean_min_temperature_C")]
```

Everything is looking good! Now I will standardize data to be scaled and centered for downstream analysis.

```{r}
# quick data inspection
head(seascapeEnv_trim)
head(expEnv_trim)

# scale
std_sea <- scale(seascapeEnv_trim[,10:13])
std_exp <- scale(expEnv_trim[,2:5])
colnames(std_sea) <- colnames(std_exp) <-  c("max_sal_scaled","min_sal_scaled","max_temp_scaled","min_temp_scaled")

# attach scaled information back to env dfs
sea_scaled <- as.data.frame(cbind(seascapeEnv_trim, std_sea))
exp_scaled <- as.data.frame(cbind(expEnv_trim, std_exp))

# keep track of scaling factors
sea_scale_env <- attr(std_sea, "scaled:scale")
sea_center_env <- attr(std_sea, "scaled:center")
exp_scale_env <- attr(std_exp, "scaled:scale")
exp_center_env <- attr(std_exp, "scaled:center")

# put those scaling factors into their own df
scaling <- as.data.frame(rbind(sea_scale_env, sea_center_env, exp_scale_env, exp_center_env))
scaling$factor_type <- rownames(scaling)
rownames(scaling) <- NULL
scaling
```

Finally I want to make sure now that my environmental data will match my genomic data by having one row in my environmental dataset for each individual in my genomic dataset. To do so, I will need to process the individual data and pair this with my environmental data.

```{r}
# need ind level data first, process in GenDatProcess
```

## Formatting for LEA

LEA requires environmental data be formatted as a .env file type.

```{r}
# prep as matrix
sea_mat <- as.matrix(sea_scaled[,c("max_sal_scaled","min_sal_scaled","max_temp_scaled","min_temp_scaled")])
exp_mat <- as.matrix(exp_scaled[,c("max_sal_scaled","min_sal_scaled","max_temp_scaled","min_temp_scaled")])

# and convert to env format
write.env(sea_mat, "data/EnvDat/sea_scaled.env")
write.env(exp_mat, "data/EnvDat/exp_scaled.env")
```


## Formatting for RDA & GF
