---
title: "CviMVP_oyster_EnvDatProcess"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Environmental Data Processing

This script is to process environmental data for both the seascape and experiment by reducing the variable set to retain uncorrelated variables, standardize data, and convert data to the correct formats for downstream analysis. 

## Setup

```{r}
# set wd
setwd("~/Documents/Github/MVP_oyster_GO")

# packages
library(vegan)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(psych)
library(LEA)
library(stringr)

# data
seascapeEnv <- read.csv("data/EnvDat/seascape/seascapeEnv_biotic_abiotic_27cq.csv")
expsal <- read.csv("data/EnvDat/exp/exp_salinity.csv")[,-1]
exptemp <- read.csv("data/EnvDat/exp/exp_temp.csv")
expdis <- read.csv("data/EnvDat/exp/sentinel_RFTM_organized_dermo.csv")[,-(6:10)]
expdis_lew <- read.csv("data/EnvDat/exp/LewisettaSentinels.csv")
expdis_yrk <- read.csv("data/EnvDat/exp/YorkRiverShorelineDisease-24March2025.csv")
peacrab <- read.csv("data/EnvDat/exp/MVP23-FieldBags - finalTissue.csv")
sitedate <- read.csv("data/EnvDat/seascape/SeascapeSamples - date.csv")
```


## Data Pre-processing

First, I need to get the experimental environmental data into a workable format - right now, the disease data & environmental quantiles are all over the place, not in a single df.

Will start by checking the environmental data structure and processing seascape data to retain only  biologically relevant & uncorrelated variables.

```{r}
# check data
seascapeEnv
colnames(seascapeEnv)

# keep seascape env quantiles and disease prevalences, remove all else (need the same variables in both sets)
seascapeEnv_trim <- seascapeEnv[,c("ID_SiteDate","Site_Order","State_Abbrev","Latitude_manual","Longitude_manual","salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90","Dermo_Prevalence","MSX_Prevalence","Pea_crab")]
```

I need to do some work to put together the experimental data frame

```{r}
# start with the disease data, need to organize averages across experimental period
# this is coming from sentinel population
lewyrk_dis_avgs <- expdis %>% 
  group_by(SITE) %>% 
  dplyr::summarize(Dermo_Prevalence = mean(P_dermo), Dermo_Weighted_Prevalence = mean(WP_dermo, na.rm = T))
lewyrk_dis_avgs <- as.data.frame(lewyrk_dis_avgs)

# get the peacrab data from finalTissue
peacrab_trim <- peacrab[,c("bagID","peaCrab")]

# figure out which sites each bag comes from
peacrab_trim$bagnum <- as.numeric(str_split_fixed(peacrab_trim$bagID, "-", 3)[,3])
peacrab_trim$site_name <- (str_split_fixed(peacrab_trim$bagID, "-", 3)[,2])
peacrab_trim$site <- ifelse(peacrab_trim$bagnum < 4, "Lewisetta", "YorkRiver")

# now calculate numbers of pea crabs
lewyrk_pea_avgs <- peacrab_trim %>% 
  group_by(site) %>% 
  dplyr::summarize(Pea_crab = sum(peaCrab == T)/(sum(peaCrab == T) + sum(peaCrab == F))) # this is the percent of oysters per pop that had pea crabs
lewyrk_pea_avgs <- as.data.frame(lewyrk_pea_avgs)

# now msx; these data are formatted very differently
summary(expdis_lew)
summary(expdis_yrk)

# expdis_yrk needs serious re-formatting
# msx data only
msx_yrk <- expdis_yrk[,-(4:18)]
colnames(msx_yrk) <- msx_yrk[1,]
msx_yrk_rename <- msx_yrk[-1,]

# spring imports is relevant group
msx_yrk_si <- msx_yrk[msx_yrk$Site == "Spring Imports",]
msx_yrk_si$msx_prev <- as.numeric(msx_yrk_si$`Inf`) / as.numeric(msx_yrk_si$n)

# get experimental period
msx_yrk_si$Date <- mdy(msx_yrk_si$Date)
msx_yrk_exp <- msx_yrk_si[msx_yrk_si$Date > as.Date("2023-01-01"),]

# use full dataset for deby disease, 2023-2024 for york river site
deby_prev_msx <- sum(as.numeric(msx_yrk_si$`Inf`)) / sum(as.numeric(msx_yrk_si$n))
yrk_prev_msx <- sum(as.numeric(msx_yrk_exp$`Inf`)) / sum(as.numeric(msx_yrk_exp$n))

# now lewisetta msx - we only have data spanning back to 2023, so lola and lew prev are the same
expdis_lew_nona <- expdis_lew[!is.na(expdis_lew$msx_intensity),]
lew_prev_msx <- 1 - (nrow(expdis_lew_nona[which(expdis_lew$msx_intensity == "N"),]) / nrow(expdis_lew_nona))

# put all msx data into single df
msx_avgs <- data.frame(site_name = c("Lewisetta","YorkRiver","S1_LOLA","S2_DEBY"),
                       MSX_Prevalence = c(lew_prev_msx, yrk_prev_msx, 
                                    lew_prev_msx, deby_prev_msx))
```


```{r}
# move on to other experimental data
# put together temp and sal
expsal_red <- expsal[,c("site_name","lat","lon","Mean_Annual_Salinity_ppt","Mean_max_Salinity_ppt","Mean_min_Salinity_ppt","salinity_quantile_10","salinity_quantile_90")]
exptemp_red <- exptemp[,c("site_name","lat","lon","Mean_Annual_Temperature_C","Mean_max_temperature_C","Mean_min_temperature_C","temp_quantile_10","temp_quantile_90")]
expenv_mg <- merge(expsal_red, exptemp_red, by = c("site_name", "lat", "lon"))

# experimental site info output
#write.csv(expenv_mg[,c("site_name", "lat", "lon")], "data/EnvDat/exp/exp_site_info.csv")

# keeping just quantiles
expenv_trim <- expenv_mg[,c("site_name","lat","lon","salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90")]

# pull out just lewisetta and york
lewyrk_quants <- expenv_trim[expenv_trim$site_name %in% c("LEW", "YRK"),]
lewyrk_quants$site_name <- ifelse(lewyrk_quants$site_name == "LEW", "Lewisetta", "YorkRiver")

# and add disease data
lewyrk_dis_avgs
lewyrk_pea_avgs
msx_avgs
names(lewyrk_dis_avgs)[names(lewyrk_dis_avgs) == "SITE"] <- names(lewyrk_pea_avgs)[names(lewyrk_pea_avgs) == "site"] <- "site_name"
lewyrk_dis <- merge(msx_avgs[msx_avgs$site_name %in% c("Lewisetta", "YorkRiver"),], lewyrk_dis_avgs[,1:2], by = c("site_name"))
lewyrk1 <- merge(lewyrk_quants, lewyrk_dis, by = c("site_name"))
lewyrkEnv_all <- merge(lewyrk1, lewyrk_pea_avgs, by = c("site_name"))

# the rest of the experimental disease info for experimental pops is in the seascape data; DEBY and LOLA are the same as the lewyrk_data (except DEBY msx), but we won't train the models with them
seascape2exp <- data.frame(site_name = c("CB_TX_2022-6-9", "Sister_Lk_LA_2022-8-9", "KP_FGR_FL_2022-8-23", "DWS_CB_VA_2022-10-7", "SR_GB_NH_2022-8-12", "DC_DR_ME_2022-10-25"), site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "W5_NH", "W6_ME"))

# merge
expEnv_all <- merge(seascape2exp, seascapeEnv_trim, by = c("site_name"))

# the VA site has no quantile data, but we did have some compiled from H2F, take from there
quantile_forVA <- read.csv("data/EnvDat/exp/quantiles_05232025.csv")[,-1]
VA_only <- quantile_forVA[quantile_forVA$site_name == "W4-VA", c("site_name","salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90")]
expEnv_all[expEnv_all$site_name == "W4_VA", c("salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90")] <- as.numeric(VA_only[,2:5])
seascapeEnv_trim[seascapeEnv_trim$site_name == "DWS_CB_VA_2022-10-7", c("salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90")] <- as.numeric(VA_only[,2:5])

# remove other nas from seascape
seascapeEnv_trim_narm <- seascapeEnv_trim[!is.na(seascapeEnv_trim$salinity_quantile_10),]

# add lola and deby
lola_deb <- expenv_trim[expenv_trim$site_name %in% c("S1_LOLA", "S2_DEBY"),]
lola_deb$ID_SiteDate <- c("LOLA", "DEBY")
lola_deb$Dermo_Prevalence <- lewyrkEnv_all$Dermo_Prevalence
lola_deb$Pea_crab <- lewyrkEnv_all$Pea_crab
lola_deb$MSX_Prevalence <- c(lewyrkEnv_all[lewyrkEnv_all$site_name == "Lewisetta",]$MSX_Prevalence, deby_prev_msx)

# rename lat long & peacrab
names(lewyrkEnv_all)[names(lewyrkEnv_all) == "lat"] <- "Latitude_manual"
names(lewyrkEnv_all)[names(lewyrkEnv_all) == "lon"] <- "Longitude_manual"
names(lola_deb)[names(lola_deb) == "lat"] <- "Latitude_manual"
names(lola_deb)[names(lola_deb) == "lon"] <- "Longitude_manual"
names(seascapeEnv_trim)[names(seascapeEnv_trim) == "ID_SiteDate"] <- "site_name"

# trim to just what we have for the other sites
seascapeEnv_ready <- seascapeEnv_trim_narm[,c("site_name","Latitude_manual","Longitude_manual","salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90","Dermo_Prevalence","Pea_crab", "MSX_Prevalence")]
expEnv_ready <- expEnv_all[,c("site_name","Latitude_manual","Longitude_manual","salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90","Dermo_Prevalence","Pea_crab", "MSX_Prevalence")]
lewyrkEnv_ready <- lewyrkEnv_all[,c("site_name","Latitude_manual","Longitude_manual","salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90","Dermo_Prevalence","Pea_crab", "MSX_Prevalence")]
lola_deb_ready <- lola_deb[,c("site_name","Latitude_manual","Longitude_manual","salinity_quantile_10","salinity_quantile_90","temp_quantile_10","temp_quantile_90","Dermo_Prevalence","Pea_crab", "MSX_Prevalence")]
```

Everything is looking good! Now I will standardize data to be scaled and centered for downstream analysis. To start this process, I will first put everything into a single data frame. One thing to consider is that I do have individual level disease data, but am standardizing data at the population level and then expanding this scaled data out - I should probably do the expansion first, then scale.

```{r}
# quick data inspection
head(seascapeEnv_ready)
head(expEnv_ready)
head(lewyrkEnv_ready)
head(lola_deb_ready)

# scale all together
allenv <- rbind(na.omit(seascapeEnv_ready[!(seascapeEnv_ready$site_name %in% seascape2exp$ID_SiteDate), c("site_name", "Latitude_manual", "Longitude_manual", "salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90", "Dermo_Prevalence", "Pea_crab", "MSX_Prevalence")]), expEnv_ready[,c("site_name", "Latitude_manual", "Longitude_manual", "salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90", "Dermo_Prevalence", "Pea_crab", "MSX_Prevalence")], lola_deb_ready[,c("site_name", "Latitude_manual", "Longitude_manual", "salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90", "Dermo_Prevalence", "Pea_crab", "MSX_Prevalence")], lewyrkEnv_ready[,c("site_name", "Latitude_manual", "Longitude_manual", "salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90", "Dermo_Prevalence", "Pea_crab", "MSX_Prevalence")])

# check corrs
pairs.panels(allenv[,4:10])

# scale the full env version
std_allenv <- scale(allenv[,4:ncol(allenv)])
colnames(std_allenv) <- c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "MSX_Prevalence_scaled")
allenv_scaled <- as.data.frame(cbind(allenv, std_allenv))

# keep track of scaling factors
all_scale_env <- attr(std_allenv, "scaled:scale")
all_center_env <- attr(std_allenv, "scaled:center")

# put those scaling factors into their own df
scaling <- as.data.frame(rbind(all_scale_env, all_center_env))
scaling$factor_type <- rownames(scaling)
rownames(scaling) <- NULL
scaling

# write out this env data
write.csv(allenv_scaled, paste0("data/EnvDat/env_scaled_",Sys.Date(),".csv"), row.names = F)
```


I want to make sure now that my environmental data will match my genomic data by having one row in my environmental dataset for each individual in my genomic dataset. To do so, I will need to process the individual data and pair this with my environmental data.

```{r}
# need ind level dat
indsSea <- readRDS("data/IndDat/20250604_seascape_indsmatrix.rds")

# keep just relevant data
indsSea_red <- indsSea[,c("clean_ID","Label_ind","Key_ind","ID_SiteDate","ID_Site","SiteDate_Key")]
levels(as.factor(indsSea_red$ID_SiteDate))

# how many inds per site?
table(indsSea_red$ID_SiteDate) # oops, we have a laguna madre sample in here, we need to remove this

# rm laguna madre
indsSea_redrm <- indsSea_red[!(indsSea_red$ID_SiteDate == "LM_TX_2022-7-10"),]

# all env scaled where sites are named for experiment
map <- list("CB_TX_2022-6-9" = "W1_TX", "Sister_Lk_LA_2022-8-9" = "W2_LA", "KP_FGR_FL_2022-8-23" = "W3_FL",  "DWS_CB_VA_2022-10-7" = "W4_VA", "SR_GB_NH_2022-8-12" = "W5_NH", "DC_DR_ME_2022-10-25" = "W6_ME")
allenv_scaled2 <- allenv_scaled %>% mutate(site_name = ifelse(site_name %in% names(map), map[site_name], site_name)) %>% as.data.frame()
allenv_scaled2_nodup <- allenv_scaled2[!(duplicated(allenv_scaled2$site_name)),]

# split the allenv df into respective datasets
sea_env_scale <- allenv_scaled[!(allenv_scaled$site_name %in% c("Lewisetta", "YorkRiver")),] # don't take out the expOrg data here, as it is the data for the seascape sites that I removed before scaling
expOrg_env_scale <- allenv_scaled2_nodup[allenv_scaled2_nodup$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "W5_NH", "W6_ME","S1_LOLA","S2_DEBY"),]
expSite_env_scale <- allenv_scaled[allenv_scaled$site_name %in% c("Lewisetta", "YorkRiver"),]

# rename the expOrg site_names that are still in the sea_env_scale df
# seascape2exp
# map2 <- list("W1_TX" = "CB_TX_2022-6-9", "W2_LA" = "Sister_Lk_LA_2022-8-9", "W3_FL" = "KP_FGR_FL_2022-8-23", "W4_VA" = "WS_CB_VA_2022-10-7", "W5_NH" = "SR_GB_NH_2022-8-12", "W6_ME" = "DC_DR_ME_2022-10-25")
# sea_env_scale_mut <- sea_env_scale %>% mutate(site_name = ifelse(site_name %in% names(map2), map2[site_name], site_name)) %>% as.data.frame()

# site_name is a list, has duplicates, has non seascape sites
sea_env_scale_mut$site_name <- as.factor(unlist(sea_env_scale_mut$site_name))
sea_env_scale_mut_nosel <- sea_env_scale_mut[!(sea_env_scale_mut$site_name %in% c("S1_LOLA", "S2_DEBY")),]
sea_env_scale_mut_nosel_nodup <- sea_env_scale_mut_nosel[!(duplicated(sea_env_scale_mut_nosel$site_name)),]
site_names_seascape <- sea_env_scale_mut_nosel_nodup$site_name

write.csv(site_names_seascape, "data/EnvDat/seascape/seascape_sitenames.csv")

# also get the site names for exp data to reattach to env files
site_names_exp <- unlist(expOrg_env_scale$site_name)
write.csv(site_names_exp, "data/EnvDat/exp_sitenames.csv")

# the experimental sites Lew & York need to be added to the experimental origin data in a single df
expSite_env_scale_yrk <- expSite_env_scale[expSite_env_scale$site_name == "YorkRiver",]
expSite_env_scale_lew <- expSite_env_scale[expSite_env_scale$site_name == "Lewisetta",]
colnames(expSite_env_scale_lew) <- c("site_name", "Latitude_manual", "Longitude_manual", "lew_salinity_quantile_10", "lew_salinity_quantile_90", "lew_temp_quantile_10", "lew_temp_quantile_90", "lew_Dermo_Prevalence", "lew_Pea_crab", "lew_MSX_Prevalence", "lew_salinity_quantile_10_scaled", "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled","lew_MSX_Prevalence_scaled")
colnames(expSite_env_scale_yrk) <- c("site_name", "Latitude_manual", "Longitude_manual", "yrk_salinity_quantile_10", "yrk_salinity_quantile_90", "yrk_temp_quantile_10", "yrk_temp_quantile_90", "yrk_Dermo_Prevalence", "yrk_Pea_crab", "yrk_MSX_Prevalence", "yrk_salinity_quantile_10_scaled", "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled", "yrk_MSX_Prevalence_scaled")

# add this info to the expOrg df
exp_env_scale <- cbind(expOrg_env_scale, (expSite_env_scale_lew[,-1]), (expSite_env_scale_yrk[,-1]))

# site name is a list again
exp_env_scale$site_name <- as.character(exp_env_scale$site_name)

# add site_name column to seascape inds
indsSea_redrm$site_name <- indsSea_redrm$ID_SiteDate

# merge
sea_scaled_expanded <- merge(as.data.frame(indsSea_redrm), as.data.frame(sea_env_scale_mut_nosel_nodup), by = c("site_name"))

# we lose a lot of rows when merging seascape, what are we losing?
dim(sea_scaled_expanded)
dim(indsSea_redrm) # 115 rows lost

levels(as.factor(indsSea_redrm[which(!(indsSea_redrm$site_name %in% sea_env_scale_mut$site_name)),]$site_name)) # we lose five seascape sites, namely "EelLake_Canada_2022-9-27", "FC_LIS_CT_2023-10-11", "NPS_PS_NC_2022-10-12", "NR_NB_RI_2022-8-19", and "PrEdIs_CN_A_2022-10-18"

# is it because these had NA's?
head(seascapeEnv_trim[seascapeEnv_trim$site_name %in% c("EelLake_Canada_2022-9-27", "FC_LIS_CT_2023-10-11", "NPS_PS_NC_2022-10-12", "NR_NB_RI_2022-8-19", "PrEdIs_CN_A_2022-10-18"),]) # yes, these sites only have disease data - they should be removed from analysis, so it's ok that they are not in the seascape scaled df

# write these out to .csv
write.csv(as.data.frame(exp_env_scale), paste0("data/EnvDat/exp/exp_scaled_", Sys.Date(), ".csv"))
write.csv(as.data.frame(sea_env_scale_mut), paste0("data/EnvDat/seascape/sea_scaled_", Sys.Date(), ".csv"))
write.csv(sea_scaled_expanded, paste0("data/EnvDat/seascape/sea_scaled_expanded_", Sys.Date(), ".csv"))
```

Because I do have individual level disease data, I should probably also integrate that into my environmental seascape data frame. I also want to ensure that this seascape data matches the order of the genetic data

```{r}
# data
sea_scaled_expanded

# remove the pop level disease and scaled quantile data
sea_exp <- sea_scaled_expanded[,c("site_name","clean_ID","Label_ind","Key_ind","ID_SiteDate","SiteDate_Key", "salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90")]

# ensure env data matches genomic data
# make sure genotype data is in the same order as individual & environmental data
sea_scaled_expanded$clean_ID %in% inds_genomat 

# ids for the VA pop are not in inds_genomat -VA pop needs to be removed
sea_scaled_expanded_nova <- sea_scaled_expanded[!sea_scaled_expanded$site_name == "DWS_CB_VA_2022-10-7",]

sea_scaled_expanded$clean_ID == inds_genomat # nope, not in the same order - need to reorder the environmental data

# reorder env data
env_id_reord <- match(inds_genomat, sea_scaled_expanded$clean_ID)
env_sea_ordered <- sea_scaled_expanded[env_id_reord,]
env_sea_ordered$clean_ID == inds_genomat

# output this file
write.csv(env_sea_ordered, paste0("data/EnvDat/seascape/sea_scaled_expanded_reord_", Sys.Date(), ".csv"))
```


## Environmental Distance

In addition to processing this environmental data, I want to calculate environmental distance for downstream analyses. I will also visualize this distance.

```{r}
# data
expenv_dist_df <- allenv_scaled

# reduce variable set
expenv_dist_red_df <- expenv_dist_df[,c("site_name", "salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90","Dermo_Prevalence","Pea_crab","MSX_Prevalence")]

# make sure site names all match up
expenv_dist_red_df[which(duplicated(expenv_dist_red_df$site_name)),]$site_name <- c("W1_TX", "W6_ME", "W4_VA", "W3_FL", "W2_LA", "W5_NH")

# calculate environmental distance
library(vegan)
envdist_std <- round(decostand(expenv_dist_red_df[,c("salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90","Dermo_Prevalence","Pea_crab","MSX_Prevalence")], method = "standardize"), 4)
envdist_std_nb <- round(decostand(expenv_dist_red_df[,c("salinity_quantile_10", "salinity_quantile_90", "temp_quantile_10", "temp_quantile_90")], method = "standardize"), 4)

env_dist <- round(vegdist(envdist_std, method = "euclidian", 
                          upper = FALSE, diag = TRUE, na.rm = F),
                  4)
env_dist_df <- as.data.frame(as.matrix(env_dist))
colnames(env_dist_df) <- rownames(env_dist_df) <- expenv_dist_red_df$site_name

env_dist_nb <- round(vegdist(envdist_std_nb, method = "euclidian", 
                          upper = FALSE, diag = TRUE, na.rm = F),
                     4)
env_dist_nb_df <- as.data.frame(as.matrix(env_dist_nb))
colnames(env_dist_nb_df) <- rownames(env_dist_nb_df) <- expenv_dist_red_df$site_name

# matrix
env_dist_mat <- cbind(rownames(env_dist_df),env_dist_df)
env_dist_mat_nb <- cbind(rownames(env_dist_nb_df),env_dist_nb_df)

# export matrix
write.csv(env_dist_mat, paste0("data/EnvDat/exp/distance_matrix_",Sys.Date(),".csv"))
write.csv(env_dist_mat_nb, paste0("data/EnvDat/exp/distance_matrix_abiotic_",Sys.Date(),".csv"))

# visualize this distance as heatmap
library(reshape2)
env_dist_longer <- melt(env_dist_mat)
env_dist_longer_nb <- melt(env_dist_mat_nb)
head(env_dist_longer)
colnames(env_dist_longer) <- colnames(env_dist_longer_nb) <- c("site1","site2","dist")

popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME", "LEWISETTA","YORKRIVER"), 
                     order = 1:10, 
                     latitude = c(28.0841, 29.24916667, 30.44116, 37.211, 37.9803, 37.247284, 43.0524, 43.986, 37.9803, 37.247284),
                     longitude = c(-97.2009, -90.9211111, -81.43908, -76.787, -76.4619, -76.499369, -70.91181, -69.55, -76.4619, -76.499369),
                     cols = c("#3e4989" ,"#31688e", "#35b779", "#6ece58", "#440154", "#482878", "#fde725", "#b5de2b","slateblue1","slateblue1"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME","Lewisetta","YorkRiver"))

popsdf1 <- popsdf
popsdf2 <- popsdf
colnames(popsdf1) <- c("pop","order1","latitude","longitude","cols","shape","site1")
colnames(popsdf2) <- c("pop","order2","latitude","longitude","cols","shape","site2")
env_dist_long_mg <- merge(popsdf1[,c("site1","order1")], env_dist_longer, by = "site1")
env_dist_long_mg2 <- merge(popsdf2[,c("site2","order2")], env_dist_long_mg, by = "site2")
env_dist_long_mg_nb <- merge(popsdf1[,c("site1","order1")], env_dist_longer_nb, by = "site1")
env_dist_long_mg2_nb <- merge(popsdf2[,c("site2","order2")], env_dist_long_mg_nb, by = "site2")

# keep this too
write.csv(env_dist_long_mg2, paste0("data/EnvDat/exp/distance_matrix_longer_", Sys.Date(), ".csv"))
write.csv(env_dist_long_mg2_nb, paste0("data/EnvDat/exp/distance_matrix_longer_abiotic_", Sys.Date(), ".csv"))

# visualize
library(viridis)
envdist_plot <- ggplot(env_dist_long_mg2, aes(x = fct_reorder(site1, order1), 
                              y = fct_reorder(site2, order2), 
                              fill = dist))+
  theme_classic()+
  geom_tile(color = "black") +
#  geom_text(aes(label = distances), color = "white", size = 4)+
  scale_fill_viridis_c(name = "Euclidean 
Distance", limits = c(0, 7))+
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top") +
  labs(title = "Environmental Distance between Sites", x = "", y = "") +
  theme(plot.title = element_text(size = 18), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top")
envdist_plot

# show just the lew and york comparisons
env_dist_long_expsite <- env_dist_long_mg2[env_dist_long_mg2$site2 %in% c("Lewisetta", "YorkRiver"),]
env_dist_long_expsite_red <- env_dist_long_expsite[!(env_dist_long_expsite$site1 %in% c("Lewisetta", "YorkRiver")),]

# plot
envdist_plot_gardens <- ggplot(env_dist_long_expsite_red, aes(x = fct_reorder(site2, order2), 
                              y = fct_reorder(site1, order1), 
                              fill = dist))+
  theme_classic()+
  geom_tile(color = "black") +
#  geom_text(aes(label = distances), color = "white", size = 4)+
  scale_fill_viridis_c(name = "Euclidean 
Distance", limits = c(0, 5.5))+
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top") +
  labs(title = "Environmental Distance between Sites", x = "", y = "") +
  theme(plot.title = element_text(size = 18), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top")
envdist_plot_gardens

# what about without the disease?
envdist_plot_nb <- ggplot(env_dist_long_mg2_nb, aes(x = fct_reorder(site1, order1), 
                              y = fct_reorder(site2, order2), 
                              fill = dist))+
  theme_classic()+
  geom_tile(color = "black") +
#  geom_text(aes(label = distances), color = "white", size = 4)+
  scale_fill_viridis_c(name = "Euclidean 
Distance", limits = c(0, 5.5))+
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top") +
  labs(title = "Environmental Distance (Abiotic Only) between Sites", x = "", y = "") +
  theme(plot.title = element_text(size = 18), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top")
envdist_plot_nb

# show just the lew and york comparisons
env_dist_long_expsite_nb <- env_dist_long_mg2_nb[env_dist_long_mg2_nb$site2 %in% c("Lewisetta", "YorkRiver"),]
env_dist_long_expsite_red_nb <- env_dist_long_expsite_nb[!(env_dist_long_expsite_nb$site1 %in% c("Lewisetta", "YorkRiver")),]

# plot
envdist_plot_nb_gardens <- ggplot(env_dist_long_expsite_red_nb, aes(x = fct_reorder(site2, order2), 
                              y = fct_reorder(site1, order1), 
                              fill = dist))+
  theme_classic()+
  geom_tile(color = "black") +
#  geom_text(aes(label = distances), color = "white", size = 4)+
  scale_fill_viridis_c(name = "Euclidean 
Distance", limits = c(0, 5.5))+
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top") +
  labs(title = "Environmental Distance (Abiotic Only) between Sites", x = "", y = "") +
  theme(plot.title = element_text(size = 18), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top")
envdist_plot_nb_gardens

# write to ppt
library(officer) # for powerpoint export/editing
library(rvg) # for converting r graphics to editable vector graphics

doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = envdist_plot), location = ph_location_type(type = "body")) %>%
  
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = envdist_plot_gardens), location = ph_location_type(type = "body")) %>%
  
add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = envdist_plot_nb), location = ph_location_type(type = "body")) %>% 
  
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = envdist_plot_nb_gardens), location = ph_location_type(type = "body"))

print(doc, target = "results/figures/EnvDist_ExpSitesGardens.pptx")
```


## Interpolated Environmental Distance

```{r}
library(sf)
library(terra)
library(stars)
library(gstat)
library(rnaturalearth)
library(ggrepel)

# check data
env_dist_mat$site_name <- env_dist_mat$`rownames(env_dist_df)`
env_dist_expsites <- merge(allenv_scaled[,c("site_name","Latitude_manual","Longitude_manual")], env_dist_mat[,c("site_name","Lewisetta","YorkRiver")])

# create sf
sea_sf_envdist <- st_as_sf(env_dist_expsites, 
                   coords = c("Longitude_manual","Latitude_manual"),
                   crs = 4326) # for WGS84 projection

# set spherical geometry
sf::sf_use_s2(FALSE)

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin =-102, xmax = -65, ymin = 24, ymax = 47.05))
coast <- ne_coastline(scale = "medium", returnclass = "sf")
coast_crop <- sf::st_crop(coast, c(xmin =-102, xmax = -65, ymin = 24, ymax = 47.05))

# start by downloading and cropping oceans data
oceans110 <- ne_download(scale=110L, type = "ocean", category = "physical")
oceans_crop <- sf::st_crop(oceans110, c(xmin = -102, xmax = -65, ymin = 24, ymax = 47.05))

# check crs
st_crs(oceans_crop)
st_crs(coast_crop)
st_crs(world_crop) # looks good

# define buffer distance
dist <- 0.25 # 0.25º difference, approximately 15 miles

# define buffer region - note we need to specify the layer, as there are lots of different coastlines
buffered_region <- st_buffer(coast_crop, dist)[51,]
plot(buffered_region, col = "lightblue2")

# crop shapefile to region of interest for interpolation
# this is the region we will plot, but we need to interpolate the entire buffered region
int_region <- st_intersection(oceans_crop, buffered_region) 

# define grid across ocean area
grid <- st_make_grid(buffered_region, cellsize = 0.1, what = "polygons")
grid_int <- st_intersection(grid, buffered_region)
grid_int_crop <- sf::st_crop(grid_int, c(xmin = -102, xmax = -65, 
                                         ymin = 24, ymax = 47.05))

# interpolate
interpolated_lew_envdist <- idw(Lewisetta~1, sea_sf_envdist, grid_int)
saveRDS(interpolated_lew_envdist, "results/lg_results/interpolated_envdist_lew.rds")

interpolated_yrk_envdist <- idw(YorkRiver~1, sea_sf_envdist, grid_int)
saveRDS(interpolated_yrk_envdist, "results/lg_results/interpolated_envdist_yrk.rds")

# empty raster
ext <- ext(st_bbox(interpolated_lew_envdist)) # get spatial extent
rast_empty <- rast(ext, resolution = 0.1, crs = "+proj=longlat +datum=WGS84")

# create a raster stack
# set up
sf_ls <- list(interpolated_lew_envdist, interpolated_yrk_envdist)
rast_ls <- list()

# loop through sf files to rasterize
for (i in seq_along(sf_ls)) {
  rast = st_rasterize(sf_ls[[i]] %>% dplyr::select(var1.pred, geometry))
  rast_ls[[i]] = rast
}

# combine
rast_all <- c(rast_ls)
names(rast_all) <- c("interpolated_lew_envdist","interpolated_yrk_envdist")

# save
saveRDS(rast_all, "results/GO_results/envdist_interpolated.RDS")

# plot
# prep seascape actual sites for plotting
envdist_sea_plotting <- na.omit(env_dist_expsites)

# add the color palette
popsdf <- data.frame(pop = c("TX", "LA", "FL", "VA", "DEBY", "LOLA", "NH", "ME", "LEWISETTA","YORKRIVER"), 
                     order = 1:10, 
                     latitude = c(28.0841, 29.24916667, 30.44116, 37.211, 37.9803, 37.247284, 43.0524, 43.986, 37.9803, 37.247284),
                     longitude = c(-97.2009, -90.9211111, -81.43908, -76.787, -76.4619, -76.499369, -70.91181, -69.55, -76.4619, -76.499369),
                     cols = c("#3e4989" ,"#31688e", "#35b779", "#6ece58", "#440154", "#482878", "#fde725", "#b5de2b","deeppink","deeppink"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME","Lewisetta","YorkRiver"))

envdist_plotting <- merge(envdist_sea_plotting, popsdf, by = c("site_name"), all = T)

# try mapping
map_lew_int_envdist <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_lew_envdist, mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = envdist_sea_plotting, aes(x = Longitude_manual, y = Latitude_manual, 
                                              fill = Lewisetta), 
             shape = 21, size = 2) +
  geom_point(data = envdist_plotting[envdist_plotting$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x = longitude, y = latitude, 
                                       fill = Lewisetta), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  geom_label_repel(data = envdist_plotting[envdist_plotting$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x=Longitude_manual, y=Latitude_manual, label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "Environmental Distance to \nLewisetta", x = "Longitude", y = "Latitude", col = "Environmental \nDistance", fill = "Environmental \nDistance")
(map_lew_int_envdist)

lew_int_envdist_inset <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_lew_envdist, mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = envdist_sea_plotting, aes(x = Longitude_manual, y = Latitude_manual, 
                                              fill = Lewisetta), 
             shape = 21, size = 5) +
  geom_point(data = envdist_plotting[envdist_plotting$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x = longitude, y = latitude, 
                                       fill = Lewisetta), 
             shape = 21, size = 9)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 6)) +
  geom_point(data = envdist_plotting[envdist_plotting$site_name == "Lewisetta",], 
             aes(x = longitude-0.2, y = latitude), fill = "deeppink", shape = 24, size = 7)+
  geom_label_repel(data = envdist_plotting[envdist_plotting$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x = longitude, y = latitude, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.5,
                  fill = "white",
                  size = 9)+
  guides(fill = "none")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
lew_int_envdist_inset

map_yrk_int_envdist <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_yrk_envdist, mapping = aes(col = var1.pred, fill = var1.pred)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = envdist_sea_plotting, aes(x = Longitude_manual, y = Latitude_manual, 
                                              fill = YorkRiver), 
             shape = 21, size = 2) +
  geom_point(data = envdist_plotting[envdist_plotting$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x = longitude, y = latitude, fill = YorkRiver), 
             shape = 21, size = 4)+
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  geom_label_repel(data = envdist_plotting[envdist_plotting$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W5_NH", "W6_ME"),], aes(x = longitude, y = latitude, 
                                       label = site_name),
                  nudge_x = -3,
                  nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = "right")+
  labs(title = "Environmental Distance to \nYork River", x = "Longitude", y = "Latitude", col = "Environmental \nDistance", fill = "Environmental \nDistance")
(map_yrk_int_envdist)

yrk_int_envdist_inset <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = interpolated_yrk_envdist, mapping = aes(fill = var1.pred, col = var1.pred)) +
  scale_colour_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0,6)) +
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = envdist_sea_plotting, aes(x = Longitude_manual, y = Latitude_manual, 
                                              fill = YorkRiver), 
             shape = 21, size = 5) +
  geom_point(data = envdist_plotting[envdist_plotting$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, fill = YorkRiver), 
             shape = 21, size = 9)+
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 6)) + # double check these values
  geom_point(data = envdist_plotting[envdist_plotting$site_name == "YorkRiver",], 
             aes(x = longitude, y = latitude+0.2), fill = "deeppink", shape = 24, size = 7)+
  geom_label_repel(data = envdist_plotting[envdist_plotting$site_name %in% c("W4_VA", "S1_LOLA", "S2_DEBY"),], aes(x=longitude, y=latitude, label = site_name),
                  nudge_x = 2.5,
                  nudge_y = 0.7,
                  fill = "white",
                  size = 9)+
  guides(fill = "none")+
  theme(plot.title = element_text(size = 24), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right') +
        xlim(-77.856476, -72.762787) +
        ylim(36.019557, 39.939163)
yrk_int_envdist_inset
```


## Formatting for LEA

LEA requires environmental data be formatted as a .env file type.

```{r}
# prep as matrix
sea_mat <- as.matrix(sea_env_scale_mut[,8:13])

env_sea_ordered_mat <- env_sea_ordered[,c("salinity_quantile_10_scaled", "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "MSX_Prevalence_scaled")]

full_exp_mat <- as.matrix(exp_env_scale[,c("salinity_quantile_10_scaled",  "salinity_quantile_90_scaled", "temp_quantile_10_scaled", "temp_quantile_90_scaled", "Dermo_Prevalence_scaled", "Pea_crab_scaled", "MSX_Prevalence_scaled", "lew_salinity_quantile_10_scaled",  "lew_salinity_quantile_90_scaled", "lew_temp_quantile_10_scaled", "lew_temp_quantile_90_scaled", "lew_Dermo_Prevalence_scaled", "lew_Pea_crab_scaled", "lew_MSX_Prevalence_scaled", "yrk_salinity_quantile_10_scaled",  "yrk_salinity_quantile_90_scaled", "yrk_temp_quantile_10_scaled", "yrk_temp_quantile_90_scaled", "yrk_Dermo_Prevalence_scaled", "yrk_Pea_crab_scaled","yrk_MSX_Prevalence_scaled")])

# and convert to env format
write.env(sea_mat, paste0("data/EnvDat/seascape/seascape_scaled_", Sys.Date(), ".env"))
write.env(env_sea_ordered_mat, paste0("data/EnvDat/seascape/seascape_scaled_expanded_reord_", Sys.Date(), ".env"))
write.env(full_exp_mat, paste0("data/EnvDat/exp/exp_scaled_", Sys.Date(), ".env"))
```


## Formatting for RDA & GF

Here, the csv format should work fine so there's no need to go further.
