---
title: "exp_env_plotting"
output: html_document
---

# Plotting Environment of Origin & Common Garden

This script is used to plot temperature and salinity for the populations included in the common garden experiments, including DEBY and LOLA, and the two common gardens themselves.

The raw data for these sites was compiled by ME, NM, and CR. Data for the selection lines and common garden sites is processed in exp_envdat.Rmd. The other data are where the quantiles used for model prediction come from, but are not processed in this repo - the scripts for processing these data are in the seascape repo and the H2F repo.

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# setwd
setwd("~/Documents/GitHub/MVP_oyster_GO/")

# libraries
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readr)
library(forcats)
library(ragg)
library(gridExtra)
library(ggrastr)
library(ggpubr)
library(rvg)
library(officer)

# read in data
w1_raw <- read.csv("data/EnvDat/exp/exp_sites_raw/W1-TX-raw.csv")
w2_raw <- read.csv("data/EnvDat/exp/exp_sites_raw/W2-LA-raw.csv")
w3s <- read.csv("data/EnvDat/exp/exp_sites_raw/W3-FL-raw_sal.csv")
w3t <- read.csv("data/EnvDat/exp/exp_sites_raw/W3-FL-raw_temp.csv")
w4_raw <- read.csv("data/EnvDat/exp/exp_sites_raw/VA-corrected.csv")
w5_raw <- read.csv("data/EnvDat/exp/exp_sites_raw/W5-NH-raw.csv")
w6_raw <- read.csv("data/EnvDat/exp/exp_sites_raw/W6-ME-raw.csv")
lew_all <- read.csv("data/EnvDat/exp/lew_raw/lewisetta_corrected.csv")
yrk_all <- read.csv("data/EnvDat/exp/yrk_raw/YRK-raw.csv")
```


## Pre-process

Start with quick data clean & merge:

```{r}
## wild pops
############
# W1-TX
w1 <- w1_raw[,c("DateTimeStamp","Temp","Sal")]
w1$site_name <- "W1_TX"

# W2-LA
w2 <- w2_raw[,c("datetime","temp","salinity")]
w2$site_name <- "W2_LA"

# W3-FL
w3 <- merge(w3t, w3s, by = "Timestamp", all = T)
w3$site_name <- "W3_FL"

# W4-VA
w4 <- w4_raw[,c("datetime","temp_corrected","salinity_corrected")]
w4$site_name <- "W4_VA"

# W5
w5 <- w5_raw[,c("DateTimeStamp","Temp","Sal")]
w5$site_name <- "W5_NH"

# W6
w6 <- w6_raw[,c("Timestamp","temperature","salinity")]
w6$site_name <- "W6_ME"

# match colnames
colnames(w1) <- colnames(w2) <- colnames(w3) <- colnames(w4) <- colnames(w5) <- colnames(w6) <- c("timestamp","temp","sal","site_name")

# ensure variables coded correctly
w1$temp <- as.numeric(w1$temp)
w2$temp <- as.numeric(w2$temp)
w3$temp <- as.numeric(w3$temp)
w4$temp <- as.numeric(w4$temp)
w5$temp <- as.numeric(w5$temp)
w6$temp <- as.numeric(w6$temp)

w1$sal <- as.numeric(w1$sal)
w2$sal <- as.numeric(w2$sal)
w3$sal <- as.numeric(w3$sal)
w4$sal <- as.numeric(w4$sal)
w5$sal <- as.numeric(w5$sal)
w6$sal <- as.numeric(w6$sal)

# convert datetimes
wild <- list(w1, w2, w3, w4, w5, w6)
wild_timeclean <- map(wild, ~ .x %>%
  mutate(datetime_clean = parse_date_time(timestamp, orders = c("mdy HM", "ymd HMS", "mdy", "ymd"), tz = "UTC"))) # datetimes 

# failing somewhere, find out where
wild_timeclean[[2]] %>% filter(is.na(datetime_clean) & !is.na(timestamp)) # coming from the LA df, these should be removed
wild_timeclean[[2]] <- wild_timeclean[[2]] %>%
  filter(!is.na(timestamp))

combined_wild <- bind_rows(wild_timeclean)
############

## selection lines & common gardens
###################################
# lola
head(lew_all)
lew_clean <- lew_all[,c("datetime","temp_corrected","sal_corrected")]

# deby
head(yrk_all)
yrk_clean <- yrk_all[,c("datetime","temp","salinity")]

# colnames
colnames(lew_clean) <- colnames(yrk_clean) <- c("timestamp","temp","sal")

# correct coding of variables
lew_clean$temp <- as.numeric(lew_clean$temp)
yrk_clean$temp <- as.numeric(yrk_clean$temp)

lew_clean$sal <- as.numeric(lew_clean$sal)
yrk_clean$sal <- as.numeric(yrk_clean$sal)

# deal with datetimes individually so we can be clear about appropriate site labels
lew_timeclean <- lew_clean %>% mutate(datetime_clean = parse_date_time(timestamp, orders = c("mdy HM", "ymd HMS", "mdy", "ymd"), tz = "UTC"))
yrk_timeclean <- yrk_clean %>% mutate(datetime_clean = parse_date_time(timestamp, orders = c("mdy HM", "ymd HMS", "mdy", "ymd"), tz = "UTC"))

# site names
lew_timeclean$site_name <- ifelse(lew_timeclean$datetime_clean < "2023-05-01 00:00:00", "S1_LOLA", "Lewisetta")
yrk_timeclean$site_name <- ifelse(yrk_timeclean$datetime_clean < "2023-05-01 00:00:00", "S2_DEBY", "YorkRiver")

# put all the raw data together
lew_timeclean <- lew_timeclean[,c("timestamp","temp","sal","site_name","datetime_clean")]
yrk_timeclean <- yrk_timeclean[,c("timestamp","temp","sal","site_name","datetime_clean")]
combined_wild

all_env_raw <- rbind(combined_wild, lew_timeclean, yrk_timeclean)

# check values make sense, i.e. temp and sal btwn 0-40
all_env_raw_clean <- all_env_raw[all_env_raw$temp >= 0 & all_env_raw$temp <= 40 & all_env_raw$sal >= 0 & all_env_raw$sal <= 40,]
```

## Plotting

Now I want to make plots showing the salinity and temperature spread at each of these sites.

```{r}
popsdf_exp <- data.frame(pop = c("TX", "LA", "FL", "VA", "LOLA", "DEBY", "NH", "ME", "LEW", "YRK"), 
                     order = 1:10, 
                     cols = c("#365C8DFF", "#277F8EFF", "#1FA187FF", "#4AC16DFF", "#440154FF", "#46337EFF", "#9FDA3AFF", "#FDE725FF","hotpink","deeppink4"),
                     shape = c(21,21,21,21,23,23,21,21,24,24),
                     site_name = c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME", "Lewisetta", "YorkRiver"),
                     site_name_plotting = c("W1-TX", "W2-LA", "W3-FL", "W4-VA", "S1-LOLA", "S2-DEBY", "W5-NH", "W6-ME", "Lewisetta", "York River"),
                     site_type = c("exp_pop","exp_pop","exp_pop","exp_pop","exp_pop","exp_pop","exp_pop","exp_pop","exp_site","exp_site"))

# put it together
env_plotting <- merge(popsdf_exp, all_env_raw_clean, by = c("site_name"))

# plotting salinity
sal <- ggplot(data = env_plotting, aes(x = fct_reorder(site_name_plotting, order), y = sal)) +
  rasterize(geom_boxplot(position=position_dodge(1.0), fill = "lightblue2"), dpi = 300, dev = "ragg_png") +  
  labs(title = "Salinity Trends", x = "Site", y = "Salinity (ppt)") +
  theme_classic()+ 
  theme(strip.background = element_blank(), 
        strip.text.x = element_text(), 
        strip.placement = "outside",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.spacing.x = unit(3, "lines"))

# plotting temp
temp <- ggplot(data = env_plotting, aes(x = fct_reorder(site_name_plotting, order), y = temp)) +
  rasterize(geom_boxplot(position=position_dodge(1.0), fill = "#FFD3A5"), dpi = 300, dev = "ragg_png") +  
  labs(title = "Temperature Trends", x = "Site", y = "Temp (ÂşC)") +
  theme_classic()+ 
  theme(strip.background = element_blank(), 
        strip.text.x = element_text(), 
        strip.placement = "outside",
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.spacing.x = unit(3, "lines"))

# combine these into single ggobj
g <- arrangeGrob(sal, temp, ncol = 1)
saltemp_plot <- as_ggplot(g)
saltemp_plot
```

I also want to make some figures with the processed data & environmental distance from CviMVP_oyster_EnvDatProcess.Rmd

```{r}
# env dist
dists <- read.csv("data/EnvDat/exp/distance_matrix_longer_2025-12-05.csv")
env_dist_long_expsite <- dists[dists$site2 %in% c("Lewisetta", "YorkRiver"),]
env_dist_long_expsite_red <- env_dist_long_expsite[!(env_dist_long_expsite$site1 %in% c("Lewisetta", "YorkRiver")),]
env_dist_long_expsite_red$site_name_plotting <- gsub("_", "-", env_dist_long_expsite_red$site1)


# plot
envdist_plot_gardens <- ggplot(env_dist_long_expsite_red, aes(x = fct_reorder(site2, order2), 
                              y = fct_reorder(site_name_plotting, order1), 
                              fill = dist))+
  theme_classic()+
  geom_tile(color = "black") +
  scale_fill_viridis_c(name = "Euclidean \nDistance", limits = c(0, 6), 
                       option = "mako", direction = -1) +
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top") +
  labs(title = "Environmental Distance \nbetween Sites", x = "", y = "") +
  coord_fixed() +
  theme(plot.title = element_text(size = 18), 
        legend.title = element_text(size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right", legend.justification = "top")
envdist_plot_gardens

# all env
env <- read.csv("data/EnvDat/env_scaled_2025-12-05.csv")
env_trim <- env[env$site_name %in% c("W1_TX", "W2_LA", "W3_FL", "W4_VA", "S1_LOLA", "S2_DEBY", "W5_NH", "W6_ME", "Lewisetta", "YorkRiver"), 1:10]

# plotting
env_plotting <- merge(popsdf_exp, env_trim, by = "site_name")

env_plotting$text_col_dermo <- ifelse(round(env_plotting$Dermo_Prevalence, digits = 2) < 0.5, "white", "black")
env_plotting$text_col_msx <- ifelse(round(env_plotting$MSX_Prevalence, digits = 2) < 0.5, "white", "black")
env_plotting$text_col_pea <- ifelse(round(env_plotting$Pea_crab, digits = 2) < 0.5, "white", "black")

# try plotting heatmaps of disease
# dermo
dermo <- ggplot(env_plotting, aes(x = fct_reorder(site_name_plotting, order), y = 1, fill = Dermo_Prevalence)) + 
  scale_fill_viridis(option = "plasma", limits = c(0,1)) +
  geom_tile() + 
  coord_fixed() +
  geom_text(aes(label = round(Dermo_Prevalence, digits = 2)), col = env_plotting$text_col_dermo) +
  labs(title = "Dermo Prevalence", x = "Site", y = "Dermo \nPrevalence", fill = "Dermo \nPrevalence") +
  theme_classic() +
  theme(axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1))

# msx
msx <- ggplot(env_plotting, aes(x = fct_reorder(site_name_plotting, order), y = 1, fill = MSX_Prevalence)) + 
  scale_fill_viridis(option = "plasma", limits = c(0,1)) +
  geom_tile() + 
  coord_fixed() +
  geom_text(aes(label = round(MSX_Prevalence, digits = 2)), col = env_plotting$text_col_msx) +
  labs(title = "MSX Prevalence", x = "Site", y = "MSX \nPrevalence", fill = "MSX \nPrevalence") +
  theme_classic() +
  theme(axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1))

# pea crab
pea <- ggplot(env_plotting, aes(x = fct_reorder(site_name_plotting, order), y = 1, fill = Pea_crab)) + 
  scale_fill_viridis(option = "plasma", limits = c(0,1)) +
  geom_tile() + 
  coord_fixed() +
  geom_text(aes(label = round(Pea_crab, digits = 2)), col = env_plotting$text_col_pea) +
  labs(title = "Pea Crab Prevalence", x = "Site", y = "Pea Crab \nPrevalence", fill = "Pea Crab \nPrevalence") +
  theme_classic() +
  theme(axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(dermo, msx, pea, ncol = 1)

# long way with all disease in one
env_long_dis <- env_plotting[,c("site_name","site_name_plotting","order","cols","site_type","Dermo_Prevalence", "MSX_Prevalence", "Pea_crab")] %>%
  pivot_longer(
    cols = c(Dermo_Prevalence, MSX_Prevalence, Pea_crab),
    names_to = "Disease",
    values_to = "Prevalence"
  )

# text colors
env_long_dis$text_col <- ifelse(env_long_dis$Prevalence > 0.5, "white", "black")

# now plot
disease_plot <- ggplot(env_long_dis, aes(x = fct_reorder(site_name_plotting, order), 
                     y = Disease, 
                     fill = Prevalence)) +
  geom_tile(color = "white") + 
  scale_fill_viridis(option = "magma", limits = c(0, 1), direction = -1) +
  geom_text(aes(label = round(Prevalence, 2)), color = env_long_dis$text_col) +
  coord_fixed() +
  scale_y_discrete(labels = c("Pea_crab" = "Pea Crab", "MSX_Prevalence" = "MSX", "Dermo_Prevalence" = "Dermo")) +
  labs(
    title = "Disease Prevalence",
    x = "Site",
    y = NULL,
    fill = "Prevalence"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
    axis.line = element_blank(),
    axis.ticks = element_blank())
disease_plot
```

### Export

```{r}
doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = saltemp_plot), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = disease_plot), location = ph_location_type(type = "body")) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = envdist_plot_gardens), location = ph_location_type(type = "body"))
print(doc, target = "results/figures/exporg_env.pptx")
```


