---
title: "exp_envdat"
author: "Camille Rumberger, Nicole Mongill0, & Madeline Eppley"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
# set wd
setwd("~/Documents/Github/MVP_oyster_GO")

# packages
library(dplyr) # data cleaning
library(tidyr) # data cleaning
library(lubridate) # dealing with dates
library(readr) # read in data
library(ggplot2) # plotting
library(vroom) # data files

# data
# YORK
# downloaded 10/08/25 from VIMS http://vecos.vims.edu/StationDetail.aspx?param=YRK005.40&program=CMON
yrk_03_22 <- vroom(list.files(path = "data/EnvDat/exp/yrk_raw", pattern = "\\.csv$", full.names = TRUE)[1:20])

yrk_23 <- read.csv("data/EnvDat/exp/yrk_raw/York_2023.csv")
yrk_24 <- read.csv("data/EnvDat/exp/yrk_raw/York_2024.csv")
yrk_25 <- read.csv("data/EnvDat/exp/yrk_raw/York_2025.csv")

# lew data comes from Willy Reas from Coan River
raw_LOLA_env21 <- read.csv("data/EnvDat/exp/lew_raw/Jess_COAN_2021.csv")
raw_LOLA_env22 <- read.csv("data/EnvDat/exp/lew_raw/Jess_COAN_2022.csv")
raw_LOLA_env25 <- read.csv("data/EnvDat/exp/lew_raw/Jess_COAN_2025.csv")
raw_LOLA_env23_24 <- read.csv("data/EnvDat/exp/lew_raw/LEW-raw.csv")
```


## Pre-processing - York River

```{r}
# put together data
raw_yrk_exp <- rbind(yrk_23, yrk_24, yrk_25)

# subset relevant data
yrk_exp0 <- raw_yrk_exp[,c("SAMPLE_DATETIME", "WTEMP", "WTEMP_A", "SALINITY", "SALINITY_A")] 
yrk_exp0$SAMPLE_DATETIME <- as.POSIXct(yrk_exp0$SAMPLE_DATETIME, "%Y-%m-%d %H:%M:%S", tz = "")
yrk_sea0 <- yrk_03_22[,c("SAMPLE_DATETIME", "WTEMP", "WTEMP_A", "SALINITY", "SALINITY_A")]

parsed <- strptime(yrk_sea0$SAMPLE_DATETIME, "%m/%d/%Y %H:%M:%S")
yrk_sea0$SAMPLE_DATETIME <- as.POSIXct(parsed, "%Y/%m/%d %H:%M:%S", tz = "")

yrk_all <- rbind(yrk_sea0, yrk_exp0)

# keep just dates of experiment
yrk_all_exp <- subset(yrk_all,
                      SAMPLE_DATETIME >= as.POSIXct("2023-05-01 00:00:00.00",
                                                    tz = "") &
                      SAMPLE_DATETIME <= as.POSIXct("2025-05-28 00:00:00.00",
                                                    tz = ""))

# temp & sal subset, no error codes
yrk_noerr <- yrk_all[(yrk_all$WTEMP_A %in% c("", NA)), ]
yrk_all_sea <- yrk_noerr[(yrk_noerr$SALINITY_A %in% c("",NA)), ]

yrk_exp_noerr <- yrk_all_exp[(yrk_all_exp$WTEMP_A %in% c("", NA)), ]
yrk_exp <- yrk_exp_noerr[(yrk_exp_noerr$SALINITY_A %in% c("",NA)), ]

# trim + rename cols
yrk_exp <- yrk_exp[,c("SAMPLE_DATETIME", "WTEMP", "SALINITY")]
colnames(yrk_exp) <- c("datetime", "temp", "salinity")

yrk_allyrs <- yrk_all_sea[,c("SAMPLE_DATETIME", "WTEMP", "SALINITY")]
colnames(yrk_allyrs) <- c("datetime", "temp", "salinity")

# remove NAs
yrk_exp <- yrk_exp %>%
    filter(!is.na(datetime))

yrk_all_years <- yrk_allyrs %>%
  filter(!is.na(datetime))

# check
summary(yrk_exp)
summary(yrk_all_years)

# also store variables that we will include in the final data frame
site_name_y <- "YRK"
download_date_y <- "2025-10-08"
source_description_y <- "VECOS"
lat_y <- 37.247284
lon_y <- -76.499369
firstyear_y <- 2023
finalyear_y <- 2025

firstyear_yall <- 2003
finalyear_yall <- 2025
```


## Pre-processing: Lewisetta

This code comes in part from Nicole Mongillo.

```{r}
raw_LOLA_env23_24_prep <- raw_LOLA_env23_24[,c("Date","EST.Time","Temp","Salinity")]
colnames(raw_LOLA_env21) <- colnames(raw_LOLA_env22) <- colnames(raw_LOLA_env23_24) <- colnames(raw_LOLA_env25) <- c("Date","EST.Time","Temp","Salinity")
raw_LOLA_env <- rbind(raw_LOLA_env21, raw_LOLA_env22, raw_LOLA_env23_24_prep, raw_LOLA_env25)

#The data include separate columns for date and time, which I want to merge into one.
raw_LOLA_env$datetime <- paste(raw_LOLA_env$Date, raw_LOLA_env$EST.Time)

#Remove Date and EST.Time columns
raw_LOLA_env <- subset(raw_LOLA_env, select = -c(Date, EST.Time))

#Reorder columns so datetime is first
raw_LOLA_env <- raw_LOLA_env[, c(3,1,2)]

# View how the data are stored. Note the variable names and the format and units that the data are stored in.  
summary(raw_LOLA_env)

#rename columns. "datetime" = date and time of data collection, "temp" = water temperature in degrees C, "salinity" = water salinity in parts per thousand (ppt)
colnames(raw_LOLA_env) <- c("datetime", "temp", "salinity")

#Convert to POSIXct format. Tell R what the current date/time format is so it knows how to convert it. Store it into a column named datetime in the data frame.
raw_LOLA_env$datetime <- as.POSIXct(raw_LOLA_env$datetime, "%m/%d/%y %H:%M:%S", tz = "")

#add site name and create new data frame with full envr data set
raw_LOLA_env$site_name <- "LOLA"
LOLA_env_full <- raw_LOLA_env

#reorder columns with site_name first
LOLA_env_full <- LOLA_env_full[, c(4, 1, 2, 3)]

#separate out salinity and temp data into new frames for other analyses
LOLA_temp_full <- LOLA_env_full[, c(1,2,3)]

LOLA_sal_full <- LOLA_env_full[, c(1,2,4)]

#Store variables that we will include in the final data frame
site_name <- "LEW"
download_date <- as.Date("2025-10-14")
source_description <- "KCB hatchery in Coan River as part of CBNERR water quality monitoring, comes from Willy Reay"
lat <- 37.9803
lon <- -76.4619
#Lat and long data are from this site(https://www.vims.edu/cbnerr/monitoring/water_quality/). Data came from the KCB hatchery in the Coan as part of the CBNERR water quality monitoring initiative.
firstyear_l <- 2021
finalyear_l <- 2025
```

## Data filtering

Filter any of the variables that have data points outside of normal range. We will use 0-40 as the accepted range for salinity (ppt) and temperature (C) values. Note, in the summer, salinity values can sometimes exceed 40.

```{r}
# qc filtering yrk
filtered_yrk_exp_sal <- yrk_exp %>%
    filter(between(salinity, 0, 40)) 
filtered_yrk_exp_env <- filtered_yrk_sal %>%
    filter(between(temp, 0, 40))

filtered_yrk_sal <- yrk_all_years %>%
    filter(between(salinity, 0, 40)) 
filtered_yrk_env <- filtered_yrk_sal %>%
    filter(between(temp, 0, 40))

# store dat in new obj for ease
yrk <- filtered_yrk_env
yrk_exp <- filtered_yrk_exp_env

# filter lew
# keep just dates of experiment
LOLA_env_full$datetime <- as.POSIXct(LOLA_env_full$datetime, "%Y-%m-%d %H:%M:%S", tz = "")
LOLA_env_exp <- subset(LOLA_env_full,
                      datetime >= as.POSIXct("2023-05-01 00:00:00.00",
                                                    tz = "") &
                      datetime <= as.POSIXct("2025-05-28 00:00:00.00",
                                                    tz = ""))

# qc filtering lew
filtered_lew_exp_sal <- LOLA_env_exp %>%
    filter(between(salinity, 0, 40)) 
filtered_lew_exp_env <- filtered_lew_exp_sal %>%
    filter(between(temp, 0, 40))

filtered_lew_sal <- LOLA_env_full %>%
    filter(between(salinity, 0, 40)) 
filtered_lew_env <- filtered_lew_sal %>%
    filter(between(temp, 0, 40))

# store dat in new obj for ease
lew_exp <- filtered_lew_exp_env
lew <- filtered_lew_env
```

## Plotting Trends

Check trends in environment by visualizing. This code comes in part from N. Mongillo.

```{r}
salplot_yrk <- ggplot(yrk, aes(x = datetime)) +
    geom_line(aes(y = salinity, color = "Salinity (ppt)")) +
    ylim(0,40) +
    labs(x = "Time", y = "Salinity ppt", title = "Salinity Plot for York River") +
    scale_color_manual(values = c("Salinity (ppt)" = "lightblue3")) +
    theme_minimal()
salplot_yrk

tempplot_yrk <- ggplot(yrk, aes(x = datetime)) +
    geom_line(aes(y = temp, color = "Temperature (C)")) +
    ylim(0, 40) +
    labs(x = "Time", y = "Temperature C", title = "Temperature Plot for York River") +
    scale_color_manual(values = c( "Temperature (C)" = "pink3")) +
    theme_minimal()
tempplot_yrk

salplot_lew <- ggplot(lew, aes(x = datetime)) +
    geom_line(aes(y = salinity, color = "Salinity (ppt)")) +
    ylim(0,40) +
    labs(x = "Time", y = "Salinity ppt", title = "Salinity Plot for Lewisetta") +
    scale_color_manual(values = c("Salinity (ppt)" = "lightblue3")) +
    theme_minimal()
salplot_lew

tempplot_lew <- ggplot(lew, aes(x = datetime)) +
    geom_line(aes(y = temp, color = "Temperature (C)")) +
    ylim(0, 40) +
    labs(x = "Time", y = "Temperature C", title = "Temperature Plot for Lewisetta") +
    scale_color_manual(values = c( "Temperature (C)" = "pink3")) +
    theme_minimal()
tempplot_lew
```

## Summary statistics

Calculate means, maxes, mins, medians, etc.

```{r}
# this chunk comes modified from environmental processing script written by M. Eppley to retain consistency in variable naming schemes

# YORK RIVER
# calculate mean, maximum, and minimum values for salinity and temperature for each month. 
yrk_env_mo <- yrk %>%
    dplyr::mutate(year = year(datetime), month = month(datetime)) %>%
    dplyr::group_by(year, month, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      length_temp = length(temp))
  
print(yrk_env_mo)


#Calculate the mean, maximum, and minimum values for salinity and temperature for each year. 
yrk_env_yr <- yrk %>%
    dplyr::mutate(year = year(datetime)) %>%
    dplyr::group_by(year, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp))

print(yrk_env_yr)

#Calculate the mean, maximum, and minimum values for salinity and temperature for each day. 
yrk_env_day <- yrk %>%
    dplyr::mutate(year = year(datetime), month = month(datetime), day = day(datetime)) %>%
    dplyr::group_by(year, month, day, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      length_temp = length(temp))
  
print(yrk_env_day)

# check plots
timeplot <- ggplot(yrk_env_mo, aes(x = year)) +
    geom_point(aes(y = month, color = length_salinity), size = 4) +
    labs(x = "Time", y = "Month", title = "Salinity Timeplot for YRK - Gloucester Point, Virginia") +
    ylim(1,12) +
    theme_minimal()
timeplot

# LEWISETTA
# calculate mean, maximum, and minimum values for salinity and temperature for each month. 
lew_env_mo <- lew %>%
    dplyr::mutate(year = year(datetime), month = month(datetime)) %>%
    dplyr::group_by(year, month, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      length_temp = length(temp))
  
print(lew_env_mo)


#Calculate the mean, maximum, and minimum values for salinity and temperature for each year. 
lew_env_yr <- lew %>%
    dplyr::mutate(year = year(datetime)) %>%
    dplyr::group_by(year, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp))

print(lew_env_yr)

#Calculate the mean, maximum, and minimum values for salinity and temperature for each day. 
lew_env_day <- lew %>%
    dplyr::mutate(year = year(datetime), month = month(datetime), day = day(datetime)) %>%
    dplyr::group_by(year, month, day, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      length_temp = length(temp))
  
print(lew_env_day)

# check plots
timeplot <- ggplot(lew_env_mo, aes(x = year)) +
    geom_point(aes(y = month, color = length_salinity), size = 4) +
    labs(x = "Time", y = "Month", title = "Salinity Timeplot for LEW - Virginia") +
    ylim(1,12) +
    theme_minimal()
timeplot


# now for experimental time frame
# York River
yrk_envexp_yr <- yrk_exp %>%
    dplyr::mutate(year = year(datetime)) %>%
    dplyr::group_by(year, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp))

yrk_envexp_day <- yrk_exp %>%
    dplyr::mutate(year = year(datetime), month = month(datetime), day = day(datetime)) %>%
    dplyr::group_by(year, month, day, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      length_temp = length(temp))

yrk_envexp_mo <- yrk_exp %>%
    dplyr::mutate(year = year(datetime), month = month(datetime)) %>%
    dplyr::group_by(year, month, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      length_temp = length(temp))


# Lewisetta
lew_envexp_yr <- lew_exp %>%
    dplyr::mutate(year = year(datetime)) %>%
    dplyr::group_by(year, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp))

lew_envexp_day <- lew_exp %>%
    dplyr::mutate(year = year(datetime), month = month(datetime), day = day(datetime)) %>%
    dplyr::group_by(year, month, day, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      length_temp = length(temp))
  
lew_envexp_mo <- lew_exp %>%
    dplyr::mutate(year = year(datetime), month = month(datetime)) %>%
    dplyr::group_by(year, month, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      length_temp = length(temp))
```


# Calculate days above and below thresholds and plot

```{r low-sal-stress}
# blank df spans all 12 months for all years of data collection for this site
# merge with observations so that we can plot over time 
complete_year_month_l <- expand.grid(
  year = unique(firstyear_l:finalyear_l),
  month = 1:12
)
complete_year_month_y <- expand.grid(
  year = unique(firstyear_yall:finalyear_yall),
  month = 1:12
)

# start with low salinity stress 
# first calculate for low salinity
yrk$low_sal_stress <- yrk$salinity < 12
lew$low_sal_stress <- lew$salinity < 12

low_sal_stress_count_y <- yrk %>%
  dplyr::mutate(year = year(datetime), 
         month = month(datetime), 
         day = day(datetime)) %>% 
  dplyr::filter(low_sal_stress == 1) %>%  
  dplyr::distinct(year, month, day) %>%  # remove dups
  dplyr::group_by(year, month) %>%
  dplyr::summarise(low_sal_stress = n(), .groups = "drop") # group all occurrences together by month rather than by numeric day

low_sal_stress_count_l <- lew %>%
  dplyr::mutate(year = year(datetime), 
         month = month(datetime), 
         day = day(datetime)) %>% 
  dplyr::filter(low_sal_stress == 1) %>%  
  dplyr::distinct(year, month, day) %>%  # remove dups
  dplyr::group_by(year, month) %>%
  dplyr::summarise(low_sal_stress = n(), .groups = "drop") # group all occurrences together by month rather than by numeric day

print(low_sal_stress_count_y) # 95 stress days
print(low_sal_stress_count_l) # 735 stress days
```

# now calculate for high salinity stress

```{r high-sal-stress}
yrk$high_sal_stress <- yrk$salinity > 35

high_sal_stress_count <- yrk %>%
  dplyr::mutate(year = year(datetime), 
         month = month(datetime), 
         day = day(datetime)) %>% 
  dplyr::filter(high_sal_stress == 1) %>%  
  dplyr::distinct(year, month, day) %>%  # remove dups
  dplyr::group_by(year, month) %>%
  dplyr::summarise(high_sal_stress = n(), .groups = "drop") # group all occurrences together by month rather than by numeric day

print(high_sal_stress_count) # no high sal stress 

lew$high_sal_stress <- lew$salinity > 35

high_sal_stress_count <- lew %>%
  dplyr::mutate(year = year(datetime), 
         month = month(datetime), 
         day = day(datetime)) %>% 
  dplyr::filter(high_sal_stress == 1) %>%  
  dplyr::distinct(year, month, day) %>%  # remove dups
  dplyr::group_by(year, month) %>%
  dplyr::summarise(high_sal_stress = n(), .groups = "drop") # group all occurrences together by month rather than by numeric day

print(high_sal_stress_count) # no high sal stress 
```


# high temp stress calculations

```{r high-temp-stress}
yrk$high_temp_stress <- yrk$temp > 28

high_temp_stress_count <- yrk %>%
  dplyr::mutate(year = year(datetime), 
         month = month(datetime), 
         day = day(datetime)) %>% 
  dplyr::filter(high_temp_stress == 1) %>%  
  dplyr::distinct(year, month, day) %>%  # remove dups
  dplyr::group_by(year, month) %>%
  dplyr::summarise(high_temp_stress = n(), .groups = "drop") # group all occurrences together by month rather than by numeric day

print(high_temp_stress_count)

lew$high_temp_stress <- lew$temp > 28

high_temp_stress_count <- lew %>%
  dplyr::mutate(year = year(datetime), 
         month = month(datetime), 
         day = day(datetime)) %>% 
  dplyr::filter(high_temp_stress == 1) %>%  
  dplyr::distinct(year, month, day) %>%  # remove dups
  dplyr::group_by(year, month) %>%
  dplyr::summarise(high_temp_stress = n(), .groups = "drop") # group all occurrences together by month rather than by numeric day

print(high_temp_stress_count)

ggplot(high_temp_stress_count, aes(x = factor(month), y = high_temp_stress, fill = factor(month))) +
  geom_bar(stat = "identity", fill = "#DD4124FF") +
  facet_wrap(~ year) +
  labs(title = "YRK: count of high temperature (> 28 C) days per month",
       x = "month",
       y = "count of days") +
  theme_minimal()

high_temp_complete_data <- complete_year_month %>%
  left_join(high_temp_stress_count, by = c("year", "month")) %>%
  mutate(high_temp_stress = ifelse(is.na(high_temp_stress), 0, high_temp_stress))

high_temp_complete_data$month <- as.numeric(high_temp_complete_data$month)

# bars
ggplot(high_temp_complete_data, aes(x = factor(month), y = high_temp_stress)) +
  geom_bar(stat = "identity", fill = "#DD4124FF") + 
  facet_wrap(~ year) +
  labs(title = "YRK: count of high temperature (> 28 C) days per month",
       x = "month",
       y = "count of days") +
  theme_minimal()

# points
ggplot(high_temp_complete_data, aes(x = factor(month), y = high_temp_stress)) +
  geom_point(color = "#DD4124FF", size = 2) + 
  facet_wrap(~ year) +  
  labs(title = "YRK: count of high temperature (> 28 C) days per month",
       x = "month",
       y = "count of days") +
  theme_minimal()

```

## Putting it all together

THIS COMES FROM MADELINE'S SCRIPT

We can now calculate a list of variables that we will have collected for all sites. This will allow us to compare sites easily. We will calculate the number of observations from each site, the mean annual, maximum annual, and minimum annual value for all variables. 

Our list of variables includes: 

- Mean_Annual_Temperature_C: average of all available data
- Mean_max_temperature_C: average of maximums for each year
- Mean_min_temperature_C: average of minimums for each year
- Temperature_st_dev: standard deviation of all available data
- Temperature_n: total number of data points
- Temperature_years: number of years in data set

- Mean_Annual_Salinity_ppt: average of all available data
- Mean_min_Salinity_ppt: average of minimums for each year
- Mean_max_Salinity_ppt: average of maximums for each year
- Salinity_st_dev: standard deviation of all available data
- Salinity_n: total number of data points
- Salinity_years: number of years in data set

### York River

Start with the temperature variables

```{r}
#Calculate temperature variables. 
Mean_Annual_Temperature_C <- mean(yrk$temp)
Mean_max_temperature_C <- mean(yrk_env_yr$max_temp)
Mean_min_temperature_C <- mean(yrk_env_yr$min_temp)
temp_quantile_10 <- quantile(yrk$temp, 0.1)
temp_quantile_90 <- quantile(yrk$temp, 0.9)
Temperature_st_dev <- sd(yrk$temp)
Temperature_n <- nrow(yrk)
Temperature_years <- nrow(yrk_env_yr)
high_temp_stress_days <- sum(high_temp_stress_count$high_temp_stress)
frac_high_temp_stress_days <- high_temp_stress_days/nrow(yrk_env_day)

# Create a data frame to store the temperature results
yrk_temp <- cbind(data.frame(dataset = "sea", site_name = "S2_DEBY"), download_date, source_description, lat, lon, firstyear, finalyear, Mean_Annual_Temperature_C, Mean_max_temperature_C, Mean_min_temperature_C, temp_quantile_10, temp_quantile_90, Temperature_st_dev, high_temp_stress_days, frac_high_temp_stress_days, Temperature_n, Temperature_years)
print(yrk_temp)

# Write to a unique new CSV file
write.csv(yrk_temp, paste0("data/EnvDat/seascape/yrk_temperature_",Sys.Date(),".csv"))
```

```{r}
#Calculate temperature variables. 
Mean_Annual_Temperature_C <- mean(yrk_exp$temp)
Mean_max_temperature_C <- mean(yrk_envexp_yr$max_temp)
Mean_min_temperature_C <- mean(yrk_envexp_yr$min_temp)
temp_quantile_10 <- quantile(yrk_exp$temp, 0.1)
temp_quantile_90 <- quantile(yrk_exp$temp, 0.9)
Temperature_st_dev <- sd(yrk_exp$temp)
Temperature_n <- nrow(yrk_exp)
Temperature_years <- nrow(yrk_exp)
high_temp_stress_days <- sum(high_temp_stress_count$high_temp_stress)
frac_high_temp_stress_days <- high_temp_stress_days/nrow(yrk_env_day)

# Create a data frame to store the temperature results
yrk_exp_temp <- cbind(data.frame(dataset = "exp", site_name = "YRK"), download_date, source_description, lat, lon, firstyear, finalyear, Mean_Annual_Temperature_C, Mean_max_temperature_C, Mean_min_temperature_C, temp_quantile_10, temp_quantile_90, Temperature_st_dev, high_temp_stress_days, frac_high_temp_stress_days, Temperature_n, Temperature_years)
print(yrk_exp_temp)

# Write to a unique new CSV file
write.csv(yrk_exp_temp, paste0("data/EnvDat/exp/yrk_temperature_",Sys.Date(),".csv"))
```

Now the salinity variables

```{r}
#Calculate the salinity variables
Mean_Annual_Salinity_ppt <- mean(yrk$salinity)
Mean_max_Salinity_ppt <- mean(yrk_env_yr$max_salinity)
Mean_min_Salinity_ppt <- mean(yrk_env_yr$min_salinity)
salinity_quantile_10 <- quantile(yrk$salinity, 0.1)
salinity_quantile_90 <- quantile(yrk$salinity, 0.9)
Salinity_st_dev <- sd(yrk$salinity)
Salinity_n <- nrow(yrk)
Salinity_years <- nrow(yrk_env_yr)
high_sal_stress_days <- sum(high_sal_stress_count$high_sal_stress)
low_sal_stress_days <- sum(low_sal_stress_count$low_sal_stress)
frac_high_sal_stress_days <- high_sal_stress_days/nrow(yrk_env_day)
frac_low_sal_stress_days <- low_sal_stress_days/nrow(yrk_env_day)

#Create a data frame to store the temperature results
yrk_salinity <- cbind(data.frame(dataset = "sea", site_name = "S2_DEBY"), download_date, source_description, lat, lon, firstyear, finalyear, Mean_Annual_Salinity_ppt, Mean_max_Salinity_ppt, Mean_min_Salinity_ppt, salinity_quantile_10, salinity_quantile_90, high_sal_stress_days,low_sal_stress_days, frac_high_sal_stress_days, frac_low_sal_stress_days, Salinity_st_dev, Salinity_n, Salinity_years)
print(yrk_salinity)

# Write all year data to a unique new CSV file
write.csv(yrk_salinity, paste0("data/EnvDat/seascape/yrk_salinity_",Sys.Date(),".csv"))
```

```{r}
#Calculate the salinity variables
Mean_Annual_Salinity_ppt <- mean(yrk_exp$salinity)
Mean_max_Salinity_ppt <- mean(yrk_envexp_yr$max_salinity)
Mean_min_Salinity_ppt <- mean(yrk_envexp_yr$min_salinity)
salinity_quantile_10 <- quantile(yrk_exp$salinity, 0.1)
salinity_quantile_90 <- quantile(yrk_exp$salinity, 0.9)
Salinity_st_dev <- sd(yrk_exp$salinity)
Salinity_n <- nrow(yrk_exp)
Salinity_years <- nrow(yrk_envexp_yr)
high_sal_stress_days <- sum(high_sal_stress_count$high_sal_stress)
low_sal_stress_days <- sum(low_sal_stress_count$low_sal_stress)
frac_high_sal_stress_days <- high_sal_stress_days/nrow(yrk_envexp_day)
frac_low_sal_stress_days <- low_sal_stress_days/nrow(yrk_envexp_day)

#Create a data frame to store the temperature results
yrkexp_salinity <- cbind(data.frame(dataset = "exp", site_name = "YRK"), download_date, source_description, lat, lon, firstyear, finalyear, Mean_Annual_Salinity_ppt, Mean_max_Salinity_ppt, Mean_min_Salinity_ppt, salinity_quantile_10, salinity_quantile_90,  high_sal_stress_days,low_sal_stress_days, frac_high_sal_stress_days, frac_low_sal_stress_days, Salinity_st_dev, Salinity_n, Salinity_years)
print(yrkexp_salinity)

# Write all year data to a unique new CSV file
write.csv(yrkexp_salinity, paste0("data/EnvDat/exp/exp_yrk_salinity_",Sys.Date(),".csv"))
```

### Lewisetta

Start with the temperature variables

```{r}
#Calculate temperature variables. 
Mean_Annual_Temperature_C <- mean(lew$temp)
Mean_max_temperature_C <- mean(lew_env_yr$max_temp)
Mean_min_temperature_C <- mean(lew_env_yr$min_temp)
temp_quantile_10 <- quantile(lew$temp, 0.1)
temp_quantile_90 <- quantile(lew$temp, 0.9)
Temperature_st_dev <- sd(lew$temp)
Temperature_n <- nrow(lew)
Temperature_years <- nrow(lew_env_yr)
high_temp_stress_days <- sum(high_temp_stress_count$high_temp_stress)
frac_high_temp_stress_days <- high_temp_stress_days/nrow(lew_env_day)

# Create a data frame to store the temperature results
lew_temp <- cbind(data.frame(dataset = "sea", site_name = "S1_LOLA"), download_date, source_description, lat, lon, firstyear, finalyear, Mean_Annual_Temperature_C, Mean_max_temperature_C, Mean_min_temperature_C, temp_quantile_10, temp_quantile_90, Temperature_st_dev, high_temp_stress_days, frac_high_temp_stress_days, Temperature_n, Temperature_years)
print(lew_temp)

# Write to a unique new CSV file
write.csv(lew_temp, paste0("data/EnvDat/seascape/lew_temperature_",Sys.Date(),".csv"))
```

```{r}
#Calculate temperature variables. 
Mean_Annual_Temperature_C <- mean(lew_exp$temp)
Mean_max_temperature_C <- mean(lew_envexp_yr$max_temp)
Mean_min_temperature_C <- mean(lew_envexp_yr$min_temp)
temp_quantile_10 <- quantile(lew$temp, 0.1)
temp_quantile_90 <- quantile(lew$temp, 0.9)
Temperature_st_dev <- sd(lew_exp$temp)
Temperature_n <- nrow(lew_exp)
Temperature_years <- nrow(lew_envexp_yr)
high_temp_stress_days <- sum(high_temp_stress_count$high_temp_stress)
frac_high_temp_stress_days <- high_temp_stress_days/nrow(lew_envexp_day)

# Create a data frame to store the temperature results
lew_exp_temp <- cbind(data.frame(dataset = "exp", site_name = "LEW"), download_date, source_description, lat, lon, firstyear, finalyear, Mean_Annual_Temperature_C, Mean_max_temperature_C, Mean_min_temperature_C, temp_quantile_10, temp_quantile_90, Temperature_st_dev, high_temp_stress_days, frac_high_temp_stress_days, Temperature_n, Temperature_years)
print(lew_exp_temp)

# Write to a unique new CSV file
write.csv(lew_exp_temp, paste0("data/EnvDat/exp/lew_temperature_",Sys.Date(),".csv"))
```

Now the salinity variables

```{r}
#Calculate the salinity variables
Mean_Annual_Salinity_ppt <- mean(lew$salinity)
Mean_max_Salinity_ppt <- mean(lew_env_yr$max_salinity)
Mean_min_Salinity_ppt <- mean(lew_env_yr$min_salinity)
salinity_quantile_10 <- quantile(lew$salinity, 0.1)
salinity_quantile_90 <- quantile(lew$salinity, 0.9)
Salinity_st_dev <- sd(lew$salinity)
Salinity_n <- nrow(lew)
Salinity_years <- nrow(lew_env_yr)
high_sal_stress_days <- sum(high_sal_stress_count$high_sal_stress)
low_sal_stress_days <- sum(low_sal_stress_count$low_sal_stress)
frac_high_sal_stress_days <- high_sal_stress_days/nrow(lew_env_day)
frac_low_sal_stress_days <- low_sal_stress_days/nrow(lew_env_day)

#Create a data frame to store the temperature results
lew_salinity <- cbind(data.frame(dataset = "sea", site_name = "S1_LOLA"), download_date, source_description, lat, lon, firstyear, finalyear, Mean_Annual_Salinity_ppt, Mean_max_Salinity_ppt, Mean_min_Salinity_ppt, salinity_quantile_10, salinity_quantile_90, high_sal_stress_days,low_sal_stress_days, frac_high_sal_stress_days, frac_low_sal_stress_days, Salinity_st_dev, Salinity_n, Salinity_years)
print(lew_salinity)

# Write all year data to a unique new CSV file
write.csv(lew_salinity, paste0("data/EnvDat/seascape/lew_salinity_",Sys.Date(),".csv"))
```

```{r}
#Calculate the salinity variables
Mean_Annual_Salinity_ppt <- mean(lew_exp$salinity)
Mean_max_Salinity_ppt <- mean(lew_envexp_yr$max_salinity)
Mean_min_Salinity_ppt <- mean(lew_envexp_yr$min_salinity)
salinity_quantile_10 <- quantile(lew_exp$salinity, 0.1)
salinity_quantile_90 <- quantile(lew_exp$salinity, 0.9)
Salinity_st_dev <- sd(lew_exp$salinity)
Salinity_n <- nrow(lew_exp)
Salinity_years <- nrow(lew_envexp_yr)
high_sal_stress_days <- sum(high_sal_stress_count$high_sal_stress)
low_sal_stress_days <- sum(low_sal_stress_count$low_sal_stress)
frac_high_sal_stress_days <- high_sal_stress_days/nrow(lew_envexp_day)
frac_low_sal_stress_days <- low_sal_stress_days/nrow(lew_envexp_day)

#Create a data frame to store the temperature results
lewexp_salinity <- cbind(data.frame(dataset = "exp", site_name = "LEW"), download_date, source_description, lat, lon, firstyear, finalyear, Mean_Annual_Salinity_ppt, Mean_max_Salinity_ppt, Mean_min_Salinity_ppt, salinity_quantile_10, salinity_quantile_90, high_sal_stress_days,low_sal_stress_days, frac_high_sal_stress_days, frac_low_sal_stress_days, Salinity_st_dev, Salinity_n, Salinity_years)
print(lewexp_salinity)

# Write all year data to a unique new CSV file
write.csv(lewexp_salinity, paste0("data/EnvDat/exp/exp_lew_salinity_",Sys.Date(),".csv"))
```

Put together the full environmental dataset with all quantiles. 

```{r}
# other experimental pops
exppop_sal <- read.csv("data/EnvDat/exp/all_salinity.csv")
exppop_temp <- read.csv("data/EnvDat/exp/all_temperature.csv")

# make datasets compatible
exppop_sal$dataset <- exppop_temp$dataset <- "exp"
exppop_sal$collection_type <- exppop_temp$collection_type <- NULL

# remove exp sites from these datasheets
exppop_sal_mg <- exppop_sal %>% dplyr::filter(!(site_name %in% c("LEW","YRK", "S1-LOLA", "S2-DEBY")))
exppop_temp_mg <- exppop_temp %>% dplyr::filter(!(site_name %in% c("LEW","YRK", "S1-LOLA", "S2-DEBY")))

# final experimental dataset
all_salinity <- rbind(exppop_sal_mg, lew_salinity, lewexp_salinity, yrk_salinity, yrkexp_salinity)
all_temp <- rbind(exppop_temp_mg, lew_temp, lew_exp_temp, yrk_temp, yrk_exp_temp)

write.csv(all_salinity, "data/EnvDat/exp/exp_salinity.csv")
write.csv(all_temp, "data/EnvDat/exp/exp_temp.csv")
```
