---
title: "exp_envdat"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
# set wd
setwd("~/Documents/Github/MVP_oyster_GO")

# packages
library(dplyr) # data cleaning
library(tidyr) # data cleaning
library(lubridate) # dealing with dates
library(readr) # read in data
library(ggplot2) # plotting

# data
# YORK
# downloaded 10/08/25 from VIMS http://vecos.vims.edu/StationDetail.aspx?param=YRK005.40&program=CMON
yrk_23 <- read.csv("data/EnvDat/exp/yrk_raw/York_2023.csv")
yrk_24 <- read.csv("data/EnvDat/exp/yrk_raw/York_2024.csv")
yrk_25 <- read.csv("data/EnvDat/exp/yrk_raw/York_2025.csv")

# lew data comes from Willy Reas from Coan River -- still need data from Nov'24-May'25
setwd("data/EnvDat/exp/lew_raw")
#merge files into one
raw_lew <- list.files(path=".") %>% 
  lapply(read.csv) %>% 
  bind_rows

# set wd
setwd("~/Documents/Github/MVP_oyster_GO")
```


## Pre-processing - York River

```{r}
# put together data
raw_yrk <- rbind(yrk_23, yrk_24, yrk_25)

# subset relevant data
yrk_all <- raw_yrk[,c("SAMPLE_DATETIME", "WTEMP", "WTEMP_A", "SALINITY", "SALINITY_A")] 

# keep just dates of experiment
yrk_all$SAMPLE_DATETIME <- as.POSIXct(yrk_all$SAMPLE_DATETIME, "%Y-%m-%d %H:%M:%S", tz = "")
yrk_all_exp <- subset(yrk_all,
                      SAMPLE_DATETIME >= as.POSIXct("2023-05-01 00:00:00.00",
                                                    tz = "") &
                      SAMPLE_DATETIME <= as.POSIXct("2025-05-28 00:00:00.00",
                                                    tz = ""))

# temp subset, no error codes
yrk_noerr <- yrk_all_exp[(yrk_all_exp$WTEMP_A %in% c("", NA)), ]

# sal subset, no error codes
yrk_exp <- yrk_noerr[(yrk_noerr$SALINITY_A %in% c("", NA)), ]

# trim + rename cols
yrk_exp <- yrk_exp[,c("SAMPLE_DATETIME", "WTEMP", "SALINITY")]
colnames(yrk_exp) <- c("datetime", "temp", "salinity")

# remove NAs
yrk_exp <- yrk_exp %>%
    filter(!is.na(datetime))

# check
summary(yrk_exp)

# also store variables that we will include in the final data frame
site_name <- "YRK"
download_date <- as.Date(2025-10-08)
source_description <- "VECOS"
lat <- 37.247284
lon <- -76.499369
firstyear <- 2023
finalyear <- 2025
```

## Data filtering

Filter any of the variables that have data points outside of normal range. We will use 0-40 as the accepted range for salinity (ppt) and temperature (C) values. Note, in the summer, salinity values can sometimes exceed 40.

```{r}
# filter 
filtered_yrk_sal <- yrk_exp %>%
    filter(between(salinity, 0, 40)) 
filtered_yrk_env <- filtered_yrk_sal %>%
    filter(between(temp, 0, 40))

# store dat in new obj for ease
yrk <- filtered_yrk_env
```

## Plotting Trends

Check trends in environment by visualizing.

```{r}
salplot <- ggplot(yrk, aes(x = datetime)) +
    geom_line(aes(y = salinity, color = "Salinity (ppt)")) +
    ylim(0,40) +
    labs(x = "Time", y = "Salinity ppt", title = "Salinity Plot for YRK - Gloucester Point, Virginia") +
    scale_color_manual(values = c("Salinity (ppt)" = "lightblue3")) +
    theme_minimal()
salplot

tempplot <- ggplot(yrk, aes(x = datetime)) +
    geom_line(aes(y = temp, color = "Temperature (C)")) +
    ylim(0, 40) +
    labs(x = "Time", y = "Temperature C", title = "Temperature Plot for YRK - Gloucester Point, Virginia") +
    scale_color_manual(values = c( "Temperature (C)" = "pink3")) +
    theme_minimal()
tempplot
```

## Summary statistics

Calculate means, maxes, mins, medians, etc.

```{r}
# this chunk comes modified from environmental processing script written by M. Eppley to retain consistency in variable naming schemes

# calculate mean, maximum, and minimum values for salinity and temperature for each month. 
yrk_env_mo <- yrk %>%
    dplyr::mutate(year = year(datetime), month = month(datetime)) %>%
    dplyr::group_by(year, month, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      length_temp = length(temp))
  
print(yrk_env_mo)


#Calculate the mean, maximum, and minimum values for salinity and temperature for each year. 
yrk_env_yr <- yrk %>%
    dplyr::mutate(year = year(datetime)) %>%
    dplyr::group_by(year, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp))

print(yrk_env_yr)

#Calculate the mean, maximum, and minimum values for salinity and temperature for each day. 
yrk_env_day <- yrk %>%
    dplyr::mutate(year = year(datetime), month = month(datetime), day = day(datetime)) %>%
    dplyr::group_by(year, month, day, .drop = FALSE) %>%
    dplyr::summarise(
      min_salinity = min(salinity),
      max_salinity = max(salinity),
      mean_salinity = mean(salinity),
      length_salinity = length(salinity),
      min_temp = min(temp),
      max_temp = max(temp),
      mean_temp = mean(temp),
      length_temp = length(temp))
  
print(yrk_env_day)

# check plots
timeplot <- ggplot(yrk_env_mo, aes(x = year)) +
    geom_point(aes(y = month, color = length_salinity), size = 4) +
    labs(x = "Time", y = "Month", title = "Salinity Timeplot for YRK - Gloucester Point, Virginia") +
    ylim(1,12) +
    theme_minimal()
timeplot
```

# Calculate days above and below thresholds and plot

```{r low-sal-stress}
# blank df spans all 12 months for all years of data collection for this site
# merge with observations so that we can plot over time 
complete_year_month <- expand.grid(
  year = unique(firstyear:finalyear),
  month = 1:12
)

# start with low salinity stress 
# first calculate for low salinity
yrk$low_sal_stress <- yrk$salinity < 12

low_sal_stress_count <- yrk %>%
  dplyr::mutate(year = year(datetime), 
         month = month(datetime), 
         day = day(datetime)) %>% 
  dplyr::filter(low_sal_stress == 1) %>%  
  dplyr::distinct(year, month, day) %>%  # remove dups
  dplyr::group_by(year, month) %>%
  dplyr::summarise(low_sal_stress = n(), .groups = "drop") # group all occurrences together by month rather than by numeric day

print(low_sal_stress_count) # no low stress days
```

# now calculate for high salinity stress

```{r high-sal-stress}
yrk$high_sal_stress <- yrk$salinity > 35

high_sal_stress_count <- yrk %>%
  dplyr::mutate(year = year(datetime), 
         month = month(datetime), 
         day = day(datetime)) %>% 
  dplyr::filter(high_sal_stress == 1) %>%  
  dplyr::distinct(year, month, day) %>%  # remove dups
  dplyr::group_by(year, month) %>%
  dplyr::summarise(high_sal_stress = n(), .groups = "drop") # group all occurrences together by month rather than by numeric day

print(high_sal_stress_count) # no high sal stress 
```


# high temp stress calculations

```{r high-temp-stress}
yrk$high_temp_stress <- yrk$temp > 28

high_temp_stress_count <- yrk %>%
  dplyr::mutate(year = year(datetime), 
         month = month(datetime), 
         day = day(datetime)) %>% 
  dplyr::filter(high_temp_stress == 1) %>%  
  dplyr::distinct(year, month, day) %>%  # remove dups
  dplyr::group_by(year, month) %>%
  dplyr::summarise(high_temp_stress = n(), .groups = "drop") # group all occurrences together by month rather than by numeric day

print(high_temp_stress_count)

ggplot(high_temp_stress_count, aes(x = factor(month), y = high_temp_stress, fill = factor(month))) +
  geom_bar(stat = "identity", fill = "#DD4124FF") +
  facet_wrap(~ year) +
  labs(title = "YRK: count of high temperature (> 28 C) days per month",
       x = "month",
       y = "count of days") +
  theme_minimal()

high_temp_complete_data <- complete_year_month %>%
  left_join(high_temp_stress_count, by = c("year", "month")) %>%
  mutate(high_temp_stress = ifelse(is.na(high_temp_stress), 0, high_temp_stress))

high_temp_complete_data$month <- as.numeric(high_temp_complete_data$month)

# bars
ggplot(high_temp_complete_data, aes(x = factor(month), y = high_temp_stress)) +
  geom_bar(stat = "identity", fill = "#DD4124FF") + 
  facet_wrap(~ year) +
  labs(title = "YRK: count of high temperature (> 28 C) days per month",
       x = "month",
       y = "count of days") +
  theme_minimal()

# points
ggplot(high_temp_complete_data, aes(x = factor(month), y = high_temp_stress)) +
  geom_point(color = "#DD4124FF", size = 2) + 
  facet_wrap(~ year) +  
  labs(title = "YRK: count of high temperature (> 28 C) days per month",
       x = "month",
       y = "count of days") +
  theme_minimal()

```

## Putting it all together

THIS COMES FROM MADELINE'S SCRIPT

We can now calculate a list of variables that we will have collected for all sites. This will allow us to compare sites easily. We will calculate the number of observations from each site, the mean annual, maximum annual, and minimum annual value for all variables. 

Our list of variables includes: 

- Mean_Annual_Temperature_C: average of all available data
- Mean_max_temperature_C: average of maximums for each year
- Mean_min_temperature_C: average of minimums for each year
- Temperature_st_dev: standard deviation of all available data
- Temperature_n: total number of data points
- Temperature_years: number of years in data set

- Mean_Annual_Salinity_ppt: average of all available data
- Mean_min_Salinity_ppt: average of minimums for each year
- Mean_max_Salinity_ppt: average of maximums for each year
- Salinity_st_dev: standard deviation of all available data
- Salinity_n: total number of data points
- Salinity_years: number of years in data set

Start with the temperature variables

```{r}
#Calculate temperature variables. 
Mean_Annual_Temperature_C <- mean(yrk$temp)
Mean_max_temperature_C <- mean(yrk_env_yr$max_temp)
Mean_min_temperature_C <- mean(yrk_env_yr$min_temp)
Temperature_st_dev <- sd(yrk$temp)
Temperature_n <- nrow(yrk)
Temperature_years <- nrow(yrk_env_yr)
high_temp_stress_days <- sum(high_temp_stress_count$high_temp_stress)
frac_high_temp_stress_days <- high_temp_stress_days/nrow(yrk_env_day)

Mean_Monthly_Temperature_C <- yrk_env_mo %>%
  filter(!is.na(month)) %>% 
  group_by(month) %>%
  summarise(Mean_Temperature = mean(mean_temp))

Mean_min_Monthly_Temperature_C <- yrk_env_mo %>%
  filter(!is.na(month)) %>% 
  group_by(month) %>%
  summarise(Mean_min_Temperature = mean(min_temp))

Mean_max_Monthly_Temperature_C <- yrk_env_mo %>%
  filter(!is.na(month)) %>% 
  group_by(month) %>%
  summarise(Mean_max_Temperature = mean(max_temp))

# Create a data frame to store the temperature results
yrk_temp <- cbind(site_name, download_date, source_description, lat, lon, firstyear, finalyear, Mean_Annual_Temperature_C, Mean_max_temperature_C, Mean_min_temperature_C, Temperature_st_dev, high_temp_stress_days, frac_high_temp_stress_days, Temperature_n, Temperature_years)
print(yrk_temp)

yrk_monthly_temp <- cbind(Mean_Monthly_Temperature_C, Mean_min_Monthly_Temperature_C, Mean_max_Monthly_Temperature_C)
yrk_monthly_temp <- yrk_monthly_temp[, !duplicated(names(yrk_monthly_temp))]
print(yrk_monthly_temp)

# Write to the combined file with all sites 
write.table(yrk_temp, paste0("data/EnvDat/exp/all_temperature",Sys.Date(),".csv"), sep = ",", append = TRUE, col.names = FALSE, row.names = FALSE) # The column names should be changed to FALSE after 1st row is added to the data frame

# Write to a unique new CSV file
write.csv(yrk_temp, paste0("data/EnvDat/exp/yrk_temperature_",Sys.Date(),".csv"))

# Write all montly data to a unique new CSV file
write.csv(yrk_monthly_temp, paste0("data/EnvDat/exp/yrk_temperature_monthly_",Sys.Date(),".csv"))
```

Now the salinity variables

```{r}
#Calculate the salinity variables
Mean_Annual_Salinity_ppt <- mean(yrk$salinity)
Mean_max_Salinity_ppt <- mean(yrk_env_yr$max_salinity)
Mean_min_Salinity_ppt <- mean(yrk_env_yr$min_salinity)
Salinity_st_dev <- sd(yrk$salinity)
Salinity_n <- nrow(yrk)
Salinity_years <- nrow(yrk_env_yr)
high_sal_stress_days <- sum(high_sal_stress_count$high_sal_stress)
low_sal_stress_days <- sum(low_sal_stress_count$low_sal_stress)
frac_high_sal_stress_days <- high_sal_stress_days/nrow(yrk_env_day)
frac_low_sal_stress_days <- low_sal_stress_days/nrow(yrk_env_day)

Mean_Monthly_Salinity <- yrk_env_mo %>%
  filter(!is.na(month)) %>%
  group_by(month) %>%
  summarise(Mean_Salinity = mean(mean_salinity))

Min_Monthly_Salinity <- yrk_env_mo %>%
  filter(!is.na(month)) %>%
  group_by(month) %>%
  summarise(Min_Salinity = mean(min_salinity))

Max_Monthly_Salinity <- yrk_env_mo %>%
  filter(!is.na(month)) %>%
  group_by(month) %>%
  summarise(Max_Salinity = mean(max_salinity))

#Create a data frame to store the temperature results
yrk_salinity <- cbind(site_name, download_date, source_description, lat, lon, firstyear, finalyear, Mean_Annual_Salinity_ppt, Mean_max_Salinity_ppt, Mean_min_Salinity_ppt, high_sal_stress_days,low_sal_stress_days, frac_high_sal_stress_days, frac_low_sal_stress_days, Salinity_st_dev, Salinity_n, Salinity_years)
print(yrk_salinity)

yrk_monthly_sal <- cbind(Mean_Monthly_Salinity, Min_Monthly_Salinity, Max_Monthly_Salinity)
yrk_monthly_sal <- yrk_monthly_sal[, !duplicated(names(yrk_monthly_sal))]
print(yrk_monthly_sal)

# Write to the combined file with all sites 
write.table(yrk_salinity, paste0("data/EnvDat/exp/all_salinity_",Sys.Date(),".csv"), sep = ",", append = TRUE, col.names = FALSE, row.names = FALSE) # The column names should be changed to FALSE after 1st row is added to the data frame

# Write all year data to a unique new CSV file
write.csv(yrk_salinity, paste0("data/EnvDat/exp/yrk_salinity_",Sys.Date(),".csv"))
# Write all monthly data to a unique new CSV file
write.csv(yrk_monthly_sal, paste0("data/EnvDat/exp/yrk_monthly_sal_",Sys.Date(),".csv"))
```
