---
title: "CviMVP_oyster_compareGOexp"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "~/Documents/GitHub/MVP_oyster_GO")
```

# Compare Predicted GO to Experimental Results

This script is for analyzing how well various genomic offset methods predict experimentally obtained fitness. I compare LEA, GF, and RDA offsets to experimental mortality, length, and condition index at the conclusion of the experiment. 

## Setup

```{r}
# setwd
setwd("~/Documents/GitHub/MVP_oyster_GO")

# packages
library(dplyr)
library(tidyr)
library(stringr)
library(vegan)
library(ggplot2)
library(plotrix)

# data
# offset predictions
lea_go <- read.csv("results/GO_results/offsets_LEA_expSites_04222025.csv")[,-1]
#rda_go <- 
#gf_go <- 

# fitness metrics
mort <- read.csv("data/ExpDat/mortality_bag_overall_12092024.csv")[,-1]
#len <- read.csv("data/ExpDat/")
```

## Data Prep

Prepare data from offsets & experiment to be directly compared.

```{r}
# inspect data
lea_go
mort

# we want to get avg mort per pop per site
mort_popsite <- mort %>% group_by(site,pop) %>%
  summarise(mean_t1 = mean(as.numeric(survival_1)), 
            sd_t1 = sd(as.numeric(survival_1)),
            se_t1 = std.error(as.numeric(survival_1)),
            mean_t2 = mean(as.numeric(survival_2)), 
            sd_t2 = sd(as.numeric(survival_2)),
            se_t2 = std.error(as.numeric(survival_2)),
            mean_t3 = mean(as.numeric(survival_3)), 
            sd_t3 = sd(as.numeric(survival_3)),
            se_t3 = std.error(as.numeric(survival_3)),
            .groups = 'drop') %>% as.data.frame()

# # split go & fitness by site
# # go
# leaLew <- lea_go[,c("site","offset_lew","order_lew")]
# leaYrk <- lea_go[,c("site","offset_lew","order_yrk")]
# # fitness
# mortlew <- mort_popsite[mort_popsite$site == "Lewisetta",]
# mortyrk <- mort_popsite[mort_popsite$site == "York River",]

# need to align naming conventions to merge
colnames(lea_go) <- c("pop", "offset_lew", "order_lew", "offset_yrk", "order_yrk")
lea_go$pop <- str_split_fixed(lea_go$pop, "-",2)[,1]
mort_popsite[mort_popsite$pop == "JR",]$pop <- "VA"

# merge gos & fitness dat
goFit <- merge(mort_popsite, lea_go, by = "pop")
```


## Compare Offsets to Fitness

Now that offsets and fitness are in a comparable format, look at how well the two match.

```{r}
# correlations between go/lew mort at each time point
# lea
corLEAt1lew <- cor(goFit[goFit$site == "Lewisetta",]$offset_lew, goFit[goFit$site == "Lewisetta",]$mean_t1)
corLEAt2lew <- cor(goFit[goFit$site == "Lewisetta",]$offset_lew, goFit[goFit$site == "Lewisetta",]$mean_t2)
corLEAt3lew <- cor(goFit[goFit$site == "Lewisetta",]$offset_lew, goFit[goFit$site == "Lewisetta",]$mean_t3)
# rda
# gf

# correlations between go/york mort at each time point
# lea
corLEAt1yrk <- cor(goFit[goFit$site == "York River",]$offset_yrk, goFit[goFit$site == "York River",]$mean_t1)
corLEAt2yrk <- cor(goFit[goFit$site == "York River",]$offset_yrk, goFit[goFit$site == "York River",]$mean_t2)
corLEAt3yrk <- cor(goFit[goFit$site == "York River",]$offset_yrk, goFit[goFit$site == "York River",]$mean_t3)
# rda
# gf

# these corrs seems crazy, they're very high but positive when we expect them to be negative if the method performs well - try kendall's tau by getting rankings

# put this all into a df with method/time point/fitness metric/site/corr

```


Plot these data sets against each other

```{r}
# 

# plot lea vs mort at each site/time point
gofit_lew_plot <- ggplot(data = goFit[goFit$site == "Lewisetta",], aes(x = mean_t1, y = offset_lew)) +
  geom_point(col = "cornflowerblue") +
  geom_point(data = goFit[goFit$site == "Lewisetta",], aes(x = mean_t2, y = offset_lew), col = "slateblue") +
  geom_point(data = goFit[goFit$site == "Lewisetta",], aes(x = mean_t3, y = offset_lew), col = "purple4") +
  theme_classic()
gofit_lew_plot

gofit_yrk_plot <- ggplot(data = goFit[goFit$site == "York River",], aes(x = mean_t1, y = offset_yrk)) +
  geom_point(col = "cornflowerblue") +
  geom_point(data = goFit[goFit$site == "York River",], aes(x = mean_t2, y = offset_yrk), col = "slateblue") +
  geom_point(data = goFit[goFit$site == "York River",], aes(x = mean_t3, y = offset_yrk), col = "purple4") +
  theme_classic()
gofit_yrk_plot
```

## Compare Offset Methods

Once I have predictions from different methods, I will look at how offset predictions made by these methods compare to each other and how different their ability is to predict experimental fitness.

```{r}

```

